### 2014-06-11 paper.tex
Initial commit

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach any log files that you could have,
% and specify the details of your LaTeX setup (which operating system and LaTeX
% version / tools you are using).

%=================================================================

% LaTeX Class File and Rendering Mode (choose one)
% You will need to save the "mdpi.cls" and "mdpi.bst" files into the same folder
% as this template file.

%=================================================================

\documentclass[energies,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, administrativesciences, aerospace, agriculture, agronomy,
% algorithms, animals, antibiotics, antibodies, antioxidants, appliedsciences,
% arts, atmosphere, atoms, axioms, behavioralsciences, bioengineering, biology,
% biomedicines, biomolecules, biosensors, brainsciences, buildings, cancers,
% catalysts, cells, challenges, chemosensors, children, chromatography, climate,
% coatings, computation, computers, cosmetics, crystals, dentistryjournal,
% diagnostics, diseases, diversity, econometrics, economies, education,
% electronics, energies, entropy, environmentalsciences, environments, fibers,
% foods, forests, futureinternet, galaxies, games, genes, geosciences,
% healthcare, humanities, hydrology, informatics, information, inorganics,
% insects, ijerph, ijfs, ijms, ijgi, jcdd, jcm, jdb, jfb, joi, jlpea, jmse,
% jpcg, jpm, jrfm, jsan, land, laws, life, lubricants, machines, marinedrugs,
% materials, mathematics, medicalsciences, membranes, metabolites, metals,
% microarrays, micromachines, microorganisms, minerals, molbank, molecules,
% nanomaterials, ncrna, nutrients, pathogens, pharmaceuticals, pharmaceutics,
% pharmacy, photonics, plants, polymers, processes, proteomes, publications,
% religions, remotesensing, resources, risks, robotics, sensors, socialsciences,
% societies, sports, sustainability, symmetry, systems, technologies, toxics,
% toxins, vaccines, veterinarysciences, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but could be replaced by using one
% of the class options:
% article, review, communication, commentary, bookreview, correction, addendum,
% editorial, changes, supfile, casereport, comment, conceptpaper,
% conferencereport, meetingreport, discussion, essay, letter, newbookreceived,
% opinion, projectreport, reply, retraction, shortnote, technicalnote, creative
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office
% when the paper is accepted. This will only make changes to the frontpage (e.g.
% the logo of the journal will get visible), the headings, and the copyright
% information. Journal info and pagination for accepted papers will also be
% assigned by the Editorial Office.
% Please insert a blank line is before and after all equation and eqnarray
% environments to ensure proper line numbering when option submit is chosen
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used.
% Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option "pdftex" is for use with pdfLaTeX only. If eps figure are used, use
% the optioin "dvipdfm", with LaTeX and dvi2pdf only.

%=================================================================
\setcounter{page}{1}
\lastpage{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2014}
\history{Received: xx / Accepted: xx / Published: xx}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================

% Add packages and commands to include here
% The amsmath, amsthm, amssymb, hyperref, caption, float and color packages are
% loaded by the MDPI class.
\usepackage{graphicx}
%\usepackage{subfigure,psfig}
\usepackage[draft]{todonotes}

%=================================================================
%% Please use the following mathematics environments:
%\theoremstyle{mdpi}
%\newcounter{thm}
%\setcounter{thm}{0}
%\newcounter{ex}
%\setcounter{ex}{0}
%\newcounter{re}
%\setcounter{re}{0}
%\newtheorem{Theorem}[thm]{Theorem}
%\newtheorem{Lemma}[thm]{Lemma}
%\newtheorem{Characterization}[thm]{Characterization}
%\newtheorem{Proposition}[thm]{Proposition}
%\newtheorem{Property}[thm]{Property}
%\newtheorem{Problem}[thm]{Problem}
%\newtheorem{Example}[ex]{Example}
%\newtheorem{Remark}[re]{Remark}
%\newtheorem{Corollary}[thm]{Corollary}
%\newtheorem{Definition}[thm]{Definition}
%% For proofs, please use the proof environment (the amsthm package is loaded by
% the MDPI class).

%=================================================================

% Full title of the paper (Capitalized)
\Title{Effects of Reynolds number on the performance and near-wake of a
cross-flow turbine}

% Authors (Add full first names)
\Author{Peter Bachant $^{1,}$* and Martin Wosnik $^{1}$}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ Center for Ocean Renewable Energy, 
University of New Hampshire, 24 Colovos Rd., Durham, NH, USA}

% Contact information of the corresponding author (Add [2] after \corres if
% there are more than one corresponding author.)
\corres{pxL3@unh.edu}

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{
This is the abstract section. The abstract should be one section and count less than 200 words.
}

% Keywords: add 3 to 10 keywords
\keyword{keyword; keyword; keyword}

% The fields PACS, MSC, and JEL may be left empty or commented out if not
% applicable
%\PACS{}
%\MSC{}
%\JEL{}

\begin{document}

\listoftodos

\section{Introduction}

\todo[inline]{Come up with a better title? Needs to be different from the METS
one, but not so blatantly the same.}

Scaled physical models are often used in science and engineering to approximate
real-world systems. In the fields of wind and marine hydrokinetic (MHK) energy,
scaled physical models are used to validate numerical models, test prototypes of
individual devices, and design or investigate arrays of devices. Despite being
significantly less expensive, there is still some concern at which scale an
experiment must be performed to be relevant to full-scale application.

For numerical models, one might obtain favorable predictions for scaled systems
due to modified physics. If the numerical model is then considered
``validated,'' there is a risk that its application at full-scale may produce
incorrect predictions. For scaled physical prototypes, it is of interest how
small---since size, or scale generally correlates inversely with cost---the
prototype can be while remaining a reliable predictor of full-scale behavior.

When designing or studying arrays, it is very common to use very small
(geometrically) scaled devices. Sometimes these are not even models of turbines,
but wake generating objects, e.g., porous disks, that are meant to replicate the
wakes of real turbines. In this case, it is of interest to determine at what
scale one might be able to realistically study wake flows in an array, and also
to evaluate the effectiveness of a wake generator, i.e., a wake generator may do
a fine job simulating a scaled turbine, but how well can it simulate a real
turbine?

Previously, Bachant and Wosnik \cite{Bachant2014_JOT} showed that a cross-flow
turbine wake's unique mean flow field is responsible for accelerated wake
recovery when compared with conventional axial-flow propeller-type turbines.

\todo[inline]{Gotta get the JOT paper published first before we can say that!}

In this study we set out to measure the effects of scale, both Reynolds and
Froude scales, on the performance of the turbine, and its near-wake
characteristics.

\section{Experimental Section}

Experiments were performed in the University of New Hampshire's tow/wave tank
turbine test bed. The turbine model is dubbed the UNH-RVAT, for Reference
Vertical Axis Turbine, which is designed to be a generic case for numerical
model testing, similar to the Sandia National Labs/US Department of Energy RM2
River Turbine. Details of the turbine are described in \cite{Bachant2014_JOT}.

\todo[inline]{Put up CAD model for citation/download on figshare.}

\subsection{Test plan}

Approximately 1500 tows were performed, each one used for
a single data point on either a performance curve or wake map.

A performance curve

\subsection{Data processing}

From each set of tows, a standard time interval was set, and each run was
analyzed to compute statistics over this interval, truncating the end slightly
to achieve and integer number of blade passages.

\todo[inline]{Put up reduced data and processing code on figshare.}

\section{Results and Discussion}

\subsection{Performance curves}

\todo[inline]{Put the performance curves here.}

\subsection{Wake characteristics}

\todo[inline]{Put wake characteristics here. Maybe create difference plots to
show absolute differences, or RMS, or something}

\section{Conclusions}

In this study we determined that, as previously shown, turbine performance
becomes essentially $Re$-independent at a turbine diameter Reynolds number $Re_D
\approx 10^6$, or an approximate blade chord Reynolds number $Re_c \approx 2
\times 10^5$.

\acknowledgements{Acknowledgements}

The authors would like to acknowledge funding through a National Science
Foundation CAREER award (PI Wosnik, NSF-CBET 1150797, program manager Dr. Ram
Gupta), a grant through the Leslie S. Hubbard Marine Program Endowment to
purchase acoustic flow measurement instrumentation, and a grant for laboratory
infrastructure upgrades through the US Department of Energy.

\authorcontributions{Author Contributions}

\todo[inline]{Is this section necessary?}

\conflictofinterests{Conflicts of Interest}

The authors declare no conflict of interest.

\bibliography{../../../Library/Library}
\bibliographystyle{mdpi}

\end{document}

### 2014-06-16 paper.tex
Make titles regular title case
Add some info to test plan, including table of Reynolds numbers

@@
 % Full title of the paper (Capitalized)
-\Title{Effects of Reynolds number on the performance and near-wake of a
-cross-flow turbine}
+\Title{Effects of Reynolds Number on the Performance and Near-Wake of a
+Cross-Flow Turbine}
 
@@
 
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+% Affiliations / Addresses (Add [1] after \address if there is only one
+% affiliation.)
 \address{%
-$^{1}$ Center for Ocean Renewable Energy, 
-University of New Hampshire, 24 Colovos Rd., Durham, NH, USA}
+$^{1}$ Center for Ocean Renewable Energy, University of New Hampshire, 24
+Colovos Rd., Durham, NH, USA}
 
@@
 % Abstract (Do not use inserted blank lines, i.e. \\)
-\abstract{
-This is the abstract section. The abstract should be one section and count less than 200 words.
-}
+\abstract{This is the abstract section. The abstract should be one section and
+count less than 200 words. }
 
@@
 
-\section{Experimental Section}
+\section{Experimental Setup}
 
@@
 model testing, similar to the Sandia National Labs/US Department of Energy RM2
-River Turbine. Details of the turbine are described in \cite{Bachant2014_JOT}.
+River Turbine. Details of the turbine and experimental setup are described in
+\cite{Bachant2014_JOT}.
 
@@
 
-\subsection{Test plan}
+\subsection{Test Plan}
 
-Approximately 1500 tows were performed, each one used for
-a single data point on either a performance curve or wake map.
+Approximately 1500 tows were performed, each one used for a single data point on
+either a performance curve or wake map. A performance curve consisted of 31
+tows, where during each tow the mean turbine tip speed ratio was held constant,
+ranging from 0.1--3.1 in 0.1 increments. Full performance curve data were
+acquired for tow speeds from 0.4 to 1.2 m/s in 0.2 m/s increments, for which
+turbine diameter and approximate blade chord Reynolds number are presented in
+Table~\ref{tab:Re}. Performance was also measured for $\lambda=1.9$ at tow
+speeds [0.3, 0.5, 0.7, 0.9, 1.1, 1.3] m/s, each of which consisted of two tows.
 
-A performance curve
+\begin{table}
+\centering
+\begin{tabular}{ccc}
+Tow speed (m/s) & $Re_D$ & $Re_c$ ($\lambda = 1.9$) \\ 
+\hline
+0.2 & $2.0 \times 10^5$ &  \\ 
+0.4 & $4.0 \times 10^5$ &  \\ 
+0.6 & $6.0 \times 10^5$ &  \\ 
+0.8 & $8.0 \times 10^5$ &  \\ 
+1.0 & $1.0 \times 10^6$ &  \\ 
+1.2 & $1.2 \times 10^6$ &  \\ 
+\end{tabular} 
+\caption{Turbine diameter and approximate blade chord Reynolds numbers for the
+tow speeds at which full performance curves and wake maps were measured.}
+\label{tab:Re}
+\end{table}
 
-\subsection{Data processing}
+A wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler Velocimeter (ADV) at
+
+\subsection{Data Processing}
 
@@
 
-\subsection{Performance curves}
+\subsection{Performance Curves}
 
@@
 
-\subsection{Wake characteristics}
+\subsection{Wake Characteristics}
 

### 2014-06-16 paper.tex
Add description of wake measurement locations

@@
 
-A wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler Velocimeter (ADV) at
+A wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
+Velocimeter (ADV) at 270 different locations, varied in the cross-stream and
+vertical directions at one turbine diameter downstream. These locations have
+vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
+the cross-stream direction $y/R = \pm 3$. These locations are show in
+Figure~\ref{fig:wake_locations}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/wake_meas_grid}
+\caption{Wake measurement coordinate system and locations.}
+\label{fig:wake_locations}
+\end{figure}
 
@@
 
-\authorcontributions{Author Contributions}
-
-\todo[inline]{Is this section necessary?}
-
 \conflictofinterests{Conflicts of Interest}

### 2014-06-17 paper.tex
Add wake characteristics subsubsections and defs for derivative mathrms

@@
 \usepackage[draft]{todonotes}
+
+\def \p{\partial}
+\def \d{\mathrm{d}}
+\def \D{\mathrm{D}}
 
@@
 
+\todo[inline]{Put asymptotic behavior here.}
+
 \subsection{Wake Characteristics}
@@
 show absolute differences, or RMS, or something}
+
+\todo[inline]{Plot asymptotic behavior for a bunch of locations all on the same
+plot.}
+
+\subsubsection{Mean Kinetic Energy Transport}
+
+The relative importance of various physical processes on mean kinetic energy
+transport/recovery in the streamwise direction $\p K / \p x$ are plotted in...
+
+\todo[inline]{Recreate the kinetic energy bar graph for all $Re$.}
 

### 2014-06-22 paper.tex
Change wake measurement grid figure to newer drawing
Add intro comments about Reynolds number dependence and CAD model reference

@@
 
+\subsection{Previous Work}
+\todo[inline]{Add lit review here}
+
+\subsection{Causes for Reynolds Number Dependence}
+
+Turbine performance is strongly dependent on the lift-to-drag ratio $l/d$ acting
+on the blades, which is a function of the blades' angles of attack $\alpha$.
+These angles of attack are a function of turbine tip speed ratio, induction
+(slowing/turning of the free stream flow prior to reaching the turbine), and
+turbine rotation or azimuthal angle $\theta$. Looking at static airfoil data, we
+see that the static stall angle---the angle just beyond that at which $l/d$ is
+maximum, characterized by flow separation near the leading edge---increases with
+blade chord Reynolds number $Re_c$. This can be explained as the transition to a
+turbulent boundary layer adjacent to the blade delaying separation. However,
+static foil performance does not tell the whole story for a cross-flow turbine.
+The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
+turbine implies the occurrence of dynamic stall at tip speed ratios at and below
+those of maximum rotor torque\cite{Para2002}. We then must understand the
+effects of Reynolds number on the dynamic stall process.
+
+As previously mentioned, the near-wake of a CFT is not very similar to a simple
+drag source, or plane shear flow. One of the mean wake's streamwise vortex pairs
+can be predicted by potential flow theory as a result of finite-span blades
+producing lift (blade end effects), but the other---caused by viscous effects,
+i.e., dynamic stall---is not. We can attempt to predict changes
+\todo[inline]{Finish this thought with Reynolds number dependence of these
+simpler flows, and set up the question whether or not we can assume these will
+combine linearly to produce the wake $Re$-dependence we see.}
+
 \section{Experimental Setup}
@@
 River Turbine. Details of the turbine and experimental setup are described in
-\cite{Bachant2014_JOT}.
-
-\todo[inline]{Put up CAD model for citation/download on figshare.}
+\cite{Bachant2014_JOT}, and a CAD model of the turbine is available from
+\cite{Bachant2014_CAD}.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/wake_meas_grid}
+\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
 \caption{Wake measurement coordinate system and locations.}

### 2014-09-17 paper.tex
Add comment about coordinate system dimensions being in meters
Add todo item for turbine model with no hubs
Add a bunch of stuff

@@
 
+Previously we took a general look at how maximum performance and some 2D wake
+profiles are affected by Reynolds number, but here we week to look much deeper.
+
 In this study we set out to measure the effects of scale, both Reynolds and
@@
 
+\todo[inline]{Make sure to use drawing of turbine with no hubs, since these were
+included in the tare drag"}
+
 \subsection{Test Plan}
@@
 \includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
-\caption{Wake measurement coordinate system and locations.}
+\caption{Wake measurement coordinate system and locations. Dimensions are in
+meters.} 
 \label{fig:wake_locations}
@@
 plot.}
+
+\subsubsection{Mean Momentum Transport}
+
+\todo[inline]{Make bar graph for all $Re$.}
+
+\subsubsection{Dominant Timescales}
+
+\todo[inline]{Look at how spectra change when turbine rotational frequency
+changes relative to inertial subrange.}
 

### 2015-03-23 paper.tex
Copy bib file locally
Fix citation keys

@@
 %------------------------------------------------------------------
-% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+% The following line should be uncommented if the LaTeX file is uploaded to
+%arXiv.org
 %\pdfoutput=1
@@
 
-Previously, Bachant and Wosnik \cite{Bachant2014_JOT} showed that a cross-flow
+Previously, Bachant and Wosnik \cite{Bachant2015-JoT} showed that a cross-flow
 turbine wake's unique mean flow field is responsible for accelerated wake
@@
 River Turbine. Details of the turbine and experimental setup are described in
-\cite{Bachant2014_JOT}, and a CAD model of the turbine is available from
-\cite{Bachant2014_CAD}.
+\cite{Bachant2015-JoT}, and a CAD model of the turbine is available from
+\cite{Bachant2014-RVAT-CAD}.
 
@@
 
-\bibliography{../../../Library/Library}
+\bibliography{library}
 \bibliographystyle{mdpi}

### 2015-03-23 paper.tex
Add 'vertical-axis' to title and start abstract

@@
 \pubvolume{xx}
-\pubyear{2014}
+\pubyear{2015}
 \history{Received: xx / Accepted: xx / Published: xx}
@@
 \Title{Effects of Reynolds Number on the Performance and Near-Wake of a
-Cross-Flow Turbine}
+Vertical-Axis Cross-Flow Turbine}
 
@@
 
-% Abstract (Do not use inserted blank lines, i.e. \\)
-\abstract{This is the abstract section. The abstract should be one section and
-count less than 200 words. }
+% Abstract (Less than 200 words)
+\abstract{The acceptable Reynolds number mismatch for scaled physical model
+testing a cross-flow turbine is investigated experimentally.}
 
@@
 \section{Introduction}
-
-\todo[inline]{Come up with a better title? Needs to be different from the METS
-one, but not so blatantly the same.}
 
@@
 recovery when compared with conventional axial-flow propeller-type turbines.
-
-\todo[inline]{Gotta get the JOT paper published first before we can say that!}
 

### 2015-03-25 paper.tex
Rearrange introduction and add to abstract

@@
 \abstract{The acceptable Reynolds number mismatch for scaled physical model
-testing a cross-flow turbine is investigated experimentally.}
+testing a cross-flow turbine is investigated experimentally, both with respect
+to prediction of full-scale performance, and wake characteristics, which will be
+drivers of overall performance and interaction between multiple turbines.}
 
@@
 
-\subsection{Previous Work}
+\subsection{Reynolds Number Effects}
 \todo[inline]{Add lit review here}
 
-\subsection{Causes for Reynolds Number Dependence}
+In this study we are focused on a turbine constructed from foils---a lift-driven
+device in contrast to drag-driven. A logical starting point for understanding
+the effects of $Re$ on overall turbine loading is to look at simplified cases
+of individual foils, with respect to their static and dynamic loading.
+
+\subsubsection{On Static Foil Behavior}
 
@@
 combine linearly to produce the wake $Re$-dependence we see.}
+
+\subsubsection{On Dynamic Foil Behavior}
+
+Dynamic effects on blade loading can be distinguished between attached and
+separated conditions.
+
+\subsubsection{On Wake Flows}
+
+\todo[inline]{What do studies in the literature tell us about $Re$ effects
+on shear flows, wakes, and vortex flows?}
 

### 2015-03-28 paper.tex
Comment out figure for now
Add citation for Re-dependence of dynamic stall
Update Re table

@@
 
+Bousman showed that dynamic stall was insensitive to Reynolds number for $Re=1.0
+\times 10^5$--$2.5 \times 10^5$ \cite{Bousman2000-evaluation}.
+
 \subsubsection{On Wake Flows}
@@
 \begin{tabular}{ccc}
-Tow speed (m/s) & $Re_D$ & $Re_c$ ($\lambda = 1.9$) \\ 
+Tow speed (m/s) & $Re_D$ & $Re_{c,\mathrm{ave}}$ ($\lambda = 1.9$) \\ 
 \hline
-0.2 & $2.0 \times 10^5$ &  \\ 
-0.4 & $4.0 \times 10^5$ &  \\ 
-0.6 & $6.0 \times 10^5$ &  \\ 
-0.8 & $8.0 \times 10^5$ &  \\ 
-1.0 & $1.0 \times 10^6$ &  \\ 
-1.2 & $1.2 \times 10^6$ &  \\ 
+0.3 & $3.0 \times 10^5$ & $8.0 \times 10^4$ \\ 
+0.4 & $4.0 \times 10^5$ & $1.1 \times 10^5$ \\ 
+0.5 & $5.0 \times 10^5$ & $1.3 \times 10^5$ \\ 
+0.6 & $6.0 \times 10^5$ & $1.6 \times 10^5$ \\ 
+0.7 & $7.0 \times 10^5$ & $1.9 \times 10^5$ \\ 
+0.8 & $0.8 \times 10^5$ & $2.1 \times 10^5$ \\ 
+0.9 & $0.9 \times 10^5$ & $2.4 \times 10^5$ \\ 
+1.0 & $1.0 \times 10^6$ & $2.7 \times 10^5$ \\ 
+1.1 & $1.1 \times 10^6$ & $2.9 \times 10^5$ \\ 
+1.2 & $1.2 \times 10^6$ & $3.2 \times 10^5$ \\ 
+1.3 & $1.3 \times 10^6$ & $3.4 \times 10^5$ \\ 
 \end{tabular} 
 \caption{Turbine diameter and approximate blade chord Reynolds numbers for the
-tow speeds at which full performance curves and wake maps were measured.}
+tow speeds used in the experiment.}
 \label{tab:Re}
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
+%\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
 \caption{Wake measurement coordinate system and locations. Dimensions are in

### 2015-03-31 paper.tex
Add comments about Re dependence of wake generation processes
Add performance description

@@
 
+We expect that lower lift on the blades at lower $Re$ will weaken tip vortex
+shedding, and decrease the levels of turbulence shed from the blade boundary
+layers. As mentioned previously, the dynamic stall vortex shedding is not
+expected to have a large Reynolds number sensitivity.
+
 \todo[inline]{What do studies in the literature tell us about $Re$ effects
@@
 \todo[inline]{Make sure to use drawing of turbine with no hubs, since these were
-included in the tare drag"}
+included in the tare drag}
 
@@
 vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
-the cross-stream direction $y/R = \pm 3$. These locations are show in
-Figure~\ref{fig:wake_locations}.
+the cross-stream direction $y/R = \pm 3$. These locations are shown in
+Figure~\ref{fig:wake-locations}.
 
@@
 meters.} 
-\label{fig:wake_locations}
+\label{fig:wake-locations}
 \end{figure}
@@
 
-\todo[inline]{Put up reduced data and processing code on figshare.}
+\todo[inline]{Cite reduced data and processing code on figshare.}
 
@@
 
-\subsection{Performance Curves}
+\subsection{Performance}
 
-\todo[inline]{Put the performance curves here.}
+Full performance curves for various Reynolds numbers are plotted in
+Figure~\ref{fig:perf-curves}.
 
-\todo[inline]{Put asymptotic behavior here.}
+\todo[inline]{Discuss drag coefficient curves.}
+
+In general, maximum $C_P$ increases with Reynolds number, which is due to an
+increase in the foil lift-to-drag ratio. The power coefficient curves also show
+a downward shift in the optimal tip speed ratio with increasing Reynolds number.
+This is caused by the stall delay from a more turbulent boundary later on the
+blade suction side.
+
+\begin{figure}[ht]
+\caption{Mean power (left) and drag (right) coefficient curves plotted for
+multiple Reynolds numbers.}
+\label{fig:perf-curves}
+\end{figure}
+
+Mean power and drag coefficients at $\lambda=1.9$ are plotted in
+Figure~\ref{fig:perf-Re-dep}
+
+\begin{figure}[ht]
+\caption{Mean power (left) and drag (right) coefficients at $\lambda=1.9$
+plotted versus Reynolds number.} 
+\label{fig:perf-Re-dep}
+\end{figure}
 

### 2015-03-31 paper.tex
Add todo item for XFOIL
Add todo item for Froude number discussion
Add point about CFD model validation and turbulent transition

@@
 
+\todo[inline]{Talk about interdependence of Froude number.}
+
 \subsection{Reynolds Number Effects}
@@
 
+\todo[inline]{Add foil characteristics computed by XFOIL here.}
+
 \subsection{Wake Characteristics}
@@
 
+These results suggest that to validate predictive engineering models, one must
+use data of at least this scale, especially high-fidelity CFD models, where
+transition to turbulence may be completely different between scaled physical
+model and prototype.
+
 \acknowledgements{Acknowledgements}

### 2015-04-01 paper.tex
Add to section on static foil coefficient model

@@
 
+\subsubsection{Understanding via Static Foil Coefficients}
+
+To help understand---and possibly predict---the $Re$-sensitivity on turbine
+performance, a series of foil coefficient datasets were computed with the XFOIL
+viscous panel method code.
+
+The turbine torque coefficient $C_T$ can be related to the blade section
+chordwise force coefficient $C_c$ by
+\begin{equation}
+C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
+\end{equation}
+where the blade section chordwise force (for zero preset pitch)
+\begin{equation}
+C_c = C_l \sin \alpha - C_d \cos \alpha.
+\end{equation}
+
 \todo[inline]{Add foil characteristics computed by XFOIL here.}
+
+Note that the common static foil performance metrics fail to show Reynolds
+number dependence as close as the max geometric torque coefficient. This means
+that the turbine kinematics play an important role, even if the exact dynamics
+of the flow are not known.
 

### 2015-04-07 getfigs.py
Get figs from QBlade data

@@ -9,11 +9,22 @@ import sys
 import shutil
 
 figdir = os.path.join(os.path.expanduser("~"), "Research",
-                      "Experiments", "2014", "RVAT Re dep", "Figures")
+                      "Experiments", "RVAT Re dep", "Figures")
 
 figlist = ["cp_vs_tsr.pdf",
            "mean_u.pdf",
            "mean_upvp.pdf"]
+
+for fig in figlist:
+    shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))
+
+    
+figdir = os.path.join(os.path.expanduser("~"), "Google Drive", "Research",
+                      "Foils", "Data", "NACAXX20_QBlade", "figures")
+
+figlist = ["all_foils_re_dep.pdf",
+           "cft_re_dep_foils.pdf",
+           "foil_kinematics_ct.pdf"]
 
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))

### 2015-04-07 getfigs.py
Change figs taken from RVAT Re dep

@@ -11,9 +11,10 @@ import shutil
 figdir = os.path.join(os.path.expanduser("~"), "Research",
                       "Experiments", "RVAT Re dep", "Figures")
 
-figlist = ["cp_vs_tsr.pdf",
-           "mean_u.pdf",
-           "mean_upvp.pdf"]
+figlist = ["perf_curves.pdf",
+           "perf_re_dep.pdf",
+           "K_trans_bar_graph.pdf",
+           "mom_bar_graph.pdf"]
 
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))

### 2015-04-08 paper.tex
Add point to conclusion regarding scaled physical model array predictions
Add figures

@@
 \centering
-%\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
+\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
 \caption{Wake measurement coordinate system and locations. Dimensions are in
@@
 \begin{figure}[ht]
+\includegraphics[width=0.95\textwidth]{figures/perf_curves}
 \caption{Mean power (left) and drag (right) coefficient curves plotted for
@@
 \begin{figure}[ht]
+\includegraphics[width=0.95\textwidth]{figures/perf_re_dep}
 \caption{Mean power (left) and drag (right) coefficients at $\lambda=1.9$
@@
 
-\todo[inline]{Add foil characteristics computed by XFOIL here.}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/all_foils_re_dep}
+\caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
+\label{fig:foil-Re-dep}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/foil_kinematics_ct}
+\caption{Geometric angle of attack (a), relative velocity (b) and torque coefficient
+(c) calculated with a NACA 0020 foil.}
+\label{fig:blade-kinematics}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.65\textwidth]{figures/cft_re_dep_foils}
+\caption{Reynolds number depenence of the peak torque coefficient calculated via
+static foil coefficients and blade kinematics.}
+\label{fig:foils-C_T-Re-dep}
+\end{figure}
 
@@
 
-\todo[inline]{Make bar graph for all $Re$.}
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/mom_bar_graph}
+\caption{Momentum transport quantities.}
+\label{fig:mom-bar-graph}
+\end{figure}
 
@@
 
-\todo[inline]{Recreate the kinetic energy bar graph for all $Re$.}
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/K_trans_bar_graph}
+\caption{Mean kinetic energy transport quantities.}
+\label{fig:K-bar-graph}
+\end{figure}
 
@@
 model and prototype.
+
+If using scaled physical models to predict array performance, it is important to
+keep all turbines in the linear regime to avoid exaggerated power deficiencies
+for downstream turbines, despite similarities in wake characteristics
 

### 2015-04-08 paper.tex
Add Chamorro et al reference for AFT wake

@@
 \subsection{Reynolds Number Effects}
-\todo[inline]{Add lit review here}
 
@@
 of individual foils, with respect to their static and dynamic loading.
+
+Chamorro et al. showed that turbulence statistics in an AFT wake became
+$Re$-independent at $Re_D \approx 1 \times 10^5$, with mean velocity slightly
+earlier at $Re_D \approx 5 \times 10^4$ \cite{Chamorro2012}.
 
@@
 \includegraphics[width=0.65\textwidth]{figures/cft_re_dep_foils}
-\caption{Reynolds number depenence of the peak torque coefficient calculated via
-static foil coefficients and blade kinematics.}
+\caption{Reynolds number dependence of the peak torque coefficient calculated using 
+only static foil coefficients and blade kinematics.}
 \label{fig:foils-C_T-Re-dep}

### 2015-04-15 getfigs.py
Add wake spectra figure

@@ -14,7 +14,8 @@ figdir = os.path.join(os.path.expanduser("~"), "Research",
 figlist = ["perf_curves.pdf",
            "perf_re_dep.pdf",
            "K_trans_bar_graph.pdf",
-           "mom_bar_graph.pdf"]
+           "mom_bar_graph.pdf",
+	   "wake_spectra.pdf"]
 
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))

### 2015-04-16 getfigs.py
Add wake transport totals plotted versus Re

@@ -15,7 +15,8 @@ figlist = ["perf_curves.pdf",
            "perf_re_dep.pdf",
            "K_trans_bar_graph.pdf",
            "mom_bar_graph.pdf",
-	   "wake_spectra.pdf"]
+	   "wake_spectra.pdf",
+	   "wake_trans_totals.pdf"]
 
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))

### 2015-04-16 paper.tex
Remove long comment block
Add analysis comment
Add wake spectra figure
Adjust float
Add wake transport totals plotted versus Re
Add placeholder for figure from McMasters and Henderson

@@
 \documentclass[energies,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
-%--------------------
-% Class Options:
-%--------------------
-% journal
-%----------
-% Choose between the following MDPI journals:
-% actuators, administrativesciences, aerospace, agriculture, agronomy,
-% algorithms, animals, antibiotics, antibodies, antioxidants, appliedsciences,
-% arts, atmosphere, atoms, axioms, behavioralsciences, bioengineering, biology,
-% biomedicines, biomolecules, biosensors, brainsciences, buildings, cancers,
-% catalysts, cells, challenges, chemosensors, children, chromatography, climate,
-% coatings, computation, computers, cosmetics, crystals, dentistryjournal,
-% diagnostics, diseases, diversity, econometrics, economies, education,
-% electronics, energies, entropy, environmentalsciences, environments, fibers,
-% foods, forests, futureinternet, galaxies, games, genes, geosciences,
-% healthcare, humanities, hydrology, informatics, information, inorganics,
-% insects, ijerph, ijfs, ijms, ijgi, jcdd, jcm, jdb, jfb, joi, jlpea, jmse,
-% jpcg, jpm, jrfm, jsan, land, laws, life, lubricants, machines, marinedrugs,
-% materials, mathematics, medicalsciences, membranes, metabolites, metals,
-% microarrays, micromachines, microorganisms, minerals, molbank, molecules,
-% nanomaterials, ncrna, nutrients, pathogens, pharmaceuticals, pharmaceutics,
-% pharmacy, photonics, plants, polymers, processes, proteomes, publications,
-% religions, remotesensing, resources, risks, robotics, sensors, socialsciences,
-% societies, sports, sustainability, symmetry, systems, technologies, toxics,
-% toxins, vaccines, veterinarysciences, viruses, water
-%---------
-% article
-%---------
-% The default type of manuscript is article, but could be replaced by using one
-% of the class options:
-% article, review, communication, commentary, bookreview, correction, addendum,
-% editorial, changes, supfile, casereport, comment, conceptpaper,
-% conferencereport, meetingreport, discussion, essay, letter, newbookreceived,
-% opinion, projectreport, reply, retraction, shortnote, technicalnote, creative
-%----------
-% submit
-%----------
-% The class option "submit" will be changed to "accept" by the Editorial Office
-% when the paper is accepted. This will only make changes to the frontpage (e.g.
-% the logo of the journal will get visible), the headings, and the copyright
-% information. Journal info and pagination for accepted papers will also be
-% assigned by the Editorial Office.
-% Please insert a blank line is before and after all equation and eqnarray
-% environments to ensure proper line numbering when option submit is chosen
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used.
-% Otherwise use the class option moreauthors.
-%---------
-% pdftex
-%---------
-% The option "pdftex" is for use with pdfLaTeX only. If eps figure are used, use
-% the optioin "dvipdfm", with LaTeX and dvi2pdf only.
 
@@
 % The following line should be uncommented if the LaTeX file is uploaded to
-%arXiv.org
+% arXiv.org
 %\pdfoutput=1
@@
 combine linearly to produce the wake $Re$-dependence we see.}
+
+\begin{figure}[ht]
+\caption{Airfoil lift to drag ratio as a function of Reynolds number, from 
+\cite{McMasters1980}.}
+\end{figure}
 
@@
 
-\subsubsection{Mean Momentum Transport}
+\subsubsection{Dominant Timescales}
 
-\begin{figure}
+\todo[inline]{Look at how spectra change when turbine rotational frequency
+changes relative to inertial subrange.}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/wake_spectra}
+\caption{Wake spectra at $z/H=0.25$, $y/R=-1.0$ (a) and $y/R=1.5$ (b). Dashed vertical
+lines indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}
+\label{fig:wake-spectra}
+\end{figure}
+
+\subsubsection{Mean Momentum and Kinetic Energy Transport}
+
+\begin{figure}[ht]
 \centering
@@
 
-\subsubsection{Dominant Timescales}
-
-\todo[inline]{Look at how spectra change when turbine rotational frequency
-changes relative to inertial subrange.}
-
-\subsubsection{Mean Kinetic Energy Transport}
-
 The relative importance of various physical processes on mean kinetic energy
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 \label{fig:K-bar-graph}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.65\textwidth]{figures/wake_trans_totals}
+\caption{Normalized transport totals from Figures~\ref{fig:mom-bar-graph} and 
+\ref{fig:K-bar-graph} plotted versus Reynolds number.}
+\label{fig:wake-trans-totals}
 \end{figure}

### 2015-04-16 paper.tex
Add comment about C_D curves

@@
 Table~\ref{tab:Re}. Performance was also measured for $\lambda=1.9$ at tow
-speeds [0.3, 0.5, 0.7, 0.9, 1.1, 1.3] m/s, each of which consisted of two tows.
+speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s, each of which consisted of two tows.
 
@@
 Full performance curves for various Reynolds numbers are plotted in
-Figure~\ref{fig:perf-curves}.
+Figure~\ref{fig:perf-curves}. There is essentially no change in the shape of the
+drag coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
 

### 2015-04-16 paper.tex
Add two paragraphs discussing performance

@@
 
-\todo[inline]{Discuss drag coefficient curves.}
-
 In general, maximum $C_P$ increases with Reynolds number, which is due to an
@@
 
-Mean power and drag coefficients at $\lambda=1.9$ are plotted in
-Figure~\ref{fig:perf-Re-dep}
+Mean power and drag coefficients at $\lambda=1.9$ are plotted for all Reynolds
+numbers in Figure~\ref{fig:perf-Re-dep}. There is a drastic improvement in $C_P$
+with increasing Reynolds number for the lower values. Power coefficient then
+becomes  essentially $Re$-independent at $Re_D = 0.8 \times 10^6$, which
+corresponds to an approximate average blade chord Reynolds number $Re_{c,
+\mathrm{ave}} = 2.1 \times 10^5$. This threshold is consistent with the behavior
+of the blade boundary layer transitioning from laminar to turbulent, thereby
+promoting reattachment of the laminar separation bubble.
+
+The behavior of the mean rotor drag coefficient is similar, though the changes
+are less dramatic. This is likely due to cross-flow turbine geometry, where
+increases in blade drag at lower $Re$ somewhat offset the reduction in lift, to
+keep total streamwise force from changing as much as the rotor torque. The
+tendency for $C_D$ to continue increasing with $Re$ may also be a consequence of
+increasing Froude number, which therefore increases free surface deformation and
+wave drag during towing without increasing flow through the turbine.
 

### 2015-04-17 paper.tex
Write section on static foil coefficients

@@
 
-\subsubsection{Understanding via Static Foil Coefficients}
+\subsubsection{Relation to Static Foil Characteristics}
 
 To help understand---and possibly predict---the $Re$-sensitivity on turbine
-performance, a series of foil coefficient datasets were computed with the XFOIL
-viscous panel method code.
+performance, a series of static foil coefficient datasets were computed with the
+XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
+open-source turbine design software QBlade. Simulations were run with the
+default settings and zero Mach number, for the approximate average blade chord
+Reynolds numbers encountered in the experiments.
 
-The turbine torque coefficient $C_T$ can be related to the blade section
-chordwise force coefficient $C_c$ by
+To investigate the effects of the blades' ``virtual camber'' due to their
+circular path \cite{Migliore1980}, the XFOIL calculations were performed for
+20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520), and 4\% (NACA 4520)
+camber about their half-chord location. A 4\% camber approximates the maximum
+distance between the blade chord line and path for the UNH-RVAT and the 2\%
+camber takes into account the reduction in virtual flow curvature from the
+non-curved inflow velocity by dividing by the tip speed ratio $\lambda=1.9$.
+
+Results from the XFOIL calculations are shown in Figure~\ref{fig:foil-Re-dep},
+where values of maximum lift coefficient, minimum drag coefficient, and maximum
+lift-to-drag ratio are normalized (for comparison between foils) and plotted
+versus $Re_c$. In general, larger camber is associated with decreased foil
+performance at lower Reynolds number.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.95\textwidth]{figures/all_foils_re_dep}
+\caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
+\label{fig:foil-Re-dep}
+\end{figure}
+
+In conjunction with the cross-flow turbine blade kinematics, the foil
+coefficients were used to approximate turbine performance by calculating the
+peak torque coefficient on the upstream half of the blade path. The turbine
+torque coefficient $C_T$ can be related to the blade section chordwise force
+coefficient $C_c$ by
 \begin{equation}
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.95\textwidth]{figures/all_foils_re_dep}
-\caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
-\label{fig:foil-Re-dep}
-\end{figure}
+The relative blade velocity was calculated by adding the free stream to the
+opposite of the blade tangential velocity. Note that this neglects any
+induction, or slowing of the free stream by the turbine forces, which would be
+present in a momentum/streamtube model. Since we were not looking to predict
+absolute performance but rather relative changes with $Re$, this method was
+deemed acceptable as it is extremely simple and fast to compute.
+
+Values for the blade angle of attack, relative velocity, and torque coefficient
+are plotted in Figure~\ref{fig:blade-kinematics}. The effects of static stall
+are clearly present in the torque coefficient plot, and by the time the angle of
+attack has decreased below that of stall, the relative velocity is so low that
+there is not much contribution to the torque coefficient.
 
@@
 
+Results for the peak torque coefficient for each foil are plotted versus $Re_c$
+in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the convergence of
+$C_{T_\mathrm{max}}$ with increasing Reynolds number is more dramatic than any of the
+common foil characteristics plotted in Figure~\ref{fig:foil-Re-dep}, meaning
+that the cross-flow turbine's unique kinematics must be taken into account when
+attempting to predict the effect of transitional Reynolds numbers on turbine
+performance.
+
+From the peak torque coefficient metric, the Reynolds number independence is
+achieved at lower values, and more dramatically. We see that the trend of the
+NACA 0020 curve matches almost perfectly at low $Re$, but the trend to continue
+increasing linearly is not matched by the experimental data, which looks to
+behave more like the cambered foil results. Though this method is very
+approximate, it nevertheless appears to be a reasonable way to predict the
+transitional Reynolds number regime for cross-flow turbine performance using
+only 2-D static airfoil characteristics
+
 \begin{figure}[ht]
@@
 \end{figure}
-
-Note that the common static foil performance metrics fail to show Reynolds
-number dependence as close as the max geometric torque coefficient. This means
-that the turbine kinematics play an important role, even if the exact dynamics
-of the flow are not known.
 

### 2015-04-17 paper.tex
Remove hyphen from 'open source'

@@
 XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
-open-source turbine design software QBlade. Simulations were run with the
+open source turbine design software QBlade. Simulations were run with the
 default settings and zero Mach number, for the approximate average blade chord

### 2015-04-17 paper.tex
Some work on introduction; still very rough

@@
 
-Previously we took a general look at how maximum performance and some 2D wake
-profiles are affected by Reynolds number, but here we week to look much deeper.
-
-In this study we set out to measure the effects of scale, both Reynolds and
-Froude scales, on the performance of the turbine, and its near-wake
-characteristics.
+Previously we took a general look at how maximum performance and some 2-D wake
+profiles are affected by Reynolds number, but here we look deeper. In this study
+we set out to measure the effects of scale, i.e. Reynolds number on the
+performance of the turbine, and its near-wake characteristics. We were looking
+to find a threshold scale beyond which a physical model study could reliably
+predict full-scale performance. We were also looking to investigate the
+sensitivity as small scales, and what the implications might be for validating
+the prediction capabilities of numerical models.
 

### 2015-04-17 paper.tex
Add to wake results section

@@
 
-\subsubsection{Dominant Timescales}
+\subsubsection{Dominant Timescales and Turbulence Spectra}
 
@@
 
-\subsubsection{Mean Momentum and Kinetic Energy Transport}
+\subsubsection{Transport of Mean Momentum and Kinetic Energy}
+
+The relative importance of various physical processes on mean streamwise
+momentum and kinetic energy transport/recovery in the streamwise direction are
+plotted in Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph},
+respectively. We note that similar to \cite{Bachant2015-JoT}, the vertical
+advection at this point in the wake is the dominant contributor to positive wake
+recovery, caused by the unique vortex pattern created by the blade tips and
+wakes.
+
+We see that in general, levels of turbulent transport are slightly lower at
+lower Reynolds numbers. The viscous diffusion and dissipation, though still
+three orders of magnitude smaller than the other terms, do increase at low
+Reynolds numbers, which is consistent with the definition of the Reynolds number
+in general.
+
+\todo[inline]{At what $Re$ will the viscous terms get large enough to start
+messing with the balance of the equation?}
 
@@
 
-The relative importance of various physical processes on mean kinetic energy
-transport/recovery in the streamwise direction $\p K / \p x$ are plotted in...
-
 \begin{figure}[ht]
@@
 \end{figure}
+
+Totals for all the wake transport terms calculated are plotted in
+Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport is
+enhanced at lower Reynolds numbers and levels off consistent with the power
+coefficient data. This is an important consideration if studying sub-scale
+models of turbine arrays, where increased levels of wake recovery will motivate
+different ideal array configurations when compared with full-scale turbines.
 

### 2015-04-17 paper.tex
Clarify tows for odd speeds
Describe spectra plots

@@
 Table~\ref{tab:Re}. Performance was also measured for $\lambda=1.9$ at tow
-speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s, each of which consisted of two tows.
+speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows each.
 
@@
 
+Spectra of the cross-stream velocity normalized by the free stream are plotted
+in Figure~\ref{fig:wake-spectra} for regions on either side of the turbine. On
+the $-y$ side of the turbine there is broadband turbulence produced by blade
+stall, and on the $+y$ side there is a clear peak in the spectra caused by the
+blade passage. We see that on both sides there is higher spectral energy at
+lower Reynolds numbers. On the $+y$ side of the turbine we notice higher
+spectral energy at the blade passage frequency's first harmonic, or $6
+f_\mathrm{turbine}$.
+
 \todo[inline]{Look at how spectra change when turbine rotational frequency
 changes relative to inertial subrange.}
+
+\todo[inline]{Make sure the spectra are not just shifting due to normalizing the
+noise.}
 
@@
 \includegraphics[width=0.95\textwidth]{figures/wake_spectra}
-\caption{Wake spectra at $z/H=0.25$, $y/R=-1.0$ (a) and $y/R=1.5$ (b). Dashed vertical
+\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$, 
+$y/R=-1.0$ (a) and $y/R=1.5$ (b). Dashed vertical
 lines indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}

### 2015-04-17 paper.tex
Add to conclusions

@@
 keep all turbines in the linear regime to avoid exaggerated power deficiencies
-for downstream turbines, despite similarities in wake characteristics
+for downstream turbines, despite similarities in wake characteristics. These
+results also suggest that low Reynolds number physical model stydies of turbine
+arrays may see exaggerated levels of wake recovery, leading to inadequate or
+inappropriate spacing or layouts.
 

### 2015-04-18 paper.tex
Change float priority
Rewording
Some work on introduction

@@
 real-world systems. In the fields of wind and marine hydrokinetic (MHK) energy,
-scaled physical models are used to validate numerical models, test prototypes of
-individual devices, and design or investigate arrays of devices. Despite being
-significantly less expensive, there is still some concern at which scale an
-experiment must be performed to be relevant to full-scale application.
+scaled physical models are used to validate predictive numerical models, predict
+full-scale performance of individual turbines, and design or investigate arrays
+of devices. Despite being significantly less expensive, there is still some
+concern at which scale an experiment must be performed to be relevant to
+full-scale application.
 
@@
 
-When designing or studying arrays, it is very common to use very small
-(geometrically) scaled devices. Sometimes these are not even models of turbines,
-but wake generating objects, e.g., porous disks, that are meant to replicate the
+When designing or studying arrays, it is common to use very small
+(geometrically) scaled devices. 
+% %
+\todo[inline]{Get a reference for this, and a sample geometric scale.} 
+% %
+Sometimes these are not even models of turbines, but
+wake generating objects, e.g., porous disks, that are meant to replicate the
 wakes of real turbines. In this case, it is of interest to determine at what
 scale one might be able to realistically study wake flows in an array, and also
-to evaluate the effectiveness of a wake generator, i.e., a wake generator may do
-a fine job simulating a scaled turbine, but how well can it simulate a real
-turbine?
+to evaluate the effectiveness of a wake generator. In other words, a wake
+generator may do a fine job simulating a scaled turbine, but how well can it
+simulate a real turbine?
 
 Previously, Bachant and Wosnik \cite{Bachant2015-JoT} showed that a cross-flow
-turbine wake's unique mean flow field is responsible for accelerated wake
-recovery when compared with conventional axial-flow propeller-type turbines.
+turbine wake's unique vortical mean flow field is responsible for accelerated
+wake recovery when compared with conventional axial-flow propeller-type
+turbines. In this study we seek to replicate the same energy balance
+considerations at multiple Reynolds numbers, to examine the implications on how
+small scale, i.e. low Reynolds number experiments may be used to study flows in
+turbine arrays.
 
@@
 of individual foils, with respect to their static and dynamic loading.
-
-Chamorro et al. showed that turbulence statistics in an AFT wake became
-$Re$-independent at $Re_D \approx 1 \times 10^5$, with mean velocity slightly
-earlier at $Re_D \approx 5 \times 10^4$ \cite{Chamorro2012}.
 
@@
 on shear flows, wakes, and vortex flows?}
+
+Chamorro et al. showed that turbulence statistics in an AFT wake became
+$Re$-independent at $Re_D \approx 1 \times 10^5$, with mean velocity slightly
+earlier at $Re_D \approx 5 \times 10^4$ \cite{Chamorro2012}.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
 \centering
@@
 Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport is
-enhanced at lower Reynolds numbers and levels off consistent with the power
-coefficient data. This is an important consideration if studying sub-scale
-models of turbine arrays, where increased levels of wake recovery will motivate
-different ideal array configurations when compared with full-scale turbines.
+enhanced at lower Reynolds numbers and levels off consistent with the behavior
+of the  turbine power coefficient. This is an important consideration if
+studying sub-scale models of turbine arrays, where increased levels of wake
+recovery will motivate different ideal array configurations when compared with
+full-scale turbines.
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
 \centering

### 2015-04-18 paper.tex
Clarify one turbine diameter downstream

@@
 Velocimeter (ADV) at 270 different locations, varied in the cross-stream and
-vertical directions at one turbine diameter downstream. These locations have
-vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
-the cross-stream direction $y/R = \pm 3$. These locations are shown in
-Figure~\ref{fig:wake-locations}.
+vertical directions at one turbine diameter downstream ($x/D=1$). These
+locations have vertical coordinates from the turbine centerline up to
+$z/H=0.625$, ranging in the cross-stream direction $y/R = \pm 3$. These
+locations are shown in Figure~\ref{fig:wake-locations}.
 

### 2015-04-18 paper.tex
Add drawing of experimental setup
Better description of the RVAT

@@
 model testing, similar to the Sandia National Labs/US Department of Energy RM2
-River Turbine. Details of the turbine and experimental setup are described in
-\cite{Bachant2015-JoT}, and a CAD model of the turbine is available from
-\cite{Bachant2014-RVAT-CAD}.
+River Turbine, but with higher solidity, or blade chord to radius ratio. Details
+of the turbine and experimental setup are described in \cite{Bachant2015-JoT},
+and a CAD model of the turbine is available from \cite{Bachant2014-RVAT-CAD}.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.75\textwidth]{figures/exp_setup_drawing}
+\caption{Drawing of the experimental setup.}
+\label{fig:exp-setup}
+\end{figure}
 

### 2015-04-18 paper.tex
Update introduction

@@
 Scaled physical models are often used in science and engineering to approximate
-real-world systems. In the fields of wind and marine hydrokinetic (MHK) energy,
-scaled physical models are used to validate predictive numerical models, predict
-full-scale performance of individual turbines, and design or investigate arrays
-of devices. Despite being significantly less expensive, there is still some
-concern at which scale an experiment must be performed to be relevant to
-full-scale application.
+real-world systems. However, the principle of dynamic similarity allows for
+geometrically scaled systems to be dynamically identical in scale if certain
+nondimensional physical parameters are matched. In the case of fluid systems,
+the most important parameter is often the Reynolds number, which quantifies the
+importance of viscosity on the flow. As such, hereafter we will use the word
+``scale'' to refer to this dynamical scale rather than the geometric scale.
+
+With regards to wind and marine hydrokinetic (MHK) turbines, scaled physical
+models are used to validate predictive numerical models, predict full-scale
+performance of individual turbines, and design or investigate arrays of devices.
+Despite being significantly less expensive, there is still some concern at which
+scale an experiment must be performed to be relevant to full-scale application.
 
@@
 small---since size, or scale generally correlates inversely with cost---the
-prototype can be while remaining a reliable predictor of full-scale behavior.
+prototype can be while remaining a reliable predictor of full-scale behavior. At
+the very least, it is necessary to test at a scale where results can be
+extrapolated, i.e. changes are occurring quasi-linearly without dramatic changes
+in the underlying physical principles governing the behavior of interest, e.g.
+power production in the case of a wind or MHK turbine.
 
@@
 
-Previously we took a general look at how maximum performance and some 2-D wake
-profiles are affected by Reynolds number, but here we look deeper. In this study
-we set out to measure the effects of scale, i.e. Reynolds number on the
-performance of the turbine, and its near-wake characteristics. We were looking
-to find a threshold scale beyond which a physical model study could reliably
+Bachant and Wosnik observed how maximum performance and 2-D near-wake profiles
+are affected by Reynolds number, but here we look deeper. In this study we set
+out to measure the effects of scale, i.e. Reynolds number, on the performance of
+the turbine, and its more detailed near-wake characteristics. We were looking to
+find a threshold scale beyond which a physical model study could reliably
 predict full-scale performance. We were also looking to investigate the
@@
 blade chord Reynolds number $Re_c$. This can be explained as the transition to a
-turbulent boundary layer adjacent to the blade delaying separation. However,
-static foil performance does not tell the whole story for a cross-flow turbine.
-The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
-turbine implies the occurrence of dynamic stall at tip speed ratios at and below
-those of maximum rotor torque\cite{Para2002}. We then must understand the
-effects of Reynolds number on the dynamic stall process.
+turbulent boundary layer adjacent to the blade delaying separation.
 
@@
 \subsubsection{On Dynamic Foil Behavior}
+
+Static foil performance does not tell the whole story for a cross-flow turbine.
+The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
+turbine implies the occurrence of dynamic stall at tip speed ratios at and below
+those of maximum rotor torque\cite{Para2002}. We then must understand the
+effects of Reynolds number on the dynamic stall process.
 

### 2015-04-18 paper.tex
Add description of Blackwell et al and Bravo et al to intro

@@
 power production in the case of a wind or MHK turbine.
+
+Blackwell et al. investigated the effects of Reynolds number on the performance
+of a 2 m diameter Darrieus vertical-axis cross-flow wind turbine with NACA 0012
+blades \cite{Blackwell1976}. By varying wind tunnel speed, the turbine was made
+to operate at approximately constant blade chord Reynolds number $Re_c$ ranging
+from $1 \times 10^5$ to $3 \times 10^5$. In this regime the turbine power
+coefficient $C_P$ was shown to be sensitive to $Re_c$, with sensitivity
+diminishing at the higher Reynolds numbers, especially for turbines of lower
+solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
+maximum radius). More recently, Bravo et al. observed power coefficient of a
+straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
+\times 10^5$.
 

### 2015-04-18 paper.tex
Refine paragraph about foil performance

@@
 Bachant and Wosnik observed how maximum performance and 2-D near-wake profiles
-are affected by Reynolds number, but here we look deeper. In this study we set
-out to measure the effects of scale, i.e. Reynolds number, on the performance of
-the turbine, and its more detailed near-wake characteristics. We were looking to
-find a threshold scale beyond which a physical model study could reliably
-predict full-scale performance. We were also looking to investigate the
+are affected by Reynolds number \cite{Bachant2014}. In this study we set out to
+analyze in greater depth the effects of scale, i.e. Reynolds number, on the
+performance of the turbine, and its more detailed near-wake characteristics. We
+were looking to find a threshold scale beyond which a physical model study could
+reliably predict full-scale performance. We were also looking to investigate the
 sensitivity as small scales, and what the implications might be for validating
@@
 
-Turbine performance is strongly dependent on the lift-to-drag ratio $l/d$ acting
-on the blades, which is a function of the blades' angles of attack $\alpha$.
-These angles of attack are a function of turbine tip speed ratio, induction
-(slowing/turning of the free stream flow prior to reaching the turbine), and
-turbine rotation or azimuthal angle $\theta$. Looking at static airfoil data, we
-see that the static stall angle---the angle just beyond that at which $l/d$ is
-maximum, characterized by flow separation near the leading edge---increases with
-blade chord Reynolds number $Re_c$. This can be explained as the transition to a
-turbulent boundary layer adjacent to the blade delaying separation.
+Blade performance is generally characterized by the lift-to-drag ratio $l/d$,
+which is a function of the foil's angle of attack $\alpha$. The angles of attack
+for cross-flow turbine blades are a function of turbine tip speed ratio,
+induction (slowing/turning of the free stream flow prior to reaching the
+turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
+airfoil data, we see that the static stall angle---the angle just beyond that at
+which $l/d$ is maximum---increases with blade chord Reynolds number $Re_c$. This
+in general is caused by the blade boundary becoming more turbulent as a fraction
+of the chord length, diminishes the negative effects of a laminar separation
+bubble occurring near the leading edge. A review of low Reynolds number airfoil
+behavior is presented in \cite{Lissaman1983}.
 

### 2015-04-18 paper.tex
Add to foil performance section in the introduction

@@
 of the chord length, diminishes the negative effects of a laminar separation
-bubble occurring near the leading edge. A review of low Reynolds number airfoil
-behavior is presented in \cite{Lissaman1983}.
-
-As previously mentioned, the near-wake of a CFT is not very similar to a simple
-drag source, or plane shear flow. One of the mean wake's streamwise vortex pairs
-can be predicted by potential flow theory as a result of finite-span blades
-producing lift (blade end effects), but the other---caused by viscous effects,
-i.e., dynamic stall---is not. We can attempt to predict changes
-\todo[inline]{Finish this thought with Reynolds number dependence of these
-simpler flows, and set up the question whether or not we can assume these will
-combine linearly to produce the wake $Re$-dependence we see.}
+bubble occurring near the leading edge. For smooth airfoils, this transition can
+cause a dramatic increase in foil performance, a shown in
+Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
+$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
+presented in \cite{Lissaman1983}.
 
@@
 \cite{McMasters1980}.}
+\label{fig:McMasters}
 \end{figure}
@@
 The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
-turbine implies the occurrence of dynamic stall at tip speed ratios at and below
-those of maximum rotor torque\cite{Para2002}. We then must understand the
-effects of Reynolds number on the dynamic stall process.
+turbine implies dynamic loading. Furthermore, the blades commonly undergo
+dynamic stall for tip speed ratios at and below those of maximum rotor
+torque\cite{Para2002}. We then must understand the effects of Reynolds number on
+the dynamic stall process.
 
@@
 separated conditions.
+
+\todo[inline]{Find reference to describe $Re$ effects on dynamic foil behavior.}
 
@@
 layers. As mentioned previously, the dynamic stall vortex shedding is not
-expected to have a large Reynolds number sensitivity.
+expected to have a large Reynolds number sensitivity, though a larger separation
+bubble at lower $Re$ may induce higher levels of turbulence as shed vortices
+become unstable and break down.
 

### 2015-04-19 getfigs.py
Add figures

@@ -15,8 +15,10 @@ figlist = ["perf_curves.pdf",
            "perf_re_dep.pdf",
            "K_trans_bar_graph.pdf",
            "mom_bar_graph.pdf",
-	   "wake_spectra.pdf",
-	   "wake_trans_totals.pdf"]
+           "wake_spectra.pdf",
+           "wake_trans_totals.pdf",
+           "meancontquiv_10.pdf",
+           "k_contours_10.pdf"]
 
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))

### 2015-04-19 paper.tex
Add wake figures for Re_D = 1 x 10^6
Add two paragraphs to wake measurements section based on JoT results

@@
 \begin{figure}[ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/McMasters-Henderson-1980}
 \caption{Airfoil lift to drag ratio as a function of Reynolds number, from 
@@
 
-\todo[inline]{Put wake characteristics here. Maybe create difference plots to
-show absolute differences, or RMS, or something}
+The wake of this turbine at $x/D=1$, $\lambda=1.9$, $Re_D = 1.0 \times 10^6$ was
+described thoroughly in \cite{Bachant2015-JoT}. Similar data were taken for the
+experiment here, the results for the mean velocity field and turbulence kinetic
+energy are shown in Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont},
+respectively. With respect to the mean velocity field, we see asymmetry and a
+mean vortex structure created by the blade tip vortex shedding. The effects of
+the tip vortices are also seen in the turbulence kinetic energy measurements,
+along with turbulence generated by the blades undergoing dynamic stall on the
+$-y$ side of the turbine.
 
-\todo[inline]{Plot asymptotic behavior for a bunch of locations all on the same
-plot.}
+These same wake maps were taken for $Re_D = 0.4 \time 10^6$, $0.6 \times 10^6$,
+$0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look strikingly
+similar, so they have not been plotted here. We will instead compare and
+contrast the wake behavior by examining spectra and wake transport terms in the
+equations that govern the downstream evolution.
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/meancontquiv_10}
+\caption{Mean velocity field at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0 \times 10^6$.}
+\label{fig:meancontquiv}
+\end{figure}
+
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/k_contours_10}
+\caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$, and 
+$Re_D=1.0 \times 10^6$.}
+\label{fig:kcont}
+\end{figure}
 

### 2015-04-19 paper.tex
Add placeholder for turbine drawing

@@
 River Turbine, but with higher solidity, or blade chord to radius ratio. Details
-of the turbine and experimental setup are described in \cite{Bachant2015-JoT},
-and a CAD model of the turbine is available from \cite{Bachant2014-RVAT-CAD}.
+of the turbine and experimental setup, shown in Figure~\ref{fig:turbine} and
+Figure~\ref{fig:exp-setup}, respectively, are described in
+\cite{Bachant2015-JoT}, and a CAD model of the turbine is available from
+\cite{Bachant2014-RVAT-CAD}.
+
+\begin{figure}[ht!]
+\caption{Drawing of the turbine model. Note that the upper and lower mounting
+flanges have been excluded as these were included in the tare drag
+measurements.} 
+\label{fig:turbine}
+\end{figure}
 
@@
 \end{figure}
-
-\todo[inline]{Make sure to use drawing of turbine with no hubs, since these were
-included in the tare drag}
 

### 2015-04-21 paper.tex
Update figure caption for turbine drawing
Add turbine drawing
Add 'Dynamics' to title
Update bibfile and add citation for figshare repo

@@
 % Full title of the paper (Capitalized)
-\Title{Effects of Reynolds Number on the Performance and Near-Wake of a
+\Title{Effects of Reynolds Number on the Performance and Near-Wake Dynamics of a
 Vertical-Axis Cross-Flow Turbine}
@@
 \begin{figure}[ht!]
-\caption{Drawing of the turbine model. Note that the upper and lower mounting
+\centering
+\includegraphics[width=0.4\textwidth]{figures/turbine}
+\caption{Turbine model drawing. Note that the upper and lower mounting
 flanges have been excluded as these were included in the tare drag
@@
 
-\todo[inline]{Cite reduced data and processing code on figshare.}
+The experimental data and processing/plotting code are available from
+\cite{Bachant2015-RVAT-Re-dep-data}.
 

### 2015-04-22 paper.tex
Add comment about lack of airfoil data; resolves #12
Adjust citation of our previous work

@@
 
-Previously, Bachant and Wosnik \cite{Bachant2015-JoT} showed that a cross-flow
-turbine wake's unique vortical mean flow field is responsible for accelerated
-wake recovery when compared with conventional axial-flow propeller-type
-turbines. In this study we seek to replicate the same energy balance
-considerations at multiple Reynolds numbers, to examine the implications on how
-small scale, i.e. low Reynolds number experiments may be used to study flows in
-turbine arrays.
+Previously, is has been observed experimentally that a cross-flow turbine wake's
+unique vortical mean flow field is responsible for accelerated wake recovery
+when compared with conventional axial-flow propeller-type turbines
+\cite{Bachant2015-JoT}. In this study we seek to replicate the same energy
+balance considerations at multiple Reynolds numbers, to examine the implications
+on how small scale, i.e. low Reynolds number experiments may be used to study
+flows in turbine arrays.
 
-Bachant and Wosnik observed how maximum performance and 2-D near-wake profiles
-are affected by Reynolds number \cite{Bachant2014}. In this study we set out to
+The Reynolds number effects on the maximum performance and 2-D near-wake
+profiles of the University of New Hampshire's Reference Vertical Axis Turbine
+(UNH-RVAT) were presented \cite{Bachant2014}. In this study we set out to
 analyze in greater depth the effects of scale, i.e. Reynolds number, on the
@@
 $10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
-presented in \cite{Lissaman1983}.
+presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
+reliable data for airfoils in this transitional regime and below. An evaluation
+of the various databases relevant to cross-flow turbines is presented in
+\cite{Bedon2014}.
 

### 2015-04-22 paper.tex
Cite Polagye and Cavagnaro micropower paper
Clarify turbine type in conclusion

@@
 solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
-maximum radius). More recently, Bravo et al. observed power coefficient of a
+maximum radius). More recently, Polagye and Cavagnaro observed high Reynolds
+number sensitivity for a high solidity helical cross-flow turbine
+\cite{Polagye2013b}, and Bravo et al. observed power coefficient of a
 straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
@@
 
-In this study we determined that, as previously shown, turbine performance
+In this study we determined that, as previously shown, performance for a
+high-solidity cross-flow turbine constructed from straight NACA 0020 blades
 becomes essentially $Re$-independent at a turbine diameter Reynolds number $Re_D

### 2015-04-23 paper.tex
Work on introduction

@@
 the most important parameter is often the Reynolds number, which quantifies the
-importance of viscosity on the flow. As such, hereafter we will use the word
-``scale'' to refer to this dynamical scale rather than the geometric scale.
+importance of viscosity on the flow physics. As such, hereafter we will use the
+word ``scale'' to refer to this dynamical scale rather than the geometric scale.
 
@@
 When designing or studying arrays, it is common to use very small
-(geometrically) scaled devices. 
-% %
-\todo[inline]{Get a reference for this, and a sample geometric scale.} 
-% %
-Sometimes these are not even models of turbines, but
-wake generating objects, e.g., porous disks, that are meant to replicate the
-wakes of real turbines. In this case, it is of interest to determine at what
+(geometrically) scaled devices \cite{Chamorro2011, Chamorro2011b}. It is
+therefore important to realize the limitations of evaluating both the power
+output of turbines and the wake recovery when Reynolds number is very far from
+full scale. Sometimes models used are not turbines, but wake generating objects,
+e.g., porous disks, that are meant to replicate the wakes of real turbines
+\cite{Goldenberg1983}. In this case, it is of interest to determine at what
 scale one might be able to realistically study wake flows in an array, and also
@@
 generator may do a fine job simulating a scaled turbine, but how well can it
-simulate a real turbine?
+simulate a full-scale device?
 

### 2015-04-23 paper.tex
Center figures
Describe data processing in more detail

@@
 
-From each set of tows, a standard time interval was set, and each run was
+From each set of tows, a standard time interval was set, which allowed the
+turbine performance and wake to reach a quasi-periodic state. Each run was
 analyzed to compute statistics over this interval, truncating the end slightly
-to achieve and integer number of blade passages.
+to achieve and integer number of blade passages. Turbine shaft angular velocity
+and tow carriage speed were calculated using a second order central differencing
+scheme on the respective position measurements. Power and drag coefficients are
+calculated as instantaneous quantities using the carriage speed as the free
+stream velocity.
 
-The experimental data and processing/plotting code are available from
+Wake velocity data were filtered for spurious ``spikes'' by removing datapoints
+8 standard deviations or 0.9 m/s from the mean. The experimental data and code
+for processing and visualization are available from
 \cite{Bachant2015-RVAT-Re-dep-data}.
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[width=0.95\textwidth]{figures/perf_curves}
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[width=0.95\textwidth]{figures/perf_re_dep}

### 2015-04-23 paper.tex
Discuss the experimental setup in more detail

@@
 model testing, similar to the Sandia National Labs/US Department of Energy RM2
-River Turbine, but with higher solidity, or blade chord to radius ratio. Details
-of the turbine and experimental setup, shown in Figure~\ref{fig:turbine} and
-Figure~\ref{fig:exp-setup}, respectively, are described in
-\cite{Bachant2015-JoT}, and a CAD model of the turbine is available from
+River Turbine, but with higher solidity, or blade chord to radius ratio. A CAD
+model of the turbine, depicted in \ref{fig:turbine}, is available from
 \cite{Bachant2014-RVAT-CAD}.
@@
 
+The turbine was mounted in a frame constructed from NACA 0020 sections. The
+turbine shaft ran up through the water surface, coupled to a permanent magnet
+servo motor with a 20:1 gearbox. This servo was controlled by the tow tank's
+main motion controller for high synchronization with the carriage motion. A
+rotary torque transducer was installed inline between the servo and the turbine,
+and the servo was also mounted on a slewing ring bearing, which allowed a
+redundant measurement of torque via an arm and load cell used to counteract the
+turbine moment. The frame was mounted to the carriage via linear guides, such
+that the total streamwise drag force was transferred to a pair of S-beam load
+cells, providing the rotor drag measurements, after a measured tare drag was
+subtracted in post processing. Similarly, a tare torque was measured by rotating
+the turbine shaft in air. Turbine angular and tow carriage linear position were
+measured using quadrature encoder signals, with $10^5$ counts-per-rev for the
+turbine and 10 $\mathrm{\mu m}$ resolution for the carriage position. These
+signals, along with the torque and drag signals, were sampled at 2 kHz.
+
 \begin{figure}[ht!]
@@
 \end{figure}
+
+Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
+(ADV), sampling at 200 Hz. The probe was mounted on an automated positioning
+system, also controlled by the tow tank's main motion controller. The ADV and
+DAQ systems' sampling times were synchronized by triggering the start of data
+acquisition via a pulse sent from the motion controller.
 

### 2015-04-23 paper.tex
Remove repeat ADV definition

@@
 
-A wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
-Velocimeter (ADV) at 270 different locations, varied in the cross-stream and
-vertical directions at one turbine diameter downstream ($x/D=1$). These
-locations have vertical coordinates from the turbine centerline up to
-$z/H=0.625$, ranging in the cross-stream direction $y/R = \pm 3$. These
-locations are shown in Figure~\ref{fig:wake-locations}.
+A wake map was generated by positioning the Vectrino probe at 270 different
+locations, varied in the cross-stream and vertical directions at one turbine
+diameter downstream ($x/D=1$). These locations have vertical coordinates from
+the turbine centerline up to $z/H=0.625$, ranging in the cross-stream direction
+$y/R = \pm 3$. These locations are shown in Figure~\ref{fig:wake-locations}.
 

### 2015-04-23 paper.tex
Add XFOIL parameters; resolves #18

@@
 XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
-open source turbine design software QBlade. Simulations were run with the
-default settings and zero Mach number, for the approximate average blade chord
-Reynolds numbers encountered in the experiments.
+open source turbine design software QBlade. Simulations were run for an angle of
+attack range of 0--40 degrees, in increments of 0.5 degrees. Solver parameters
+used a 100 panels, a fixed speed, zero Mach number, \texttt{NCrit=9} (default
+$\mathrm{e}^n$ transition criteria parameter for an average wind tunnel), and no
+forced boundary layer transition. Characteristics for computed for the
+approximate average blade chord Reynolds numbers encountered in the tow tank
+experiments.
 

### 2015-04-23 paper.tex
Add reference for Re effects on dynamic stall
Add to conclusions regarding XFOIL calculations

@@
 Bousman showed that dynamic stall was insensitive to Reynolds number for $Re=1.0
-\times 10^5$--$2.5 \times 10^5$ \cite{Bousman2000-evaluation}.
+\times 10^5$--$2.5 \times 10^5$ \cite{Bousman2000-evaluation}, Though Singleton
+and Yeager Jr. state that Reynolds number effects on dynamics stall remains an
+unsolved question \cite{Singleton2000}.
 
@@
 
+We propose a method for predicting Reynolds number dependence of a cross-flow
+turbine using static airfoil characteristics and turbine blade kinematics. It is
+shown that the peak torque coefficient computed this way shows similar Reynolds
+number sensitivity---a better predictor than conventional quantifications of
+airfoil performance, e.g., lift-to-drag ratio. Is it also shown from this metric
+that foils with higher camber are more Reynolds number dependent but have
+smaller slopes in the linear regime.
+
 If using scaled physical models to predict array performance, it is important to
@@
 for downstream turbines, despite similarities in wake characteristics. These
-results also suggest that low Reynolds number physical model stydies of turbine
+results also suggest that low Reynolds number physical model studies of turbine
 arrays may see exaggerated levels of wake recovery, leading to inadequate or

### 2015-06-29 paper.tex
Fix typo
Split perf curve subplots into their own figures and update since resizing

@@
 \usepackage[draft]{todonotes}
+\usepackage{lineno}
 
@@
 \begin{document}
-
+%\linenumbers
 \listoftodos
@@
 
-Full performance curves for various Reynolds numbers are plotted in
-Figure~\ref{fig:perf-curves}. There is essentially no change in the shape of the
-drag coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
+Full power and rotor drag coefficient curves for various Reynolds numbers are
+plotted in Figure~\ref{fig:cp-curves} and Figure~\ref{fig:cd-curves},
+respectively. There is essentially no change in the shape of the drag
+coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
 
@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/perf_curves}
-\caption{Mean power (left) and drag (right) coefficient curves plotted for
+\includegraphics[width=0.65\textwidth]{figures/cp_curves}
+\caption{Mean power coefficient curves plotted for
 multiple Reynolds numbers.}
-\label{fig:perf-curves}
+\label{fig:cp-curves}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.65\textwidth]{figures/cd_curves}
+\caption{Mean rotor drag coefficient curves plotted for
+multiple Reynolds numbers.}
+\label{fig:cd-curves}
 \end{figure}
@@
 
-These same wake maps were taken for $Re_D = 0.4 \time 10^6$, $0.6 \times 10^6$,
+These same wake maps were taken for $Re_D = 0.4 \times 10^6$, $0.6 \times 10^6$,
 $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look strikingly
@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/wake_spectra}
+\includegraphics[width=0.85\textwidth]{figures/wake_spectra}
 \caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$, 

### 2015-06-29 getfigs.py

@@ -11,7 +11,8 @@ import shutil
 figdir = os.path.join(os.path.expanduser("~"), "Research",
                       "Experiments", "RVAT Re dep", "Figures")
 
-figlist = ["perf_curves.pdf",
+figlist = ["cp_curves.pdf",
+           "cd_curves.pdf",
            "perf_re_dep.pdf",
            "K_trans_bar_graph.pdf",
            "mom_bar_graph.pdf",
@@ -23,7 +24,7 @@ figlist = ["perf_curves.pdf",
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))
 
-    
+
 figdir = os.path.join(os.path.expanduser("~"), "Google Drive", "Research",
                       "Foils", "Data", "NACAXX20_QBlade", "figures")
 

### 2015-06-30 paper.tex
Resize figures and remove titles for differing quantities; resolves #22 and #28

@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/all_foils_re_dep}
+\includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
 \caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/foil_kinematics_ct}
-\caption{Geometric angle of attack (a), relative velocity (b) and torque coefficient
-(c) calculated with a NACA 0020 foil.}
+\includegraphics[width=0.9\textwidth]{figures/foil_kinematics_ct}
+\caption{Geometric angle of attack, relative velocity and torque coefficient
+calculated with a NACA 0020 foil.}
 \label{fig:blade-kinematics}

### 2015-09-03 paper.tex
Scale down figure
changeto secind sentence of Introduction (test edit & commit)

@@
 Scaled physical models are often used in science and engineering to approximate
-real-world systems. However, the principle of dynamic similarity allows for
-geometrically scaled systems to be dynamically identical in scale if certain
-nondimensional physical parameters are matched. In the case of fluid systems,
-the most important parameter is often the Reynolds number, which quantifies the
+real-world systems. The principle of dynamic similarity allows for geometrically
+scaled systems to be dynamically similar in scale if certain nondimensional
+physical parameters are matched. In the case of fluid systems, the most
+important parameter is often the Reynolds number, which quantifies the
 importance of viscosity on the flow physics. As such, hereafter we will use the
@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/perf_re_dep}
+\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
 \caption{Mean power (left) and drag (right) coefficients at $\lambda=1.9$

### 2015-09-03 paper.tex
changeto secind sentence of Introduction (test edit & commit)

@@
 real-world systems. The principle of dynamic similarity allows for geometrically
-scaled systems to be dynamically similar in scale if certain nondimensional
+scaled systems to be dynamically identical in scale if certain nondimensional
 physical parameters are matched. In the case of fluid systems, the most
-important parameter is often the Reynolds number, which quantifies the
-importance of viscosity on the flow physics. As such, hereafter we will use the
-word ``scale'' to refer to this dynamical scale rather than the geometric scale.
+important nondimensional parameter is often the Reynolds number, which
+quantifies the importance of inertial forces over viscous forces on the flow
+physics. It should be noted that in general the geometric scale of a physical
+model is not necessarily the same as its dynamical scale. In the case of
+geometrically similar physical models of turbines and propellers operating under
+kinematic similarity, i.e. at the same tip speed or advance ratio, however, the
+dynamical scale can be (is) the same as the geometric scale. It is the dynamical
+scale of the model which we are concerned with hereafter.
+
+\todo[inline]{Need to find a good way to succintly introduce scaling 
+subleties.}
 

### 2015-10-10 paper.tex
The last commit comment should have been: Introduction - edits for first paragraph
Add line numbers and remove list of todos; resolves #20
All edits by MW from last 2 weeks. need to address To-Dos.

--- 2015-06-30 paper.tex
@@
 \usepackage[draft]{todonotes}
-\usepackage{lineno}
 
@@
 % there are more than one corresponding author.)
-\corres{pxL3@unh.edu}
+\corres[2]{pxL3@unh.edu, martin.wosnik@unh.edu}
 
 % Abstract (Less than 200 words)
-\abstract{The acceptable Reynolds number mismatch for scaled physical model
-testing a cross-flow turbine is investigated experimentally, both with respect
-to prediction of full-scale performance, and wake characteristics, which will be
-drivers of overall performance and interaction between multiple turbines.}
+\abstract{The performance and wake characteristics of cross-flow (vertical 
+axis) turbines depend on many parameters: solidity, blade profile, blade pitch, 
+number of blades, strut drag, operational parameters such as tip speed ratio 
+--- and Reynolds number. If numerical models are validated with physical model 
+data that was obtained at insufficiently high Reynolds numbers, it cannot be 
+determined whether problems with model predictions are caused by Reynolds 
+number effects, other turbine parameters or issues with the numerical model.
+%It is uncertain whether numerical models validated with physical model data 
+%obtained at low Reynolds number should be considered validated at all, since 
+%the scale at which the model will be applied is orders of magnitude larger. 
+Experiments were performed with a large laboratory-scale cross flow turbine of 
+to investigate Reynolds number dependence of performance and wake transport 
+characteristics, 
+and to determine what an acceptable Reynolds number mismatch is for scaled 
+physical model tests of cross-flow turbines. It was demonstrated that the 
+performance of the cross-flow turbine becomes essentially $Re$-independent at a 
+Reynolds number based on turbine 
+diameter of $Re_D \approx 10^6$, or an approximate Reynolds number based on 
+blade chord of $Re_c \approx 2 \times 10^5$. 
+A simple model that calculates peak torque coefficient from 
+static foil data and cross-flow turbine kinematics was shown to be a reasonable 
+estimator for predicting Reynolds number dependence of an actual cross flow 
+turbine operating under dynamic conditions.}
 
 % Keywords: add 3 to 10 keywords
-\keyword{keyword; keyword; keyword}
+\keyword{Reynolds number; cross-flow turbine; turbine performance; marine 
+hydrokinetic energy; wind energy}
 
@@
 \begin{document}
-%\linenumbers
+\todo[inline]{Rewrite abstract so it is more of a summary of what was done - 
+tone down the background information a bit. (Note: some background information 
+was moved to Introduction)} 
+% %
+\todo[inline]{Include a List of Symbols?} 
+% %
+
+% %
 \listoftodos
+% %
 
 \section{Introduction}
-
 Scaled physical models are often used in science and engineering to approximate 
-real-world systems. However, the principle of dynamic similarity allows for
+real-world systems. The principle of dynamic similarity allows for
 geometrically scaled systems to be dynamically identical in scale if certain
@@
 the most important parameter is often the Reynolds number, which quantifies the
-importance of viscosity on the flow physics. As such, hereafter we will use the
-word ``scale'' to refer to this dynamical scale rather than the geometric scale.
+importance of viscosity on the flow. As such, hereafter we will use the word
+``scale'' to refer to this dynamical scale rather than the geometric scale.
 
@@
 performance of individual turbines, and design or investigate arrays of devices.
-Despite being significantly less expensive, there is still some concern at which
+Scaled models have the benefit of being significantly less expensive, however, 
+there still is concern at which
 scale an experiment must be performed to be relevant to full-scale application. 
+This is particularly important for cross-flow turbines, where Reynolds number 
+effects for experiments conducted at low $Re$ are often difficult to 
+distinguish from the effects on performance of several other turbine 
+parameters.
 
-For numerical models, one might obtain favorable predictions for scaled systems
-due to modified physics. If the numerical model is then considered
+The performance and wake characteristics of cross-flow (vertical 
+axis) turbines depend on turbine solidity, blade profile (lift/drag, dynamic 
+stall at reduced frequency of turbine rotation, symmetry, thickness, camber), 
+blade pitch, number of blades, strut drag and operational parameters such as 
+tip speed ratio --- {\em and} on Reynolds number. 
+Note that an average blade chord Reynolds number, $Re_{c,\mathrm{avg}} \approx 
+\lambda U_\infty c/ \nu$, can be expressed in terms of tip speed ratio $\lambda 
+\equiv \omega R/ U_\infty$; the value of $\lambda$ at which a turbine reaches 
+peak performance is again a function of turbine solidity $Nc/\pi D$.
+
+If numerical models are validated with physical model data that
+was obtained at insufficiently high Reynolds numbers, it cannot be determined
+whether problems with model predictions are caused by Reynolds number effects, 
+other turbine parameters or issues with the numerical model.
+One might also obtain different predictions of scaled systems
+due to modified physics at lower Reynolds numbers, e.g., laminar boundary 
+layers and transitional flow regimes. If the numerical model is then considered
 ``validated,'' there is a risk that its application at full-scale may produce
-incorrect predictions. For scaled physical prototypes, it is of interest how
+incorrect predictions. It is uncertain whether numerical models validated with 
+physical model data obtained at low Reynolds number should be considered 
+validated at all, since the scale at which the model will be applied is orders 
+of magnitude larger.
+
+This can be overcome by showing that a scaled physical model test has become 
+Reynolds number independent, and therefore validation efforts should be 
+relevant at full-scale.  For scaled physical models, it is of interest how
 small---since size, or scale generally correlates inversely with cost---the
-prototype can be while remaining a reliable predictor of full-scale behavior. At
+model can be while remaining a reliable predictor of full-scale behavior. At
 the very least, it is necessary to test at a scale where results can be
-extrapolated, i.e. changes are occurring quasi-linearly without dramatic changes
-in the underlying physical principles governing the behavior of interest, e.g.
-power production in the case of a wind or MHK turbine.
+extrapolated, i.e., changes are occurring quasi-linearly without dramatic 
+changes in the underlying physical principles governing the behavior of 
+interest, e.g., power production in the case of a wind or MHK turbine.
 
-Blackwell et al. investigated the effects of Reynolds number on the performance
+Blackwell et al. investigated the effects of Reynolds number on the 
+performance
 of a 2 m diameter Darrieus vertical-axis cross-flow wind turbine with NACA 0012
@@
 solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
-maximum radius). More recently, Polagye and Cavagnaro observed high Reynolds
-number sensitivity for a high solidity helical cross-flow turbine
-\cite{Polagye2013b}, and Bravo et al. observed power coefficient of a
+maximum radius). More recently, Bravo et al. observed power coefficient of a
 straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
 \times 10^5$.
+\todo[inline]{Add reference for Bravo et al.} 
+\todo[inline]{Why did you take Polagye and Cavagnaro citation out - not 
+relevant?} 
+\todo[inline]{bring back the examples of studies at small scale? (too small)}
 
-When designing or studying arrays, it is common to use very small
-(geometrically) scaled devices \cite{Chamorro2011, Chamorro2011b}. It is
-therefore important to realize the limitations of evaluating both the power
-output of turbines and the wake recovery when Reynolds number is very far from
-full scale. Sometimes models used are not turbines, but wake generating objects,
-e.g., porous disks, that are meant to replicate the wakes of real turbines
-\cite{Goldenberg1983}. In this case, it is of interest to determine at what
+When designing or studying turbine arrays, it is common to use very small
+(geometrically) scaled devices. 
+% %
+\todo[inline]{Get a reference for this, and a sample geometric scale.} 
+% %
+Sometimes these are not even models of turbines, but
+wake generating objects, e.g., porous disks, that are meant to replicate the
+wakes of real turbines. In this case, it is of interest to determine at what
 scale one might be able to realistically study wake flows in an array, and also
-to evaluate the effectiveness of a wake generator. In other words, a wake
-generator may do a fine job simulating a scaled turbine, but how well can it
-simulate a full-scale device?
+to evaluate the effectiveness of a wake generator. In other words, a scale 
+model wake generator may do a fine job simulating a model turbine of the same 
+scale, but how well do either of them simulate a real turbine?
+\todo[inline]{examples of porous disk papers, e.g. Aubrun et al?}
 
-Previously, is has been observed experimentally that a cross-flow turbine wake's
-unique vortical mean flow field is responsible for accelerated wake recovery
-when compared with conventional axial-flow propeller-type turbines
-\cite{Bachant2015-JoT}. In this study we seek to replicate the same energy
-balance considerations at multiple Reynolds numbers, to examine the implications
-on how small scale, i.e. low Reynolds number experiments may be used to study
-flows in turbine arrays.
+Previously, Bachant and Wosnik \cite{Bachant2015-JoT} showed that a cross-flow
+turbine wake's unique vertical mean flow field is responsible for accelerated
+wake recovery when compared with conventional axial-flow propeller-type
+turbines. In this study we seek to replicate the same energy balance
+considerations at multiple Reynolds numbers, to examine the implications on how
+small scale, i.e. low Reynolds number experiments may be used to study flows in
+turbine arrays.
 
-The Reynolds number effects on the maximum performance and 2-D near-wake
-profiles of the University of New Hampshire's Reference Vertical Axis Turbine
-(UNH-RVAT) were presented \cite{Bachant2014}. In this study we set out to
-analyze in greater depth the effects of scale, i.e. Reynolds number, on the
-performance of the turbine, and its more detailed near-wake characteristics. We
-were looking to find a threshold scale beyond which a physical model study could
-reliably predict full-scale performance. We were also looking to investigate the
-sensitivity as small scales, and what the implications might be for validating
-the prediction capabilities of numerical models.
+Bachant and Wosnik \cite{Bachant2014} observed that maximum performance and 2-D 
+near-wake profiles are affected by Reynolds number. In the present study we set 
+out to analyze in greater depth the effects of scale, i.e. Reynolds number, on 
+the performance of the turbine and its detailed near-wake 
+characteristics. 
+The goal of this study was to find a threshold scale beyond which a physical 
+model study could reliably predict full-scale performance. 
+%We were also looking to investigate the sensitivity as small scales, and what 
+%%%the implications might be for validating
+%the prediction capabilities of numerical models.
+Experiments were performed that investigated Reynolds number dependence for 
+scaled physical model tests of a cross-flow turbine with respect
+to prediction of full-scale performance, and wake characteristics, which will be
+drivers of overall performance and interaction between multiple 
+turbines.
 
-\todo[inline]{Talk about interdependence of Froude number.}
+\todo[inline]{Talk about interdependence of Froude number. whatis teh Froude 
+number at 1 m/s?}
 
@@
 
-In this study we are focused on a turbine constructed from foils---a lift-driven
-device in contrast to drag-driven. A logical starting point for understanding
-the effects of $Re$ on overall turbine loading is to look at simplified cases
-of individual foils, with respect to their static and dynamic loading.
+In this study we are using a simple cross-flow turbine constructed from 
+hydrofoils, i.e., the 
+turbine is a lift-driven device in contrast to drag-driven. A logical starting 
+point for understanding the effects of $Re$ on overall turbine loading is to 
+first consider simplified cases of individual foils, and their static and 
+dynamic loading.
 
@@
 
-Blade performance is generally characterized by the lift-to-drag ratio $l/d$,
-which is a function of the foil's angle of attack $\alpha$. The angles of attack
+Hydro- or airfoil section performance is generally characterized by the 
+lift-to-drag 
+force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ),
+which is primarily a function of the foil's angle of attack $\alpha$. The 
+angles of attack
 for cross-flow turbine blades are a function of turbine tip speed ratio,
@@
 airfoil data, we see that the static stall angle---the angle just beyond that at
-which $l/d$ is maximum---increases with blade chord Reynolds number $Re_c$. This
-in general is caused by the blade boundary becoming more turbulent as a fraction
-of the chord length, diminishes the negative effects of a laminar separation
-bubble occurring near the leading edge. For smooth airfoils, this transition can
-cause a dramatic increase in foil performance, a shown in
+which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$ 
+(reference?). This
+is caused by the boundary layer on the foil becoming more turbulent and more 
+energetic, which enables it to advance further downstream against the adverse 
+pressure gradient on the suction side and delay separation to higher angels of 
+attack, as well as diminish the negative effects of a laminar separation
+bubble occurring near the leading edge (reference?). For smooth airfoils, this 
+transition can
+cause a dramatic increase in foil performance, as seen in
 Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
 $10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
-presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
-reliable data for airfoils in this transitional regime and below. An evaluation
-of the various databases relevant to cross-flow turbines is presented in
-\cite{Bedon2014}.
+presented in \cite{Lissaman1983} and corroborates these trends.
+
+\todo[inline]{Laminar separation bubble - is it always there? alse need 
+reference}
 
@@
 The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
-turbine implies dynamic loading. Furthermore, the blades commonly undergo
+turbine implies dynamic loading. Furthermore, since the range of angle of 
+attack encountered by a rotating blade increases with decreasing tip speed 
+ratio, the blades commonly undergo
 dynamic stall for tip speed ratios at and below those of maximum rotor
@@
 Bousman showed that dynamic stall was insensitive to Reynolds number for $Re=1.0
-\times 10^5$--$2.5 \times 10^5$ \cite{Bousman2000-evaluation}, Though Singleton
-and Yeager Jr. state that Reynolds number effects on dynamics stall remains an
-unsolved question \cite{Singleton2000}.
+\times 10^5$--$2.5 \times 10^5$ \cite{Bousman2000-evaluation}.
+
+\todo[inline]{Bousman: explain in what type of system, with what type of foil 
+this was done.}
+\todo[inline]{you took quite a few references out of this section...?}
+\todo[inline]{additional data in helicopter literature?}
 
@@
 \todo[inline]{What do studies in the literature tell us about $Re$ effects
-on shear flows, wakes, and vortex flows?}
+on shear flows, wakes, and vortex flows? Note-MW I'm looking into Re 
+dependence in general wake studies.}
 
-Chamorro et al. showed that turbulence statistics in an AFT wake became
-$Re$-independent at $Re_D \approx 1 \times 10^5$, with mean velocity slightly
-earlier at $Re_D \approx 5 \times 10^4$ \cite{Chamorro2012}.
-
+Chamorro et al. \cite{Chamorro2012} showed that turbulence statistics in an AFT 
+wake became
+$Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean velocity 
+profiles became $Re$-independent slightly earlier at $Re_D \approx 5 \times 
+10^4$.
+\todo[inline]{What was Chamorro's blade chord Re?}
 \section{Experimental Setup}
@@
 Experiments were performed in the University of New Hampshire's tow/wave tank
-turbine test bed. The turbine model is dubbed the UNH-RVAT, for Reference
-Vertical Axis Turbine, which is designed to be a generic case for numerical
-model testing, similar to the Sandia National Labs/US Department of Energy RM2
-River Turbine, but with higher solidity, or blade chord to radius ratio. A CAD
-model of the turbine, depicted in \ref{fig:turbine}, is available from
+in a turbine test bed specifically designed for cross-flow turbines. The 
+turbine model used in this study was the UNH Reference Vertical Axis Turbine 
+(RVAT), which was designed to be a generic test case for numerical
+model validation, similar to the Sandia National Labs/US Department of Energy 
+RM2
+River Turbine, but with somewhat higher solidity, or blade chord to radius 
+ratio.
+
+\todo[inline]{add reference: V. Neary, M. Previsic, R. Jepsen, M. Lawson, Y. 
+Yu, A. Copping, A. Fontaine, K. Hallett, D. Murray (2014). Methodology for 
+Design and Economic Analysis of Marine Energy Conversion (MEC) Technologies. 
+SAND2014-9040}
+
+The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 
+0.140 m chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and 
+the turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f. 
+Figure~\ref{fig:turbine}. The turbine test bed allows precise control and 
+measurement of the turbine shaft angle and angular velocity, and therefore tip 
+speed ratio, through a permanent magnet servo motor.
+The primary torque measurement was provided by an in-line torque cell, with a 
+redundant measurement by a moment arm and load cell and an indirect measurement 
+through the motor via its torque constant.
+The turbine was mounted with in a low-drag frame constructed from NACA 0020 
+foil sections.  The frame is attached to the tow carriage by four linear 
+bearings, allowing the total streamwise drag force to be measured by a pair of 
+S-beam load cells. Turbine rotor drag is calculated by subtracting the test bed 
+tare drag.
+A schematic of the turbine test bed is shown in 
+Figure~\ref{fig:exp-setup}.
+Additional details of the turbine and experimental setup are described in
+\cite{Bachant2015-JoT}, and a CAD model of the turbine is available from
 \cite{Bachant2014-RVAT-CAD}.
@@
 \includegraphics[width=0.4\textwidth]{figures/turbine}
-\caption{Turbine model drawing. Note that the upper and lower mounting
+\caption{UNH-RVAT turbine model. Turbine blades and struts made from NACA 0020 
+profiles with 0.140 m chord. Note that the upper and lower mounting
 flanges have been excluded as these were included in the tare drag
@@
 
-The turbine was mounted in a frame constructed from NACA 0020 sections. The
-turbine shaft ran up through the water surface, coupled to a permanent magnet
-servo motor with a 20:1 gearbox. This servo was controlled by the tow tank's
-main motion controller for high synchronization with the carriage motion. A
-rotary torque transducer was installed inline between the servo and the turbine,
-and the servo was also mounted on a slewing ring bearing, which allowed a
-redundant measurement of torque via an arm and load cell used to counteract the
-turbine moment. The frame was mounted to the carriage via linear guides, such
-that the total streamwise drag force was transferred to a pair of S-beam load
-cells, providing the rotor drag measurements, after a measured tare drag was
-subtracted in post processing. Similarly, a tare torque was measured by rotating
-the turbine shaft in air. Turbine angular and tow carriage linear position were
-measured using quadrature encoder signals, with $10^5$ counts-per-rev for the
-turbine and 10 $\mathrm{\mu m}$ resolution for the carriage position. These
-signals, along with the torque and drag signals, were sampled at 2 kHz.
-
 \begin{figure}[ht!]
@@
 \includegraphics[width=0.75\textwidth]{figures/exp_setup_drawing}
-\caption{Drawing of the experimental setup.}
+\caption{Experimental setup: turbine test bed installed in UNH tow tank.}
 \label{fig:exp-setup}
@@
 
-Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
-(ADV), sampling at 200 Hz. The probe was mounted on an automated positioning
-system, also controlled by the tow tank's main motion controller. The ADV and
-DAQ systems' sampling times were synchronized by triggering the start of data
-acquisition via a pulse sent from the motion controller.
-
 \subsection{Test Plan}
 
-Approximately 1500 tows were performed, each one used for a single data point on
-either a performance curve or wake map. A performance curve consisted of 31
+Approximately 1500 tows were conducted for the study reported here, each tow 
+was used for a single data 
+point on either a performance curve or wake map. 
+Each performance curve consisted of 31
 tows, where during each tow the mean turbine tip speed ratio was held constant,
@@
 
-A wake map was generated by positioning the Vectrino probe at 270 different
-locations, varied in the cross-stream and vertical directions at one turbine
-diameter downstream ($x/D=1$). These locations have vertical coordinates from
-the turbine centerline up to $z/H=0.625$, ranging in the cross-stream direction
-$y/R = \pm 3$. These locations are shown in Figure~\ref{fig:wake-locations}.
+Each wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
+Velocimeter (ADV), measuring three components of velocity at 200 Hz sampling 
+frequency, at 270 different locations, varied in the cross-stream and
+vertical directions at one turbine diameter downstream ($x/D=1$). These
+locations have vertical coordinates from the turbine centerline up to
+$z/H=0.625$, ranging in the cross-stream direction $y/R = \pm 3$. These
+locations are shown in Figure~\ref{fig:wake-locations}.
 
@@
 
+\todo[inline]{Add a thicker dashed line at $y= -0.5m$, and maybe a dashed 1m 
+circle in the top view of the turbine.}
+
 \subsection{Data Processing}
 
-From each set of tows, a standard time interval was set, which allowed the
-turbine performance and wake to reach a quasi-periodic state. Each run was
+For each set of tows, a standard time interval was set, and each run was
 analyzed to compute statistics over this interval, truncating the end slightly
-to achieve and integer number of blade passages. Turbine shaft angular velocity
-and tow carriage speed were calculated using a second order central differencing
-scheme on the respective position measurements. Power and drag coefficients are
-calculated as instantaneous quantities using the carriage speed as the free
-stream velocity.
+to achieve an integer number of blade passages.
 
-Wake velocity data were filtered for spurious ``spikes'' by removing datapoints
-8 standard deviations or 0.9 m/s from the mean. The experimental data and code
-for processing and visualization are available from
+The experimental data and processing/plotting code are available from
 \cite{Bachant2015-RVAT-Re-dep-data}.
@@
 
-Full power and rotor drag coefficient curves for various Reynolds numbers are
-plotted in Figure~\ref{fig:cp-curves} and Figure~\ref{fig:cd-curves},
-respectively. There is essentially no change in the shape of the drag
-coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
-
-In general, maximum $C_P$ increases with Reynolds number, which is due to an
+Full performance and turbine rotor drag )thrust) curves for various Reynolds 
+numbers are plotted in Figure~\ref{fig:perf-curves}. 
+In general, maximum $C_P$ increases with Reynolds number, from about 0.20 at 
+$Re_D=0.4 \times 10^6$, to 0.26 at higher Reynolds numbers, which is due to an
 increase in the foil lift-to-drag ratio. The power coefficient curves also show
-a downward shift in the optimal tip speed ratio with increasing Reynolds number.
+a slight downward shift in the optimal tip speed ratio (peak performance) with 
+increasing Reynolds 
+number, from about $\lambda \approx$ 2.0 to 1.9.
 This is caused by the stall delay from a more turbulent boundary later on the
 blade suction side.
+There is essentially no change in the shape of the
+drag coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
 
 \begin{figure}[ht]
-\centering
-\includegraphics[width=0.65\textwidth]{figures/cp_curves}
-\caption{Mean power coefficient curves plotted for
-multiple Reynolds numbers.}
-\label{fig:cp-curves}
+\includegraphics[width=0.95\textwidth]{figures/perf_curves}
+\caption{UNH-RVAT experiments: mean power (left) and rotor drag (right) 
+coefficient curves plotted for multiple Reynolds numbers.}
+\label{fig:perf-curves}
 \end{figure}
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.65\textwidth]{figures/cd_curves}
-\caption{Mean rotor drag coefficient curves plotted for
-multiple Reynolds numbers.}
-\label{fig:cd-curves}
-\end{figure}
-
-Mean power and drag coefficients at $\lambda=1.9$ are plotted for all Reynolds
-numbers in Figure~\ref{fig:perf-Re-dep}. There is a drastic improvement in $C_P$
-with increasing Reynolds number for the lower values. Power coefficient then
+Mean power and drag coefficients at $\lambda=1.9$ are plotted vs.\ Reynolds
+number in Figure~\ref{fig:perf-Re-dep}. There is a drastic improvement in $C_P$
+with increasing Reynolds number at lower values of Reynolds number. The power 
+coefficient then
 becomes  essentially $Re$-independent at $Re_D = 0.8 \times 10^6$, which
@@
 promoting reattachment of the laminar separation bubble.
+\todo[inline]{Is there still a laminar separation bubble near the leading edge 
+at these Reynolds numbers?}
 
@@
 \begin{figure}[ht]
-\centering
 \includegraphics[width=0.95\textwidth]{figures/perf_re_dep}
-\caption{Mean power (left) and drag (right) coefficients at $\lambda=1.9$
-plotted versus Reynolds number.} 
+\caption{UNH-RVAT experiments: mean power (left) and rotor drag (right) 
+coefficients at $\lambda=1.9$ plotted versus Reynolds number.} 
 \label{fig:perf-Re-dep}
@@
 XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
-open source turbine design software QBlade. Simulations were run for an angle of
-attack range of 0--40 degrees, in increments of 0.5 degrees. Solver parameters
-used a 100 panels, a fixed speed, zero Mach number, \texttt{NCrit=9} (default
-$\mathrm{e}^n$ transition criteria parameter for an average wind tunnel), and no
-forced boundary layer transition. Characteristics for computed for the
-approximate average blade chord Reynolds numbers encountered in the tow tank
-experiments.
+open source turbine design software QBlade. Simulations were run with the
+default settings and zero Mach number, for the range of blade chord
+Reynolds numbers encountered in the experiments.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
+\includegraphics[width=0.95\textwidth]{figures/all_foils_re_dep}
 \caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
@@
 \end{equation}
-where the blade section chordwise force (for zero preset pitch)
+where the blade section chordwise force (for zero preset pitch) is given by
 \begin{equation}
@@
 
-The relative blade velocity was calculated by adding the free stream to the
-opposite of the blade tangential velocity. Note that this neglects any
-induction, or slowing of the free stream by the turbine forces, which would be
-present in a momentum/streamtube model. Since we were not looking to predict
-absolute performance but rather relative changes with $Re$, this method was
-deemed acceptable as it is extremely simple and fast to compute.
+The relative blade velocity $U_\mathrm{rel}$ was calculated by vector addition 
+of the free 
+stream velocity and the opposite of the blade tangential velocity. Note that 
+this 
+neglects any induction, or slowing of the free stream by the turbine forces, 
+which would be present in a momentum/streamtube model. Since the goal of this 
+approach was not to predict absolute performance but rather gain insight into 
+relative changes with $Re$, 
+this method was deemed acceptable as it is extremely simple and fast to 
+compute.
 
 Values for the blade angle of attack, relative velocity, and torque coefficient
-are plotted in Figure~\ref{fig:blade-kinematics}. The effects of static stall
+are plotted in Figure~\ref{fig:blade-kinematics} for the upstream half of 
+rotation of the turbine. The effects of static stall
 are clearly present in the torque coefficient plot, and by the time the angle of
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/foil_kinematics_ct}
-\caption{Geometric angle of attack, relative velocity and torque coefficient
-calculated with a NACA 0020 foil.}
+\includegraphics[width=0.95\textwidth]{figures/foil_kinematics_ct}
+\caption{Geometric angle of attack (a), relative velocity (b) and torque coefficient
+(c) calculated with a NACA 0020 foil.}
 \label{fig:blade-kinematics}
@@
 
-Results for the peak torque coefficient for each foil are plotted versus $Re_c$
+Results for the normalized peak torque coefficient for each foil are plotted 
+versus $Re_c$
 in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the convergence of
 $C_{T_\mathrm{max}}$ with increasing Reynolds number is more dramatic than any of the
-common foil characteristics plotted in Figure~\ref{fig:foil-Re-dep}, meaning
+common foil performance characteristics plotted in 
+Figure~\ref{fig:foil-Re-dep}, meaning
 that the cross-flow turbine's unique kinematics must be taken into account when
@@
 achieved at lower values, and more dramatically. We see that the trend of the
-NACA 0020 curve matches almost perfectly at low $Re$, but the trend to continue
-increasing linearly is not matched by the experimental data, which looks to
-behave more like the cambered foil results. Though this method is very
+(non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1 
+\times 10^5$, but then continues to
+increase linearly with a small slope with $Re$. This is not matched by the 
+experimental data, which at higher $Re$ looks more like the cambered foil 
+results. Though this method is very
 approximate, it nevertheless appears to be a reasonable way to predict the
@@
 \includegraphics[width=0.65\textwidth]{figures/cft_re_dep_foils}
-\caption{Reynolds number dependence of the peak torque coefficient calculated using 
-only static foil coefficients and blade kinematics.}
+\caption{Reynolds number dependence of the normalized peak torque coefficient 
+calculated from static foil coefficients and blade kinematics, compared to 
+experimental data from a cross-flow turbine.}
 \label{fig:foils-C_T-Re-dep}
@@
 
-The wake of this turbine at $x/D=1$, $\lambda=1.9$, $Re_D = 1.0 \times 10^6$ was
-described thoroughly in \cite{Bachant2015-JoT}. Similar data were taken for the
-experiment here, the results for the mean velocity field and turbulence kinetic
-energy are shown in Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont},
-respectively. With respect to the mean velocity field, we see asymmetry and a
-mean vortex structure created by the blade tip vortex shedding. The effects of
+The near-wake of this turbine at $x/D=1$, $\lambda=1.9$, and momentum and 
+kinetic energy balances at a $Re_D = 1.0 \times 
+10^6$ were discussed in \cite{Bachant2015-JoT}. Similar data were taken for the
+experiment here, and results for the mean velocity field and turbulence kinetic
+energy calculated from wake maps of 270 individual measurements (tows) are 
+shown in Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont},
+respectively, looking upstream towards the turbine. With respect to the mean 
+velocity field, we see asymmetry and a
+mean vortex structure created by blade tip vortex shedding. The effects of
 the tip vortices are also seen in the turbulence kinetic energy measurements,
@@
 
-These same wake maps were taken for $Re_D = 0.4 \times 10^6$, $0.6 \times 10^6$,
-$0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look strikingly
+These same wake maps were measured for $Re_D = 0.4 \time 10^6$, $0.6 \times 
+10^6$,
+$0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these wake maps look 
+strikingly
 similar, so they have not been plotted here. We will instead compare and
 contrast the wake behavior by examining spectra and wake transport terms in the
-equations that govern the downstream evolution.
+equations that govern the downstream evolution of mean momentum and kinetic 
+energy.
 
@@
 \centering
-\includegraphics[width=0.85\textwidth]{figures/wake_spectra}
-\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$, 
-$y/R=-1.0$ (a) and $y/R=1.5$ (b). Dashed vertical
+\includegraphics[width=0.95\textwidth]{figures/wake_spectra}
+\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$ 
+and $y/R=-1.0$ (or: -0.5 m in Figure~\ref{fig:wake-locations}), and  $y/R=1.5$ 
+(or: +0.75 m in Figure~\ref{fig:wake-locations} (b). Dashed vertical
 lines indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}
@@
 three orders of magnitude smaller than the other terms, do increase at low
-Reynolds numbers, which is consistent with the definition of the Reynolds number
+Reynolds numbers, which is consistent with the physical meaning of the Reynolds 
+number
 in general.
@@
 
-Totals for all the wake transport terms calculated are plotted in
-Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport is
+\todo[inline]{Any additional insight we can gainfrom these figures?}
+
+The totals for all the wake transport terms calculated in 
+Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph} are plotted in
+Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport of 
+both mean streamwise momentum and kinetic energy is
 enhanced at lower Reynolds numbers and levels off consistent with the behavior
@@
 studying sub-scale models of turbine arrays, where increased levels of wake
-recovery will motivate different ideal array configurations when compared with
-full-scale turbines.
+recovery could motivate different ideal array configurations when compared with
+full-scale turbines if the scale model Reynolds number is too low.
 
@@
 
-In this study we determined that, as previously shown, performance for a
-high-solidity cross-flow turbine constructed from straight NACA 0020 blades
-becomes essentially $Re$-independent at a turbine diameter Reynolds number $Re_D
-\approx 10^6$, or an approximate blade chord Reynolds number $Re_c \approx 2
-\times 10^5$.
+In this study it was demonstrated that the performance of a cross-flow turbine
+becomes essentially $Re$-independent at a Reynolds number based on turbine 
+diameter of $Re_D \approx 10^6$, or an approximate Reynolds number based on 
+blade chord of $Re_c \approx 2 \times 10^5$.
+
+\todo[inline]{Should we discuss that these results depend on turbine type, and 
+solidity? Lower solidity turbines may only achieve a "linear" regime, not a 
+truly Re-indept regime. Additional issues such as blade termination/strength of 
+tip vortices?}
+
+The Reynolds number dependence of the peak torque coefficient calculated from 
+static foil data and cross-flow turbine kinematics was shown to be a reasonable 
+indicator for predicting Reynolds number dependence of an actual cross flow 
+turbine operating under dynamic conditions, including dynamic stall. The simple 
+peak torque coefficient model also predicted Reynolds number independence for  
+Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$.
 
 These results suggest that to validate predictive engineering models, one must
-use data of at least this scale, especially high-fidelity CFD models, where
-transition to turbulence may be completely different between scaled physical
-model and prototype.
-
-We propose a method for predicting Reynolds number dependence of a cross-flow
-turbine using static airfoil characteristics and turbine blade kinematics. It is
-shown that the peak torque coefficient computed this way shows similar Reynolds
-number sensitivity---a better predictor than conventional quantifications of
-airfoil performance, e.g., lift-to-drag ratio. Is it also shown from this metric
-that foils with higher camber are more Reynolds number dependent but have
-smaller slopes in the linear regime.
+use data of at least this scale, especially for validation of high-fidelity CFD 
+models, where transition to turbulence may be completely different between 
+scaled physical model and full-scale prototype.
 
 If using scaled physical models to predict array performance, it is important to
-keep all turbines in the linear regime to avoid exaggerated power deficiencies
+keep all turbines in the regime where the power coefficient varies linearly to 
+avoid exaggerated power deficiencies
 for downstream turbines, despite similarities in wake characteristics. These

### 2015-10-10 paper.tex

@@
 \begin{document}
+\linenumbers
+\listoftodos
+
 \todo[inline]{Rewrite abstract so it is more of a summary of what was done -
@@
 \todo[inline]{Include a List of Symbols?}
-% %
-
-% %
-\listoftodos
-% %
 
 \section{Introduction}
+
 Scaled physical models are often used in science and engineering to approximate
-real-world systems. The principle of dynamic similarity allows for
-geometrically scaled systems to be dynamically identical in scale if certain
-nondimensional physical parameters are matched. In the case of fluid systems,
-the most important parameter is often the Reynolds number, which quantifies the
-importance of viscosity on the flow. As such, hereafter we will use the word
-``scale'' to refer to this dynamical scale rather than the geometric scale.
+real-world systems. The principle of dynamic similarity allows for geometrically
+scaled systems to be dynamically identical in scale if certain nondimensional
+physical parameters are matched. In the case of fluid systems, the most
+important nondimensional parameter is often the Reynolds number, which
+quantifies the importance of inertial forces over viscous forces on the flow
+physics. It should be noted that in general the geometric scale of a physical
+model is not necessarily the same as its dynamical scale. As such, hereafter we
+will use the word ``scale'' to refer to this dynamical scale rather than the
+geometric scale.
+
+\todo[inline]{Kinematic scaling effects? Matching TSR?}
 
@@
 
-When designing or studying turbine arrays, it is common to use very small
-(geometrically) scaled devices. 
-% %
-\todo[inline]{Get a reference for this, and a sample geometric scale.} 
-% %
-Sometimes these are not even models of turbines, but
-wake generating objects, e.g., porous disks, that are meant to replicate the
-wakes of real turbines. In this case, it is of interest to determine at what
+When designing or studying arrays, it is common to use very small
+(geometrically) scaled devices \cite{Chamorro2011, Chamorro2011b}. It is
+therefore important to realize the limitations of evaluating both the power
+output of turbines and the wake recovery when Reynolds number is very far from
+full scale. Sometimes models used are not turbines, but wake generating objects,
+e.g., porous disks, that are meant to replicate the wakes of real turbines
+\cite{Goldenberg1983}. In this case, it is of interest to determine at what
 scale one might be able to realistically study wake flows in an array, and also
-to evaluate the effectiveness of a wake generator. In other words, a scale 
-model wake generator may do a fine job simulating a model turbine of the same 
-scale, but how well do either of them simulate a real turbine?
-\todo[inline]{examples of porous disk papers, e.g. Aubrun et al?}
+to evaluate the effectiveness of a wake generator. In other words, a wake
+generator may do a fine job simulating a scaled turbine, but how well can it
+simulate a full-scale device?
 
-Previously, Bachant and Wosnik \cite{Bachant2015-JoT} showed that a cross-flow
-turbine wake's unique vertical mean flow field is responsible for accelerated
-wake recovery when compared with conventional axial-flow propeller-type
-turbines. In this study we seek to replicate the same energy balance
-considerations at multiple Reynolds numbers, to examine the implications on how
-small scale, i.e. low Reynolds number experiments may be used to study flows in
-turbine arrays.
+Previously, is has been observed experimentally that a cross-flow turbine wake's
+unique vortical mean flow field is responsible for accelerated wake recovery
+when compared with conventional axial-flow propeller-type turbines
+\cite{Bachant2015-JoT}. In this study we seek to replicate the same energy
+balance considerations at multiple Reynolds numbers, to examine the implications
+on how small scale, i.e. low Reynolds number experiments may be used to study
+flows in turbine arrays.
 
-Bachant and Wosnik \cite{Bachant2014} observed that maximum performance and 2-D 
-near-wake profiles are affected by Reynolds number. In the present study we set 
-out to analyze in greater depth the effects of scale, i.e. Reynolds number, on 
-the performance of the turbine and its detailed near-wake 
-characteristics. 
-The goal of this study was to find a threshold scale beyond which a physical 
-model study could reliably predict full-scale performance. 
-%We were also looking to investigate the sensitivity as small scales, and what 
-%%%the implications might be for validating
-%the prediction capabilities of numerical models.
-Experiments were performed that investigated Reynolds number dependence for 
-scaled physical model tests of a cross-flow turbine with respect
-to prediction of full-scale performance, and wake characteristics, which will be
-drivers of overall performance and interaction between multiple 
-turbines.
+The Reynolds number effects on the maximum performance and 2-D near-wake
+profiles of the University of New Hampshire's Reference Vertical Axis Turbine
+(UNH-RVAT) were presented \cite{Bachant2014}. In this study we set out to
+analyze in greater depth the effects of scale, i.e. Reynolds number, on the
+performance of the turbine, and its more detailed near-wake characteristics. We
+were looking to find a threshold scale beyond which a physical model study could
+reliably predict full-scale performance. We were also looking to investigate the
+sensitivity as small scales, and what the implications might be for validating
+the prediction capabilities of numerical models.
 
-\todo[inline]{Talk about interdependence of Froude number. whatis teh Froude 
+\todo[inline]{Talk about interdependence of Froude number. What is the Froude
 number at 1 m/s?}
@@
 $10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
-presented in \cite{Lissaman1983} and corroborates these trends.
+presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
+reliable data for airfoils in this transitional regime and below. An evaluation
+of the various databases relevant to cross-flow turbines is presented in
+\cite{Bedon2014}.
 
-\todo[inline]{Laminar separation bubble - is it always there? alse need 
+\todo[inline]{Laminar separation bubble - is it always there? also need
 reference}
@@
 
-Experiments were performed in the University of New Hampshire's tow/wave tank
-in a turbine test bed specifically designed for cross-flow turbines. The 
+Experiments were performed in a turbine test bed specifically designd for cross-flow turbines. The
 turbine model used in this study was the UNH Reference Vertical Axis Turbine
-(RVAT), which was designed to be a generic test case for numerical
-model validation, similar to the Sandia National Labs/US Department of Energy 
-RM2
-River Turbine, but with somewhat higher solidity, or blade chord to radius 
-ratio.
-
-\todo[inline]{add reference: V. Neary, M. Previsic, R. Jepsen, M. Lawson, Y. 
-Yu, A. Copping, A. Fontaine, K. Hallett, D. Murray (2014). Methodology for 
-Design and Economic Analysis of Marine Energy Conversion (MEC) Technologies. 
-SAND2014-9040}
+(RVAT), which was designed to be a generic case for numerical
+model testing, similar to the Sandia National Labs/US Department of Energy RM2
+River Turbine, but with higher solidity, or blade chord to radius ratio.
 
@@
 
+\todo[inline]{add reference: V. Neary, M. Previsic, R. Jepsen, M. Lawson, Y.
+Yu, A. Copping, A. Fontaine, K. Hallett, D. Murray (2014). Methodology for
+Design and Economic Analysis of Marine Energy Conversion (MEC) Technologies.
+SAND2014-9040}
+
 \begin{figure}[ht!]
@@
 
-For each set of tows, a standard time interval was set, and each run was
+From each set of tows, a standard time interval was set, which allowed the
+turbine performance and wake to reach a quasi-periodic state. Each run was
 analyzed to compute statistics over this interval, truncating the end slightly
-to achieve an integer number of blade passages.
+to achieve and integer number of blade passages. Turbine shaft angular velocity
+and tow carriage speed were calculated using a second order central differencing
+scheme on the respective position measurements. Power and drag coefficients are
+calculated as instantaneous quantities using the carriage speed as the free
+stream velocity.
 
-The experimental data and processing/plotting code are available from
+Wake velocity data were filtered for spurious ``spikes'' by removing datapoints
+8 standard deviations or 0.9 m/s from the mean. The experimental data and code
+for processing and visualization are available from
 \cite{Bachant2015-RVAT-Re-dep-data}.
@@
 
-Full performance and turbine rotor drag )thrust) curves for various Reynolds 
-numbers are plotted in Figure~\ref{fig:perf-curves}. 
-In general, maximum $C_P$ increases with Reynolds number, from about 0.20 at 
-$Re_D=0.4 \times 10^6$, to 0.26 at higher Reynolds numbers, which is due to an
+Full power and rotor drag (a.k.a. thrust) coefficient curves for various Reynolds numbers are
+plotted in Figure~\ref{fig:cp-curves} and Figure~\ref{fig:cd-curves},
+respectively. There is essentially no change in the shape of the drag
+coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
+
+In general, maximum $C_P$ increases with Reynolds number, which is due to an
 increase in the foil lift-to-drag ratio. The power coefficient curves also show
@@
 \begin{figure}[ht]
-\includegraphics[width=0.95\textwidth]{figures/perf_curves}
-\caption{UNH-RVAT experiments: mean power (left) and rotor drag (right) 
-coefficient curves plotted for multiple Reynolds numbers.}
-\label{fig:perf-curves}
+\centering
+\includegraphics[width=0.65\textwidth]{figures/cp_curves}
+\caption{Mean power coefficient curves plotted for
+multiple Reynolds numbers.}
+\label{fig:cp-curves}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.65\textwidth]{figures/cd_curves}
+\caption{Mean rotor drag coefficient curves plotted for
+multiple Reynolds numbers.}
+\label{fig:cd-curves}
 \end{figure}
@@
 \begin{figure}[ht]
-\includegraphics[width=0.95\textwidth]{figures/perf_re_dep}
-\caption{UNH-RVAT experiments: mean power (left) and rotor drag (right) 
-coefficients at $\lambda=1.9$ plotted versus Reynolds number.} 
+\centering
+\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
+\caption{UNH-RVAT measured mean power (left) and drag (right) coefficients at $\lambda=1.9$
+plotted versus Reynolds number.}
 \label{fig:perf-Re-dep}
@@
 XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
-open source turbine design software QBlade. Simulations were run with the
-default settings and zero Mach number, for the range of blade chord
-Reynolds numbers encountered in the experiments.
+open source turbine design software QBlade. Simulations were run for an angle of
+attack range of 0--40 degrees, in increments of 0.5 degrees. Solver parameters
+used a 100 panels, a fixed speed, zero Mach number, \texttt{NCrit=9} (default
+$\mathrm{e}^n$ transition criteria parameter for an average wind tunnel), and no
+forced boundary layer transition. Characteristics for computed for the
+approximate average blade chord Reynolds numbers encountered in the tow tank
+experiments.
 
@@
 
-These same wake maps were measured for $Re_D = 0.4 \time 10^6$, $0.6 \times 
-10^6$,
-$0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these wake maps look 
-strikingly
+These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times 10^6$,
+$0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look strikingly
 similar, so they have not been plotted here. We will instead compare and
@@
 \centering
-\includegraphics[width=0.95\textwidth]{figures/wake_spectra}
-\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$ 
-and $y/R=-1.0$ (or: -0.5 m in Figure~\ref{fig:wake-locations}), and  $y/R=1.5$ 
-(or: +0.75 m in Figure~\ref{fig:wake-locations} (b). Dashed vertical
+\includegraphics[width=0.85\textwidth]{figures/wake_spectra}
+\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$,
+$y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (a) and $y/R=1.5$ ($+0.75$ m in Figure~\ref{fig:wake-locations}) (b). Dashed vertical
 lines indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}
@@
 
-The Reynolds number dependence of the peak torque coefficient calculated from 
-static foil data and cross-flow turbine kinematics was shown to be a reasonable 
-indicator for predicting Reynolds number dependence of an actual cross flow 
-turbine operating under dynamic conditions, including dynamic stall. The simple 
-peak torque coefficient model also predicted Reynolds number independence for  
-Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$.
-
 These results suggest that to validate predictive engineering models, one must
@@
 scaled physical model and full-scale prototype.
+
+We propose a method for predicting Reynolds number dependence of a cross-flow
+turbine using static airfoil characteristics and turbine blade kinematics. It is
+shown that the peak torque coefficient computed this way shows similar Reynolds
+number sensitivity---a better predictor than conventional quantifications of
+airfoil performance, e.g., lift-to-drag ratio.
+The simple peak geometric torque coefficient model predicted Reynolds number independence for
+Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$.
+Is it also shown from this metric
+that foils with higher camber are more Reynolds number dependent but have
+smaller slopes in the linear regime. The Reynolds number dependence of the peak torque coefficient calculated from
+static foil data and cross-flow turbine kinematics was shown to be a reasonable
+indicator for predicting Reynolds number dependence of an actual cross-flow
+turbine operating under dynamic conditions, including dynamic stall.
 

### 2015-10-10 paper.tex
Remove erroneous comment
Text wrapping

@@
 \todo[inline]{Add reference for Bravo et al.}
-\todo[inline]{Why did you take Polagye and Cavagnaro citation out - not
-relevant?}
+
 \todo[inline]{bring back the examples of studies at small scale? (too small)}
@@
 
-Experiments were performed in a turbine test bed specifically designd for cross-flow turbines. The
-turbine model used in this study was the UNH Reference Vertical Axis Turbine
-(RVAT), which was designed to be a generic case for numerical
-model testing, similar to the Sandia National Labs/US Department of Energy RM2
-River Turbine, but with higher solidity, or blade chord to radius ratio.
+Experiments were performed in a turbine test bed specifically designd for
+cross-flow turbines. The turbine model used in this study was the UNH Reference
+Vertical Axis Turbine (RVAT), which was designed to be a generic case for
+numerical model testing, similar to the Sandia National Labs/US Department of
+Energy RM2 River Turbine, but with higher solidity, or blade chord to radius
+ratio.
 
-The UNH-RVAT turbine has three blades made from NACA 0020 profiles with
-0.140 m chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and
-the turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
+The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.140 m
+chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and the
+turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
 Figure~\ref{fig:turbine}. The turbine test bed allows precise control and
 measurement of the turbine shaft angle and angular velocity, and therefore tip
-speed ratio, through a permanent magnet servo motor.
-The primary torque measurement was provided by an in-line torque cell, with a
-redundant measurement by a moment arm and load cell and an indirect measurement
-through the motor via its torque constant.
-The turbine was mounted with in a low-drag frame constructed from NACA 0020
-foil sections.  The frame is attached to the tow carriage by four linear
-bearings, allowing the total streamwise drag force to be measured by a pair of
-S-beam load cells. Turbine rotor drag is calculated by subtracting the test bed
-tare drag.
-A schematic of the turbine test bed is shown in
-Figure~\ref{fig:exp-setup}.
-Additional details of the turbine and experimental setup are described in
-\cite{Bachant2015-JoT}, and a CAD model of the turbine is available from
-\cite{Bachant2014-RVAT-CAD}.
+speed ratio, through a permanent magnet servo motor. The primary torque
+measurement was provided by an in-line torque cell, with a redundant measurement
+by a moment arm and load cell and an indirect measurement through the motor via
+its torque constant. The turbine was mounted with in a low-drag frame
+constructed from NACA 0020 foil sections.  The frame is attached to the tow
+carriage by four linear bearings, allowing the total streamwise drag force to be
+measured by a pair of S-beam load cells. Turbine rotor drag is calculated by
+subtracting the test bed tare drag. A schematic of the turbine test bed is shown
+in Figure~\ref{fig:exp-setup}. Additional details of the turbine and
+experimental setup are described in \cite{Bachant2015-JoT}, and a CAD model of
+the turbine is available from \cite{Bachant2014-RVAT-CAD}.
 
@@
 
-Approximately 1500 tows were conducted for the study reported here, each tow
-was used for a single data
-point on either a performance curve or wake map.
-Each performance curve consisted of 31
-tows, where during each tow the mean turbine tip speed ratio was held constant,
-ranging from 0.1--3.1 in 0.1 increments. Full performance curve data were
-acquired for tow speeds from 0.4 to 1.2 m/s in 0.2 m/s increments, for which
-turbine diameter and approximate blade chord Reynolds number are presented in
-Table~\ref{tab:Re}. Performance was also measured for $\lambda=1.9$ at tow
-speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows each.
+Approximately 1500 tows were conducted for the study reported here, each tow was
+used for a single data point on either a performance curve or wake map. Each
+performance curve consisted of 31 tows, where during each tow the mean turbine
+tip speed ratio was held constant, ranging from 0.1--3.1 in 0.1 increments. Full
+performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2
+m/s increments, for which turbine diameter and approximate blade chord Reynolds
+number are presented in Table~\ref{tab:Re}. Performance was also measured for
+$\lambda=1.9$ at tow speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows
+each.
 
@@
 Velocimeter (ADV), measuring three components of velocity at 200 Hz sampling
-frequency, at 270 different locations, varied in the cross-stream and
-vertical directions at one turbine diameter downstream ($x/D=1$). These
-locations have vertical coordinates from the turbine centerline up to
-$z/H=0.625$, ranging in the cross-stream direction $y/R = \pm 3$. These
-locations are shown in Figure~\ref{fig:wake-locations}.
+frequency, at 270 different locations, varied in the cross-stream and vertical
+directions at one turbine diameter downstream ($x/D=1$). These locations have
+vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
+the cross-stream direction $y/R = \pm 3$. These locations are shown in
+Figure~\ref{fig:wake-locations}.
 
@@
 to achieve and integer number of blade passages. Turbine shaft angular velocity
-and tow carriage speed were calculated using a second order central differencing
-scheme on the respective position measurements. Power and drag coefficients are
-calculated as instantaneous quantities using the carriage speed as the free
-stream velocity.
+and tow carriage speed were calculated using as second order central
+differencing scheme on the respective position measurements. Power and drag
+coefficients are calculated as instantaneous quantities using the carriage speed
+as the free stream velocity.
 
@@
 
-Full power and rotor drag (a.k.a. thrust) coefficient curves for various Reynolds numbers are
-plotted in Figure~\ref{fig:cp-curves} and Figure~\ref{fig:cd-curves},
-respectively. There is essentially no change in the shape of the drag
-coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
+Full power and rotor drag (a.k.a. thrust) coefficient curves for various
+Reynolds numbers are plotted in Figure~\ref{fig:cp-curves} and
+Figure~\ref{fig:cd-curves}, respectively. There is essentially no change in the
+shape of the drag coefficient curves---merely an upward shift in $C_D$ with
+increasing $Re$.
 
@@
 a slight downward shift in the optimal tip speed ratio (peak performance) with
-increasing Reynolds
-number, from about $\lambda \approx$ 2.0 to 1.9.
-This is caused by the stall delay from a more turbulent boundary later on the
-blade suction side.
-There is essentially no change in the shape of the
-drag coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
+increasing Reynolds number, from about $\lambda \approx$ 2.0 to 1.9. This is
+caused by the stall delay from a more turbulent boundary later on the blade
+suction side. There is essentially no change in the shape of the drag
+coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
 
@@
 with increasing Reynolds number at lower values of Reynolds number. The power
-coefficient then
-becomes  essentially $Re$-independent at $Re_D = 0.8 \times 10^6$, which
-corresponds to an approximate average blade chord Reynolds number $Re_{c,
-\mathrm{ave}} = 2.1 \times 10^5$. This threshold is consistent with the behavior
-of the blade boundary layer transitioning from laminar to turbulent, thereby
-promoting reattachment of the laminar separation bubble.
+coefficient then becomes  essentially $Re$-independent at $Re_D = 0.8 \times
+10^6$, which corresponds to an approximate average blade chord Reynolds number
+$Re_{c, \mathrm{ave}} = 2.1 \times 10^5$. This threshold is consistent with the
+behavior of the blade boundary layer transitioning from laminar to turbulent,
+thereby promoting reattachment of the laminar separation bubble.
 \todo[inline]{Is there still a laminar separation bubble near the leading edge
@@
 \centering
+
 \includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
-\caption{UNH-RVAT measured mean power (left) and drag (right) coefficients at $\lambda=1.9$
-plotted versus Reynolds number.}
+
+\caption{UNH-RVAT measured mean power (left) and drag (right) coefficients at
+    $\lambda=1.9$ plotted versus Reynolds number.}
+
 \label{fig:perf-Re-dep}
@@
 The relative blade velocity $U_\mathrm{rel}$ was calculated by vector addition
-of the free
-stream velocity and the opposite of the blade tangential velocity. Note that
-this
-neglects any induction, or slowing of the free stream by the turbine forces,
-which would be present in a momentum/streamtube model. Since the goal of this
-approach was not to predict absolute performance but rather gain insight into
-relative changes with $Re$,
-this method was deemed acceptable as it is extremely simple and fast to
-compute.
+of the free stream velocity and the opposite of the blade tangential velocity.
+Note that this neglects any induction, or slowing of the free stream by the
+turbine forces, which would be present in a momentum/streamtube model. Since the
+goal of this approach was not to predict absolute performance but rather gain
+insight into relative changes with $Re$, this method was deemed acceptable as it
+is extremely simple and fast to compute.
 
@@
 are plotted in Figure~\ref{fig:blade-kinematics} for the upstream half of
-rotation of the turbine. The effects of static stall
-are clearly present in the torque coefficient plot, and by the time the angle of
-attack has decreased below that of stall, the relative velocity is so low that
-there is not much contribution to the torque coefficient.
+rotation of the turbine. The effects of static stall are clearly present in the
+torque coefficient plot, and by the time the angle of attack has decreased below
+that of stall, the relative velocity is so low that there is not much
+contribution to the torque coefficient.
 
@@
 Results for the normalized peak torque coefficient for each foil are plotted
-versus $Re_c$
-in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the convergence of
-$C_{T_\mathrm{max}}$ with increasing Reynolds number is more dramatic than any of the
-common foil performance characteristics plotted in
-Figure~\ref{fig:foil-Re-dep}, meaning
-that the cross-flow turbine's unique kinematics must be taken into account when
-attempting to predict the effect of transitional Reynolds numbers on turbine
-performance.
+versus $Re_c$ in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the
+convergence of $C_{T_\mathrm{max}}$ with increasing Reynolds number is more
+dramatic than any of the common foil performance characteristics plotted in
+Figure~\ref{fig:foil-Re-dep}, meaning that the cross-flow turbine's unique
+kinematics must be taken into account when attempting to predict the effect of
+transitional Reynolds numbers on turbine performance.
 
@@
 (non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
-\times 10^5$, but then continues to
-increase linearly with a small slope with $Re$. This is not matched by the
-experimental data, which at higher $Re$ looks more like the cambered foil
-results. Though this method is very
-approximate, it nevertheless appears to be a reasonable way to predict the
-transitional Reynolds number regime for cross-flow turbine performance using
-only 2-D static airfoil characteristics
+\times 10^5$, but then continues to increase linearly with a small slope with
+$Re$. This is not matched by the experimental data, which at higher $Re$ looks
+more like the cambered foil results. Though this method is very approximate, it
+nevertheless appears to be a reasonable way to predict the transitional Reynolds
+number regime for cross-flow turbine performance using only 2-D static airfoil
+characteristics
 
@@
 The near-wake of this turbine at $x/D=1$, $\lambda=1.9$, and momentum and
-kinetic energy balances at a $Re_D = 1.0 \times
-10^6$ were discussed in \cite{Bachant2015-JoT}. Similar data were taken for the
-experiment here, and results for the mean velocity field and turbulence kinetic
-energy calculated from wake maps of 270 individual measurements (tows) are
-shown in Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont},
-respectively, looking upstream towards the turbine. With respect to the mean
-velocity field, we see asymmetry and a
-mean vortex structure created by blade tip vortex shedding. The effects of
-the tip vortices are also seen in the turbulence kinetic energy measurements,
-along with turbulence generated by the blades undergoing dynamic stall on the
-$-y$ side of the turbine.
+kinetic energy balances at a $Re_D = 1.0 \times 10^6$ were discussed in
+\cite{Bachant2015-JoT}. Similar data were taken for the experiment here, and
+results for the mean velocity field and turbulence kinetic energy calculated
+from wake maps of 270 individual measurements (tows) are shown in
+Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont}, respectively, looking
+upstream towards the turbine. With respect to the mean velocity field, we see
+asymmetry and a mean vortex structure created by blade tip vortex shedding. The
+effects of the tip vortices are also seen in the turbulence kinetic energy
+measurements, along with turbulence generated by the blades undergoing dynamic
+stall on the $-y$ side of the turbine.
 
-These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times 10^6$,
-$0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look strikingly
-similar, so they have not been plotted here. We will instead compare and
-contrast the wake behavior by examining spectra and wake transport terms in the
-equations that govern the downstream evolution of mean momentum and kinetic
+These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times
+10^6$, $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look
+strikingly similar, so they have not been plotted here. We will instead compare
+and contrast the wake behavior by examining spectra and wake transport terms in
+the equations that govern the downstream evolution of mean momentum and kinetic
 energy.
@@
 \centering
+
 \includegraphics[width=0.85\textwidth]{figures/wake_spectra}
+
 \caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$,
-$y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (a) and $y/R=1.5$ ($+0.75$ m in Figure~\ref{fig:wake-locations}) (b). Dashed vertical
-lines indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}
+    $y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (a) and $y/R=1.5$
+    ($+0.75$ m in Figure~\ref{fig:wake-locations}) (b). Dashed vertical lines
+    indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}
+
 \label{fig:wake-spectra}
@@
 Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport of
-both mean streamwise momentum and kinetic energy is
-enhanced at lower Reynolds numbers and levels off consistent with the behavior
-of the  turbine power coefficient. This is an important consideration if
-studying sub-scale models of turbine arrays, where increased levels of wake
-recovery could motivate different ideal array configurations when compared with
-full-scale turbines if the scale model Reynolds number is too low.
+both mean streamwise momentum and kinetic energy is enhanced at lower Reynolds
+numbers and levels off consistent with the behavior of the  turbine power
+coefficient. This is an important consideration if studying sub-scale models of
+turbine arrays, where increased levels of wake recovery could motivate different
+ideal array configurations when compared with full-scale turbines if the scale
+model Reynolds number is too low.
 
@@
 number sensitivity---a better predictor than conventional quantifications of
-airfoil performance, e.g., lift-to-drag ratio.
-The simple peak geometric torque coefficient model predicted Reynolds number independence for
-Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$.
-Is it also shown from this metric
-that foils with higher camber are more Reynolds number dependent but have
-smaller slopes in the linear regime. The Reynolds number dependence of the peak torque coefficient calculated from
-static foil data and cross-flow turbine kinematics was shown to be a reasonable
-indicator for predicting Reynolds number dependence of an actual cross-flow
-turbine operating under dynamic conditions, including dynamic stall.
+airfoil performance, e.g., lift-to-drag ratio. The simple peak geometric torque
+coefficient model predicted Reynolds number independence for Reynolds numbers
+based on blade chord of $Re_c > 2 \times 10^5$. Is it also shown from this
+metric that foils with higher camber are more Reynolds number dependent but have
+smaller slopes in the linear regime. The Reynolds number dependence of the peak
+torque coefficient calculated from static foil data and cross-flow turbine
+kinematics was shown to be a reasonable indicator for predicting Reynolds number
+dependence of an actual cross-flow turbine operating under dynamic conditions,
+including dynamic stall.
 
@@
 keep all turbines in the regime where the power coefficient varies linearly to
-avoid exaggerated power deficiencies
-for downstream turbines, despite similarities in wake characteristics. These
-results also suggest that low Reynolds number physical model studies of turbine
-arrays may see exaggerated levels of wake recovery, leading to inadequate or
-inappropriate spacing or layouts.
+avoid exaggerated power deficiencies for downstream turbines, despite
+similarities in wake characteristics. These results also suggest that low
+Reynolds number physical model studies of turbine arrays may see exaggerated
+levels of wake recovery, leading to inadequate or inappropriate spacing or
+layouts.
 

### 2015-10-11 paper.tex
Add todo item
Turn off line numbers until submission time
Add 'in'

@@
 \begin{document}
-\linenumbers
+%\linenumbers
 \listoftodos
@@
 profiles of the University of New Hampshire's Reference Vertical Axis Turbine
-(UNH-RVAT) were presented \cite{Bachant2014}. In this study we set out to
+(UNH-RVAT) were presented in \cite{Bachant2014}. In this study we set out to
 analyze in greater depth the effects of scale, i.e. Reynolds number, on the
@@
 \centering
+
 \includegraphics[width=0.65\textwidth]{figures/cft_re_dep_foils}
+
 \caption{Reynolds number dependence of the normalized peak torque coefficient
@@
 experimental data from a cross-flow turbine.}
+
 \label{fig:foils-C_T-Re-dep}
 \end{figure}
+
+\todo[inline]{Is this the peak torque coefficient or the mean power coefficient?}
 

### 2015-10-11 paper.tex
Minor wording
Add references for AFT performance Re-independence; resolves #16

@@
 straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
-\times 10^5$.
-\todo[inline]{Add reference for Bravo et al.}
+\times 10^5$ \cite{Bravo2007}.
+
+Krogstad and Adaramola observed Reynolds number independence of the performance for a 0.9 m
+diameter axial-flow turbine rotor at $Re_D \approx 5 \times 10^5$ in wind tunnel
+tests \cite{Krogstad2012a}.
+Their turbine blades had an S826 profile along their entire span, a profile chosen for its $Re$-independence.
+Walker et al.
+\cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow turbine tests in a towing tank, which were performed at a mid span Reynolds number $Re_c = 4 \times 10^5$.
 
@@
 profiles of the University of New Hampshire's Reference Vertical Axis Turbine
-(UNH-RVAT) were presented in \cite{Bachant2014}. In this study we set out to
+(UNH-RVAT) were presented in \cite{Bachant2014}. Here we set out to
 analyze in greater depth the effects of scale, i.e. Reynolds number, on the

### 2015-10-11 paper.tex
Add high solidity to title
Add energy conversion to title
Add examples of XFOIL usage and cite QBlade; resolves #27

@@
 % Full title of the paper (Capitalized)
-\Title{Effects of Reynolds Number on the Performance and Near-Wake Dynamics of a
-Vertical-Axis Cross-Flow Turbine}
+\Title{Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a
+High Solidity Vertical-Axis Cross-Flow Turbine}
 
@@
 performance, a series of static foil coefficient datasets were computed with the
-XFOIL viscous panel method code \cite{Drela1989}, implemented as part of the
-open source turbine design software QBlade. Simulations were run for an angle of
+XFOIL viscous panel method code \cite{Drela1989}, a commonly used tool for airfoil analysis, e.g. \cite{Castelli2011, Walker2014}, implemented as part of the
+open source turbine design software QBlade \cite{Marten2013}. Simulations were run for an angle of
 attack range of 0--40 degrees, in increments of 0.5 degrees. Solver parameters

### 2015-10-11 paper.tex
Angel --> angle; fixes #30

@@
 energetic, which enables it to advance further downstream against the adverse
-pressure gradient on the suction side and delay separation to higher angels of
+pressure gradient on the suction side and delay separation to higher angles of
 attack, as well as diminish the negative effects of a laminar separation

### 2015-10-11 paper.tex
Em dash spacing

@@
 axis) turbines depend on many parameters: solidity, blade profile, blade pitch,
-number of blades, strut drag, operational parameters such as tip speed ratio
---- and Reynolds number. If numerical models are validated with physical model
+number of blades, strut drag, operational parameters such as tip speed ratio---and Reynolds number. If numerical models are validated with physical model
 data that was obtained at insufficiently high Reynolds numbers, it cannot be
@@
 blade pitch, number of blades, strut drag and operational parameters such as
-tip speed ratio --- {\em and} on Reynolds number.
+tip speed ratio---{\em and} on Reynolds number.
 Note that an average blade chord Reynolds number, $Re_{c,\mathrm{avg}} \approx

### 2015-10-11 paper.tex
Abstract formatting
Less background and more specifics in abstract

@@
 % Abstract (Less than 200 words)
-\abstract{The performance and wake characteristics of cross-flow (vertical
-axis) turbines depend on many parameters: solidity, blade profile, blade pitch,
-number of blades, strut drag, operational parameters such as tip speed ratio---and Reynolds number. If numerical models are validated with physical model
-data that was obtained at insufficiently high Reynolds numbers, it cannot be
-determined whether problems with model predictions are caused by Reynolds
-number effects, other turbine parameters or issues with the numerical model.
-%It is uncertain whether numerical models validated with physical model data
-%obtained at low Reynolds number should be considered validated at all, since
-%the scale at which the model will be applied is orders of magnitude larger.
-Experiments were performed with a large laboratory-scale cross flow turbine of
-to investigate Reynolds number dependence of performance and wake transport
-characteristics,
-and to determine what an acceptable Reynolds number mismatch is for scaled
-physical model tests of cross-flow turbines. It was demonstrated that the
-performance of the cross-flow turbine becomes essentially $Re$-independent at a
-Reynolds number based on turbine
-diameter of $Re_D \approx 10^6$, or an approximate Reynolds number based on
-blade chord of $Re_c \approx 2 \times 10^5$.
-A simple model that calculates peak torque coefficient from
-static foil data and cross-flow turbine kinematics was shown to be a reasonable
+\abstract{The performance and wake characteristics of cross-flow (vertical axis)
+    turbines depend on many parameters: solidity, blade profile, blade pitch, number
+    of blades, strut drag, operational parameters such as tip speed ratio---and
+    Reynolds number. Tolerable mismatch between physical model Reynolds number and
+    full-scale Reynolds number was investigated from both the perspective of
+    predicting full scale performance and flow dynamics, as well as numerical model
+    validation. Experiments were performed with a large laboratory-scale cross flow
+    turbine of to investigate Reynolds number dependence of performance and wake
+    transport characteristics, and to determine what an acceptable Reynolds number
+    mismatch is for scaled physical model tests of cross-flow turbines. It was
+    demonstrated that the performance of the cross-flow turbine becomes essentially
+    $Re$-independent at a Reynolds number based on turbine diameter of $Re_D \approx
+    10^6$, or an approximate Reynolds number based on blade chord of $Re_c \approx 2
+    \times 10^5$. A simple model that calculates peak torque coefficient from static
+    foil data and cross-flow turbine kinematics was shown to be a reasonable
     estimator for predicting Reynolds number dependence of an actual cross flow
@@
 
-\todo[inline]{Rewrite abstract so it is more of a summary of what was done -
-tone down the background information a bit. (Note: some background information
-was moved to Introduction)}
-% %
 \todo[inline]{Include a List of Symbols?}

### 2015-10-11 paper.tex
Ensure hyphens in cross-flow

@@
     predicting full scale performance and flow dynamics, as well as numerical model
-    validation. Experiments were performed with a large laboratory-scale cross flow
+    validation. Experiments were performed with a large laboratory-scale cross-flow
     turbine of to investigate Reynolds number dependence of performance and wake
@@
     foil data and cross-flow turbine kinematics was shown to be a reasonable
-    estimator for predicting Reynolds number dependence of an actual cross flow
+    estimator for predicting Reynolds number dependence of an actual cross-flow
     turbine operating under dynamic conditions.}

### 2015-10-11 paper.tex
Move issue to GitHub
Text wrapping
Wrapping
More text wrapping

@@
 
-\todo[inline]{Include a List of Symbols?}
-
 \section{Introduction}
@@
 Scaled models have the benefit of being significantly less expensive, however,
-there still is concern at which
-scale an experiment must be performed to be relevant to full-scale application.
-This is particularly important for cross-flow turbines, where Reynolds number
-effects for experiments conducted at low $Re$ are often difficult to
-distinguish from the effects on performance of several other turbine
-parameters.
+there still is concern at which scale an experiment must be performed to be
+relevant to full-scale application. This is particularly important for
+cross-flow turbines, where Reynolds number effects for experiments conducted at
+low $Re$ are often difficult to distinguish from the effects on performance of
+several other turbine parameters.
 
-The performance and wake characteristics of cross-flow (vertical
-axis) turbines depend on turbine solidity, blade profile (lift/drag, dynamic
-stall at reduced frequency of turbine rotation, symmetry, thickness, camber),
-blade pitch, number of blades, strut drag and operational parameters such as
-tip speed ratio---{\em and} on Reynolds number.
-Note that an average blade chord Reynolds number, $Re_{c,\mathrm{avg}} \approx
-\lambda U_\infty c/ \nu$, can be expressed in terms of tip speed ratio $\lambda
-\equiv \omega R/ U_\infty$; the value of $\lambda$ at which a turbine reaches
-peak performance is again a function of turbine solidity $Nc/\pi D$.
+The performance and wake characteristics of cross-flow (vertical axis) turbines
+depend on turbine solidity, blade profile (lift/drag, dynamic stall at reduced
+frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
+of blades, strut drag and operational parameters such as tip speed ratio---{\em
+    and} on Reynolds number. Note that an average blade chord Reynolds number,
+$Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be expressed in terms
+of tip speed ratio $\lambda \equiv \omega R/ U_\infty$; the value of $\lambda$
+at which a turbine reaches peak performance is again a function of turbine
+solidity $Nc/\pi D$.
 
-If numerical models are validated with physical model data that
-was obtained at insufficiently high Reynolds numbers, it cannot be determined
-whether problems with model predictions are caused by Reynolds number effects,
-other turbine parameters or issues with the numerical model.
-One might also obtain different predictions of scaled systems
-due to modified physics at lower Reynolds numbers, e.g., laminar boundary
-layers and transitional flow regimes. If the numerical model is then considered
-``validated,'' there is a risk that its application at full-scale may produce
-incorrect predictions. It is uncertain whether numerical models validated with
-physical model data obtained at low Reynolds number should be considered
-validated at all, since the scale at which the model will be applied is orders
-of magnitude larger.
+If numerical models are validated with physical model data that was obtained at
+insufficiently high Reynolds numbers, it cannot be determined whether problems
+with model predictions are caused by Reynolds number effects, other turbine
+parameters or issues with the numerical model. One might also obtain different
+predictions of scaled systems due to modified physics at lower Reynolds numbers,
+e.g., laminar boundary layers and transitional flow regimes. If the numerical
+model is then considered ``validated,'' there is a risk that its application at
+full-scale may produce incorrect predictions. It is uncertain whether numerical
+models validated with physical model data obtained at low Reynolds number should
+be considered validated at all, since the scale at which the model will be
+applied is orders of magnitude larger.
 

### 2015-10-11 paper.tex
Remove todo item on TSR effects
Text wrapping
Add info about McTavish et al (2013); resolves #26

@@
 
-\todo[inline]{Kinematic scaling effects? Matching TSR?}
-
 With regards to wind and marine hydrokinetic (MHK) turbines, scaled physical
@@
 
-The performance and wake characteristics of cross-flow (vertical axis) turbines
+The performance and wake characteristics of cross-flow (vertical-axis) turbines
 depend on turbine solidity, blade profile (lift/drag, dynamic stall at reduced
@@
 This can be overcome by showing that a scaled physical model test has become
-Reynolds number independent, and therefore validation efforts should be
-relevant at full-scale.  For scaled physical models, it is of interest how
-small---since size, or scale generally correlates inversely with cost---the
-model can be while remaining a reliable predictor of full-scale behavior. At
-the very least, it is necessary to test at a scale where results can be
-extrapolated, i.e., changes are occurring quasi-linearly without dramatic
-changes in the underlying physical principles governing the behavior of
-interest, e.g., power production in the case of a wind or MHK turbine.
+Reynolds number independent, and therefore validation efforts should be relevant
+at full-scale.  For scaled physical models, it is of interest how small---since
+size, or scale generally correlates inversely with cost---the model can be while
+remaining a reliable predictor of full-scale behavior. At the very least, it is
+necessary to test at a scale where results can be extrapolated, i.e., changes
+are occurring quasi-linearly without dramatic changes in the underlying physical
+principles governing the behavior of interest, e.g., power production in the
+case of a wind or MHK turbine.
 
-Blackwell et al. investigated the effects of Reynolds number on the
-performance
+Blackwell et al. investigated the effects of Reynolds number on the performance
 of a 2 m diameter Darrieus vertical-axis cross-flow wind turbine with NACA 0012
@@
 
-Krogstad and Adaramola observed Reynolds number independence of the performance for a 0.9 m
-diameter axial-flow turbine rotor at $Re_D \approx 5 \times 10^5$ in wind tunnel
-tests \cite{Krogstad2012a}.
-Their turbine blades had an S826 profile along their entire span, a profile chosen for its $Re$-independence.
-Walker et al.
-\cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow turbine tests in a towing tank, which were performed at a mid span Reynolds number $Re_c = 4 \times 10^5$.
+Krogstad and Adaramola observed Reynolds number independence of the performance
+for a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5 \times 10^5$ in
+wind tunnel tests \cite{Krogstad2012a}. Their turbine blades had an S826 profile
+along their entire span, a profile chosen for its $Re$-independence. Walker et
+al. \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
+turbine tests in a towing tank, which were performed at a mid span Reynolds
+number $Re_c = 4 \times 10^5$. McTavish et al. \cite{McTavish2013} showed how
+the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
+numbers, concluding that physical models should be designed with
+$Re$-independence in mind if they are to be run at low $Re$.
 
-\todo[inline]{bring back the examples of studies at small scale? (too small)}
+\todo[inline]{Examples of very low $Re$ CFT work, e.g. Brochier et al.?}
 
@@
 
-Previously, is has been observed experimentally that a cross-flow turbine wake's
-unique vortical mean flow field is responsible for accelerated wake recovery
-when compared with conventional axial-flow propeller-type turbines
-\cite{Bachant2015-JoT}. In this study we seek to replicate the same energy
-balance considerations at multiple Reynolds numbers, to examine the implications
-on how small scale, i.e. low Reynolds number experiments may be used to study
-flows in turbine arrays.
+Previously, is has been observed experimentally that a high solidity cross-flow
+turbine wake's unique vortical mean flow field is responsible for accelerated
+wake recovery when compared with conventional axial-flow propeller-type turbines
+\cite{Bachant2015-JoT}. In this study we seek to replicate the same momentum and
+energy balance considerations at multiple Reynolds numbers, to examine the
+implications on how small scale, i.e. low Reynolds number experiments may be
+used to study flows in turbine arrays.
 

### 2015-10-11 paper.tex
Text wrapping
A little cleanup

@@
 In this study we are using a simple cross-flow turbine constructed from
-hydrofoils, i.e., the
-turbine is a lift-driven device in contrast to drag-driven. A logical starting
-point for understanding the effects of $Re$ on overall turbine loading is to
-first consider simplified cases of individual foils, and their static and
-dynamic loading.
+hydrofoils, i.e., the turbine is a lift-driven device in contrast to
+drag-driven. A logical starting point for understanding the effects of $Re$ on
+overall turbine loading is to first consider simplified cases of individual
+foils, and their static and dynamic loading.
 
@@
 Hydro- or airfoil section performance is generally characterized by the
-lift-to-drag
-force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ),
-which is primarily a function of the foil's angle of attack $\alpha$. The
-angles of attack
-for cross-flow turbine blades are a function of turbine tip speed ratio,
+lift-to-drag force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ), which
+is primarily a function of the foil's angle of attack $\alpha$. The angles of
+attack for cross-flow turbine blades are a function of turbine tip speed ratio,
 induction (slowing/turning of the free stream flow prior to reaching the
@@
 which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
-(reference?). This
-is caused by the boundary layer on the foil becoming more turbulent and more
-energetic, which enables it to advance further downstream against the adverse
-pressure gradient on the suction side and delay separation to higher angles of
-attack, as well as diminish the negative effects of a laminar separation
-bubble occurring near the leading edge (reference?). For smooth airfoils, this
-transition can
-cause a dramatic increase in foil performance, as seen in
-Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
-$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
-presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
-reliable data for airfoils in this transitional regime and below. An evaluation
-of the various databases relevant to cross-flow turbines is presented in
-\cite{Bedon2014}.
+(reference?). This is caused by the boundary layer on the foil becoming more
+turbulent and more energetic, which enables it to advance further downstream
+against the adverse pressure gradient on the suction side and delay separation
+to higher angles of attack, as well as diminish the negative effects of a
+laminar separation bubble occurring near the leading edge (reference?). For
+smooth airfoils, this transition can cause a dramatic increase in foil
+performance, as seen in Figure~\ref{fig:McMasters} at a blade chord Reynolds
+number on the order of $10^5$ \cite{McMasters1980}. A review of low Reynolds
+number airfoil behavior is presented in \cite{Lissaman1983}. Note that there is
+a distinct lack of highly reliable data for airfoils in this transitional regime
+and below. An evaluation of the various databases relevant to cross-flow
+turbines is presented in \cite{Bedon2014}.
 
@@
 \centering
+
 \includegraphics[width=0.7\textwidth]{figures/McMasters-Henderson-1980}
+
 \caption{Airfoil lift to drag ratio as a function of Reynolds number, from
 \cite{McMasters1980}.}
+
 \label{fig:McMasters}
@@
 The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
-turbine implies dynamic loading. Furthermore, since the range of angle of
-attack encountered by a rotating blade increases with decreasing tip speed
-ratio, the blades commonly undergo
-dynamic stall for tip speed ratios at and below those of maximum rotor
-torque\cite{Para2002}. We then must understand the effects of Reynolds number on
-the dynamic stall process.
+turbine implies dynamic loading. Furthermore, since the range of angle of attack
+encountered by a rotating blade increases with decreasing tip speed ratio, the
+blades commonly undergo dynamic stall for tip speed ratios at and below those of
+maximum rotor torque\cite{Para2002}. We then must understand the effects of
+Reynolds number on the dynamic stall process.
 
@@
 Chamorro et al. \cite{Chamorro2012} showed that turbulence statistics in an AFT
-wake became
-$Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean velocity
-profiles became $Re$-independent slightly earlier at $Re_D \approx 5 \times
-10^4$.
+wake became $Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean
+velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
+\times 10^4$.
+
 \todo[inline]{What was Chamorro's blade chord Re?}
+
 \section{Experimental Setup}
@@
 \centering
+
 \includegraphics[width=0.4\textwidth]{figures/turbine}
+
 \caption{UNH-RVAT turbine model. Turbine blades and struts made from NACA 0020
-profiles with 0.140 m chord. Note that the upper and lower mounting
-flanges have been excluded as these were included in the tare drag
-measurements.}
+    profiles with 0.140 m chord. Note that the upper and lower mounting flanges have
+    been excluded as these were included in the tare drag measurements.}
+
 \label{fig:turbine}

### 2015-10-11 paper.tex
Clean up dynamic foil section a bit

@@
 blades commonly undergo dynamic stall for tip speed ratios at and below those of
-maximum rotor torque\cite{Para2002}. We then must understand the effects of
+maximum rotor torque \cite{Para2002}. We then must understand the effects of
 Reynolds number on the dynamic stall process.
 
-Dynamic effects on blade loading can be distinguished between attached and
-separated conditions.
+Bousman \cite{Bousman2000-evaluation} states that dynamic stall is relatively
+insensitive to Reynolds number for $Re=1.0 \times 10^5$--$2.5 \times
+10^5$---judging from measurements on a pitching VR-7 foil in a wind and water
+tunnel---since the loading is dominated by vortex shedding. However, Singleton
+and Yeager Jr. \cite{Singleton2000} state that Reynolds number effects on
+dynamic stall remains an unsolved question.
 
-\todo[inline]{Find reference to describe $Re$ effects on dynamic foil behavior.}
-
-Bousman showed that dynamic stall was insensitive to Reynolds number for $Re=1.0
-\times 10^5$--$2.5 \times 10^5$ \cite{Bousman2000-evaluation}, Though Singleton
-and Yeager Jr. state that Reynolds number effects on dynamics stall remains an
-unsolved question \cite{Singleton2000}.
-
-\todo[inline]{Bousman: explain in what type of system, with what type of foil
-this was done.}
-\todo[inline]{you took quite a few references out of this section...?}
 \todo[inline]{additional data in helicopter literature?}

### 2015-10-11 paper.tex
Cite Neary et al paper

@@
 numerical model testing, similar to the Sandia National Labs/US Department of
-Energy RM2 River Turbine, but with higher solidity, or blade chord to radius
-ratio.
+Energy RM2 River Turbine \cite{Neary2014}, but with higher solidity, or blade
+chord to radius ratio.
 
@@
 the turbine is available from \cite{Bachant2014-RVAT-CAD}.
-
-\todo[inline]{add reference: V. Neary, M. Previsic, R. Jepsen, M. Lawson, Y.
-Yu, A. Copping, A. Fontaine, K. Hallett, D. Murray (2014). Methodology for
-Design and Economic Analysis of Marine Energy Conversion (MEC) Technologies.
-SAND2014-9040}
 

### 2015-10-11 paper.tex
Move todo item to GH issues
Misc cleanup

@@
 
-\todo[inline]{Add a thicker dashed line at $y= -0.5m$, and maybe a dashed 1m
-circle in the top view of the turbine.}
-
 \subsection{Data Processing}
@@
 
-Mean power and drag coefficients at $\lambda=1.9$ are plotted vs.\ Reynolds
+Mean power and drag coefficients at $\lambda=1.9$ are plotted versus Reynolds
 number in Figure~\ref{fig:perf-Re-dep}. There is a drastic improvement in $C_P$
@@
 thereby promoting reattachment of the laminar separation bubble.
+
 \todo[inline]{Is there still a laminar separation bubble near the leading edge
@@
 performance, a series of static foil coefficient datasets were computed with the
-XFOIL viscous panel method code \cite{Drela1989}, a commonly used tool for airfoil analysis, e.g. \cite{Castelli2011, Walker2014}, implemented as part of the
-open source turbine design software QBlade \cite{Marten2013}. Simulations were run for an angle of
-attack range of 0--40 degrees, in increments of 0.5 degrees. Solver parameters
-used a 100 panels, a fixed speed, zero Mach number, \texttt{NCrit=9} (default
-$\mathrm{e}^n$ transition criteria parameter for an average wind tunnel), and no
-forced boundary layer transition. Characteristics for computed for the
-approximate average blade chord Reynolds numbers encountered in the tow tank
-experiments.
+XFOIL viscous panel method code \cite{Drela1989}, a commonly used tool for
+airfoil analysis, e.g. \cite{Castelli2011, Walker2014}, implemented as part of
+the open source turbine design software QBlade \cite{Marten2013}. Simulations
+were run for an angle of attack range of 0--40 degrees, in increments of 0.5
+degrees. Solver parameters used a 100 panels, a fixed speed, zero Mach number,
+\texttt{NCrit=9} (default $\mathrm{e}^n$ transition criteria parameter for an
+average wind tunnel), and no forced boundary layer transition. Characteristics
+for computed for the approximate average blade chord Reynolds numbers
+encountered in the tow tank experiments.
 

### 2015-10-11 paper.tex
Fix logic
High --> significant

@@
 solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
-maximum radius). More recently, Polagye and Cavagnaro observed high Reynolds
-number sensitivity for a high solidity helical cross-flow turbine
+maximum radius). More recently, Polagye and Cavagnaro observed significant
+Reynolds number sensitivity for a high solidity helical cross-flow turbine
 \cite{Polagye2013b}, and Bravo et al. observed power coefficient of a
@@
 We see that in general, levels of turbulent transport are slightly lower at
-lower Reynolds numbers. The viscous diffusion and dissipation, though still
+larger Reynolds numbers. The viscous diffusion and dissipation, though still
 three orders of magnitude smaller than the other terms, do increase at low
 Reynolds numbers, which is consistent with the physical meaning of the Reynolds
-number
-in general.
+number in general.
 

### 2015-10-11 paper.tex
Reword sentence

@@
 smooth airfoils, this transition can cause a dramatic increase in foil
-performance, as seen in Figure~\ref{fig:McMasters} at a blade chord Reynolds
-number on the order of $10^5$ \cite{McMasters1980}. A review of low Reynolds
-number airfoil behavior is presented in \cite{Lissaman1983}. Note that there is
-a distinct lack of highly reliable data for airfoils in this transitional regime
-and below. An evaluation of the various databases relevant to cross-flow
-turbines is presented in \cite{Bedon2014}.
+performance, at a blade chord Reynolds number on the order of $10^5$
+\cite{McMasters1980}, shown in Figure~\ref{fig:McMasters}. A review of low
+Reynolds number airfoil behavior is presented in \cite{Lissaman1983}. Note that
+there is a distinct lack of highly reliable data for airfoils in this
+transitional regime and below. An evaluation of the various databases relevant
+to cross-flow turbines is presented in \cite{Bedon2014}.
 

### 2015-10-12 paper.tex
Better summary for Re effects on static foil perf

@@
 which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
-(reference?). This is caused by the boundary layer on the foil becoming more
-turbulent and more energetic, which enables it to advance further downstream
-against the adverse pressure gradient on the suction side and delay separation
-to higher angles of attack, as well as diminish the negative effects of a
-laminar separation bubble occurring near the leading edge (reference?). For
-smooth airfoils, this transition can cause a dramatic increase in foil
-performance, at a blade chord Reynolds number on the order of $10^5$
-\cite{McMasters1980}, shown in Figure~\ref{fig:McMasters}. A review of low
-Reynolds number airfoil behavior is presented in \cite{Lissaman1983}. Note that
-there is a distinct lack of highly reliable data for airfoils in this
-transitional regime and below. An evaluation of the various databases relevant
-to cross-flow turbines is presented in \cite{Bedon2014}.
+\cite{Jacobs1937}.
 
-\todo[inline]{Laminar separation bubble - is it always there? also need
-reference}
+A review of Reynolds number effects on airfoil behavior is presented in
+\cite{Lissaman1983}. In general, airfoil performance is enhanced as the boundary
+layer on the foil transitions to turbulence closer to the leading edge, which
+enables it to advance further downstream against the adverse pressure gradient
+on the suction side, delaying separation to higher angles of attack. For smooth
+airfoils, this transition can cause a dramatic increase in foil performance, at
+a blade chord Reynolds number on the order of $10^5$ \cite{McMasters1980}, shown
+in Figure~\ref{fig:McMasters}. Note that there is a distinct lack of highly
+reliable data for airfoils in this transitional regime and below. An evaluation
+of the various databases relevant to cross-flow turbines is presented in
+\cite{Bedon2014}.
 

### 2015-10-12 paper.tex
Estimate Chamorro Re_c
Address separation bubble comment

@@
 velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
-\times 10^4$.
-
-\todo[inline]{What was Chamorro's blade chord Re?}
+\times 10^4$. Note that in this study the tip chord Reynolds number was not
+reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
+10^5$.
 
@@
 behavior of the blade boundary layer transitioning from laminar to turbulent,
-thereby promoting reattachment of the laminar separation bubble.
-
-\todo[inline]{Is there still a laminar separation bubble near the leading edge
-    at these Reynolds numbers?}
+thereby promoting either the suppression or reattachment of the laminar
+separation bubble \cite{Lissaman1983}.
 

### 2015-10-12 paper.tex
Use high solidity in conclusion

@@
 
-In this study it was demonstrated that the performance of a cross-flow turbine
-becomes essentially $Re$-independent at a Reynolds number based on turbine
-diameter of $Re_D \approx 10^6$, or an approximate Reynolds number based on
-blade chord of $Re_c \approx 2 \times 10^5$.
+In this study it was demonstrated that the performance of a high solidity
+($c/R=0.28$) cross-flow turbine becomes essentially $Re$-independent at a
+Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
+approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
+10^5$.
 
-\todo[inline]{Should we discuss that these results depend on turbine type, and
-solidity? Lower solidity turbines may only achieve a "linear" regime, not a
-truly Re-indept regime. Additional issues such as blade termination/strength of
-tip vortices?}
+\todo[inline]{Issues such as blade termination/strength of tip vortices?}
 

### 2015-10-12 paper.tex
Address todo item regarding experimental results

@@
     calculated from static foil coefficients and blade kinematics, compared to
-experimental data from a cross-flow turbine.}
+    experimental data from a cross-flow turbine. Note that the experimental data
+    represents the mean torque coefficient, not the maximum.}
 
@@
 \end{figure}
-
-\todo[inline]{Is this the peak torque coefficient or the mean power coefficient?}
 

### 2015-10-12 paper.tex
Adjust caption
Lift and drag coefficient notation formatting

@@
 Hydro- or airfoil section performance is generally characterized by the
-lift-to-drag force ratio $L/D$  (or lift vs drag coefficient, $c_L/c_D$ ), which
+lift-to-drag force ratio $L/D$  (or lift vs drag coefficient, $C_l/C_d$ ), which
 is primarily a function of the foil's angle of attack $\alpha$. The angles of
@@
 
-\caption{Airfoil lift to drag ratio as a function of Reynolds number, from
+\caption{Airfoil maximum lift to drag ratio versus Reynolds number, from
     \cite{McMasters1980}.}

### 2015-10-12 paper.tex
Froude number is mentioned later

@@
 
-\todo[inline]{Talk about interdependence of Froude number. What is the Froude
-number at 1 m/s?}
-
 \subsection{Reynolds Number Effects}
@@
 Reynolds numbers are plotted in Figure~\ref{fig:cp-curves} and
-Figure~\ref{fig:cd-curves}, respectively. There is essentially no change in the
-shape of the drag coefficient curves---merely an upward shift in $C_D$ with
-increasing $Re$.
-
-In general, maximum $C_P$ increases with Reynolds number, which is due to an
-increase in the foil lift-to-drag ratio. The power coefficient curves also show
-a slight downward shift in the optimal tip speed ratio (peak performance) with
-increasing Reynolds number, from about $\lambda \approx$ 2.0 to 1.9. This is
-caused by the stall delay from a more turbulent boundary later on the blade
-suction side. There is essentially no change in the shape of the drag
-coefficient curves---merely an upward shift in $C_D$ with increasing $Re$.
+Figure~\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
+with Reynolds number, which is due to an increase in the foil lift-to-drag
+ratio. The power coefficient curves also show a slight downward shift in the
+optimal tip speed ratio (peak performance) with increasing Reynolds number, from
+about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall delay from a
+more turbulent boundary later on the blade suction side. There is essentially no
+change in the shape of the drag coefficient curves---merely an upward shift in
+$C_D$ with increasing $Re$.
 
@@
 are less dramatic. This is likely due to cross-flow turbine geometry, where
-increases in blade drag at lower $Re$ somewhat offset the reduction in lift, to
+increases in blade drag at lower $Re$ somewhat offsets the reduction in lift, to
 keep total streamwise force from changing as much as the rotor torque. The

### 2015-10-12 paper.tex
Add TSR to foil geometric considerations figure
Make estimate on viscous effects

@@
 \centering
+
 \includegraphics[width=0.9\textwidth]{figures/foil_kinematics_ct}
+
 \caption{Geometric angle of attack, relative velocity and torque coefficient
-calculated with a NACA 0020 foil.}
+    calculated with a NACA 0020 foil operating at $\lambda=1.9$.}
+
 \label{fig:blade-kinematics}
@@
 Reynolds numbers, which is consistent with the physical meaning of the Reynolds
-number in general.
-
-\todo[inline]{At what $Re$ will the viscous terms get large enough to start
-messing with the balance of the equation?}
+number in general. From these results one might expect that the viscous effects
+will become significant to the wake dynamics around $Re_D \sim 10^4$.
 

### 2015-10-13 paper.tex
Remove spectra noise todo item
Offer explanation on the low energy in the 6f_b frequency
Typo

@@
 average wind tunnel), and no forced boundary layer transition. Characteristics
-for computed for the approximate average blade chord Reynolds numbers
+were computed for the approximate average blade chord Reynolds numbers
 encountered in the tow tank experiments.
@@
 spectral energy at the blade passage frequency's first harmonic, or $6
-f_\mathrm{turbine}$.
-
-\todo[inline]{Look at how spectra change when turbine rotational frequency
-changes relative to inertial subrange.}
-
-\todo[inline]{Make sure the spectra are not just shifting due to normalizing the
-noise.}
+f_\mathrm{turbine}$. This is likely due to the blade's shed vorticity being less
+stable at higher $Re$.
 

### 2015-10-13 paper.tex
Add repo for XFOIL coeffs; resolves #19
Add references on CFT wake studies

@@
 
+The wake of a cross-flow turbine in dynamic stall has been studied via laser
+Doppler velocimetry by Brochier et al. \cite{Brochier1986} and with particle
+image velocimetry (PIV) by Fujisawa and Shibuya \cite{Fujisawa2001}. However,
+both of these studies were performed at very low Reynolds numbers---$Re_D =
+10^4$ and $10^3$, respectively. Tescione et al. \cite{Tescione2014} studied the
+wake of a vertical-axis wind turbine in a wind tunnel using stereo PIV at an
+approximate blade chord Reynolds number $Re_c = 1.7 \times 10^5$; very close to
+the $Re$-independent state reported in \cite{Bachant2014}.
+
 Krogstad and Adaramola observed Reynolds number independence of the performance
@@
 $Re$-independence in mind if they are to be run at low $Re$.
-
-\todo[inline]{Examples of very low $Re$ CFT work, e.g. Brochier et al.?}
 
@@
 versus $Re_c$. In general, larger camber is associated with decreased foil
-performance at lower Reynolds number.
+performance at lower Reynolds number. The data and processing code for these
+calculations is available from \cite{Bachant2015-NACAXX20-XFOIL}.
 

### 2015-10-13 paper.tex
Move todo item to GH issues
Remove todo item that already exists as #11
Move todo item to #38

@@
 
-\todo[inline]{additional data in helicopter literature?}
-
 \subsubsection{On Wake Flows}
@@
 become unstable and break down.
-
-\todo[inline]{What do studies in the literature tell us about $Re$ effects
-on shear flows, wakes, and vortex flows? Note-MW I'm looking into Re
-dependence in general wake studies.}
 
@@
 
-\todo[inline]{Any additional insight we can gainfrom these figures?}
-
 The totals for all the wake transport terms calculated in

### 2015-10-13 paper.tex
Move todo item to #39
Remove list of todos and add line numbers
Minor fix in abstract

@@
     predicting full scale performance and flow dynamics, as well as numerical model
-    validation. Experiments were performed with a large laboratory-scale cross-flow
-    turbine of to investigate Reynolds number dependence of performance and wake
-    transport characteristics, and to determine what an acceptable Reynolds number
-    mismatch is for scaled physical model tests of cross-flow turbines. It was
-    demonstrated that the performance of the cross-flow turbine becomes essentially
-    $Re$-independent at a Reynolds number based on turbine diameter of $Re_D \approx
-    10^6$, or an approximate Reynolds number based on blade chord of $Re_c \approx 2
-    \times 10^5$. A simple model that calculates peak torque coefficient from static
-    foil data and cross-flow turbine kinematics was shown to be a reasonable
-    estimator for predicting Reynolds number dependence of an actual cross-flow
-    turbine operating under dynamic conditions.}
+    validation. Experiments were performed with a large laboratory-scale high
+    solidity cross-flow turbine to investigate Reynolds number dependence of
+    performance and wake transport characteristics, and to determine what an
+    acceptable Reynolds number mismatch is for scaled physical model tests of
+    cross-flow turbines. It was demonstrated that the performance of the cross-flow
+    turbine becomes essentially $Re$-independent at a Reynolds number based on
+    turbine diameter of $Re_D \approx 10^6$, or an approximate Reynolds number based
+    on blade chord of $Re_c \approx 2 \times 10^5$. A simple model that calculates
+    peak torque coefficient from static foil data and cross-flow turbine kinematics
+    was shown to be a reasonable estimator for predicting Reynolds number dependence
+    of an actual cross-flow turbine operating under dynamic conditions.}
 
@@
 \begin{document}
-%\linenumbers
-\listoftodos
+\linenumbers
 
@@
 
-\todo[inline]{Issues such as blade termination/strength of tip vortices?}
-
 These results suggest that to validate predictive engineering models, one must

### 2015-10-13 paper.tex
Add VAWT keyword
Define some more symbols

@@
 \keyword{Reynolds number; cross-flow turbine; turbine performance; marine
-hydrokinetic energy; wind energy}
+hydrokinetic energy; wind energy; VAWT}
 
@@
 $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be expressed in terms
-of tip speed ratio $\lambda \equiv \omega R/ U_\infty$; the value of $\lambda$
-at which a turbine reaches peak performance is again a function of turbine
-solidity $Nc/\pi D$.
+of tip speed ratio $\lambda \equiv \omega R/ U_\infty$, where $U_\infty$ is the
+free stream velocity, $c$ is the blade chord length, $\nu$ is the fluid
+kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$ is the
+rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
+is again a function of turbine solidity $Nc/\pi D$, which allows for the use of
+a simpler Reynolds number based on turbine diameter $Re_D \equiv U_\infty
+D/\nu$.
 

### 2015-10-13 paper.tex
Clean up paragraph in introduction

@@
 with model predictions are caused by Reynolds number effects, other turbine
-parameters or issues with the numerical model. One might also obtain different
-predictions of scaled systems due to modified physics at lower Reynolds numbers,
-e.g., laminar boundary layers and transitional flow regimes. If the numerical
-model is then considered ``validated,'' there is a risk that its application at
-full-scale may produce incorrect predictions. It is uncertain whether numerical
-models validated with physical model data obtained at low Reynolds number should
-be considered validated at all, since the scale at which the model will be
-applied is orders of magnitude larger.
+parameters or issues with the numerical model itself. One might also obtain
+different predictions of scaled systems due to modified physics at lower
+Reynolds numbers, e.g., laminar boundary layers and transitional flow regimes.
+If the numerical model is then considered ``validated,'' there is a risk that
+its application at full-scale may produce incorrect predictions. As such, it is
+questionable whether numerical models validated with physical model data
+obtained at Reynolds numbers orders of magnitude away from prototype scale
+should be considered validated at all.
 

### 2015-10-13 paper.tex
Clean up more intro writing

@@
 
-This can be overcome by showing that a scaled physical model test has become
-Reynolds number independent, and therefore validation efforts should be relevant
-at full-scale.  For scaled physical models, it is of interest how small---since
-size, or scale generally correlates inversely with cost---the model can be while
-remaining a reliable predictor of full-scale behavior. At the very least, it is
-necessary to test at a scale where results can be extrapolated, i.e., changes
-are occurring quasi-linearly without dramatic changes in the underlying physical
-principles governing the behavior of interest, e.g., power production in the
-case of a wind or MHK turbine.
+Uncertainty from scale effects can potentially be overcome by showing that a
+scaled physical model test has become Reynolds number independent, and therefore
+validation efforts should be relevant at full scale. It is therefore of interest
+how small---since size, or scale generally correlates inversely with cost---the
+model can be while remaining a reliable predictor of full-scale behavior. At the
+very least, it is necessary to test at a scale where results can be
+extrapolated, i.e., changes are occurring quasi-linearly without dramatic
+changes in the underlying physical principles governing the behavior of
+interest, e.g., power production in the case of a wind or MHK turbine.
 

### 2015-10-13 paper.tex
Clarification about dynamic stall regime

@@
 
-The wake of a cross-flow turbine in dynamic stall has been studied via laser
-Doppler velocimetry by Brochier et al. \cite{Brochier1986} and with particle
-image velocimetry (PIV) by Fujisawa and Shibuya \cite{Fujisawa2001}. However,
-both of these studies were performed at very low Reynolds numbers---$Re_D =
-10^4$ and $10^3$, respectively. Tescione et al. \cite{Tescione2014} studied the
-wake of a vertical-axis wind turbine in a wind tunnel using stereo PIV at an
-approximate blade chord Reynolds number $Re_c = 1.7 \times 10^5$; very close to
-the $Re$-independent state reported in \cite{Bachant2014}.
+The wake of a cross-flow turbine in the dynamic stall regime has been studied
+via laser Doppler velocimetry by Brochier et al. \cite{Brochier1986} and with
+particle image velocimetry (PIV) by Fujisawa and Shibuya \cite{Fujisawa2001}.
+However, both of these studies were performed at very low Reynolds
+numbers---$Re_D = 10^4$ and $10^3$, respectively. Tescione et al.
+\cite{Tescione2014} studied the wake of a vertical-axis wind turbine in a wind
+tunnel using stereo PIV at an approximate blade chord Reynolds number $Re_c =
+1.7 \times 10^5$; very close to the $Re$-independent state reported in
+\cite{Bachant2014}.
 

### 2015-10-13 paper.tex
For --> of

@@
 Krogstad and Adaramola observed Reynolds number independence of the performance
-for a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5 \times 10^5$ in
+of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5 \times 10^5$ in
 wind tunnel tests \cite{Krogstad2012a}. Their turbine blades had an S826 profile

### 2015-10-13 paper.tex
Mid-span

@@
 al. \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
-turbine tests in a towing tank, which were performed at a mid span Reynolds
+turbine tests in a towing tank, which were performed at a mid-span Reynolds
 number $Re_c = 4 \times 10^5$. McTavish et al. \cite{McTavish2013} showed how

### 2015-10-13 paper.tex
Scale --> Re

@@
 \cite{Goldenberg1983}. In this case, it is of interest to determine at what
-scale one might be able to realistically study wake flows in an array, and also
-to evaluate the effectiveness of a wake generator. In other words, a wake
-generator may do a fine job simulating a scaled turbine, but how well can it
-simulate a full-scale device?
+Reynolds number one might be able to realistically study wake flows in an array,
+and also to evaluate the effectiveness of a wake generator. In other words, a
+wake generator may do a fine job simulating a scaled turbine, but how well can
+it simulate a full-scale device?
 

### 2015-10-13 paper.tex
Clarify paragraph on near-wake stuff

@@
 
-Previously, is has been observed experimentally that a high solidity cross-flow
-turbine wake's unique vortical mean flow field is responsible for accelerated
-wake recovery when compared with conventional axial-flow propeller-type turbines
-\cite{Bachant2015-JoT}. In this study we seek to replicate the same momentum and
-energy balance considerations at multiple Reynolds numbers, to examine the
-implications on how small scale, i.e. low Reynolds number experiments may be
-used to study flows in turbine arrays.
+Vertical-axis wind turbine array field experiments have revealed that
+accelerated wake recovery allows for closer spacing when compared with
+conventional axial-flow propeller-type turbines \cite{Dabiri2011, Kinzel2012}.
+It was observed experimentally that a high solidity vertical-axis cross-flow
+turbine's near-wake produces a unique vertical mean velocity field---generated
+by blade tip vortex shedding---the advection by which is the largest contributor
+to streamwise momentum and energy transport or recovery \cite{Bachant2015-JoT}.
+In this study we seek to replicate those same momentum and energy balance
+considerations at multiple Reynolds numbers, to examine the implications on how
+small scale, i.e. low Reynolds number experiments may be used to study flows in
+turbine arrays.
 

### 2015-10-13 paper.tex
Remove space
Clean up cl/cd explanation

@@
 Hydro- or airfoil section performance is generally characterized by the
-lift-to-drag force ratio $L/D$  (or lift vs drag coefficient, $C_l/C_d$ ), which
-is primarily a function of the foil's angle of attack $\alpha$. The angles of
-attack for cross-flow turbine blades are a function of turbine tip speed ratio,
-induction (slowing/turning of the free stream flow prior to reaching the
-turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
-airfoil data, we see that the static stall angle---the angle just beyond that at
-which $L/D$ is maximum---increases with blade chord Reynolds number $Re_c$
-\cite{Jacobs1937}.
+lift-to-drag force ratio $L/D$ (or the ratio of lift and drag coefficients,
+$C_l/C_d$), which is primarily a function of the foil's angle of attack
+$\alpha$. The angles of attack for cross-flow turbine blades are a function of
+turbine tip speed ratio, induction (slowing/turning of the free stream flow
+prior to reaching the turbine), and turbine rotation or azimuthal angle
+$\theta$. Looking at static airfoil data, we see that the static stall
+angle---the angle just beyond that at which $L/D$ is maximum---increases with
+blade chord Reynolds number $Re_c$ \cite{Jacobs1937}.
 

### 2015-10-13 paper.tex
Fix wording

@@
 prior to reaching the turbine), and turbine rotation or azimuthal angle
-$\theta$. Looking at static airfoil data, we see that the static stall
+$\theta$. In general, static airfoil data shows that the static stall
 angle---the angle just beyond that at which $L/D$ is maximum---increases with

### 2015-10-13 paper.tex
Typically instead of in general

@@
 prior to reaching the turbine), and turbine rotation or azimuthal angle
-$\theta$. In general, static airfoil data shows that the static stall
+$\theta$. Typically, static airfoil data shows that the static stall
 angle---the angle just beyond that at which $L/D$ is maximum---increases with

### 2015-10-13 paper.tex
Hyphenate lift-to-drag
Typo

@@
 
-\caption{Airfoil maximum lift to drag ratio versus Reynolds number, from
+\caption{Airfoil maximum lift-to-drag ratio versus Reynolds number, from
     \cite{McMasters1980}.}
@@
 
-Experiments were performed in a turbine test bed specifically designd for
+Experiments were performed in a turbine test bed specifically designed for
 cross-flow turbines. The turbine model used in this study was the UNH Reference

### 2015-10-13 paper.tex
Add whitespace
Remove early explanation for increased C_P at higher Re
Whitespace in figures
Clean up redundancy in exp setup section; resolves #40

@@
 
-The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.140 m
+The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.14 m
 chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and the
 turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
-Figure~\ref{fig:turbine}. The turbine test bed allows precise control and
-measurement of the turbine shaft angle and angular velocity, and therefore tip
-speed ratio, through a permanent magnet servo motor. The primary torque
-measurement was provided by an in-line torque cell, with a redundant measurement
-by a moment arm and load cell and an indirect measurement through the motor via
-its torque constant. The turbine was mounted with in a low-drag frame
-constructed from NACA 0020 foil sections.  The frame is attached to the tow
-carriage by four linear bearings, allowing the total streamwise drag force to be
-measured by a pair of S-beam load cells. Turbine rotor drag is calculated by
-subtracting the test bed tare drag. A schematic of the turbine test bed is shown
-in Figure~\ref{fig:exp-setup}. Additional details of the turbine and
-experimental setup are described in \cite{Bachant2015-JoT}, and a CAD model of
-the turbine is available from \cite{Bachant2014-RVAT-CAD}.
+Figure~\ref{fig:turbine}. The rotor has a relatively high solidity $Nc/(\pi D) =
+0.13$ and large chord-to-radius ratio $c/R = 0.28$. A CAD model of the turbine
+is available from \cite{Bachant2014-RVAT-CAD}.
 
@@
 
-The turbine was mounted in a frame constructed from NACA 0020 sections. The
-turbine shaft ran up through the water surface, coupled to a permanent magnet
-servo motor with a 20:1 gearbox. This servo was controlled by the tow tank's
-main motion controller for high synchronization with the carriage motion. A
-rotary torque transducer was installed inline between the servo and the turbine,
-and the servo was also mounted on a slewing ring bearing, which allowed a
-redundant measurement of torque via an arm and load cell used to counteract the
-turbine moment. The frame was mounted to the carriage via linear guides, such
-that the total streamwise drag force was transferred to a pair of S-beam load
-cells, providing the rotor drag measurements, after a measured tare drag was
-subtracted in post processing. Similarly, a tare torque was measured by rotating
-the turbine shaft in air. Turbine angular and tow carriage linear position were
-measured using quadrature encoder signals, with $10^5$ counts-per-rev for the
-turbine and 10 $\mathrm{\mu m}$ resolution for the carriage position. These
-signals, along with the torque and drag signals, were sampled at 2 kHz.
+The turbine was mounted in a frame constructed from NACA 0020 sections, shown in
+Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
+coupled to a permanent magnet servo motor with a 20:1 gearbox, providing precise
+control over shaft RPM. This servo was controlled by the tow tank's main motion
+controller for high synchronization with the carriage motion, thereby giving
+precise measurement and control of tip speed ratio. A rotary torque transducer
+was installed inline between the servo and the turbine, and the servo was also
+mounted on a slewing ring bearing, which allowed a redundant measurement of
+torque via an arm and load cell used to counteract the turbine moment. The frame
+was mounted to the carriage via linear guides, such that the total streamwise
+drag force was transferred to a pair of S-beam load cells, providing the rotor
+drag measurements, after a measured tare drag was subtracted in post processing.
+Similarly, a tare torque was measured by rotating the turbine shaft in air.
+Turbine angular and tow carriage linear position were measured using quadrature
+encoder signals, with $10^5$ counts-per-rev for the turbine and 10 $\mathrm{\mu
+    m}$ resolution for the carriage position. These signals, along with the torque
+and drag signals, were sampled at 2 kHz.
 
@@
 DAQ systems' sampling times were synchronized by triggering the start of data
-acquisition via a pulse sent from the motion controller.
+acquisition via a pulse sent from the motion controller. Additional details of
+the turbine and experimental setup are described in \cite{Bachant2015-JoT}.
 
@@
 \centering
+
 \includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
+
 \caption{Wake measurement coordinate system and locations. Dimensions are in
     meters.}
+
 \label{fig:wake-locations}
@@
 Figure~\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
-with Reynolds number, which is due to an increase in the foil lift-to-drag
-ratio. The power coefficient curves also show a slight downward shift in the
-optimal tip speed ratio (peak performance) with increasing Reynolds number, from
-about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall delay from a
-more turbulent boundary later on the blade suction side. There is essentially no
-change in the shape of the drag coefficient curves---merely an upward shift in
-$C_D$ with increasing $Re$.
+with Reynolds number. The power coefficient curves also show a slight downward
+shift in the optimal tip speed ratio (peak performance) with increasing Reynolds
+number, from about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall
+delay from a more turbulent boundary later on the blade suction side. There is
+essentially no change in the shape of the drag coefficient curves---merely an
+upward shift in $C_D$ with increasing $Re$.
 
@@
 \centering
+
 \includegraphics[width=0.65\textwidth]{figures/cp_curves}
-\caption{Mean power coefficient curves plotted for
-multiple Reynolds numbers.}
+
+\caption{Mean power coefficient curves plotted for multiple Reynolds numbers.}
+
 \label{fig:cp-curves}
@@
 \centering
+
 \includegraphics[width=0.65\textwidth]{figures/cd_curves}
-\caption{Mean rotor drag coefficient curves plotted for
-multiple Reynolds numbers.}
+
+\caption{Mean rotor drag coefficient curves plotted for multiple Reynolds
+    numbers.}
+
 \label{fig:cd-curves}

### 2015-10-14 paper.tex
Fix up first intro paragraph

@@
 real-world systems. The principle of dynamic similarity allows for geometrically
-scaled systems to be dynamically identical in scale if certain nondimensional
+scaled systems to be dynamically identical if certain nondimensional
 physical parameters are matched. In the case of fluid systems, the most
@@
 quantifies the importance of inertial forces over viscous forces on the flow
-physics. It should be noted that in general the geometric scale of a physical
-model is not necessarily the same as its dynamical scale. As such, hereafter we
-will use the word ``scale'' to refer to this dynamical scale rather than the
-geometric scale.
+physics:
+\begin{equation}
+    \frac{Ul}{\nu},
+    \label{eq:Re}
+\end{equation}
+where $U$ and $l$ are characteristic velocity and length scales, respectively,
+and $\nu$ is the fluid kinematic viscosity. In general the geometric scale of a
+physical model is not necessarily the same as its dynamical scale. As such,
+hereafter we will use the word ``scale'' to refer to this dynamical scale or
+Reynolds number.
 

### 2015-10-14 paper.tex
Clean up introduction some more

@@
 with model predictions are caused by Reynolds number effects, other turbine
-parameters or issues with the numerical model itself. One might also obtain
-different predictions of scaled systems due to modified physics at lower
-Reynolds numbers, e.g., laminar boundary layers and transitional flow regimes.
-If the numerical model is then considered ``validated,'' there is a risk that
-its application at full-scale may produce incorrect predictions. As such, it is
-questionable whether numerical models validated with physical model data
-obtained at Reynolds numbers orders of magnitude away from prototype scale
-should be considered validated at all.
+geometrical or operating parameters, or issues with the numerical model itself.
+One might also obtain different predictions for scaled systems due to modified
+physics at lower Reynolds numbers, e.g., laminar boundary layers and
+transitional flow regimes. If the numerical model is then considered
+``validated,'' there is a risk that its application at full-scale may produce
+incorrect predictions. As such, it is questionable whether numerical models
+validated with physical model data obtained at Reynolds numbers orders of
+magnitude away from prototype scale should be considered validated at all.
 
@@
 validation efforts should be relevant at full scale. It is therefore of interest
-how small---since size, or scale generally correlates inversely with cost---the
-model can be while remaining a reliable predictor of full-scale behavior. At the
-very least, it is necessary to test at a scale where results can be
-extrapolated, i.e., changes are occurring quasi-linearly without dramatic
+how small---since physical model size or facility flow speed generally correlate
+with cost---the model can be while remaining a reliable predictor of full-scale
+behavior. At the very least, it is necessary to test at a scale where results
+can be extrapolated, i.e., changes are occurring quasi-linearly without dramatic
 changes in the underlying physical principles governing the behavior of

### 2015-10-14 paper.tex
Cite METS paper earlier
Cite METS wake stuff earlier too

@@
 straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
-\times 10^5$ \cite{Bravo2007}.
+\times 10^5$ \cite{Bravo2007}. Bachant and Wosnik \cite{Bachant2014} observed
+Reynolds number independence of power coefficient at optimal tip speed ratio for
+a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
+$Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.
 
@@
 1.7 \times 10^5$; very close to the $Re$-independent state reported in
-\cite{Bachant2014}.
+\cite{Bachant2014}, where 2-D near-wake profiles were measured for a high
+solidity cross-flow turbine at optimal tip speed ratio.
 

### 2015-10-14 paper.tex
Simplify wording
Remove last paragraph in first section of intro since information existed elsewhere

@@
 considerations at multiple Reynolds numbers, to examine the implications on how
-small scale, i.e. low Reynolds number experiments may be used to study flows in
+small scale, i.e., low Reynolds number experiments may be used to study flows in
 turbine arrays.
-
-The Reynolds number effects on the maximum performance and 2-D near-wake
-profiles of the University of New Hampshire's Reference Vertical Axis Turbine
-(UNH-RVAT) were presented in \cite{Bachant2014}. Here we set out to
-analyze in greater depth the effects of scale, i.e. Reynolds number, on the
-performance of the turbine, and its more detailed near-wake characteristics. We
-were looking to find a threshold scale beyond which a physical model study could
-reliably predict full-scale performance. We were also looking to investigate the
-sensitivity as small scales, and what the implications might be for validating
-the prediction capabilities of numerical models.
 
@@
 In this study we are using a simple cross-flow turbine constructed from
-hydrofoils, i.e., the turbine is a lift-driven device in contrast to
-drag-driven. A logical starting point for understanding the effects of $Re$ on
-overall turbine loading is to first consider simplified cases of individual
-foils, and their static and dynamic loading.
+hydrofoils, i.e., the rotor is lift-driven in contrast to drag-driven. A logical
+starting point for understanding the effects of $Re$ on overall turbine loading
+is to first consider simplified cases of individual foils, and their static and
+dynamic loading.
 

### 2015-10-14 paper.tex
Figure text wrapping
Fix up conclusion

@@
 \centering
+
 \includegraphics[width=0.8\textwidth]{figures/meancontquiv_10}
-\caption{Mean velocity field at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0 \times 10^6$.}
+
+\caption{Mean velocity field at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0 \times
+    10^6$.}
+
 \label{fig:meancontquiv}
@@
 \centering
+
 \includegraphics[width=0.8\textwidth]{figures/k_contours_10}
-\caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$, and
-$Re_D=1.0 \times 10^6$.}
+
+\caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0
+    \times 10^6$.}
+
 \label{fig:kcont}
@@
 approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
-10^5$.
-
-These results suggest that to validate predictive engineering models, one must
-use data of at least this scale, especially for validation of high-fidelity CFD
-models, where transition to turbulence may be completely different between
-scaled physical model and full-scale prototype.
+10^5$. The left side of the $C_P$--$\lambda$ curve is further diminished at low
+$Re$ due to deeper stalling. The $Re$-independent threshold corresponds to that
+at which the foil suction surface boundary layer becomes turbulent before having
+to recover static pressure against the adverse pressure gradient, highlighted in
+\cite{Lissaman1983, McMasters1980, Carmichael1981}.
 
@@
 
-If using scaled physical models to predict array performance, it is important to
-keep all turbines in the regime where the power coefficient varies linearly to
-avoid exaggerated power deficiencies for downstream turbines, despite
-similarities in wake characteristics. These results also suggest that low
-Reynolds number physical model studies of turbine arrays may see exaggerated
-levels of wake recovery, leading to inadequate or inappropriate spacing or
-layouts.
+At lower Reynolds numbers we observed lower levels of turbulence in the wake at
+$x/D=1$, along with lower levels of turbulent transport with respect to mean
+streamwise momentum and mean kinetic energy recovery in the streamwise
+direction. Vertical advection---the largest transport mechanism
+measured---showed little $Re$-dependence, whereas the negative effects of
+cross-stream advection was enhanced at high $Re$. Despite being order of
+magnitude smaller than other transport processes, viscous diffusion increased
+rapidly with decreasing $Re$. Overall, the total wake transport measured leveled
+off at essentially the same Reynolds number that the performance did.
+
+From these results we can make the recommendation that physical model tests be
+on the order $Re_D \sim 10^6$ to provide predictions of full-scale performance.
+This threshold also applies to the validation of predictive engineering models,
+especially for high-fidelity CFD models, where the boundary layer is to be
+resolved, since transition to turbulence plays an important role in overall
+blade loading.
+
+If using scaled physical models to predict array performance, where turbine
+power output is measured, it may also be important to keep all turbines in the
+regime where the power coefficient varies linearly to avoid exaggerated power
+deficiencies for downstream turbines, despite similarities in wake
+characteristics. Results here also suggest that low Reynolds number physical
+model studies of turbine arrays may see exaggerated levels of wake recovery,
+leading to inadequate or inappropriate spacing or layouts.
 

### 2015-10-31 paper.tex
Wosnik edits

@@
 \abstract{The performance and wake characteristics of cross-flow (vertical axis)
-    turbines depend on many parameters: solidity, blade profile, blade pitch, number
-    of blades, strut drag, operational parameters such as tip speed ratio---and
-    Reynolds number. Tolerable mismatch between physical model Reynolds number and
-    full-scale Reynolds number was investigated from both the perspective of
-    predicting full scale performance and flow dynamics, as well as numerical model
-    validation. Experiments were performed with a large laboratory-scale high
+    turbines depend on many parameters: solidity, blade profile, blade pitch, 
+    number of blades, strut drag, operational parameters such as tip speed 
+    ratio---and Reynolds number.  
+    Experiments were performed with a large laboratory-scale high
     solidity cross-flow turbine to investigate Reynolds number dependence of
-    performance and wake transport characteristics, and to determine what an
+    performance and wake transport characteristics. A key question was what an
     acceptable Reynolds number mismatch is for scaled physical model tests of
-    cross-flow turbines. It was demonstrated that the performance of the cross-flow
+    cross-flow turbines, both from the perspective of predicting full scale 
+    performance and flow dynamics, as well as numerical model validation.
+    It was demonstrated that the performance of the cross-flow
     turbine becomes essentially $Re$-independent at a Reynolds number based on
@@
 physical parameters are matched. In the case of fluid systems, the most
-important nondimensional parameter is often the Reynolds number, which
+important nondimensional parameter is often the Reynolds number, $Re$, which
 quantifies the importance of inertial forces over viscous forces on the flow
@@
 \begin{equation}
-    \frac{Ul}{\nu},
+    Re = \frac{Ul}{\nu},
     \label{eq:Re}
@@
 straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
-\times 10^5$ \cite{Bravo2007}. Bachant and Wosnik \cite{Bachant2014} observed
+\times 10^5$ \cite{Bravo2007}. Bachant and Wosnik \cite{Bachant2014} first 
+observed
 Reynolds number independence of power coefficient at optimal tip speed ratio for
@@
 \hline
-0.3 & $3.0 \times 10^5$ & $8.0 \times 10^4$ \\
-0.4 & $4.0 \times 10^5$ & $1.1 \times 10^5$ \\
-0.5 & $5.0 \times 10^5$ & $1.3 \times 10^5$ \\
-0.6 & $6.0 \times 10^5$ & $1.6 \times 10^5$ \\
-0.7 & $7.0 \times 10^5$ & $1.9 \times 10^5$ \\
-0.8 & $0.8 \times 10^5$ & $2.1 \times 10^5$ \\
-0.9 & $0.9 \times 10^5$ & $2.4 \times 10^5$ \\
+0.3 & $0.3 \times 10^6$ & $0.8 \times 10^5$ \\
+0.4 & $0.4 \times 10^6$ & $1.1 \times 10^5$ \\
+0.5 & $0.5 \times 10^6$ & $1.3 \times 10^5$ \\
+0.6 & $0.6 \times 10^6$ & $1.6 \times 10^5$ \\
+0.7 & $0.7 \times 10^6$ & $1.9 \times 10^5$ \\
+0.8 & $0.8 \times 10^6$ & $2.1 \times 10^5$ \\
+0.9 & $0.9 \times 10^6$ & $2.4 \times 10^5$ \\
 1.0 & $1.0 \times 10^6$ & $2.7 \times 10^5$ \\
@@
 keep total streamwise force from changing as much as the rotor torque. The
-tendency for $C_D$ to continue increasing with $Re$ may also be a consequence of
-increasing Froude number, which therefore increases free surface deformation and
+tendency for $C_D$ to continue increasing with $Re$ may also include the 
+effects of an 
+increasing Froude number, which increases free surface deformation and
 wave drag during towing without increasing flow through the turbine.
@@
 based on blade chord of $Re_c > 2 \times 10^5$. Is it also shown from this
-metric that foils with higher camber are more Reynolds number dependent but have
+metric that foils with higher camber are more Reynolds number dependent at 
+lower Reynolds number but have
 smaller slopes in the linear regime. The Reynolds number dependence of the peak
@@
 magnitude smaller than other transport processes, viscous diffusion increased
-rapidly with decreasing $Re$. Overall, the total wake transport measured leveled
+rapidly with decreasing $Re$, and is expected to become significant to the wake 
+dynamics around $Re_D \sim 10^4$.
+Overall, the total wake transport measured leveled
 off at essentially the same Reynolds number that the performance did.
@@
 The authors would like to acknowledge funding through a National Science
-Foundation CAREER award (PI Wosnik, NSF-CBET 1150797, program manager Dr. Ram
-Gupta), a grant through the Leslie S. Hubbard Marine Program Endowment to
+Foundation CAREER award (PI Wosnik, NSF-CBET 1150797, program manager Dr. 
+Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine Program 
+Endowment to
 purchase acoustic flow measurement instrumentation, and a grant for laboratory

### 2015-10-31 paper.tex
Dependence -> effects

@@
 \abstract{The performance and wake characteristics of cross-flow (vertical axis)
-    turbines depend on many parameters: solidity, blade profile, blade pitch, 
-    number of blades, strut drag, operational parameters such as tip speed 
-    ratio---and Reynolds number.  
-    Experiments were performed with a large laboratory-scale high
-    solidity cross-flow turbine to investigate Reynolds number dependence of
+    turbines depend on many parameters: solidity, blade profile, blade pitch, number
+    of blades, strut drag, operational parameters such as tip speed ratio---and
+    Reynolds number. Experiments were performed with a large laboratory-scale high
+    solidity cross-flow turbine to investigate Reynolds number effects on
     performance and wake transport characteristics. A key question was what an
@@
     cross-flow turbines, both from the perspective of predicting full scale
-    performance and flow dynamics, as well as numerical model validation.
-    It was demonstrated that the performance of the cross-flow
-    turbine becomes essentially $Re$-independent at a Reynolds number based on
-    turbine diameter of $Re_D \approx 10^6$, or an approximate Reynolds number based
-    on blade chord of $Re_c \approx 2 \times 10^5$. A simple model that calculates
-    peak torque coefficient from static foil data and cross-flow turbine kinematics
-    was shown to be a reasonable estimator for predicting Reynolds number dependence
-    of an actual cross-flow turbine operating under dynamic conditions.}
+    performance and flow dynamics, as well as numerical model validation. It was
+    demonstrated that the performance of the cross-flow turbine becomes essentially
+    $Re$-independent at a Reynolds number based on turbine diameter of $Re_D \approx
+    10^6$, or an approximate Reynolds number based on blade chord of $Re_c \approx 2
+    \times 10^5$. A simple model that calculates peak torque coefficient from static
+    foil data and cross-flow turbine kinematics was shown to be a reasonable
+    estimator for predicting Reynolds number dependence of an actual cross-flow
+    turbine operating under dynamic conditions.}
 

### 2015-10-31 paper.tex
Text wrapping

@@
 % Abstract (Less than 200 words)
-\abstract{The performance and wake characteristics of cross-flow (vertical axis)
-    turbines depend on many parameters: solidity, blade profile, blade pitch, number
-    of blades, strut drag, operational parameters such as tip speed ratio---and
-    Reynolds number. Experiments were performed with a large laboratory-scale high
-    solidity cross-flow turbine to investigate Reynolds number effects on
-    performance and wake transport characteristics. A key question was what an
-    acceptable Reynolds number mismatch is for scaled physical model tests of
-    cross-flow turbines, both from the perspective of predicting full scale
+\abstract{The performance and wake characteristics of cross-flow (often
+    vertical-axis) turbines depend on many parameters: solidity, blade profile,
+    blade pitch, number of blades, strut drag, operational parameters such as tip
+    speed ratio---and Reynolds number. Experiments were performed with a large
+    laboratory-scale high solidity cross-flow turbine to investigate Reynolds number
+    effects on performance and wake transport characteristics. A key question was
+    what an acceptable Reynolds number mismatch is for scaled physical model tests
+    of cross-flow turbines, both from the perspective of predicting full scale
     performance and flow dynamics, as well as numerical model validation. It was
@@
 \times 10^5$ \cite{Bravo2007}. Bachant and Wosnik \cite{Bachant2014} first
-observed
-Reynolds number independence of power coefficient at optimal tip speed ratio for
-a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
+observed Reynolds number independence of power coefficient at optimal tip speed
+ratio for a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
 $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.

### 2015-10-31 paper.tex
Wording tweak

@@
 rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
-is again a function of turbine solidity $Nc/\pi D$, which allows for the use of
-a simpler Reynolds number based on turbine diameter $Re_D \equiv U_\infty
+in general decreases with turbine solidity $Nc/\pi D$, which allows for the use
+of a simpler Reynolds number based on turbine diameter $Re_D \equiv U_\infty
 D/\nu$.

### 2015-10-31 paper.tex
Move around intro a bit

@@
 
+If numerical models are validated with physical model data that was obtained at
+insufficiently high Reynolds numbers, it cannot be determined whether problems
+with model predictions matching full scale observations are caused by Reynolds
+number effects or issues with the numerical model itself. One might also obtain
+different predictions for scaled systems due to modified physics at lower
+Reynolds numbers, e.g., laminar boundary layers and transitional flow regimes.
+If the numerical model is then considered ``validated,'' there is a risk that
+its application at full-scale may produce incorrect predictions. As such, it is
+questionable whether numerical models validated with physical model data
+obtained at Reynolds numbers orders of magnitude away from prototype scale
+should be considered validated at all.
+
+Uncertainty from scale effects can potentially be overcome by showing that a
+scaled physical model test has become Reynolds number independent, and therefore
+validation efforts should be relevant at full scale. It is therefore of interest
+how small---since physical model size or facility flow speed generally correlate
+with cost---the model can be while remaining a reliable predictor of full-scale
+behavior. At the very least, it is necessary to test at a scale where results
+can be extrapolated, i.e., changes are occurring quasi-linearly without dramatic
+changes in the underlying physical principles governing the behavior of
+interest, e.g., power production in the case of a wind or MHK turbine.
+
 The performance and wake characteristics of cross-flow (vertical-axis) turbines
@@
 D/\nu$.
-
-If numerical models are validated with physical model data that was obtained at
-insufficiently high Reynolds numbers, it cannot be determined whether problems
-with model predictions are caused by Reynolds number effects, other turbine
-geometrical or operating parameters, or issues with the numerical model itself.
-One might also obtain different predictions for scaled systems due to modified
-physics at lower Reynolds numbers, e.g., laminar boundary layers and
-transitional flow regimes. If the numerical model is then considered
-``validated,'' there is a risk that its application at full-scale may produce
-incorrect predictions. As such, it is questionable whether numerical models
-validated with physical model data obtained at Reynolds numbers orders of
-magnitude away from prototype scale should be considered validated at all.
-
-Uncertainty from scale effects can potentially be overcome by showing that a
-scaled physical model test has become Reynolds number independent, and therefore
-validation efforts should be relevant at full scale. It is therefore of interest
-how small---since physical model size or facility flow speed generally correlate
-with cost---the model can be while remaining a reliable predictor of full-scale
-behavior. At the very least, it is necessary to test at a scale where results
-can be extrapolated, i.e., changes are occurring quasi-linearly without dramatic
-changes in the underlying physical principles governing the behavior of
-interest, e.g., power production in the case of a wind or MHK turbine.
 

### 2015-10-31 paper.tex
Streamline intro a bit

@@
 there still is concern at which scale an experiment must be performed to be
-relevant to full-scale application. This is particularly important for
-cross-flow turbines, where Reynolds number effects for experiments conducted at
-low $Re$ are often difficult to distinguish from the effects on performance of
-several other turbine parameters.
-
-If numerical models are validated with physical model data that was obtained at
-insufficiently high Reynolds numbers, it cannot be determined whether problems
-with model predictions matching full scale observations are caused by Reynolds
-number effects or issues with the numerical model itself. One might also obtain
-different predictions for scaled systems due to modified physics at lower
-Reynolds numbers, e.g., laminar boundary layers and transitional flow regimes.
-If the numerical model is then considered ``validated,'' there is a risk that
-its application at full-scale may produce incorrect predictions. As such, it is
-questionable whether numerical models validated with physical model data
-obtained at Reynolds numbers orders of magnitude away from prototype scale
-should be considered validated at all.
+relevant to full-scale application. One might obtain different predictions for
+scaled systems due to modified physics at lower Reynolds numbers, e.g., laminar
+boundary layers and transitional flow regimes. If the numerical model is then
+considered ``validated,'' there is a risk that its application at full-scale may
+produce incorrect predictions. As such, it is questionable whether numerical
+models validated with physical model data obtained at Reynolds numbers orders of
+magnitude away from prototype scale should be considered validated at all.
 

### 2015-10-31 paper.tex
Re sensitivity

@@
 layers. As mentioned previously, the dynamic stall vortex shedding is not
-expected to have a large Reynolds number sensitivity, though a larger separation
+expected to be highly sensitive to Reynolds number, though a larger separation
 bubble at lower $Re$ may induce higher levels of turbulence as shed vortices

### 2015-11-02 scripts/zip.py
Add script to zip up source files

@@ -1,0 +1,22 @@
+#!/usr/bin/env python
+"""
+This script zips up all the necessary source files to compile the paper.
+"""
+
+from zipfile import ZipFile
+from subprocess import check_output
+import os
+
+# Describe version via git
+version = check_output(["git", "describe"]).strip().decode()
+
+figures = ["figures/" + f for f in os.listdir("figures")
+           if f not in ["SolidWorks", "McMasters-Henderson-1980"]]
+
+files = ["paper.tex", "library.bib", "mdpi.bst", "mdpi.cls"]
+
+with ZipFile("archive/paper-{}.zip".format(version), "w") as f:
+    for fig in figures:
+        f.write(fig)
+    for file in files:
+        f.write(file)

### 2015-12-11 paper.tex
Wrap title
Add recovery term formulas; progress on #59

@@
 % Full title of the paper (Capitalized)
-\Title{Effects of Reynolds Number on the Energy Conversion and Near-Wake Dynamics of a
-High Solidity Vertical-Axis Cross-Flow Turbine}
+\Title{Effects of Reynolds Number on the Energy Conversion and Near-Wake
+    Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine}
 
@@
 The relative importance of various physical processes on mean streamwise
-momentum and kinetic energy transport/recovery in the streamwise direction are
+momentum and kinetic energy transport/recovery in the streamwise direction were
+assessed by examining the governing equations rearranged to isolate the
+streamwise partial derivative ($\partial / \partial x$) for the quantities of
+interest \cite{Bachant2015-JoT}, i.e. momentum:
+
+\begin{equation}
+	\begin{split}
+		\frac{\p U}{\p x}  =  
+		\frac{1}{U} \bigg{[}
+		& - V\frac{\p U}{\p y}
+		- W\frac{\p U}{\p z} \\
+		& -\frac{1}{\rho}\frac{\p P}{\p x} \\
+		& - \frac{\p}{\p x} \overline{u'u'}
+		- \frac{\p}{\p y} \overline{u'v'}
+		- \frac{\p}{\p z} \overline{u'w'} \\
+		& + \nu\left(\frac{\p^2 U}{\p x^2}
+		+ \frac{\p^2 U}{\p y^2}
+		+ \frac{\p^2 U}{\p z^2} \right)
+		\bigg{]}.
+	\end{split}
+	\label{eq:momentum}
+\end{equation}
+and mean kinetic energy:
+\begin{equation}
+	\begin{split}
+		\frac{\p K}{\p x}
+		=
+		\frac{1}{U}
+		\bigg{[}
+		& - \underbrace{V \frac{\p K}{\p y}}_{y\text{-adv.}}
+		- \underbrace{W \frac{\p K}{\p z}}_{z\text{-adv.}}
+		% Pressure work:
+		- \frac{1}{\rho}\frac{\p}{\p x_j} P U_i \delta_{ij}
+		% Work by viscous forces
+		+ \frac{\p}{\p x_j} 2 \nu U_i S_{ij} % Not sure if that's capital S...
+		% Turbulent transport of K
+		- \underbrace{
+		\frac{1}{2}\frac{\p}{\p x_j} \overline{u_i' u_j'} U_i
+		}_{\text{Turb. trans.}} \\
+		% Production of k
+		& + 
+		\underbrace{
+		\overline{u_i' u_j'} \frac{\p U_i}{\p x_j}
+		}_{k\text{-prod.}}
+		% Mean dissipation
+		- 
+		\underbrace{
+		2 \nu S_{ij}S_{ij}
+		}_{\text{Mean diss.}}
+		\bigg{]}.
+	\label{eq:K-full}
+	\end{split}
+\end{equation}
+
+Results from the terms available (i.e. excluding $\partial / \partial x$) from
+the right hand sides of Equations~\ref{eq:momentum} and \ref{eq:K-full} are
 plotted in Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph},

### 2015-12-12 paper.tex
Add discussion of wake recovery calculations, blockage; resolves #59

@@
 plotted in Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph},
-respectively. We note that similar to \cite{Bachant2015-JoT}, the vertical
-advection at this point in the wake is the dominant contributor to positive wake
-recovery, caused by the unique vortex pattern created by the blade tips and
-wakes.
+respectively. Derivatives were computed using a second order finite difference
+scheme, with central differencing for interior points, and inward facing schemes
+for the boundaries. A weighted average for each term was then calculated based
+on the grid spacing. We note that similar to \cite{Bachant2015-JoT}, the
+vertical advection at this point in the wake is the dominant contributor to
+positive wake recovery, caused by the unique vortex pattern created by the blade
+tips and wakes.
 
@@
 
+Transport due to cross-stream advection appears to become more negative at
+higher $Re$. This could be a consequence of increasing free surface deformation
+(higher $Fr$), which then decreases the effective flow cross-sectional area
+downstream of the turbine, forcing more flow to accelerate around the sides of
+the turbine.
+
 \begin{figure}[ht!]
     \centering
+
     \includegraphics[width=0.9\textwidth]{figures/mom_bar_graph}
-\caption{Momentum transport quantities.}
+
+    \caption{Normalized momentum transport quantities computed as weighted
+        averages from Equation~\ref{eq:momentum}.}
+
     \label{fig:mom-bar-graph}
@@
     \centering
+
     \includegraphics[width=0.9\textwidth]{figures/K_trans_bar_graph}
-\caption{Mean kinetic energy transport quantities.}
+
+    \caption{Normalized mean kinetic energy transport quantities computed as
+        weighted averages based on Equation~\ref{eq:K-full}. Note that directions of
+        the turbulent transport terms refer to the directions of their partial
+        derivatives.}
+
     \label{fig:K-bar-graph}
@@
 based on blade chord of $Re_c > 2 \times 10^5$. Is it also shown from this
-metric that foils with higher camber are more Reynolds number dependent at 
-lower Reynolds number but have
-smaller slopes in the linear regime. The Reynolds number dependence of the peak
-torque coefficient calculated from static foil data and cross-flow turbine
-kinematics was shown to be a reasonable indicator for predicting Reynolds number
-dependence of an actual cross-flow turbine operating under dynamic conditions,
-including dynamic stall.
+metric that foils with higher camber are more Reynolds number dependent at lower
+Reynolds number but have smaller slopes in the linear regime. The Reynolds
+number dependence of the peak torque coefficient calculated from static foil
+data and cross-flow turbine kinematics was shown to be a reasonable indicator
+for predicting Reynolds number dependence of an actual cross-flow turbine
+operating under dynamic conditions, including dynamic stall.
 
@@
 rapidly with decreasing $Re$, and is expected to become significant to the wake
-dynamics around $Re_D \sim 10^4$.
-Overall, the total wake transport measured leveled
-off at essentially the same Reynolds number that the performance did.
+dynamics around $Re_D \sim 10^4$. Overall, the total wake transport measured
+leveled off at essentially the same Reynolds number that the performance did.
 

### 2015-12-12 paper.tex
Reference figure 10 for peak C_T metric; resolves #66
Add explicit algorithm to peak C_T metric; resolves #67

@@
 C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
+\label{eq:ct}
 \end{equation}
@@
 C_c = C_l \sin \alpha - C_d \cos \alpha.
+\label{eq:cc}
 \end{equation}
@@
 
-From the peak torque coefficient metric, the Reynolds number independence is
-achieved at lower values, and more dramatically. We see that the trend of the
+From the peak torque coefficient metric plotted in
+Figure~\ref{fig:foils-C_T-Re-dep}, the Reynolds number independence is achieved
+at lower values, and more dramatically. We see that the trend of the
 (non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
@@
 We propose a method for predicting Reynolds number dependence of a cross-flow
-turbine using static airfoil characteristics and turbine blade kinematics. It is
-shown that the peak torque coefficient computed this way shows similar Reynolds
-number sensitivity---a better predictor than conventional quantifications of
-airfoil performance, e.g., lift-to-drag ratio. The simple peak geometric torque
-coefficient model predicted Reynolds number independence for Reynolds numbers
-based on blade chord of $Re_c > 2 \times 10^5$. Is it also shown from this
-metric that foils with higher camber are more Reynolds number dependent at lower
-Reynolds number but have smaller slopes in the linear regime. The Reynolds
-number dependence of the peak torque coefficient calculated from static foil
-data and cross-flow turbine kinematics was shown to be a reasonable indicator
-for predicting Reynolds number dependence of an actual cross-flow turbine
-operating under dynamic conditions, including dynamic stall.
+turbine using static airfoil characteristics and turbine blade kinematics:
+\begin{enumerate}
+    \item Acquire static lift and drag coefficient data for the desired blade
+    profile.
+
+    \item For azimuthal angles of 0--180 degrees and a given tip speed ratio,
+    calculate the geometric angle of attack and relative velocity magnitude from
+    the blade and undisturbed free stream velocity ($\equiv 1$) vectors.
+
+    \item With arrays of geometric angle of attack and relative velocity
+    magnitude, calculate the blade chordwise force and then rotor torque
+    coefficient from Equations~\ref{eq:cc} and \ref{eq:ct}, respectively.
+
+    \item Extract the maximum value of $C_T$, and repeat this process for static
+    foil data at multiple $Re$.
+\end{enumerate}
+
+It was shown that the peak torque coefficient computed this way shows similar
+Reynolds number sensitivity---a better predictor than conventional
+quantifications of airfoil performance, e.g., lift-to-drag ratio. The simple
+peak geometric torque coefficient model predicted Reynolds number independence
+for Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$. Is it also
+shown from this metric that foils with higher camber are more Reynolds number
+dependent at lower Reynolds number but have smaller slopes in the linear regime.
+The Reynolds number dependence of the peak torque coefficient calculated from
+static foil data and cross-flow turbine kinematics was shown to be a reasonable
+indicator for predicting Reynolds number dependence of an actual cross-flow
+turbine operating under dynamic conditions, including dynamic stall.
 

### 2015-12-12 paper.tex
Note that peak C_T method is not restricted to standard symmetrical airfoils; resolves #68
Text wrapping
Define Re in terms of Fr; resolves #62

@@
 real-world systems. The principle of dynamic similarity allows for geometrically
-scaled systems to be dynamically identical if certain nondimensional
-physical parameters are matched. In the case of fluid systems, the most
-important nondimensional parameter is often the Reynolds number, $Re$, which
-quantifies the importance of inertial forces over viscous forces on the flow
-physics:
-\begin{equation}
-    Re = \frac{Ul}{\nu},
-    \label{eq:Re}
-\end{equation}
-where $U$ and $l$ are characteristic velocity and length scales, respectively,
-and $\nu$ is the fluid kinematic viscosity. In general the geometric scale of a
-physical model is not necessarily the same as its dynamical scale. As such,
-hereafter we will use the word ``scale'' to refer to this dynamical scale or
-Reynolds number.
+scaled systems to be dynamically identical if certain nondimensional physical
+parameters are matched. In the case of fluid systems, the most important
+nondimensional parameter is often the Reynolds number, $Re$, which quantifies
+the importance of inertial forces over viscous forces on the flow physics: $Re
+\equiv Ul/\nu$, where $U$ and $l$ are characteristic velocity and length scales,
+respectively, and $\nu$ is the fluid kinematic viscosity. The other common
+dynamical scale---for systems with a free surface---is the Froude number $Fr
+\equiv U/\sqrt{gl}$, where $g$ is the gravitational acceleration. Matching both
+Froude and Reynolds number is difficult for the same fluid, which is illustrated
+by the relation $Re = l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with
+$l$, or the geometric scale. In general the geometric scale of a physical model
+is not necessarily the same as its dynamical scale. As such, hereafter we will
+use the word ``scale'' to refer to this dynamical scale or Reynolds number and
+assume Froude number is small enough to be negligible.
 
@@
 keep total streamwise force from changing as much as the rotor torque. The
-tendency for $C_D$ to continue increasing with $Re$ may also include the 
-effects of an 
-increasing Froude number, which increases free surface deformation and
+tendency for $C_D$ to continue increasing with $Re$ may also include the effects
+of an increasing Froude number, which increases free surface deformation and
 wave drag during towing without increasing flow through the turbine.
@@
 Reynolds number sensitivity---a better predictor than conventional
-quantifications of airfoil performance, e.g., lift-to-drag ratio. The simple
-peak geometric torque coefficient model predicted Reynolds number independence
-for Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$. Is it also
+quantifications of airfoil performance, e.g., lift-to-drag ratio. Note that this
+method is not restricted to standard symmetrical NACA foils. The simple peak
+geometric torque coefficient model predicted Reynolds number independence for
+Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$. Is it also
 shown from this metric that foils with higher camber are more Reynolds number

### 2015-12-12 paper.tex
Add parentheses to solidity def

@@
 rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
-in general decreases with turbine solidity $Nc/\pi D$, which allows for the use
-of a simpler Reynolds number based on turbine diameter $Re_D \equiv U_\infty
+in general decreases with turbine solidity $Nc/(\pi D)$, which allows for the
+use of a simpler Reynolds number based on turbine diameter $Re_D \equiv U_\infty
 D/\nu$.

### 2015-12-12 paper.tex
Qualify what high solidity means; resolves #63

--- 2015-12-12 paper.tex
@@
 rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
-in general decreases with turbine solidity $Nc/\pi D$, which allows for the use
-of a simpler Reynolds number based on turbine diameter $Re_D \equiv U_\infty
-D/\nu$.
+in general decreases with turbine solidity $Nc/(\pi D)$ \cite{Templin1974},
+which allows for the use of a simpler Reynolds number based on turbine diameter
+$Re_D \equiv U_\infty D/\nu$.
+
+Solidity often directly correlates with the chord-to-radius ratio $c/R$. Rotors
+with $c/R \ge 0.1$ are considered high solidity \cite{Fiedler2009}, for which
+so-called flow curvature or virtual camber effects become significant
+\cite{Migliore1980}. These effects arise from the blade sections' circular
+paths, and can complicate the comparison with behavior of an airfoil in a linear
+flow.
 

### 2015-12-13 paper.tex
Citations directly follow names; resolves #64

@@
 
-Blackwell et al. investigated the effects of Reynolds number on the performance
-of a 2 m diameter Darrieus vertical-axis cross-flow wind turbine with NACA 0012
-blades \cite{Blackwell1976}. By varying wind tunnel speed, the turbine was made
-to operate at approximately constant blade chord Reynolds number $Re_c$ ranging
-from $1 \times 10^5$ to $3 \times 10^5$. In this regime the turbine power
-coefficient $C_P$ was shown to be sensitive to $Re_c$, with sensitivity
+Blackwell et al. \cite{Blackwell1976} investigated the effects of Reynolds
+number on the performance of a 2 m diameter Darrieus vertical-axis cross-flow
+wind turbine with NACA 0012 blades. By varying wind tunnel speed, the turbine
+was made to operate at approximately constant blade chord Reynolds number $Re_c$
+ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime the turbine
+power coefficient $C_P$ was shown to be sensitive to $Re_c$, with sensitivity
 diminishing at the higher Reynolds numbers, especially for turbines of lower
 solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
-maximum radius). More recently, Polagye and Cavagnaro observed significant
-Reynolds number sensitivity for a high solidity helical cross-flow turbine
-\cite{Polagye2013b}, and Bravo et al. observed power coefficient of a
-straight-bladed VAWT to become Reynolds number independent at $Re_c \approx 4
-\times 10^5$ \cite{Bravo2007}. Bachant and Wosnik \cite{Bachant2014} first
-observed Reynolds number independence of power coefficient at optimal tip speed
-ratio for a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
-$Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.
+maximum radius). More recently, Polagye and Cavagnaro \cite{Polagye2013b}
+observed significant Reynolds number sensitivity for a high solidity helical
+cross-flow turbine, and Bravo et al. \cite{Bravo2007} observed power coefficient
+of a straight-bladed VAWT to become Reynolds number independent at $Re_c \approx
+4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} first observed Reynolds
+number independence of power coefficient at optimal tip speed ratio for a high
+solidity cross-flow turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}}
+\approx 2 \times 10^5$.
 
@@
 
-Krogstad and Adaramola observed Reynolds number independence of the performance
-of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5 \times 10^5$ in
-wind tunnel tests \cite{Krogstad2012a}. Their turbine blades had an S826 profile
-along their entire span, a profile chosen for its $Re$-independence. Walker et
-al. \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
-turbine tests in a towing tank, which were performed at a mid-span Reynolds
-number $Re_c = 4 \times 10^5$. McTavish et al. \cite{McTavish2013} showed how
-the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
-numbers, concluding that physical models should be designed with
-$Re$-independence in mind if they are to be run at low $Re$.
+Krogstad and Adaramola \cite{Krogstad2012a} observed Reynolds number
+independence of the performance of a 0.9 m diameter axial-flow turbine rotor at
+$Re_D \approx 5 \times 10^5$ in wind tunnel tests. Their turbine blades had an
+S826 profile along their entire span, a profile chosen for its
+$Re$-independence. Walker et al. \cite{Walker2014} similarly chose a NACA 63-618
+foil for their axial-flow turbine tests in a towing tank, which were performed
+at a mid-span Reynolds number $Re_c = 4 \times 10^5$. McTavish et al.
+\cite{McTavish2013} showed how the near-wake expansion for an axial-flow rotor
+is increased at higher Reynolds numbers, concluding that physical models should
+be designed with $Re$-independence in mind if they are to be run at low $Re$.
 

### 2015-12-13 paper.tex
Add uncertainty analysis section; resolves #48

@@
 \cite{Bachant2015-RVAT-Re-dep-data}.
+
+\subsection{Uncertainty Analysis}
+
+Uncertainty was considered from a combination of systematic and random errors.
+The random error was inferred from the sample standard deviation (on a
+per-revolution basis) and the systematic from the sensor calibrations or
+datasheets. Combining both sources of error, along with their propagation into
+quantities derived from multiple measurements, followed the procedures outlined
+in Coleman and Steele \cite{ColemanSteele}, described below.
+
+An expanded uncertainty interval with 95\% confidence was computed for
+$C_P$, $C_D$, and mean wake velocities
+\begin{equation}
+    U_{95} = t_{95} u_c,
+\end{equation}
+where $t_{95}$ is the value from the Student-$t$ distribution for a 95\%
+confidence interval and $u_c$ is the combined standard uncertainty. Combined
+standard uncertainty for a given quantity $X$ is calculated by
+\begin{equation}
+    u_X^2 = s_{\bar{X}}^2 + b_X^2,
+\end{equation}
+where $s_{\bar{X}}$ is the sample standard deviation of the mean per turbine
+revolution, and $b_X$ is the systematic uncertainty, computed by
+\begin{equation}
+    b_{X}^2 = \sum_{i=1}^J \left( \frac{\partial X}{\partial x_i} \right)^2
+    b_{x_i}^2,
+\end{equation}
+where $x_i$ is a primitive quantity used to calculate $X$ (e.g. $T$, $\omega$,
+and $U_\infty$ for calculating $C_P$), and $b_{x_i}$ is the primitive quantity's
+systematic uncertainty, estimated as half the value listed on the sensor
+manufacturer's documentation.
+
+Selecting $t_{95}$ requires an estimate for degrees of freedom $\nu_X$, which
+was obtained using the Welch--Satterthwaite formula
+\begin{equation}
+    \nu_X = \frac{\left(s_X^2 + \sum_{k=1}^M b_k^2 \right)^2} {s_X^4/\nu_{s_X} +
+    \sum_{k=1}^M b_k^4/\nu_{b_k}},
+\end{equation}
+where $\nu_{s_X}$ is the number of degrees of freedom associated with $s_X$ and
+$\nu_{b_k}$ is the number of degrees of freedom associated with $b_k$.
+$\nu_{s_X}$ is assumed to be $(N-1)$, where $N$ is the number of independent
+samples (or turbine revolutions). $\nu_{b_k}$ was estimated as
+\begin{equation}
+    \nu_{b_k} = \frac{1}{2} \left( \frac{\Delta b_k}{b_k} \right)^{-2},
+\end{equation}
+where the quantity in parentheses is the relative uncertainty of $b_k$, assumed
+to be 0.25.
+
+For the Reynolds number dependence of turbine performance, error bars are
+included on the plots. Expanded uncertainty estimates for mean wake velocities
+(averaged over all runs) are listed in Table~\ref{tab:vel-unc}.
+
+\begin{table}
+    \centering
+\begin{tabular}{cccc}
+$U_\infty$ (m/s) &                $U$ &                $V$ &                $W$ \\
+\hline
+           $0.4$ & $1 \times 10^{-2}$ & $7 \times 10^{-3}$ & $6 \times 10^{-3}$ \\
+           $0.6$ & $1 \times 10^{-2}$ & $8 \times 10^{-3}$ & $8 \times 10^{-3}$ \\
+           $0.8$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
+           $1.0$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
+           $1.2$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
+\end{tabular}
+\caption{Average expanded uncertainty estimates (with 95\% confidence) for mean
+    velocity measurements at each tow speed.}
+
+\label{tab:vel-unc}
+\end{table}
 

### 2015-12-13 paper.tex
Add exp setup details; closes #55

@@
 Experiments were performed in a turbine test bed specifically designed for
-cross-flow turbines. The turbine model used in this study was the UNH Reference
-Vertical Axis Turbine (RVAT), which was designed to be a generic case for
-numerical model testing, similar to the Sandia National Labs/US Department of
-Energy RM2 River Turbine \cite{Neary2014}, but with higher solidity, or blade
-chord to radius ratio.
+cross-flow turbines. The test bed was integrated as part of the UNH tow tank,
+which is a 30 m long facility with a 3.66 m wide and 2.44 m deep cross-section.
+The turbine model used in this study was the UNH Reference Vertical Axis Turbine
+(RVAT), which was designed to be a generic case for numerical model testing,
+similar to the Sandia National Labs/US Department of Energy RM2 River Turbine
+\cite{Neary2014}, but with higher solidity, or blade chord-to-radius ratio.
 
@@
 Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
-coupled to a permanent magnet servo motor with a 20:1 gearbox, providing precise
-control over shaft RPM. This servo was controlled by the tow tank's main motion
-controller for high synchronization with the carriage motion, thereby giving
-precise measurement and control of tip speed ratio. A rotary torque transducer
-was installed inline between the servo and the turbine, and the servo was also
-mounted on a slewing ring bearing, which allowed a redundant measurement of
-torque via an arm and load cell used to counteract the turbine moment. The frame
-was mounted to the carriage via linear guides, such that the total streamwise
-drag force was transferred to a pair of S-beam load cells, providing the rotor
-drag measurements, after a measured tare drag was subtracted in post processing.
+coupled to a Kollmorgen AKM permanent magnet servo motor with a 20:1 gearbox,
+providing precise control over shaft RPM. This servo was controlled by the tow
+tank's main motion controller for high synchronization with the carriage motion,
+thereby giving precise measurement and control of tip speed ratio. An Interface
+T8 200 Nm capacity rotary torque transducer was installed inline between the
+servo and the turbine, and the servo was also mounted on a slewing ring bearing,
+which allowed a redundant measurement of torque via an arm and load cell used to
+counteract the turbine moment. The frame was mounted to the carriage via linear
+guides, such that the total streamwise drag force was transferred to a pair of
+Sentran ZB3 500 lbf capacity S-beam load cells, providing the rotor drag
+measurements, after a measured tare drag was subtracted in post processing.
 Similarly, a tare torque was measured by rotating the turbine shaft in air.
@@
     \centering
-\includegraphics[width=0.75\textwidth]{figures/exp_setup_drawing}
+
+    \includegraphics[width=0.8\textwidth]{figures/exp-setup-photo}
+
+    \includegraphics[width=0.8\textwidth]{figures/exp_setup_drawing}
+
     \caption{Experimental setup: turbine test bed installed in UNH tow tank.}
+
     \label{fig:exp-setup}
@@
 Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
-(ADV), sampling at 200 Hz. The probe was mounted on an automated positioning
-system, also controlled by the tow tank's main motion controller. The ADV and
-DAQ systems' sampling times were synchronized by triggering the start of data
-acquisition via a pulse sent from the motion controller. Additional details of
-the turbine and experimental setup are described in \cite{Bachant2015-JoT}.
+(ADV), which has a 6 mm diameter sampling volume, and sampled at 200 Hz. The
+probe was mounted on an automated positioning system, also controlled by the tow
+tank's main motion controller. The ADV and DAQ systems' sampling times were
+synchronized by triggering the start of data acquisition via a pulse sent from
+the motion controller. Additional details of the turbine and experimental setup
+are described in \cite{Bachant2015-JoT}.
 

### 2015-12-13 paper.tex
Find reference for stronger tip vortices at lower Re
Add citation for Re definition

@@
 nondimensional parameter is often the Reynolds number, $Re$, which quantifies
-the importance of inertial forces over viscous forces on the flow physics: $Re
-\equiv Ul/\nu$, where $U$ and $l$ are characteristic velocity and length scales,
-respectively, and $\nu$ is the fluid kinematic viscosity. The other common
-dynamical scale---for systems with a free surface---is the Froude number $Fr
-\equiv U/\sqrt{gl}$, where $g$ is the gravitational acceleration. Matching both
-Froude and Reynolds number is difficult for the same fluid, which is illustrated
-by the relation $Re = l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with
-$l$, or the geometric scale. In general the geometric scale of a physical model
-is not necessarily the same as its dynamical scale. As such, hereafter we will
-use the word ``scale'' to refer to this dynamical scale or Reynolds number and
-assume Froude number is small enough to be negligible.
+the importance of inertial forces over viscous forces on the flow physics
+\cite{Acheson1990}: $Re \equiv Ul/\nu$, where $U$ and $l$ are characteristic
+velocity and length scales, respectively, and $\nu$ is the fluid kinematic
+viscosity. The other common dynamical scale---for systems with a free
+surface---is the Froude number $Fr \equiv U/\sqrt{gl}$, where $g$ is the
+gravitational acceleration. Matching both Froude and Reynolds number is
+difficult for the same fluid, which is illustrated by the relation $Re = l^{3/2}
+g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$, or the geometric scale.
+In general the geometric scale of a physical model is not necessarily the same
+as its dynamical scale. As such, hereafter we will use the word ``scale'' to
+refer to this dynamical scale or Reynolds number and assume Froude number is
+small enough to be negligible.
 
@@
 
-We expect that lower lift on the blades at lower $Re$ will weaken tip vortex
-shedding, and decrease the levels of turbulence shed from the blade boundary
-layers. As mentioned previously, the dynamic stall vortex shedding is not
-expected to be highly sensitive to Reynolds number, though a larger separation
-bubble at lower $Re$ may induce higher levels of turbulence as shed vortices
-become unstable and break down.
+Despite lower lift on the blades at lower $Re$, we expect stronger tip vortex
+shedding \cite{Yoon2005}. As mentioned previously, the dynamic stall vortex
+shedding is not expected to be highly sensitive to Reynolds number, though a
+larger separation bubble at lower $Re$ may induce higher levels of turbulence as
+shed vortices become unstable and break down.
 

### 2015-12-13 paper.tex
General reference for CFT parameters

@@
 of blades, strut drag and operational parameters such as tip speed ratio---{\em
-    and} on Reynolds number. Note that an average blade chord Reynolds number,
-$Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be expressed in terms
-of tip speed ratio $\lambda \equiv \omega R/ U_\infty$, where $U_\infty$ is the
-free stream velocity, $c$ is the blade chord length, $\nu$ is the fluid
-kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$ is the
-rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
-in general decreases with turbine solidity $Nc/(\pi D)$ \cite{Templin1974},
-which allows for the use of a simpler Reynolds number based on turbine diameter
-$Re_D \equiv U_\infty D/\nu$.
+    and} on Reynolds number \cite{Para2002}. Note that an average blade chord
+Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be
+expressed in terms of tip speed ratio $\lambda \equiv \omega R/ U_\infty$, where
+$U_\infty$ is the free stream velocity, $c$ is the blade chord length, $\nu$ is
+the fluid kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$
+is the rotor radius. The value of $\lambda$ at which a turbine reaches peak
+performance in general decreases with turbine solidity $Nc/(\pi D)$
+\cite{Templin1974}, which allows for the use of a simpler Reynolds number based
+on turbine diameter $Re_D \equiv U_\infty D/\nu$.
 

### 2015-12-13 paper.tex
Tweak CFT wake studies in introduction

@@
 
-The wake of a cross-flow turbine in the dynamic stall regime has been studied
-via laser Doppler velocimetry by Brochier et al. \cite{Brochier1986} and with
-particle image velocimetry (PIV) by Fujisawa and Shibuya \cite{Fujisawa2001}.
-However, both of these studies were performed at very low Reynolds
-numbers---$Re_D = 10^4$ and $10^3$, respectively. Tescione et al.
-\cite{Tescione2014} studied the wake of a vertical-axis wind turbine in a wind
-tunnel using stereo PIV at an approximate blade chord Reynolds number $Re_c =
-1.7 \times 10^5$; very close to the $Re$-independent state reported in
-\cite{Bachant2014}, where 2-D near-wake profiles were measured for a high
-solidity cross-flow turbine at optimal tip speed ratio.
+The wake of a 2-D cross-flow turbine in the dynamic stall regime has been
+studied via laser Doppler velocimetry by Brochier et al. \cite{Brochier1986} and
+in 3-D with particle image velocimetry (PIV) by Fujisawa and Shibuya
+\cite{Fujisawa2001}. However, both of these studies were performed at very low
+Reynolds numbers---$Re_D = 10^4$ and $10^3$, respectively. Tescione et al.
+\cite{Tescione2014} studied the wake of a vertical-axis wind turbine at optimal
+tip speed ratio in a wind tunnel using stereo PIV at an approximate blade chord
+Reynolds number $Re_c = 1.7 \times 10^5$; very close to the $Re$-independence
+criteria reported in \cite{Bachant2014}, though it was not confirmed if this was
+indeed a $Re$-independent state. These studies motivate the systematic
+investigation of scale effects undertaken here.
 

### 2015-12-13 paper.tex
Add comment about Re for porous disks
Remove the word 'first'

@@
 of a straight-bladed VAWT to become Reynolds number independent at $Re_c \approx
-4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} first observed Reynolds
-number independence of power coefficient at optimal tip speed ratio for a high
-solidity cross-flow turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}}
-\approx 2 \times 10^5$.
+4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} observed Reynolds number
+independence of power coefficient at optimal tip speed ratio for a high solidity
+cross-flow turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2
+\times 10^5$.
 
@@
 wake generator may do a fine job simulating a scaled turbine, but how well can
-it simulate a full-scale device?
+it simulate a full-scale device? Note that for a porous disk, the Reynolds
+number of interest is diameter-based, since it is the scaling of the far-wake
+dynamics that matters.
 

### 2015-12-13 paper.tex
Better symbol definitions; closes #51

@@
 
-The behavior of the mean rotor drag coefficient is similar, though the changes
-are less dramatic. This is likely due to cross-flow turbine geometry, where
-increases in blade drag at lower $Re$ somewhat offsets the reduction in lift, to
-keep total streamwise force from changing as much as the rotor torque. The
-tendency for $C_D$ to continue increasing with $Re$ may also include the effects
-of an increasing Froude number, which increases free surface deformation and
-wave drag during towing without increasing flow through the turbine.
+The behavior of the mean rotor drag coefficient $C_D$ is similar, though the
+changes are less dramatic. This is likely due to cross-flow turbine geometry,
+where increases in blade drag at lower $Re$ somewhat offsets the reduction in
+lift, to keep total streamwise force from changing as much as the rotor torque.
+The tendency for $C_D$ to continue increasing with $Re$ may also include the
+effects of an increasing Froude number, which increases free surface deformation
+and wave drag during towing without increasing flow through the turbine.
 
@@
 
-Results from the XFOIL calculations are shown in Figure~\ref{fig:foil-Re-dep},
-where values of maximum lift coefficient, minimum drag coefficient, and maximum
+Results from the XFOIL calculations for the airfoil profiles are shown in
+Figure~\ref{fig:foil-Re-dep}, where values of maximum lift coefficient
+$C_{l_{\max}}$, minimum drag coefficient $C_{d_{\min}}$, and maximum
 lift-to-drag ratio are normalized (for comparison between foils) and plotted

### 2015-12-13 paper.tex
Compare and contrast CFT wake studies; closes #53

@@
 criteria reported in \cite{Bachant2014}, though it was not confirmed if this was
-indeed a $Re$-independent state. These studies motivate the systematic
-investigation of scale effects undertaken here.
+indeed a $Re$-independent state. The two lower $Re$ experiments tended to focus
+on the effects of dynamic stall, whereas the higher $Re$ experiment mostly
+concerned the mean velocity and tip vortex effects. Despite their value
+elucidating the complex wake dynamics of cross-flow turbines, these studies
+motivate the systematic investigation of scale effects undertaken here.
 

### 2015-12-14 paper.tex
Add spectra formulation; closes #58

@@
 
-Spectra of the cross-stream velocity normalized by the free stream are plotted
-in Figure~\ref{fig:wake-spectra} for regions on either side of the turbine. On
-the $-y$ side of the turbine there is broadband turbulence produced by blade
-stall, and on the $+y$ side there is a clear peak in the spectra caused by the
-blade passage. We see that on both sides there is higher spectral energy at
-lower Reynolds numbers. On the $+y$ side of the turbine we notice higher
-spectral energy at the blade passage frequency's first harmonic, or $6
-f_\mathrm{turbine}$. This is likely due to the blade's shed vorticity being less
-stable at higher $Re$.
+Spectra of the cross-stream velocity normalized by the free stream were
+computing using a fast Fourier transform-based method, applying a Hanning window
+and averaging over four adjacent frequency bands to decrease confidence
+intervals. These spectra are plotted in Figure~\ref{fig:wake-spectra} for
+regions on either side of the turbine. On the $-y$ side of the turbine there is
+broadband turbulence produced by blade stall, and on the $+y$ side there is a
+clear peak in the spectra caused by the blade passage. We see that on both sides
+there is higher spectral energy at lower Reynolds numbers. On the $+y$ side of
+the turbine we notice higher spectral energy at the blade passage frequency's
+first harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's
+shed vorticity being less stable at higher $Re$.
 
@@
     ($+0.75$ m in Figure~\ref{fig:wake-locations}) (b). Dashed vertical lines
-    indicate $[1, 3, 6, 9]$ times the turbine rotational frequency.}
+    indicate $[1, 3, 6, 9]$ times the turbine rotational frequency. Shaded regions
+    indicate 95\% confidence intervals assuming a $\chi^2$ distribution.}
 

### 2015-12-14 paper.tex
Shorten intro section on Re dependence; resolves #47

@@
 
-\subsection{Reynolds Number Effects}
+\subsection{Modes of Reynolds Number Dependence}
 
-In this study we are using a simple cross-flow turbine constructed from
-hydrofoils, i.e., the rotor is lift-driven in contrast to drag-driven. A logical
-starting point for understanding the effects of $Re$ on overall turbine loading
-is to first consider simplified cases of individual foils, and their static and
-dynamic loading.
-
-\subsubsection{On Static Foil Behavior}
-
-Hydro- or airfoil section performance is generally characterized by the
-lift-to-drag force ratio $L/D$ (or the ratio of lift and drag coefficients,
-$C_l/C_d$), which is primarily a function of the foil's angle of attack
-$\alpha$. The angles of attack for cross-flow turbine blades are a function of
-turbine tip speed ratio, induction (slowing/turning of the free stream flow
-prior to reaching the turbine), and turbine rotation or azimuthal angle
-$\theta$. Typically, static airfoil data shows that the static stall
-angle---the angle just beyond that at which $L/D$ is maximum---increases with
-blade chord Reynolds number $Re_c$ \cite{Jacobs1937}.
-
-A review of Reynolds number effects on airfoil behavior is presented in
-\cite{Lissaman1983}. In general, airfoil performance is enhanced as the boundary
-layer on the foil transitions to turbulence closer to the leading edge, which
-enables it to advance further downstream against the adverse pressure gradient
-on the suction side, delaying separation to higher angles of attack. For smooth
-airfoils, this transition can cause a dramatic increase in foil performance, at
-a blade chord Reynolds number on the order of $10^5$ \cite{McMasters1980}, shown
-in Figure~\ref{fig:McMasters}. Note that there is a distinct lack of highly
-reliable data for airfoils in this transitional regime and below. An evaluation
-of the various databases relevant to cross-flow turbines is presented in
-\cite{Bedon2014}.
+We are looking at how Reynolds number scaling affects both the blade loading,
+i.e., turbine performance, and the near-wake. Typically, static airfoil data
+shows that the static stall angle---the angle just beyond that at which the
+lift-to-drag ratio is maximum---increases with blade chord Reynolds number
+$Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil
+behavior is presented in \cite{Lissaman1983}. In general, airfoil performance is
+enhanced as the boundary layer on the foil transitions to turbulence closer to
+the leading edge, which enables it to advance further downstream against the
+adverse pressure gradient on the suction side, delaying separation to higher
+angles of attack. For smooth airfoils, this transition can cause a dramatic
+increase in foil performance, at a blade chord Reynolds number on the order of
+$10^5$ \cite{McMasters1980}, shown in Figure~\ref{fig:McMasters}. Note that
+there is a distinct lack of highly reliable data for airfoils in this
+transitional regime and below. An evaluation of the various databases relevant
+to cross-flow turbines is presented in \cite{Bedon2014}.
 
@@
 
-\subsubsection{On Dynamic Foil Behavior}
-
 Static foil performance does not tell the whole story for a cross-flow turbine.
 The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
-turbine implies dynamic loading. Furthermore, since the range of angle of attack
-encountered by a rotating blade increases with decreasing tip speed ratio, the
-blades commonly undergo dynamic stall for tip speed ratios at and below those of
-maximum rotor torque \cite{Para2002}. We then must understand the effects of
-Reynolds number on the dynamic stall process.
-
-Bousman \cite{Bousman2000-evaluation} states that dynamic stall is relatively
+turbine implies dynamic loading, encountering dynamic stall for tip speed ratios
+near and below those of maximum rotor torque \cite{Para2002}. Bousman
+\cite{Bousman2000-evaluation} states that dynamic stall is relatively
 insensitive to Reynolds number for $Re=1.0 \times 10^5$--$2.5 \times
@@
 dynamic stall remains an unsolved question.
-
-\subsubsection{On Wake Flows}
 
@@
 reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
-10^5$.
+10^5$. It is therefore our objective to observe similar scaling relationships
+for a cross-flow turbine near-wake.
 

### 2015-12-14 paper.tex
Tighten up first three paragraphs; resolves #45

@@
 there still is concern at which scale an experiment must be performed to be
-relevant to full-scale application. One might obtain different predictions for
-scaled systems due to modified physics at lower Reynolds numbers, e.g., laminar
-boundary layers and transitional flow regimes. If the numerical model is then
-considered ``validated,'' there is a risk that its application at full-scale may
-produce incorrect predictions. As such, it is questionable whether numerical
-models validated with physical model data obtained at Reynolds numbers orders of
-magnitude away from prototype scale should be considered validated at all.
-
-Uncertainty from scale effects can potentially be overcome by showing that a
-scaled physical model test has become Reynolds number independent, and therefore
-validation efforts should be relevant at full scale. It is therefore of interest
-how small---since physical model size or facility flow speed generally correlate
-with cost---the model can be while remaining a reliable predictor of full-scale
-behavior. At the very least, it is necessary to test at a scale where results
-can be extrapolated, i.e., changes are occurring quasi-linearly without dramatic
-changes in the underlying physical principles governing the behavior of
-interest, e.g., power production in the case of a wind or MHK turbine.
+relevant to full-scale application. Similarly, it is questionable whether
+numerical models validated with physical model data obtained orders of magnitude
+away from prototype scale should be considered validated at all.
 

### 2015-12-14 paper.tex
Discuss blockage ratio in experimental setup; towards #49

@@
 turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
-Figure~\ref{fig:turbine}. The rotor has a relatively high solidity $Nc/(\pi D) =
-0.13$ and large chord-to-radius ratio $c/R = 0.28$. A CAD model of the turbine
-is available from \cite{Bachant2014-RVAT-CAD}.
+Figure~\ref{fig:turbine}. The blockage ratio produced by the rotor's frontal
+area is 11\%, which we have decided not to correct for, as reliable methods are
+not yet agreed upon for CFTs \cite{Cavagnaro2014}. As such, blockage should be
+taken into account when using these data for model validation, i.e., the
+experimental domain should be mimicked. The rotor has a relatively high solidity
+$Nc/(\pi D) = 0.13$ and large chord-to-radius ratio $c/R = 0.28$. A CAD model of
+the turbine is available from \cite{Bachant2014-RVAT-CAD}.
 

### 2015-12-14 paper.tex
Discuss blockage in conclusions; resolves #49
Fix up conclusions a bit

@@
 quantifications of airfoil performance, e.g., lift-to-drag ratio. Note that this
-method is not restricted to standard symmetrical NACA foils. The simple peak
-geometric torque coefficient model predicted Reynolds number independence for
-Reynolds numbers based on blade chord of $Re_c > 2 \times 10^5$. Is it also
-shown from this metric that foils with higher camber are more Reynolds number
-dependent at lower Reynolds number but have smaller slopes in the linear regime.
-The Reynolds number dependence of the peak torque coefficient calculated from
-static foil data and cross-flow turbine kinematics was shown to be a reasonable
-indicator for predicting Reynolds number dependence of an actual cross-flow
-turbine operating under dynamic conditions, including dynamic stall.
+method is not restricted to standard symmetrical NACA foils, though those
+evaluated here with higher camber were more sensitive at lower $Re$ but had
+smaller slopes in the linear regime.
 
@@
 measured---showed little $Re$-dependence, whereas the negative effects of
-cross-stream advection was enhanced at high $Re$. Despite being order of
-magnitude smaller than other transport processes, viscous diffusion increased
-rapidly with decreasing $Re$, and is expected to become significant to the wake
-dynamics around $Re_D \sim 10^4$. Overall, the total wake transport measured
-leveled off at essentially the same Reynolds number that the performance did.
+cross-stream advection was enhanced at high $Re$, which may be due to larger
+free surface deformation reducing the near-wake's cross-sectional area. Despite
+being order of magnitude smaller than other transport processes, viscous
+diffusion increased rapidly with decreasing $Re$, and is expected to become
+significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the total
+wake transport measured leveled off at essentially the same Reynolds number that
+the performance did.
 
@@
 resolved, since transition to turbulence plays an important role in overall
-blade loading.
+blade loading. Note that in our case blockage, though reasonably low, may be
+increasing flow through the turbine compared with a free case, which could mean
+$Re$ thresholds for free flows could be slightly higher, though a reliable
+blockage correction algorithm has not yet been agreed upon for CFTs
+\cite{Cavagnaro2014}.
 
@@
 Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine Program
-Endowment to
-purchase acoustic flow measurement instrumentation, and a grant for laboratory
-infrastructure upgrades through the US Department of Energy.
+Endowment to purchase acoustic flow measurement instrumentation, and a grant for
+laboratory infrastructure upgrades through the US Department of Energy.
 

### 2015-12-14 paper.tex
Move scale mismatch phrase from abstract to intro

@@
     laboratory-scale high solidity cross-flow turbine to investigate Reynolds number
-    effects on performance and wake transport characteristics. A key question was
-    what an acceptable Reynolds number mismatch is for scaled physical model tests
-    of cross-flow turbines, both from the perspective of predicting full scale
-    performance and flow dynamics, as well as numerical model validation. It was
+    effects on performance and wake characteristics. It was
     demonstrated that the performance of the cross-flow turbine becomes essentially
@@
 performance of individual turbines, and design or investigate arrays of devices.
-Scaled models have the benefit of being significantly less expensive, however,
-there still is concern at which scale an experiment must be performed to be
-relevant to full-scale application. Similarly, it is questionable whether
-numerical models validated with physical model data obtained orders of magnitude
-away from prototype scale should be considered validated at all.
+Scaled models have the benefit of being significantly less expensive, however, a
+key question is whether or not an acceptable scale mismatch exists. Similarly,
+it is questionable whether numerical models validated with physical model data
+obtained orders of magnitude away from prototype scale should be considered
+validated at all.
 
@@
 
-At lower Reynolds numbers we observed lower levels of turbulence in the wake at
+At higher Reynolds numbers we observed lower levels of turbulence in the wake at
 $x/D=1$, along with lower levels of turbulent transport with respect to mean

### 2015-12-14 paper.tex
Add wake results to abstract; resolves #52

@@
 % Abstract (Less than 200 words)
-\abstract{The performance and wake characteristics of cross-flow (often
-    vertical-axis) turbines depend on many parameters: solidity, blade profile,
-    blade pitch, number of blades, strut drag, operational parameters such as tip
-    speed ratio---and Reynolds number. Experiments were performed with a large
-    laboratory-scale high solidity cross-flow turbine to investigate Reynolds number
-    effects on performance and wake characteristics. It was
-    demonstrated that the performance of the cross-flow turbine becomes essentially
-    $Re$-independent at a Reynolds number based on turbine diameter of $Re_D \approx
-    10^6$, or an approximate Reynolds number based on blade chord of $Re_c \approx 2
-    \times 10^5$. A simple model that calculates peak torque coefficient from static
-    foil data and cross-flow turbine kinematics was shown to be a reasonable
-    estimator for predicting Reynolds number dependence of an actual cross-flow
-    turbine operating under dynamic conditions.}
+\abstract{Experiments were performed with a large laboratory-scale high solidity
+    cross-flow turbine to investigate Reynolds number effects on performance and
+    wake characteristics, to seek scale thresholds for physical and numerical
+    modeling of individual devices and arrays. It was demonstrated that the
+    performance of the cross-flow turbine becomes essentially $Re$-independent at a
+    Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
+    approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
+    10^5$. A simple model that calculates peak torque coefficient from static foil
+    data and cross-flow turbine kinematics was shown to be a reasonable estimator
+    for predicting Reynolds number dependence of an actual cross-flow turbine
+    operating under dynamic conditions. Mean velocity and turbulence measurements in
+    the near-wake showed very subtle differences. However, when transport terms for
+    the streamwise momentum and mean kinetic energy were calculated, a similar $Re$
+    threshold was revealed. These results imply that physical model studies of
+    cross-flow turbines should be $Re_D \sim 10^6$ to properly approximate both the
+    performance and wake dynamics of full scale devices and arrays.}
 

### 2015-12-14 paper.tex
Show wake profiles to quantify wake differences; resolves #57

@@
 These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times
-10^6$, $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look
-strikingly similar, so they have not been plotted here. We will instead compare
-and contrast the wake behavior by examining spectra and wake transport terms in
-the equations that govern the downstream evolution of mean momentum and kinetic
-energy.
+10^6$, $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look very
+similar, so they have not been plotted here, though profiles at $z/H=0.0$ are
+shown in Figure~\ref{fig:profiles} to illustrate the subtlety of the
+differences. We will instead compare and contrast the wake behavior by examining
+spectra and wake transport terms in the equations that govern the downstream
+evolution of mean momentum and kinetic energy.
 
@@
     \label{fig:kcont}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+
+    \includegraphics[width=0.85\textwidth]{figures/mean_u_k_profiles}
+
+    \caption{Mean streamwise velocity (left) and turbulence kinetic energy
+        (right) profiles at $z/H=0.0$. Legend indicates $Re_D$.}
+
+    \label{fig:profiles}
 \end{figure}

### 2015-12-14 getfigs.py

@@ -1,6 +1,5 @@
 #!/usr/bin/env python
-"""
-This script copies figures from the RVAT Reynolds number dependence
+"""This script copies figures from the RVAT Reynolds number dependence
 experiment directory.
 """
 
@@ -19,7 +18,8 @@ figlist = ["cp_curves.pdf",
            "wake_spectra.pdf",
            "wake_trans_totals.pdf",
            "meancontquiv_10.pdf",
-           "k_contours_10.pdf"]
+           "k_contours_10.pdf",
+           "mean_u_k_profiles.pdf"]
 
 for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))

### 2015-12-15 paper.tex
Martin is also a corresponding author
Find example where extrapolating small scale experiments causes errors; resolves #61
Better explain why HAWT research is relevant; closes #46

@@
 % Authors (Add full first names)
-\Author{Peter Bachant $^{1,}$* and Martin Wosnik $^{1}$}
+\Author{Peter Bachant $^{1,}$* and Martin Wosnik $^{1}$*}
 
@@
 obtained orders of magnitude away from prototype scale should be considered
-validated at all.
+validated at all. An example of the errors that can result from the
+extrapolation of small scale experiments can be found in \cite{Baker1991}.
 
@@
 
-Krogstad and Adaramola \cite{Krogstad2012a} observed Reynolds number
-independence of the performance of a 0.9 m diameter axial-flow turbine rotor at
-$Re_D \approx 5 \times 10^5$ in wind tunnel tests. Their turbine blades had an
-S826 profile along their entire span, a profile chosen for its
-$Re$-independence. Walker et al. \cite{Walker2014} similarly chose a NACA 63-618
-foil for their axial-flow turbine tests in a towing tank, which were performed
-at a mid-span Reynolds number $Re_c = 4 \times 10^5$. McTavish et al.
-\cite{McTavish2013} showed how the near-wake expansion for an axial-flow rotor
-is increased at higher Reynolds numbers, concluding that physical models should
-be designed with $Re$-independence in mind if they are to be run at low $Re$.
+Scale effects on axial-flow or horizontal-axis wind turbines are better
+understood, and investigators have methods for compensating. Krogstad and
+Adaramola \cite{Krogstad2012a} observed Reynolds number independence of the
+performance of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5
+\times 10^5$ in wind tunnel tests. Their turbine blades had an S826 profile
+along their entire span, a profile chosen for its $Re$-independence. Walker et
+al. \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
+turbine tests in a towing tank, which were performed at a mid-span Reynolds
+number $Re_c = 4 \times 10^5$. McTavish et al. \cite{McTavish2013} showed how
+the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
+numbers, concluding that physical models should be designed with
+$Re$-independence in mind if they are to be run at low $Re$. It is uncertain,
+however, if these methods work equally well for cross-flow turbines.
 

### 2015-12-15 paper.tex
Minor wording change

@@
 Reynolds numbers, which is consistent with the physical meaning of the Reynolds
-number in general. From these results one might expect that the viscous effects
-will become significant to the wake dynamics around $Re_D \sim 10^4$.
+number itself. From these results one might expect that the viscous effects will
+become significant to the wake dynamics around $Re_D \sim 10^4$.
 

### 2015-12-16 scripts/diff.sh
Add script and instructions for creating diff

@@ -1,0 +1,6 @@
+#!/usr/bin/env bash
+
+rev=$1
+git show $rev:paper.tex > Archive/paper-$rev.tex
+latexdiff Archive/paper-$rev.tex paper.tex > paper-$rev-diff.tex
+# latexmk paper-$rev-diff.tex -pdf -f

### 2015-12-16 paper.tex
Use UTF-8 encoding and add linenumbers before abstract
Note link between higher drag coefficient and cross-stream advection; resolves #70
Wosnik edits to abstract

@@
 \usepackage{lineno}
+\usepackage[utf8]{inputenc}
 
@@
 
+\linenumbers
+
 % Abstract (Less than 200 words)
@@
     cross-flow turbine to investigate Reynolds number effects on performance and
-    wake characteristics, to seek scale thresholds for physical and numerical
+    wake characteristics, to establish scale thresholds for physical and numerical
     modeling of individual devices and arrays. It was demonstrated that the
@@
     Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
-    approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
-    10^5$. A simple model that calculates peak torque coefficient from static foil
-    data and cross-flow turbine kinematics was shown to be a reasonable estimator
-    for predicting Reynolds number dependence of an actual cross-flow turbine
-    operating under dynamic conditions. Mean velocity and turbulence measurements in
-    the near-wake showed very subtle differences. However, when transport terms for
-    the streamwise momentum and mean kinetic energy were calculated, a similar $Re$
-    threshold was revealed. These results imply that physical model studies of
-    cross-flow turbines should be $Re_D \sim 10^6$ to properly approximate both the
-    performance and wake dynamics of full scale devices and arrays.}
+    approximate average Reynolds number based on blade chord of $Re_c \approx 2
+    \times 10^5$. A simple model that calculates peak torque coefficient from static
+    foil data and cross-flow turbine kinematics was shown to be a reasonable
+    estimator for predicting Reynolds number dependence of an actual cross-flow
+    turbine operating under dynamic conditions. Mean velocity and turbulence
+    measurements in the near-wake showed subtle differences over the range of $Re$
+    investigated. However, when transport terms for the streamwise momentum and mean
+    kinetic energy were calculated, a similar $Re$ threshold was revealed. These
+    results imply that physical model studies of cross-flow turbines should achieve
+    $Re_D \sim 10^6$ to properly approximate both the performance and wake dynamics
+    of full scale devices and arrays.}
 
@@
 \begin{document}
-\linenumbers
 
@@
 downstream of the turbine, forcing more flow to accelerate around the sides of
-the turbine.
+the turbine. Note that we also see higher drag coefficients at higher $Re$
+(potentially for the same reason), which also helps explain larger negative
+values of cross-stream advection.
 

### 2015-12-16 paper.tex
Add keyword
Minor Wosnik edits in introduction

@@
 \keyword{Reynolds number; cross-flow turbine; turbine performance; marine
-hydrokinetic energy; wind energy; VAWT}
+hydrokinetic energy; wind energy; VAWT; scale model}
 
@@
 on the effects of dynamic stall, whereas the higher $Re$ experiment mostly
-concerned the mean velocity and tip vortex effects. Despite their value
-elucidating the complex wake dynamics of cross-flow turbines, these studies
-motivate the systematic investigation of scale effects undertaken here.
+concerned the mean velocity and tip vortex effects. Whereas the value of these
+studies was in elucidating the complex wake dynamics of cross-flow turbines,
+they also motivated the more systematic investigation of scale effects
+undertaken here.
 
@@
 
-Vertical-axis wind turbine array field experiments have revealed that
-accelerated wake recovery allows for closer spacing when compared with
-conventional axial-flow propeller-type turbines \cite{Dabiri2011, Kinzel2012}.
-It was observed experimentally that a high solidity vertical-axis cross-flow
-turbine's near-wake produces a unique vertical mean velocity field---generated
-by blade tip vortex shedding---the advection by which is the largest contributor
-to streamwise momentum and energy transport or recovery \cite{Bachant2015-JoT}.
-In this study we seek to replicate those same momentum and energy balance
-considerations at multiple Reynolds numbers, to examine the implications on how
-small scale, i.e., low Reynolds number experiments may be used to study flows in
-turbine arrays.
+Vertical-axis wind turbine array field experiments have revealed that improved
+wake recovery allows for closer spacing when compared to conventional axial-flow
+propeller-type turbines \cite{Dabiri2011, Kinzel2012}. It was observed
+experimentally that a high solidity vertical-axis cross-flow turbine's near-wake
+produces a unique vertical mean velocity field---generated by blade tip vortex
+shedding---the advection by which is the largest contributor to streamwise
+momentum and energy transport or recovery \cite{Bachant2015-JoT}. In this study
+we seek to replicate those same momentum and energy balance considerations at
+multiple Reynolds numbers, to examine the implications on how scaled, i.e., low
+Reynolds number experiments may be used to study flows in turbine arrays.
 

### 2015-12-16 paper.tex
Remove static stall angle definition

@@
 i.e., turbine performance, and the near-wake. Typically, static airfoil data
-shows that the static stall angle---the angle just beyond that at which the
-lift-to-drag ratio is maximum---increases with blade chord Reynolds number
+shows that the static stall angle increases with blade chord Reynolds number
 $Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil

### 2015-12-16 paper.tex
Plurality
Correct tank length

@@
 tunnel---since the loading is dominated by vortex shedding. However, Singleton
-and Yeager Jr. \cite{Singleton2000} state that Reynolds number effects on
+and Yeager Jr. \cite{Singleton2000} state that the effect of Reynolds number on
 dynamic stall remains an unsolved question.
@@
 cross-flow turbines. The test bed was integrated as part of the UNH tow tank,
-which is a 30 m long facility with a 3.66 m wide and 2.44 m deep cross-section.
+which is a 36 m long facility with a 3.66 m wide and 2.44 m deep cross-section.
 The turbine model used in this study was the UNH Reference Vertical Axis Turbine

### 2015-12-16 paper.tex
Minor Wosnik edits to exp setup section

@@
 Figure~\ref{fig:turbine}. The blockage ratio produced by the rotor's frontal
-area is 11\%, which we have decided not to correct for, as reliable methods are
+area is 11\%, which we decided not to correct for, as reliable methods are
 not yet agreed upon for CFTs \cite{Cavagnaro2014}. As such, blockage should be

### 2015-12-16 paper.tex
Note that tare drag was measured separately

@@
 Sentran ZB3 500 lbf capacity S-beam load cells, providing the rotor drag
-measurements, after a measured tare drag was subtracted in post processing.
-Similarly, a tare torque was measured by rotating the turbine shaft in air.
-Turbine angular and tow carriage linear position were measured using quadrature
-encoder signals, with $10^5$ counts-per-rev for the turbine and 10 $\mathrm{\mu
-    m}$ resolution for the carriage position. These signals, along with the torque
-and drag signals, were sampled at 2 kHz.
+measurements, after a separately measured tare drag was subtracted in post
+processing. Similarly, a tare torque was measured by rotating the turbine shaft
+in air. Turbine angular and tow carriage linear position were measured using
+quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10
+$\mathrm{\mu m}$ resolution for the carriage position. These signals, along with
+the torque and drag signals, were sampled at 2 kHz.
 

### 2015-12-16 paper.tex
Fix typo and add the word 'slight'

@@
 number, from about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall
-delay from a more turbulent boundary later on the blade suction side. There is
-essentially no change in the shape of the drag coefficient curves---merely an
-upward shift in $C_D$ with increasing $Re$.
+delay from a more turbulent boundary layer on the blade suction side. There is
+essentially no change in the shape of the drag coefficient curves---merely a
+slight upward shift in $C_D$ with increasing $Re$.
 

### 2015-12-16 paper.tex
Explain larger error bars at low speed; resolves #71

@@
 Mean power and drag coefficients at $\lambda=1.9$ are plotted versus Reynolds
-number in Figure~\ref{fig:perf-Re-dep}. There is a drastic improvement in $C_P$
-with increasing Reynolds number at lower values of Reynolds number. The power
-coefficient then becomes  essentially $Re$-independent at $Re_D = 0.8 \times
-10^6$, which corresponds to an approximate average blade chord Reynolds number
-$Re_{c, \mathrm{ave}} = 2.1 \times 10^5$. This threshold is consistent with the
-behavior of the blade boundary layer transitioning from laminar to turbulent,
-thereby promoting either the suppression or reattachment of the laminar
-separation bubble \cite{Lissaman1983}.
+number in Figure~\ref{fig:perf-Re-dep}. Note that the large error bars for $C_P$
+at low $Re$ are dominated by systematic error estimates for the torque
+transducer, since torque values are at the lower end of the measurement range.
+However, the uncertainty due to random error or repeatability remains relatively
+low. There is a drastic improvement in $C_P$ with increasing Reynolds number at
+lower values of Reynolds number. The power coefficient then becomes  essentially
+$Re$-independent at $Re_D = 0.8 \times 10^6$, which corresponds to an
+approximate average blade chord Reynolds number $Re_{c, \mathrm{ave}} = 2.1
+\times 10^5$. This threshold is consistent with the behavior of the blade
+boundary layer transitioning from laminar to turbulent, thereby promoting either
+the suppression or reattachment of the laminar separation bubble
+\cite{Lissaman1983}.
 
@@
 \caption{UNH-RVAT measured mean power (left) and drag (right) coefficients at
-    $\lambda=1.9$ plotted versus Reynolds number.}
+    $\lambda=1.9$ plotted versus Reynolds number. Error bars indicate expanded
+    uncertainty estimates for 95\% confidence, which for $C_P$ is dominated by
+    systematic error estimates from the torque transducer operating at the lower end
+    of its measurement range.}
 

### 2015-12-16 paper.tex
Reword C_D results

@@
 where increases in blade drag at lower $Re$ somewhat offsets the reduction in
-lift, to keep total streamwise force from changing as much as the rotor torque.
-The tendency for $C_D$ to continue increasing with $Re$ may also include the
-effects of an increasing Froude number, which increases free surface deformation
-and wave drag during towing without increasing flow through the turbine.
+lift, causing the total streamwise force to vary less than the rotor torque. The
+tendency for $C_D$ to continue increasing with $Re$ may also include the effects
+of an increasing Froude number, which increases free surface deformation and
+wave drag during towing without increasing flow through the turbine.
 

### 2015-12-16 paper.tex
Typo

@@
 were run for an angle of attack range of 0--40 degrees, in increments of 0.5
-degrees. Solver parameters used a 100 panels, a fixed speed, zero Mach number,
-\texttt{NCrit=9} (default $\mathrm{e}^n$ transition criteria parameter for an
-average wind tunnel), and no forced boundary layer transition. Characteristics
-were computed for the approximate average blade chord Reynolds numbers
-encountered in the tow tank experiments.
+degrees. Solver parameters used were 100 panels, a fixed speed, zero Mach
+number, \texttt{NCrit=9} (default $\mathrm{e}^n$ transition criteria parameter
+for an average wind tunnel), and no forced boundary layer transition.
+Characteristics were computed for the approximate average blade chord Reynolds
+numbers encountered in the tow tank experiments.
 

### 2015-12-16 paper.tex
Some rewording of wake profile paragraph

@@
 similar, so they have not been plotted here, though profiles at $z/H=0.0$ are
-shown in Figure~\ref{fig:profiles} to illustrate the subtlety of the
-differences. We will instead compare and contrast the wake behavior by examining
-spectra and wake transport terms in the equations that govern the downstream
-evolution of mean momentum and kinetic energy.
+compared in Figure~\ref{fig:profiles} to illustrate the subtlety of the
+differences at different $Re$. We will instead compare and contrast the wake
+behavior by examining spectra and wake transport terms in the equations that
+govern the downstream evolution of mean momentum and kinetic energy.
 

### 2015-12-16 paper.tex
Terms that can be computed from the experimental data
Drag coefficients are slightly higher

@@
 
-Results from the terms available (i.e. excluding $\partial / \partial x$) from
-the right hand sides of Equations~\ref{eq:momentum} and \ref{eq:K-full} are
-plotted in Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph},
-respectively. Derivatives were computed using a second order finite difference
-scheme, with central differencing for interior points, and inward facing schemes
-for the boundaries. A weighted average for each term was then calculated based
-on the grid spacing. We note that similar to \cite{Bachant2015-JoT}, the
-vertical advection at this point in the wake is the dominant contributor to
-positive wake recovery, caused by the unique vortex pattern created by the blade
-tips and wakes.
+Results from the terms that can be computed from the experimental data (i.e.
+excluding $\partial / \partial x$) from the right hand sides of
+Equations~\ref{eq:momentum} and \ref{eq:K-full} are plotted in
+Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph}, respectively.
+Derivatives were computed using a second order finite difference scheme, with
+central differencing for interior points, and inward facing schemes for the
+boundaries. A weighted average for each term was then calculated based on the
+grid spacing. We note that similar to \cite{Bachant2015-JoT}, the vertical
+advection at this point in the wake is the dominant contributor to positive wake
+recovery, caused by the unique vortex pattern created by the blade tips and
+wakes.
 
@@
 downstream of the turbine, forcing more flow to accelerate around the sides of
-the turbine. Note that we also see higher drag coefficients at higher $Re$
-(potentially for the same reason), which also helps explain larger negative
+the turbine. Note that we also see slightly higher drag coefficients at higher
+$Re$ (potentially for the same reason), which also helps explain larger negative
 values of cross-stream advection.

### 2015-12-16 paper.tex
Reword sentence a bit

@@
 approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
-10^5$. The left side of the $C_P$--$\lambda$ curve is further diminished at low
-$Re$ due to deeper stalling. The $Re$-independent threshold corresponds to that
-at which the foil suction surface boundary layer becomes turbulent before having
-to recover static pressure against the adverse pressure gradient, highlighted in
-\cite{Lissaman1983, McMasters1980, Carmichael1981}.
+10^5$. The power coefficient values on the left side of the $C_P$--$\lambda$
+curve are further reduced at low $Re$ due to deeper stalling. The
+$Re$-independent threshold corresponds to that at which the foil suction surface
+boundary layer becomes turbulent before having to recover static pressure
+against the adverse pressure gradient, highlighted in \cite{Lissaman1983,
+    McMasters1980, Carmichael1981}.
 

### 2015-12-16 paper.tex
Reword first sentence of wake paragraph in conclusions

@@
 
-At higher Reynolds numbers we observed lower levels of turbulence in the wake at
-$x/D=1$, along with lower levels of turbulent transport with respect to mean
-streamwise momentum and mean kinetic energy recovery in the streamwise
-direction. Vertical advection---the largest transport mechanism
+In the near-wake of the turbine we observed lower levels of turbulence with
+increasing Reynolds number, along with lower levels of turbulent transport with
+respect to mean streamwise momentum and mean kinetic energy recovery in the
+streamwise direction. Vertical advection---the largest transport mechanism
 measured---showed little $Re$-dependence, whereas the negative effects of

### 2015-12-16 paper.tex
Be explicit about wake recovery Re

@@
 measured---showed little $Re$-dependence, whereas the negative effects of
-cross-stream advection was enhanced at high $Re$, which may be due to larger
+cross-stream advection were enhanced at high $Re$, which may be due to larger
 free surface deformation reducing the near-wake's cross-sectional area. Despite
-being order of magnitude smaller than other transport processes, viscous
+being orders of magnitude smaller than other transport processes, viscous
 diffusion increased rapidly with decreasing $Re$, and is expected to become
@@
 wake transport measured leveled off at essentially the same Reynolds number that
-the performance did.
+the performance did---$Re_D \approx 0.8 \time 10^6$.
 

### 2015-12-16 paper.tex
Minor wording in the last paragraph

@@
 
-From these results we can make the recommendation that physical model tests be
-on the order $Re_D \sim 10^6$ to provide predictions of full-scale performance.
-This threshold also applies to the validation of predictive engineering models,
-especially for high-fidelity CFD models, where the boundary layer is to be
-resolved, since transition to turbulence plays an important role in overall
-blade loading. Note that in our case blockage, though reasonably low, may be
-increasing flow through the turbine compared with a free case, which could mean
-$Re$ thresholds for free flows could be slightly higher, though a reliable
-blockage correction algorithm has not yet been agreed upon for CFTs
+From these results we can make the recommendation that physical model tests
+should be on the order $Re_D \sim 10^6$ to provide predictions of full-scale
+performance. This threshold also applies to the validation of predictive
+engineering models, especially for high-fidelity CFD models where the boundary
+layer is to be resolved, since transition to turbulence plays an important role
+in overall blade loading. Note that in our case blockage, though reasonably low,
+may be increasing flow through the turbine compared with a free case, which
+could mean $Re$ thresholds for free flows could be slightly higher, though a
+reliable blockage correction algorithm has not yet been agreed upon for CFTs
 \cite{Cavagnaro2014}.

### 2015-12-16 paper.tex
Simplify wording in last paragraph

@@
 
-From these results we can make the recommendation that physical model tests
-should be on the order $Re_D \sim 10^6$ to provide predictions of full-scale
-performance. This threshold also applies to the validation of predictive
-engineering models, especially for high-fidelity CFD models where the boundary
-layer is to be resolved, since transition to turbulence plays an important role
-in overall blade loading. Note that in our case blockage, though reasonably low,
-may be increasing flow through the turbine compared with a free case, which
-could mean $Re$ thresholds for free flows could be slightly higher, though a
-reliable blockage correction algorithm has not yet been agreed upon for CFTs
+From these results we recommend that physical model tests be performed at $Re_D
+\sim 10^6$ to provide reasonable predictions of full-scale performance. This
+threshold also applies to the validation of predictive engineering models,
+especially for high-fidelity CFD models where the boundary layer is to be
+resolved, since transition to turbulence plays an important role in overall
+blade loading. Note that in our case blockage, though reasonably low, may be
+increasing flow through the turbine compared with a free case, which could mean
+$Re$ thresholds for free flows could be slightly higher, though a reliable
+blockage correction algorithm has not yet been agreed upon for CFTs
 \cite{Cavagnaro2014}.

### 2015-12-16 scripts/diff.sh
Diff script working

@@ -1,6 +1,11 @@
 #!/usr/bin/env bash
+# This script creates a diff version comparing the recent with whatever revision
+# is supplied as the first argument. This can be any valid Git identifier, e.g.,
+# a commit hash, tag, etc.
 
 rev=$1
 git show $rev:paper.tex > Archive/paper-$rev.tex
-latexdiff Archive/paper-$rev.tex paper.tex > paper-$rev-diff.tex
-# latexmk paper-$rev-diff.tex -pdf -f
+latexdiff --exclude-textcmd="section,subsection,subsubsection" --math-markup=off Archive/paper-$rev.tex Archive/paper-unix.tex > paper-$rev-diff.tex
+latexmk paper-$rev-diff.tex -pdf -f
+mv paper-$rev-diff.pdf Archive
+rm paper-$rev*

### 2015-12-16 scripts/diff.sh
Archive dir case

@@ -4,8 +4,8 @@
 # a commit hash, tag, etc.
 
 rev=$1
-git show $rev:paper.tex > Archive/paper-$rev.tex
-latexdiff --exclude-textcmd="section,subsection,subsubsection" --math-markup=off Archive/paper-$rev.tex Archive/paper-unix.tex > paper-$rev-diff.tex
+git show $rev:paper.tex > archive/paper-$rev.tex
+latexdiff --exclude-textcmd="section,subsection,subsubsection" --math-markup=off archive/paper-$rev.tex archive/paper-unix.tex > paper-$rev-diff.tex
 latexmk paper-$rev-diff.tex -pdf -f
-mv paper-$rev-diff.pdf Archive
+mv paper-$rev-diff.pdf archive
 rm paper-$rev*

### 2015-12-23 paper.tex
Remove McMasters figure

@@
 $Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil
-behavior is presented in \cite{Lissaman1983}. In general, airfoil performance is
+behavior is presented in \cite{Lissaman1983}. In general, airfoil
+performance---often characterized by the profile's lift-to-drag ratio---is
 enhanced as the boundary layer on the foil transitions to turbulence closer to
@@
 angles of attack. For smooth airfoils, this transition can cause a dramatic
-increase in foil performance, at a blade chord Reynolds number on the order of
-$10^5$ \cite{McMasters1980}, shown in Figure~\ref{fig:McMasters}. Note that
-there is a distinct lack of highly reliable data for airfoils in this
-transitional regime and below. An evaluation of the various databases relevant
-to cross-flow turbines is presented in \cite{Bedon2014}.
-
-\begin{figure}[ht]
-\centering
-
-\includegraphics[width=0.7\textwidth]{figures/McMasters-Henderson-1980}
-
-\caption{Airfoil maximum lift-to-drag ratio versus Reynolds number, from
-    \cite{McMasters1980}.}
-
-\label{fig:McMasters}
-\end{figure}
+increase in foil performance at a blade chord Reynolds number on the order of
+$10^5$ \cite{McMasters1980}. Note that there is a distinct lack of highly
+reliable data for airfoils in this transitional regime and below. An evaluation
+of the various databases relevant to cross-flow turbines is presented in
+\cite{Bedon2014}.
 

### 2015-12-23 paper.tex
Note in K bar chart caption that x-derivatives were not measured or computed

@@
     \caption{Normalized mean kinetic energy transport quantities computed as
-        weighted averages based on Equation~\ref{eq:K-full}. Note that directions of
-        the turbulent transport terms refer to the directions of their partial
-        derivatives.}
+        weighted averages based on Equation~\ref{eq:K-full}, omitting the
+        non-measured streamwise derivatives. Note that directions of the turbulent
+        transport terms refer to the directions of their partial derivatives.}
 

### 2015-12-24 paper.tex
Add comment about rearranged transport equations

@@
 		+ \frac{\p^2 U}{\p z^2} \right)
-		\bigg{]}.
+		\bigg{]},
 	\end{split}
@@
 \end{equation}
+Note that Equations~\ref{eq:momentum} and \ref{eq:K-full} only differ from the
+typical forms of the mean momentum and mean kinetic energy equations by a factor
+of $1/U$, but importantly, the other convective terms ( $\partial U/\partial y$,
+$\partial U/\partial z$, and $\partial K/\partial y$, $\partial K/\partial z$ )
+are moved to the right-hand-sides of the equations to illustrate with what sign
+they add to the streamwise recovery of streamwise momentum or mean kinetic
+energy.
 

### 2015-12-24 scripts/diff.sh
Fix diff script

@@ -5,7 +5,7 @@
 
 rev=$1
 git show $rev:paper.tex > archive/paper-$rev.tex
-latexdiff --exclude-textcmd="section,subsection,subsubsection" --math-markup=off archive/paper-$rev.tex archive/paper-unix.tex > paper-$rev-diff.tex
+latexdiff --exclude-textcmd="section,subsection,subsubsection" --math-markup=off archive/paper-$rev.tex paper.tex > paper-$rev-diff.tex
 latexmk paper-$rev-diff.tex -pdf -f
 mv paper-$rev-diff.pdf archive
 rm paper-$rev*

### 2016-01-14 paper.tex
Wosnik edits part 1

@@
     cross-flow turbine to investigate Reynolds number effects on performance and
-    wake characteristics, to establish scale thresholds for physical and numerical
-    modeling of individual devices and arrays. It was demonstrated that the
-    performance of the cross-flow turbine becomes essentially $Re$-independent at a
-    Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
+    wake characteristics, and to establish scale thresholds for physical and
+    numerical modeling of individual devices and arrays. It was demonstrated that
+    the performance of the cross-flow turbine becomes essentially $Re$-independent
+    at a Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
     approximate average Reynolds number based on blade chord of $Re_c \approx 2
@@
 surface---is the Froude number $Fr \equiv U/\sqrt{gl}$, where $g$ is the
-gravitational acceleration. Matching both Froude and Reynolds number is
-difficult for the same fluid, which is illustrated by the relation $Re = l^{3/2}
-g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$, or the geometric scale.
-In general the geometric scale of a physical model is not necessarily the same
-as its dynamical scale. As such, hereafter we will use the word ``scale'' to
-refer to this dynamical scale or Reynolds number and assume Froude number is
-small enough to be negligible.
+gravitational acceleration. Matching both Froude and Reynolds number is not
+possible for a scaled model in a given fluid, which is illustrated by the
+relation $Re = l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$,
+or the geometric scale. In general the geometric scale of a physical model is
+not necessarily the same as its dynamical scale. As such, hereafter we will use
+the word ``scale'' to refer to this dynamical scale or Reynolds number and
+assume Froude number is small enough to be negligible.
 
@@
 observed significant Reynolds number sensitivity for a high solidity helical
-cross-flow turbine, and Bravo et al. \cite{Bravo2007} observed power coefficient
-of a straight-bladed VAWT to become Reynolds number independent at $Re_c \approx
-4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} observed Reynolds number
-independence of power coefficient at optimal tip speed ratio for a high solidity
-cross-flow turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2
-\times 10^5$.
+cross-flow turbine, and Bravo et al. \cite{Bravo2007} observed the power
+coefficient of a straight-bladed VAWT to become Reynolds number independent at
+$Re_c \approx 4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} observed
+Reynolds number independence of power coefficient at optimal tip speed ratio for
+a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
+$Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.
 
@@
 Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
-(ADV), which has a 6 mm diameter sampling volume, and sampled at 200 Hz. The
-probe was mounted on an automated positioning system, also controlled by the tow
-tank's main motion controller. The ADV and DAQ systems' sampling times were
-synchronized by triggering the start of data acquisition via a pulse sent from
-the motion controller. Additional details of the turbine and experimental setup
-are described in \cite{Bachant2015-JoT}.
+(ADV), which has an approximately 6 mm diameter sampling volume, and sampled at
+200 Hz. The probe was mounted on an automated positioning system, also
+controlled by the tow tank's main motion controller. The ADV and DAQ systems'
+sampling times were synchronized by triggering the start of data acquisition via
+a pulse sent from the motion controller. Additional details of the turbine and
+experimental setup are described in \cite{Bachant2015-JoT}.
 
@@
 analyzed to compute statistics over this interval, truncating the end slightly
-to achieve and integer number of blade passages. Turbine shaft angular velocity
-and tow carriage speed were calculated using as second order central
-differencing scheme on the respective position measurements. Power and drag
-coefficients are calculated as instantaneous quantities using the carriage speed
-as the free stream velocity.
+to achieve an integer number of blade passages. Turbine shaft angular velocity
+and tow carriage speed were calculated using a second order central differencing
+scheme on the respective position measurements. Power and drag coefficients were
+calculated as instantaneous quantities using the carriage speed as the free
+stream velocity.
 
 Wake velocity data were filtered for spurious ``spikes'' by removing datapoints
-8 standard deviations or 0.9 m/s from the mean. The experimental data and code
+8 standard deviations, or 0.9 m/s from the mean. The experimental data and code
 for processing and visualization are available from
@@
 The random error was inferred from the sample standard deviation (on a
-per-revolution basis) and the systematic from the sensor calibrations or
+per-revolution basis) and the systematic error from the sensor calibrations or
 datasheets. Combining both sources of error, along with their propagation into
@@
 \begin{tabular}{cccc}
-$U_\infty$ (m/s) &                $U$ &                $V$ &                $W$ \\
+$U_\infty$ (m/s) &          $U$ (m/s) &          $V$ (m/s) &          $W$ (m/s) \\
 \hline
@@
 at low $Re$ are dominated by systematic error estimates for the torque
-transducer, since torque values are at the lower end of the measurement range.
+transducer, since torque values are at the lower end of its measurement range.
 However, the uncertainty due to random error or repeatability remains relatively
@@
 tendency for $C_D$ to continue increasing with $Re$ may also include the effects
-of an increasing Froude number, which increases free surface deformation and
-wave drag during towing without increasing flow through the turbine.
+of an increasing Froude number (from faster tow speeds), which increases free
+surface deformation and wave drag during towing without increasing flow through
+the turbine.
 

### 2016-01-14 paper.tex
Wosnik edits part 2

@@
 \end{equation}
-where the blade section chordwise force (for zero preset pitch) is given by
+where the blade section chordwise force coefficient (for zero preset pitch) is
+given by
 \begin{equation}
@@
 (non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
-\times 10^5$, but then continues to increase linearly with a small slope with
-$Re$. This is not matched by the experimental data, which at higher $Re$ looks
-more like the cambered foil results. Though this method is very approximate, it
-nevertheless appears to be a reasonable way to predict the transitional Reynolds
-number regime for cross-flow turbine performance using only 2-D static airfoil
-characteristics
+\times 10^5$, but then continues to increase slowly and linearly with $Re$. This
+is not matched by the experimental data, which at higher $Re$ looks more like
+the cambered foil results. Though this method does not provide absolute
+predictions of turbine performance, it predicts the transitional Reynolds number
+regime for cross-flow turbine performance using only 2-D static airfoil
+characteristics, i.e., Reynolds number scaling of the peak $C_T$ computed this
+way behaves much like that of the measured turbine performance.
 
@@
 Spectra of the cross-stream velocity normalized by the free stream were
-computing using a fast Fourier transform-based method, applying a Hanning window
+computed using a fast Fourier transform-based method, applying a Hanning window
 and averaging over four adjacent frequency bands to decrease confidence
@@
 recovery, caused by the unique vortex pattern created by the blade tips and
-wakes.
+blade wakes.
 
@@
 number itself. From these results one might expect that the viscous effects will
-become significant to the wake dynamics around $Re_D \sim 10^4$.
+become significant to the wake dynamics as Reynolds number is decreased below
+$Re_D \sim 10^4$.
 
@@
 downstream of the turbine, forcing more flow to accelerate around the sides of
-the turbine. Note that we also see slightly higher drag coefficients at higher
-$Re$ (potentially for the same reason), which also helps explain larger negative
-values of cross-stream advection.
+the turbine. Note that we also observe slightly higher drag coefficients at
+higher $Re$ (potentially for the same reason), which also helps explain larger
+negative values of cross-stream advection.
 
@@
 approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
-10^5$. The power coefficient values on the left side of the $C_P$--$\lambda$
-curve are further reduced at low $Re$ due to deeper stalling. The
-$Re$-independent threshold corresponds to that at which the foil suction surface
-boundary layer becomes turbulent before having to recover static pressure
-against the adverse pressure gradient, highlighted in \cite{Lissaman1983,
-    McMasters1980, Carmichael1981}.
+10^5$. The power coefficient values on the left hand side of the
+$C_P$--$\lambda$ curve are reduced more than those on the right hand side at low
+$Re$ due to deeper stalling. The $Re$-independent threshold corresponds to that
+at which the foil suction surface boundary layer becomes turbulent before having
+to recover static pressure against the adverse pressure gradient, highlighted in
+\cite{Lissaman1983, McMasters1980, Carmichael1981}.
 
@@
 cross-stream advection were enhanced at high $Re$, which may be due to larger
-free surface deformation reducing the near-wake's cross-sectional area. Despite
-being orders of magnitude smaller than other transport processes, viscous
-diffusion increased rapidly with decreasing $Re$, and is expected to become
-significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the total
-wake transport measured leveled off at essentially the same Reynolds number that
-the performance did---$Re_D \approx 0.8 \time 10^6$.
+free surface deformation effectively reducing the near-wake's cross-sectional
+area. Despite being orders of magnitude smaller than other transport processes,
+viscous diffusion increased rapidly with decreasing $Re$, and is expected to
+become significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the
+total wake transport measured leveled off at essentially the same Reynolds
+number that the performance did---$Re_D \approx 0.8 \times 10^6$.
 
-From these results we recommend that physical model tests be performed at $Re_D
-\sim 10^6$ to provide reasonable predictions of full-scale performance. This
-threshold also applies to the validation of predictive engineering models,
-especially for high-fidelity CFD models where the boundary layer is to be
-resolved, since transition to turbulence plays an important role in overall
-blade loading. Note that in our case blockage, though reasonably low, may be
-increasing flow through the turbine compared with a free case, which could mean
-$Re$ thresholds for free flows could be slightly higher, though a reliable
-blockage correction algorithm has not yet been agreed upon for CFTs
-\cite{Cavagnaro2014}.
+From these results we recommend that physical model tests of cross-flow turbines
+be performed at $Re_D \sim 10^6$ or greater to provide reasonable predictions of
+full-scale performance. This threshold also applies to the validation of
+predictive engineering models, especially for high-fidelity CFD models where the
+boundary layer is to be resolved, since transition to turbulence plays an
+important role in overall blade loading. Note that in our case blockage, though
+reasonably low, may be increasing flow through the turbine compared with a free
+case, which could mean $Re$ thresholds for free flows could be slightly higher,
+though a reliable blockage correction algorithm has not yet been agreed upon for
+CFTs \cite{Cavagnaro2014}.
 

### 2016-01-20 paper.tex
Suggested changes from editor

@@
 
-\documentclass[energies,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
+\documentclass[energies,article,accept,moreauthors,pdftex,10pt,a4paper]{mdpi}
 
 %=================================================================
-\setcounter{page}{1}
-\lastpage{x}
+\firstpage{1} %%MDPI internal note: new layout%%
+\lastpage{\pageref*{LastPage}}
+\makeatletter %%MDPI internal note: new layout%%
+\setcounter{page}{\@firstpage} %%MDPI internal note: new layout%%
+\makeatother %%MDPI internal note: new layout%%
+\articlenumber{0000}
 \doinum{10.3390/------}
-\pubvolume{xx}
-\pubyear{2015}
-\history{Received: xx / Accepted: xx / Published: xx}
+\pubvolume{9}
+\pubyear{2016}
+\externaleditor{Academic Editor: Sukanta Basu}
+\history{Received: 1 November 2015; Accepted: 15 January 2016; Published: date}
 %------------------------------------------------------------------
@@
 \usepackage[utf8]{inputenc}
-
+\usepackage{booktabs} 
+\usepackage{multirow}
+\usepackage{soul}
+\usepackage{microtype}
 \def \p{\partial}
@@
 \def \D{\mathrm{D}}
+\setitemize{parsep=6pt,itemsep=0pt,leftmargin=*,labelsep=5mm}
+\setenumerate{parsep=6pt,itemsep=0pt,leftmargin=*,labelsep=5mm}
+\setlist[description]{itemsep=0mm}
 
@@
 %% Please use the following mathematics environments:
-%\theoremstyle{mdpi}
-%\newcounter{thm}
-%\setcounter{thm}{0}
-%\newcounter{ex}
-%\setcounter{ex}{0}
-%\newcounter{re}
-%\setcounter{re}{0}
-%\newtheorem{Theorem}[thm]{Theorem}
-%\newtheorem{Lemma}[thm]{Lemma}
-%\newtheorem{Characterization}[thm]{Characterization}
-%\newtheorem{Proposition}[thm]{Proposition}
-%\newtheorem{Property}[thm]{Property}
-%\newtheorem{Problem}[thm]{Problem}
-%\newtheorem{Example}[ex]{Example}
-%\newtheorem{Remark}[re]{Remark}
-%\newtheorem{Corollary}[thm]{Corollary}
-%\newtheorem{Definition}[thm]{Definition}
+\theoremstyle{mdpi}
+\newcounter{thm}
+\setcounter{thm}{0}
+\newcounter{ex}
+\setcounter{ex}{0}
+\newcounter{re}
+\setcounter{re}{0}
+\newtheorem{Theorem}[thm]{Theorem}
+\newtheorem{Lemma}[thm]{Lemma}
+\newtheorem{Characterization}[thm]{Characterization}
+\newtheorem{Proposition}[thm]{Proposition}
+\newtheorem{Property}[thm]{Property}
+\newtheorem{Problem}[thm]{Problem}
+\newtheorem{Example}[ex]{Example}
+\newtheorem{Remark}[re]{Remark}
+\newtheorem{Corollary}[thm]{Corollary}
+\newtheorem{Definition}[thm]{Definition}
 %% For proofs, please use the proof environment (the amsthm package is loaded by
@@
 % Authors (Add full first names)
-\Author{Peter Bachant $^{1,}$* and Martin Wosnik $^{1}$*}
+\Author{Peter Bachant * and Martin Wosnik *}
 
@@
 % affiliation.)
-\address{%
-$^{1}$ Center for Ocean Renewable Energy, University of New Hampshire, 24
-Colovos Rd., Durham, NH, USA}
+\address[1]{%
+ Center for Ocean Renewable Energy, University of New Hampshire, 24
+Colovos Rd., \mbox{Durham, NH \hl{zip code}, USA}}
 
@@
 % there are more than one corresponding author.)
-\corres[2]{pxL3@unh.edu, martin.wosnik@unh.edu}
-
-\linenumbers
+\corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.); martin.wosnik@unh.edu (M.W.); \linebreak Tel.: \hl{(P.B.); (M.W.)}}
+
+%\linenumbers
 
@@
  cross-flow turbine to investigate Reynolds number effects on performance and
-    wake characteristics, and to establish scale thresholds for physical and
-    numerical modeling of individual devices and arrays. It was demonstrated that
+ wake characteristics and to establish scale thresholds for physical and
+ numerical modeling of individual devices and arrays.~It was demonstrated that
  the performance of the cross-flow turbine becomes essentially $Re$-independent
-    at a Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
-    approximate average Reynolds number based on blade chord of $Re_c \approx 2
-    \times 10^5$. A simple model that calculates peak torque coefficient from static
+ at a Reynolds number based on a turbine diameter of $Re_D \approx 10^6$ or an
+ approximate average Reynolds number based on a blade chord of $Re_c \approx 2
+ \times 10^5$.~A simple model that calculates the peak torque coefficient from static
  foil data and cross-flow turbine kinematics was shown to be a reasonable
-    estimator for predicting Reynolds number dependence of an actual cross-flow
+ estimator for predicting the Reynolds number dependence of an actual cross-flow
  turbine operating under dynamic conditions. Mean velocity and turbulence
  measurements in the near-wake showed subtle differences over the range of $Re$
-    investigated. However, when transport terms for the streamwise momentum and mean
+ investigated.~However, when transport terms for the streamwise momentum and mean
  kinetic energy were calculated, a similar $Re$ threshold was revealed. These
@@
  $Re_D \sim 10^6$ to properly approximate both the performance and wake dynamics
-    of full scale devices and arrays.}
+ of full-scale devices and arrays.}
 
 % Keywords: add 3 to 10 keywords
-\keyword{Reynolds number; cross-flow turbine; turbine performance; marine
-hydrokinetic energy; wind energy; VAWT; scale model}
+\keyword{\textls[-10]{Reynolds number; cross-flow turbine; turbine performance; marine
+hydrokinetic energy}; wind energy; {VAWT}; scale model}
 
@@
 Scaled physical models are often used in science and engineering to approximate
-real-world systems. The principle of dynamic similarity allows for geometrically
-scaled systems to be dynamically identical if certain nondimensional physical
+real-world systems. The principle of dynamic similarity allows for geometrically-scaled systems to be dynamically identical if certain nondimensional physical
 parameters are matched. In the case of fluid systems, the most important
@@
 the importance of inertial forces over viscous forces on the flow physics
-\cite{Acheson1990}: $Re \equiv Ul/\nu$, where $U$ and $l$ are characteristic
+\cite{Acheson1990}: \hl {$Re \equiv Ul/\nu$, } 
+%please check the symbol ``\equiv '' and make sure it is correct
+where $U$ and $l$ are characteristic
 velocity and length scales, respectively, and $\nu$ is the fluid kinematic
-viscosity. The other common dynamical scale---for systems with a free
-surface---is the Froude number $Fr \equiv U/\sqrt{gl}$, where $g$ is the
-gravitational acceleration. Matching both Froude and Reynolds number is not
+viscosity. The~other common dynamical scale, for systems with a free
+surface, is the Froude number \hl {$Fr \equiv U/\sqrt{gl}$,}
+%please check the symbol ``\equiv '' and make sure it is correct
+  where $g$ is the
+gravitational acceleration. Matching both the Froude and Reynolds number is not
 possible for a scaled model in a given fluid, which is illustrated by the
 relation $Re = l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$,
-or the geometric scale. In general the geometric scale of a physical model is
-not necessarily the same as its dynamical scale. As such, hereafter we will use
-the word ``scale'' to refer to this dynamical scale or Reynolds number and
-assume Froude number is small enough to be negligible.
+or the geometric scale. In general, the geometric scale of a physical model is
+not necessarily the same as its dynamical scale. As such, hereafter, we will use
+the word ``scale'' to refer to this dynamical scale or the Reynolds number and
+assume the Froude number is small enough to be~ negligible.
 
@@
 models are used to validate predictive numerical models, predict full-scale
-performance of individual turbines, and design or investigate arrays of devices.
-Scaled models have the benefit of being significantly less expensive, however, a
-key question is whether or not an acceptable scale mismatch exists. Similarly,
-it is questionable whether numerical models validated with physical model data
-obtained orders of magnitude away from prototype scale should be considered
+performance of individual turbines and design or investigate arrays of devices.
+Scaled models have the benefit of being significantly less expensive; however, a
+key question is whether or not an acceptable scale mismatch exists. Similarly,~it is questionable whether numerical models validated with physical model data
+obtained orders of magnitude away from the prototype scale should be considered
 validated at all. An example of the errors that can result from the
-extrapolation of small scale experiments can be found in \cite{Baker1991}.
+extrapolation of small-scale experiments can be found in \cite{Baker1991}.
 
@@
 frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
-of blades, strut drag and operational parameters such as tip speed ratio---{\em
-    and} on Reynolds number \cite{Para2002}. Note that an average blade chord
+of blades, strut drag and operational parameters, such as tip speed ratio, {
+ and} on the Reynolds number \cite{Para2002}. Note that an average blade chord
 Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be
-expressed in terms of tip speed ratio $\lambda \equiv \omega R/ U_\infty$, where
+expressed in terms of tip speed ratio \hl {$\lambda \equiv \omega R/ U_\infty$, } 
+%please check the symbol ``\equiv '' and make sure it is correct
+where
 $U_\infty$ is the free stream velocity, $c$ is the blade chord length, $\nu$ is
-the fluid kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$
+the fluid kinematic viscosity, $\omega$ is the rotor's angular velocity and $R$
 is the rotor radius. The value of $\lambda$ at which a turbine reaches peak
@@
 Solidity often directly correlates with the chord-to-radius ratio $c/R$. Rotors
-with $c/R \ge 0.1$ are considered high solidity \cite{Fiedler2009}, for which
+with $c/R \ge 0.1$ are considered to have high solidity \cite{Fiedler2009}, for which
 so-called flow curvature or virtual camber effects become significant
 \cite{Migliore1980}. These effects arise from the blade sections' circular
-paths, and can complicate the comparison with behavior of an airfoil in a linear
+paths and can complicate the comparison with the behavior of an airfoil in a linear
 flow.
 
-Blackwell et al. \cite{Blackwell1976} investigated the effects of Reynolds
-number on the performance of a 2 m diameter Darrieus vertical-axis cross-flow
+Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of Reynolds
+number on the performance of a 2-m diameter Darrieus vertical-axis cross-flow
 wind turbine with NACA 0012 blades. By varying wind tunnel speed, the turbine
-was made to operate at approximately constant blade chord Reynolds number $Re_c$
-ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime the turbine
+was made to operate at an approximately constant blade chord Reynolds number $Re_c$
+ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime, the turbine
 power coefficient $C_P$ was shown to be sensitive to $Re_c$, with sensitivity
@@
 observed significant Reynolds number sensitivity for a high solidity helical
-cross-flow turbine, and Bravo et al. \cite{Bravo2007} observed the power
+cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007} observed the power
 coefficient of a straight-bladed VAWT to become Reynolds number independent at
 $Re_c \approx 4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} observed
-Reynolds number independence of power coefficient at optimal tip speed ratio for
+Reynolds number independence of the power coefficient at the optimal tip speed ratio for
 a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
@@
 
-The wake of a 2-D cross-flow turbine in the dynamic stall regime has been
-studied via laser Doppler velocimetry by Brochier et al. \cite{Brochier1986} and
-in 3-D with particle image velocimetry (PIV) by Fujisawa and Shibuya
+The wake of a 2D cross-flow turbine in the dynamic stall regime has been
+studied via laser Doppler velocimetry by Brochier \emph{et al.} \cite{Brochier1986} and
+in 3D with particle image velocimetry (PIV) by Fujisawa and Shibuya
 \cite{Fujisawa2001}. However, both of these studies were performed at very low
-Reynolds numbers---$Re_D = 10^4$ and $10^3$, respectively. Tescione et al.
-\cite{Tescione2014} studied the wake of a vertical-axis wind turbine at optimal
+Reynolds numbers: $Re_D = 10^4$ and $10^3$, respectively. Tescione \emph{et al.}
+\cite{Tescione2014} studied the wake of a vertical-axis wind turbine at the optimal
 tip speed ratio in a wind tunnel using stereo PIV at an approximate blade chord
@@
 Adaramola \cite{Krogstad2012a} observed Reynolds number independence of the
-performance of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5
+performance of a 0.9-m diameter axial-flow turbine rotor at $Re_D \approx 5
 \times 10^5$ in wind tunnel tests. Their turbine blades had an S826 profile
-along their entire span, a profile chosen for its $Re$-independence. Walker et
-al. \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
+along their entire span, a profile chosen for its $Re$-independence. Walker \emph{et
+al.} \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
 turbine tests in a towing tank, which were performed at a mid-span Reynolds
-number $Re_c = 4 \times 10^5$. McTavish et al. \cite{McTavish2013} showed how
+number $Re_c = 4 \times 10^5$. McTavish \emph{et al.} \cite{McTavish2013} showed how
 the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
@@
 When designing or studying arrays, it is common to use very small
-(geometrically) scaled devices \cite{Chamorro2011, Chamorro2011b}. It is
+(geometrically) scaled \mbox{devices \cite{Chamorro2011, Chamorro2011b}}. It is
 therefore important to realize the limitations of evaluating both the power
-output of turbines and the wake recovery when Reynolds number is very far from
-full scale. Sometimes models used are not turbines, but wake generating objects,
+output of turbines and the wake recovery when the Reynolds number is very far from
+full scale. Sometimes,~the models used are not turbines, but wake-generating objects,
 e.g., porous disks, that are meant to replicate the wakes of real turbines
 \cite{Goldenberg1983}. In this case, it is of interest to determine at what
-Reynolds number one might be able to realistically study wake flows in an array,
+Reynolds number one might be able to realistically study wake flows in an array
 and also to evaluate the effectiveness of a wake generator. In other words, a
@@
 it simulate a full-scale device? Note that for a porous disk, the Reynolds
-number of interest is diameter-based, since it is the scaling of the far-wake
-dynamics that matters.
+number of interest is diameter based, since it is the scaling of the far-wake
+dynamics that~ matters.
 
@@
 wake recovery allows for closer spacing when compared to conventional axial-flow
-propeller-type turbines \cite{Dabiri2011, Kinzel2012}. It was observed
+propeller-type \mbox{turbines \cite{Dabiri2011, Kinzel2012}}. It~was observed
 experimentally that a high solidity vertical-axis cross-flow turbine's near-wake
-produces a unique vertical mean velocity field---generated by blade tip vortex
-shedding---the advection by which is the largest contributor to streamwise
-momentum and energy transport or recovery \cite{Bachant2015-JoT}. In this study
+produces a unique vertical mean velocity field, generated by blade tip vortex
+shedding, the advection by which is the largest contributor to streamwise
+momentum and energy transport or recovery \cite{Bachant2015-JoT}. In~this study,
 we seek to replicate those same momentum and energy balance considerations at
-multiple Reynolds numbers, to examine the implications on how scaled, i.e., low
-Reynolds number experiments may be used to study flows in turbine arrays.
+multiple Reynolds numbers, to examine the implications on how scaled, \emph{i.e.}, low
+Reynolds number experiments, may be used to study flows in turbine arrays.
 
@@
 We are looking at how Reynolds number scaling affects both the blade loading,
-i.e., turbine performance, and the near-wake. Typically, static airfoil data
-shows that the static stall angle increases with blade chord Reynolds number
+\emph{i.e.}, turbine performance, and the near-wake. Typically, static airfoil data
+show that the static stall angle increases with blade chord Reynolds number
 $Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil
 behavior is presented in \cite{Lissaman1983}. In general, airfoil
-performance---often characterized by the profile's lift-to-drag ratio---is
+performance, often characterized by the profile's lift-to-drag ratio, is
 enhanced as the boundary layer on the foil transitions to turbulence closer to
@@
 increase in foil performance at a blade chord Reynolds number on the order of
-$10^5$ \cite{McMasters1980}. Note that there is a distinct lack of highly
+\mbox{$10^5$ \cite{McMasters1980}}. Note~that there is a distinct lack of highly
 reliable data for airfoils in this transitional regime and below. An evaluation
@@
 
-Static foil performance does not tell the whole story for a cross-flow turbine.
-The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
-turbine implies dynamic loading, encountering dynamic stall for tip speed ratios
+\textls[-10]{Static foil performance does not tell the whole story for a cross-flow turbine.
+The azimuthal, and} therefore, temporal, variation of $\alpha$ in a cross-flow
+turbine implies dynamic loading, encountering \textls[-10]{dynamic stall for tip speed ratios
 near and below those of maximum rotor torque \cite{Para2002}. Bousman
-\cite{Bousman2000-evaluation} states that dynamic stall is relatively
-insensitive to Reynolds number for $Re=1.0 \times 10^5$--$2.5 \times
-10^5$---judging from measurements on a pitching VR-7 foil in a wind and water
-tunnel---since the loading is dominated by vortex shedding. However, Singleton
+\cite{Bousman2000-evaluation}} states that dynamic stall is relatively
+insensitive to Reynolds number for $Re=1.0 \times 10^5$ to $2.5 \times
+10^5$, judging from measurements on a pitching VR-7 foil in a wind and water
+tunnel, since the loading is dominated by vortex shedding. However, Singleton
 and Yeager Jr. \cite{Singleton2000} state that the effect of Reynolds number on
@@
 Despite lower lift on the blades at lower $Re$, we expect stronger tip vortex
-shedding \cite{Yoon2005}. As mentioned previously, the dynamic stall vortex
+shedding \cite{Yoon2005}. As~mentioned previously, the dynamic stall vortex
 shedding is not expected to be highly sensitive to Reynolds number, though a
@@
 
-Chamorro et al. \cite{Chamorro2012} showed that turbulence statistics in an AFT
-wake became $Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean
+Chamorro \emph{et al.} \cite{Chamorro2012} showed that turbulence statistics in an AFT%please define
+wake became $Re$-independent at $Re_D \approx 1 \times 10^5$ and that mean
 velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
-\times 10^4$. Note that in this study the tip chord Reynolds number was not
+\times 10^4$. Note that in this study, the tip chord Reynolds number was not
 reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
@@
 Experiments were performed in a turbine test bed specifically designed for
-cross-flow turbines. The test bed was integrated as part of the UNH tow tank,
-which is a 36 m long facility with a 3.66 m wide and 2.44 m deep cross-section.
+cross-flow turbines. The test bed was integrated as part of the \hl {University of New Hampshire (UNH)} 
+ tow tank,
+which is a 36 m-long facility with a 3.66 m-wide and 2.44 m-deep cross-section.
 The turbine model used in this study was the UNH Reference Vertical Axis Turbine
 (RVAT), which was designed to be a generic case for numerical model testing,
-similar to the Sandia National Labs/US Department of Energy RM2 River Turbine
-\cite{Neary2014}, but with higher solidity, or blade chord-to-radius ratio.
-
-The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.14 m
-chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and the
-turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
+similar to the Sandia National Labs/U.S. Department of Energy RM2%please define
+ River Turbine
+\cite{Neary2014}, but with a higher solidity or blade chord-to-radius ratio.
+
+The UNH-RVAT turbine has three blades made from \hl {NACA} 0020 profiles
+%please define NACA and add the information of Manufacturer, City, Country
+  with a 0.14 m
+chord. The~blades are mounted at mid-chord w.r.t. the turbine axis, and the
+turbine has a height (blade span) of 1 m and a diameter of 1 m; \textit{cf}.
 Figure~\ref{fig:turbine}. The blockage ratio produced by the rotor's frontal
 area is 11\%, which we decided not to correct for, as reliable methods are
-not yet agreed upon for CFTs \cite{Cavagnaro2014}. As such, blockage should be
-taken into account when using these data for model validation, i.e., the
+not yet agreed upon for \hl {CFTs} %please define
+\cite{Cavagnaro2014}. As such, blockage should be
+taken into account when using these data for model validation, \emph{i.e.}, the
 experimental domain should be mimicked. The rotor has a relatively high solidity
-$Nc/(\pi D) = 0.13$ and large chord-to-radius ratio $c/R = 0.28$. A CAD model of
+$Nc/(\pi D) = 0.13$ and a large chord-to-radius ratio $c/R = 0.28$. A CAD model of
 the turbine is available from \cite{Bachant2014-RVAT-CAD}.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
 \centering
@@
 
-\caption{UNH-RVAT turbine model. Turbine blades and struts made from NACA 0020
-    profiles with 0.140 m chord. Note that the upper and lower mounting flanges have
-    been excluded as these were included in the tare drag measurements.}
+\caption{\hl {University of New Hampshire (UNH)}-Reference Vertical Axis Turbine (RVAT) turbine model. Turbine blades and struts made from NACA 0020
+ profiles with 0.140-m chord. Note that the upper and lower mounting flanges have
+ been excluded, as these were included in the tare drag measurements.}
 
@@
 Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
-coupled to a Kollmorgen AKM permanent magnet servo motor with a 20:1 gearbox,
-providing precise control over shaft RPM. This servo was controlled by the tow
+coupled to a Kollmorgen \hl {AKM}
+%please define AKM and RPM
+  permanent magnet servo \hl {motor (Manufacturer, City, Country)} with a 20:1 gearbox,
+providing precise control over shaft \hl {RPM}. This servo was controlled by the tow
 tank's main motion controller for high synchronization with the carriage motion,
-thereby giving precise measurement and control of tip speed ratio. An Interface
-T8 200 Nm capacity rotary torque transducer was installed inline between the
+thereby giving precise measurement and control of the tip speed ratio. An Interface
+T8 200-Nm capacity rotary torque \hl {transducer (Manufacturer, City, Country)} was installed inline between the
 servo and the turbine, and the servo was also mounted on a slewing ring bearing,
@@
 guides, such that the total streamwise drag force was transferred to a pair of
-Sentran ZB3 500 lbf capacity S-beam load cells, providing the rotor drag
-measurements, after a separately measured tare drag was subtracted in post
-processing. Similarly, a tare torque was measured by rotating the turbine shaft
+Sentran \hl {ZB3 500-lbf} %please define/check the unit
+ capacity S-beam \hl {load cells (Manufacturer, City, Country)}, providing the rotor drag
+measurements, after a separately measured tare drag was subtracted in post-processing. Similarly, a tare torque was measured by rotating the turbine shaft
 in air. Turbine angular and tow carriage linear position were measured using
-quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10
-$\mathrm{\mu m}$ resolution for the carriage position. These signals, along with
+quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10 ${\mu}$m resolution for the carriage position. These~signals, along with
 the torque and drag signals, were sampled at 2 kHz.
 
-\begin{figure}[ht!]
-    \centering
-
-    \includegraphics[width=0.8\textwidth]{figures/exp-setup-photo}
-
-    \includegraphics[width=0.8\textwidth]{figures/exp_setup_drawing}
-
-    \caption{Experimental setup: turbine test bed installed in UNH tow tank.}
-
-    \label{fig:exp-setup}
-\end{figure}
-
-Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
-(ADV), which has an approximately 6 mm diameter sampling volume, and sampled at
+Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler \hl {velocimeter
+(ADV)  (Manufacturer, City, Country)}, which has an approximately 6-mm diameter sampling volume and sampled at
 200 Hz. The probe was mounted on an automated positioning system, also
-controlled by the tow tank's main motion controller. The ADV and DAQ systems'
+controlled by the tow tank's main motion controller. The ADV and \hl {DAQ (please define)} systems'
 sampling times were synchronized by triggering the start of data acquisition via
@@
 
+\begin{figure}[H]
+ \centering
+
+ \includegraphics[width=0.7\textwidth]{figures/exp-setup-photo}
+
+ \includegraphics[width=0.7\textwidth]{figures/exp_setup_drawing}
+
+ \caption{\hl {Experimental} setup: turbine test bed installed in the UNH tow tank.}
+ %please mark (a) and (b) in the figure and label (a) and (b) in the figure caption.
+
+ \label{fig:exp-setup}
+\end{figure}
+
 \subsection{Test Plan}
 
-Approximately 1500 tows were conducted for the study reported here, each tow was
+Approximately 1500 tows were conducted for the study reported here; each tow was
 used for a single data point on either a performance curve or wake map. Each
-performance curve consisted of 31 tows, where during each tow the mean turbine
-tip speed ratio was held constant, ranging from 0.1--3.1 in 0.1 increments. Full
-performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2
-m/s increments, for which turbine diameter and approximate blade chord Reynolds
+performance curve consisted of 31~tows, where during each tow, the mean turbine
+tip speed ratio was held constant, ranging from 0.1 to 3.1 in 0.1 increments. Full
+performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2-m/s increments, for which the turbine diameter and approximate blade chord Reynolds
 number are presented in Table~\ref{tab:Re}. Performance was also measured for
@@
 
-\begin{table}
+Each wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
+Velocimeter (ADV), measuring three components of velocity at a 200-Hz sampling
+frequency, at 270 different locations, varied in the cross-stream and vertical
+directions at one turbine diameter downstream ($x/D=1$). These locations have
+vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
+the cross-stream direction $y/R = \pm 3$. These locations are shown in
+Figure~\ref{fig:wake-locations}.
+
+\begin{table}[H]
 \centering
 \begin{tabular}{ccc}
-Tow speed (m/s) & $Re_D$ & $Re_{c,\mathrm{ave}}$ ($\lambda = 1.9$) \\
-\hline
+\toprule  
+\textbf{Tow Speed (m/s)} & \boldmath{$Re_D$} & \boldmath{$Re_{c,\mathrm{ave}}$} \textbf{(}\boldmath{$\lambda = 1.9$}\textbf{)} \\
+\midrule
 0.3 & $0.3 \times 10^6$ & $0.8 \times 10^5$ \\
@@
 1.3 & $1.3 \times 10^6$ & $3.4 \times 10^5$ \\
+ \bottomrule
 \end{tabular}
@@
 \end{table}
-
-Each wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
-Velocimeter (ADV), measuring three components of velocity at 200 Hz sampling
-frequency, at 270 different locations, varied in the cross-stream and vertical
-directions at one turbine diameter downstream ($x/D=1$). These locations have
-vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
-the cross-stream direction $y/R = \pm 3$. These locations are shown in
-Figure~\ref{fig:wake-locations}.
-
-\begin{figure}
+\unskip
+
+\begin{figure}[H]
 \centering
@@
 
-Wake velocity data were filtered for spurious ``spikes'' by removing datapoints
-8 standard deviations, or 0.9 m/s from the mean. The experimental data and code
+Wake velocity data were filtered for spurious ``spikes'' by removing data points
+8 standard deviations, or 0.9 m/s, from the mean. The experimental data and code
 for processing and visualization are available from
@@
 An expanded uncertainty interval with 95\% confidence was computed for
-$C_P$, $C_D$, and mean wake velocities
+$C_P$, $C_D$, and mean wake velocities:
 \begin{equation}
@@
 \end{equation}
-where $t_{95}$ is the value from the Student-$t$ distribution for a 95\%
+where $t_{95}$ is the value from the Student $t$-distribution for a 95\%
 confidence interval and $u_c$ is the combined standard uncertainty. Combined
-standard uncertainty for a given quantity $X$ is calculated by
+standard uncertainty for a given quantity $X$ is calculated~by:
 \begin{equation}
@@
 where $s_{\bar{X}}$ is the sample standard deviation of the mean per turbine
-revolution, and $b_X$ is the systematic uncertainty, computed by
+revolution, and $b_X$ is the systematic uncertainty, computed by:
 \begin{equation}
@@
 \end{equation}
-where $x_i$ is a primitive quantity used to calculate $X$ (e.g. $T$, $\omega$,
+where $x_i$ is a primitive quantity used to calculate $X$ (e.g., $T$, $\omega$,
 and $U_\infty$ for calculating $C_P$), and $b_{x_i}$ is the primitive quantity's
@@
 Selecting $t_{95}$ requires an estimate for degrees of freedom $\nu_X$, which
-was obtained using the Welch--Satterthwaite formula
+was obtained using the Welch--Satterthwaite formula:
 \begin{equation}
@@
 $\nu_{s_X}$ is assumed to be $(N-1)$, where $N$ is the number of independent
-samples (or turbine revolutions). $\nu_{b_k}$ was estimated as
+samples (or turbine revolutions). $\nu_{b_k}$ was estimated as:
 \begin{equation}
@@
 
-\begin{table}
+\begin{table}[H]
  \centering
 \begin{tabular}{cccc}
-$U_\infty$ (m/s) &          $U$ (m/s) &          $V$ (m/s) &          $W$ (m/s) \\
-\hline
+\toprule  
+\boldmath{$U_\infty$} \textbf{(m/s)} &   \boldmath{$U$} \textbf{(m/s)} &   \boldmath{$V$} \textbf{(m/s)} &   \boldmath{$W$} \textbf{(m/s)} \\
+ \midrule
    $0.4$ & $1 \times 10^{-2}$ & $7 \times 10^{-3}$ & $6 \times 10^{-3}$ \\
@@
    $1.2$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
+\bottomrule
 \end{tabular}
-\caption{Average expanded uncertainty estimates (with 95\% confidence) for mean
-    velocity measurements at each tow speed.}
+\caption{\textls[-10]{Average expanded uncertainty estimates (with 95\% confidence) for mean
+ velocity measurements} at each tow speed.}
 
@@
 
-Full power and rotor drag (a.k.a. thrust) coefficient curves for various
-Reynolds numbers are plotted in Figure~\ref{fig:cp-curves} and
-Figure~\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
+Full-power and rotor drag (also known as thrust) coefficient curves for various
+Reynolds numbers are plotted in Figures~\ref{fig:cp-curves} and
+\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
 with Reynolds number. The power coefficient curves also show a slight downward
-shift in the optimal tip speed ratio (peak performance) with increasing Reynolds
+shift in the optimal tip speed ratio (peak~performance) with increasing Reynolds
 number, from about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall
 delay from a more turbulent boundary layer on the blade suction side. There is
-essentially no change in the shape of the drag coefficient curves---merely a
-slight upward shift in $C_D$ with increasing $Re$.
-
-\begin{figure}[ht]
+essentially no change in the shape of the drag coefficient curves, merely a
+slight upward shift in $C_D$ with increasing~ $Re$.
+
+\begin{figure}[H]
 \centering
@@
 
-\caption{Mean power coefficient curves plotted for multiple Reynolds numbers.}
-
+\caption{\hl {Mean} power coefficient curves plotted for multiple Reynolds numbers.}
+%is it better to mark different color to the lines?
 \label{fig:cp-curves}
 \end{figure}
-
-\begin{figure}[ht]
+\unskip
+
+\begin{figure}[H]
 \centering
@@
 
-\caption{Mean rotor drag coefficient curves plotted for multiple Reynolds
+\caption{\hl {Mean} rotor drag coefficient curves plotted for multiple Reynolds
  numbers.}
+ % is it better to mark different color to the lines?
 
@@
 
-Mean power and drag coefficients at $\lambda=1.9$ are plotted versus Reynolds
+Mean power and drag coefficients at $\lambda=1.9$ are plotted \textit{versus} the Reynolds
 number in Figure~\ref{fig:perf-Re-dep}. Note that the large error bars for $C_P$
@@
 low. There is a drastic improvement in $C_P$ with increasing Reynolds number at
-lower values of Reynolds number. The power coefficient then becomes  essentially
+lower values of the Reynolds number. The power coefficient then becomes essentially
 $Re$-independent at $Re_D = 0.8 \times 10^6$, which corresponds to an
@@
 
+\begin{figure}[H]
+\centering
+
+\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
+
+\caption{UNH-RVAT measured mean power (\hl {left}) and drag (right) coefficients at
+ $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error bars indicate expanded
+ uncertainty estimates for 95\% confidence, which for $C_P$ is dominated by
+ systematic error estimates from the torque transducer operating at the lower end
+ of its measurement range.}
+
+%please use (a) and (b) to instead left and right and mark (a) and (b) in the figure.
+
+\label{fig:perf-Re-dep}
+\end{figure}
+
 The behavior of the mean rotor drag coefficient $C_D$ is similar, though the
 changes are less dramatic. This is likely due to cross-flow turbine geometry,
-where increases in blade drag at lower $Re$ somewhat offsets the reduction in
-lift, causing the total streamwise force to vary less than the rotor torque. The
-tendency for $C_D$ to continue increasing with $Re$ may also include the effects
+where increases in blade drag at lower $Re$ somewhat offset the reduction in
+lift, causing the total streamwise force to vary less than the rotor torque. The~tendency for $C_D$ to continue increasing with $Re$ may also include the effects
 of an increasing Froude number (from faster tow speeds), which increases free
@@
 
-\begin{figure}[ht]
-\centering
-
-\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
-
-\caption{UNH-RVAT measured mean power (left) and drag (right) coefficients at
-    $\lambda=1.9$ plotted versus Reynolds number. Error bars indicate expanded
-    uncertainty estimates for 95\% confidence, which for $C_P$ is dominated by
-    systematic error estimates from the torque transducer operating at the lower end
-    of its measurement range.}
-
-\label{fig:perf-Re-dep}
-\end{figure}
-
 \subsubsection{Relation to Static Foil Characteristics}
 
-To help understand---and possibly predict---the $Re$-sensitivity on turbine
+To help understand, and possibly predict, the $Re$-sensitivity on turbine
 performance, a series of static foil coefficient datasets were computed with the
-XFOIL viscous panel method code \cite{Drela1989}, a commonly used tool for
-airfoil analysis, e.g. \cite{Castelli2011, Walker2014}, implemented as part of
+\hl {XFOIL (please define)} viscous panel method code \cite{Drela1989}, a commonly-used tool for
+airfoil analysis, e.g., \cite{Castelli2011, Walker2014}, implemented as part of
 the open source turbine design software QBlade \cite{Marten2013}. Simulations
-were run for an angle of attack range of 0--40 degrees, in increments of 0.5
-degrees. Solver parameters used were 100 panels, a fixed speed, zero Mach
-number, \texttt{NCrit=9} (default $\mathrm{e}^n$ transition criteria parameter
-for an average wind tunnel), and no forced boundary layer transition.
+were run for an angle of attack range of 0$^{\circ}$
+ to 40$^{\circ}$, in increments of 0.5$^{\circ}$. Solver parameters used were 100 panels, a fixed speed, zero Mach
+number, \texttt{NCrit = 9} %please check the use of \texttt here
+ (default $\mathrm{e}^n$ transition criteria parameter
+for an average wind tunnel) and no forced boundary layer transition.
 Characteristics were computed for the approximate average blade chord Reynolds
@@
 circular path \cite{Migliore1980}, the XFOIL calculations were performed for
-20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520), and 4\% (NACA 4520)
+20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520) and 4\% (NACA 4520)
 camber about their half-chord location. A 4\% camber approximates the maximum
-distance between the blade chord line and path for the UNH-RVAT and the 2\%
+distance between the blade chord line and path for the UNH-RVAT, and the 2\%
 camber takes into account the reduction in virtual flow curvature from the
@@
 Results from the XFOIL calculations for the airfoil profiles are shown in
-Figure~\ref{fig:foil-Re-dep}, where values of maximum lift coefficient
-$C_{l_{\max}}$, minimum drag coefficient $C_{d_{\min}}$, and maximum
-lift-to-drag ratio are normalized (for comparison between foils) and plotted
-versus $Re_c$. In general, larger camber is associated with decreased foil
+Figure~\ref{fig:foil-Re-dep}, where values of the maximum lift coefficient
+$C_{l_{\max}}$, minimum drag coefficient $C_{d_{\min}}$ and maximum
+lift-to-drag ratio are normalized (for the comparison between foils) and plotted
+\textit{versus} $Re_c$. In general, larger camber is associated with decreased foil
 performance at lower Reynolds number. The data and processing code for these
@@
 
-\begin{figure}[ht!]
+\begin{figure}[H]
 \centering
@@
 torque coefficient $C_T$ can be related to the blade section chordwise force
-coefficient $C_c$ by
+coefficient $C_c$ by:
 \begin{equation}
@@
 where the blade section chordwise force coefficient (for zero preset pitch) is
-given by
+given by:
 \begin{equation}
@@
 of the free stream velocity and the opposite of the blade tangential velocity.
-Note that this neglects any induction, or slowing of the free stream by the
-turbine forces, which would be present in a momentum/streamtube model. Since the
-goal of this approach was not to predict absolute performance but rather gain
-insight into relative changes with $Re$, this method was deemed acceptable as it
-is extremely simple and fast to compute.
-
-Values for the blade angle of attack, relative velocity, and torque coefficient
+Note that this neglects any induction or slowing of the free stream by the
+turbine forces, which would be present in a momentum/streamtube model. Since~the
+goal of this approach was not to predict absolute performance, but rather gain
+insight into relative changes with $Re$, this method was deemed acceptable, as it
+is extremely simple and fast to~compute.
+
+Values for the blade angle of attack, relative velocity and torque coefficient
 are plotted in Figure~\ref{fig:blade-kinematics} for the upstream half of
@@
 torque coefficient plot, and by the time the angle of attack has decreased below
-that of stall, the relative velocity is so low that there is not much
+that of the stall, the relative velocity is so low that there is not much
 contribution to the torque coefficient.
@@
 
-\caption{Geometric angle of attack, relative velocity and torque coefficient
+\caption{\hl {Geometric} angle of attack, relative velocity and torque coefficient
  calculated with a NACA 0020 foil operating at $\lambda=1.9$.}
-
+%please mark (a) and (b), (c) in the figure and label them in the caption
 \label{fig:blade-kinematics}
@@
 Results for the normalized peak torque coefficient for each foil are plotted
-versus $Re_c$ in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the
+\textit{versus} $Re_c$ in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the
 convergence of $C_{T_\mathrm{max}}$ with increasing Reynolds number is more
@@
 Figure~\ref{fig:foils-C_T-Re-dep}, the Reynolds number independence is achieved
-at lower values, and more dramatically. We see that the trend of the
+at lower values and more dramatically. We see that the trend of the
 (non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
@@
 predictions of turbine performance, it predicts the transitional Reynolds number
-regime for cross-flow turbine performance using only 2-D static airfoil
-characteristics, i.e., Reynolds number scaling of the peak $C_T$ computed this
+regime for cross-flow turbine performance using only 2D static airfoil
+characteristics, \emph{i.e.}, the Reynolds number scaling of the peak $C_T$ computed this
 way behaves much like that of the measured turbine performance.
 
-\begin{figure}[ht!]
-\centering
-
-\includegraphics[width=0.65\textwidth]{figures/cft_re_dep_foils}
-
-\caption{Reynolds number dependence of the normalized peak torque coefficient
+\begin{figure}[H]
+\centering
+
+\includegraphics[width=0.8\textwidth]{figures/cft_re_dep_foils}
+
+\caption{\hl {Reynolds} number dependence of the normalized peak torque coefficient
  calculated from static foil coefficients and blade kinematics, compared to
@@
  represents the mean torque coefficient, not the maximum.}
+ %please pay attention to the value ``1e5'' in the figure.
+is it ``\times 10^5''.
+If yes, please revise it to \times 10 ^5 in the figure.
+meanwhile, is it better to make different color with the lines?
 
@@
 
-The near-wake of this turbine at $x/D=1$, $\lambda=1.9$, and momentum and
+The near-wake of this turbine at $x/D=1$, $\lambda=1.9$ and momentum and
 kinetic energy balances at a $Re_D = 1.0 \times 10^6$ were discussed in
-\cite{Bachant2015-JoT}. Similar data were taken for the experiment here, and
+\cite{Bachant2015-JoT}. Similar data were taken for the experiment here, and the
 results for the mean velocity field and turbulence kinetic energy calculated
 from wake maps of 270 individual measurements (tows) are shown in
-Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont}, respectively, looking
+Figures~\ref{fig:meancontquiv} and \ref{fig:kcont}, respectively, looking
 upstream towards the turbine. With respect to the mean velocity field, we see
@@
 These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times
-10^6$, $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look very
+10^6$, $0.8 \times 10^6$ and $1.2 \times 10^6$. Qualitatively, these look very
 similar, so they have not been plotted here, though profiles at $z/H=0.0$ are
@@
 
-\begin{figure}[ht!]
+\begin{figure}[H]
  \centering
@@
 
-    \caption{Mean velocity field at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0 \times
+ \caption{Mean velocity field at $x/D=1$, $\lambda=1.9$ and $Re_D=1.0 \times
  10^6$.}
@@
 \end{figure}
-
-\begin{figure}[ht!]
+\unskip
+
+\begin{figure}[H]
  \centering
@@
 
-    \caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0
+ \caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$ and $Re_D=1.0
  \times 10^6$.}
@@
 
-    \caption{Mean streamwise velocity (left) and turbulence kinetic energy
-        (right) profiles at $z/H=0.0$. Legend indicates $Re_D$.}
-
+ \caption{Mean streamwise velocity \hl {(left)} and turbulence kinetic energy
+ (right) profiles at $z/H=0.0$. The legend indicates $Re_D$.}
+ %please use (a) and (b) to instead left and right and mark (a) and (b) in the figure.
  \label{fig:profiles}
@@
 intervals. These spectra are plotted in Figure~\ref{fig:wake-spectra} for
-regions on either side of the turbine. On the $-y$ side of the turbine there is
-broadband turbulence produced by blade stall, and on the $+y$ side there is a
-clear peak in the spectra caused by the blade passage. We see that on both sides
+regions on either side of the turbine. On the $-y$ side of the turbine, there is
+broadband turbulence produced by blade stall, and on the $+y$ side, there is a
+clear peak in the spectra caused by the blade passage. We see that on both sides,
 there is higher spectral energy at lower Reynolds numbers. On the $+y$ side of
-the turbine we notice higher spectral energy at the blade passage frequency's
+the turbine, we notice higher spectral energy at the blade passage frequency's
 first harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's
@@
 \caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$,
-    $y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (a) and $y/R=1.5$
-    ($+0.75$ m in Figure~\ref{fig:wake-locations}) (b). Dashed vertical lines
+ $y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (\textbf{a}) and $y/R=1.5$
+ ($+0.75$ m in Figure~\ref{fig:wake-locations}) (\textbf{b}). Dashed vertical lines
  indicate $[1, 3, 6, 9]$ times the turbine rotational frequency. Shaded regions
-    indicate 95\% confidence intervals assuming a $\chi^2$ distribution.}
+ indicate 95\% confidence intervals assuming a $\chi^2$~distribution.}
 
@@
 streamwise partial derivative ($\partial / \partial x$) for the quantities of
-interest \cite{Bachant2015-JoT}, i.e. momentum:
+interest~\cite{Bachant2015-JoT}, \emph{i.e.}, momentum:
 
@@
 		\bigg{[}
-		& - \underbrace{V \frac{\p K}{\p y}}_{y\text{-adv.}}
+		& - \underbrace{V \frac{\p K}{\p y}}_{y\text{-adv.%please define/check the convention
+		}}
 		- \underbrace{W \frac{\p K}{\p z}}_{z\text{-adv.}}
@@
 		\frac{1}{2}\frac{\p}{\p x_j} \overline{u_i' u_j'} U_i
-		}_{\text{Turb. trans.}} \\
+		}_{\text{Turb.%please define/check the convention
+		trans.%please define/check the convention
+		}} \\
 		% Production of k
@@
 		\overline{u_i' u_j'} \frac{\p U_i}{\p x_j}
-		}_{k\text{-prod.}}
+		}_{k\text{-prod.%please define/check the convention
+		}}
 		% Mean dissipation
@@
 		2 \nu S_{ij}S_{ij}
-		}_{\text{Mean diss.}}
+		}_{\text{Mean diss.%please define/check the convention
+		}}
 		\bigg{]}.
@@
 \end{equation}
-Note that Equations~\ref{eq:momentum} and \ref{eq:K-full} only differ from the
+
+Note that Equations~(\ref{eq:momentum}) and (\ref{eq:K-full}) only differ from the
 typical forms of the mean momentum and mean kinetic energy equations by a factor
 of $1/U$, but importantly, the other convective terms ( $\partial U/\partial y$,
-$\partial U/\partial z$, and $\partial K/\partial y$, $\partial K/\partial z$ )
-are moved to the right-hand-sides of the equations to illustrate with what sign
+$\partial U/\partial z$ and $\partial K/\partial y$, $\partial K/\partial z$ )
+are moved to the right-hand sides of the equations to illustrate with what sign
 they add to the streamwise recovery of streamwise momentum or mean kinetic
@@
 
-Results from the terms that can be computed from the experimental data (i.e.
-excluding $\partial / \partial x$) from the right hand sides of
-Equations~\ref{eq:momentum} and \ref{eq:K-full} are plotted in
+Results from the terms that can be computed from the experimental data (\emph{i.e.},
+excluding $\partial / \partial x$) from the right-hand sides of
+Equations~(\ref{eq:momentum}) and (\ref{eq:K-full}) are plotted in
 Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph}, respectively.
 Derivatives were computed using a second order finite difference scheme, with
-central differencing for interior points, and inward facing schemes for the
+central differencing for interior points and inward facing schemes for the
 boundaries. A weighted average for each term was then calculated based on the
@@
 
+\begin{figure}[H]
+ \centering
+
+ \includegraphics[width=0.9\textwidth]{figures/mom_bar_graph}
+
+ \caption{Normalized momentum transport quantities computed as weighted
+ averages from Equation~(\ref{eq:momentum}).}
+
+ \label{fig:mom-bar-graph}
+\end{figure}
+
+\begin{figure}[H]
+ \centering
+
+ \includegraphics[width=0.9\textwidth]{figures/K_trans_bar_graph}
+
+ \caption{Normalized mean kinetic energy transport quantities computed as
+ weighted averages based on Equation~(\ref{eq:K-full}), omitting the
+ non-measured streamwise derivatives. Note that directions of the turbulent
+ transport terms refer to the directions of their partial derivatives.}
+
+ \label{fig:K-bar-graph}
+\end{figure}
+
 We see that in general, levels of turbulent transport are slightly lower at
@@
 Reynolds numbers, which is consistent with the physical meaning of the Reynolds
-number itself. From these results one might expect that the viscous effects will
-become significant to the wake dynamics as Reynolds number is decreased below
+number itself. From these results, one might expect that the viscous effects will
+become significant to the wake dynamics as the Reynolds number is decreased below
 $Re_D \sim 10^4$.
@@
 Transport due to cross-stream advection appears to become more negative at
-higher $Re$. This could be a consequence of increasing free surface deformation
+higher $Re$. This~could be a consequence of increasing free surface deformation
 (higher $Fr$), which then decreases the effective flow cross-sectional area
@@
 
-\begin{figure}[ht!]
-    \centering
-
-    \includegraphics[width=0.9\textwidth]{figures/mom_bar_graph}
-
-    \caption{Normalized momentum transport quantities computed as weighted
-        averages from Equation~\ref{eq:momentum}.}
-
-    \label{fig:mom-bar-graph}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-
-    \includegraphics[width=0.9\textwidth]{figures/K_trans_bar_graph}
-
-    \caption{Normalized mean kinetic energy transport quantities computed as
-        weighted averages based on Equation~\ref{eq:K-full}, omitting the
-        non-measured streamwise derivatives. Note that directions of the turbulent
-        transport terms refer to the directions of their partial derivatives.}
-
-    \label{fig:K-bar-graph}
-\end{figure}
-
-The totals for all the wake transport terms calculated in
+The totals for all of the wake transport terms calculated in
 Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph} are plotted in
-Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport of
+Figure~\ref{fig:wake-trans-totals}. We see that in general, the wake transport of
 both mean streamwise momentum and kinetic energy is enhanced at lower Reynolds
@@
 turbine arrays, where increased levels of wake recovery could motivate different
-ideal array configurations when compared with full-scale turbines if the scale
+ideal array configurations when compared to full-scale turbines if the scale
 model Reynolds number is too low.
 
-\begin{figure}[ht!]
+\begin{figure}[H]
 \centering
@@
 \caption{Normalized transport totals from Figures~\ref{fig:mom-bar-graph} and
-\ref{fig:K-bar-graph} plotted versus Reynolds number.}
+\ref{fig:K-bar-graph} plotted \textit{versus} the Reynolds number.}
 \label{fig:wake-trans-totals}
 \end{figure}
+%\unskip
 
@@
 
-In this study it was demonstrated that the performance of a high solidity
+In this study, it was demonstrated that the performance of a high solidity
 ($c/R=0.28$) cross-flow turbine becomes essentially $Re$-independent at a
-Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
-approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
-10^5$. The power coefficient values on the left hand side of the
-$C_P$--$\lambda$ curve are reduced more than those on the right hand side at low
+Reynolds number based on a turbine diameter of $Re_D \approx 10^6$ or an
+approximate Reynolds number based on a blade chord of $Re_c \approx 2 \times
+10^5$. The power coefficient values on the left-hand side of the
+$C_P$-$\lambda$ curve are reduced more than those on the right-hand side at low
 $Re$ due to deeper stalling. The $Re$-independent threshold corresponds to that
@@
 
-We propose a method for predicting Reynolds number dependence of a cross-flow
+We propose a method for predicting the Reynolds number dependence of a cross-flow
 turbine using static airfoil characteristics and turbine blade kinematics:
@@
  profile.
-
-    \item For azimuthal angles of 0--180 degrees and a given tip speed ratio,
+\item For azimuthal angles of 0 to 180 degrees and a given tip speed ratio,
  calculate the geometric angle of attack and relative velocity magnitude from
-    the blade and undisturbed free stream velocity ($\equiv 1$) vectors.
-
+ the blade and undisturbed free stream velocity \hl {($\equiv 1$)} vectors.
+ %please check this symbol
 \item With arrays of geometric angle of attack and relative velocity
-    magnitude, calculate the blade chordwise force and then rotor torque
-    coefficient from Equations~\ref{eq:cc} and \ref{eq:ct}, respectively.
-
+ magnitude, calculate the blade chordwise force and then the rotor torque
+ coefficient from Equations~(\ref{eq:ct}) and (\ref{eq:cc}), respectively.
 \item Extract the maximum value of $C_T$, and repeat this process for static
@@
 It was shown that the peak torque coefficient computed this way shows similar
-Reynolds number sensitivity---a better predictor than conventional
-quantifications of airfoil performance, e.g., lift-to-drag ratio. Note that this
+Reynolds number sensitivity, a better predictor than conventional
+quantifications of airfoil performance, e.g., the lift-to-drag ratio. Note that this
 method is not restricted to standard symmetrical NACA foils, though those
-evaluated here with higher camber were more sensitive at lower $Re$ but had
-smaller slopes in the linear regime.
-
-In the near-wake of the turbine we observed lower levels of turbulence with
+evaluated here with higher camber were more sensitive at lower $Re$, but had
+smaller slopes in the linear~regime.
+
+In the near-wake of the turbine, we observed lower levels of turbulence with
 increasing Reynolds number, along with lower levels of turbulent transport with
 respect to mean streamwise momentum and mean kinetic energy recovery in the
-streamwise direction. Vertical advection---the largest transport mechanism
-measured---showed little $Re$-dependence, whereas the negative effects of
+streamwise direction. Vertical advection, the largest transport mechanism
+measured, showed little $Re$-dependence, whereas the negative effects of
 cross-stream advection were enhanced at high $Re$, which may be due to larger
@@
 area. Despite being orders of magnitude smaller than other transport processes,
-viscous diffusion increased rapidly with decreasing $Re$, and is expected to
+viscous diffusion increased rapidly with decreasing $Re$ and is expected to
 become significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the
 total wake transport measured leveled off at essentially the same Reynolds
-number that the performance did---$Re_D \approx 0.8 \times 10^6$.
-
-From these results we recommend that physical model tests of cross-flow turbines
+number that the performance did: $Re_D \approx 0.8 \times 10^6$.
+
+From these results, we recommend that physical model tests of cross-flow turbines
 be performed at $Re_D \sim 10^6$ or greater to provide reasonable predictions of
@@
 boundary layer is to be resolved, since transition to turbulence plays an
-important role in overall blade loading. Note that in our case blockage, though
-reasonably low, may be increasing flow through the turbine compared with a free
+important role in overall blade loading. Note that in our case, blockage, though
+reasonably low, may be increasing flow through the turbine compared to a free
 case, which could mean $Re$ thresholds for free flows could be slightly higher,
 though a reliable blockage correction algorithm has not yet been agreed upon for
-CFTs \cite{Cavagnaro2014}.
+CFTs~ \cite{Cavagnaro2014}.
 
@@
 model studies of turbine arrays may see exaggerated levels of wake recovery,
-leading to inadequate or inappropriate spacing or layouts.
-
-\acknowledgements{Acknowledgements}
-
-The authors would like to acknowledge funding through a National Science
-Foundation CAREER award (PI Wosnik, NSF-CBET 1150797, program manager Dr.
-Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine Program
-Endowment to purchase acoustic flow measurement instrumentation, and a grant for
-laboratory infrastructure upgrades through the US Department of Energy.
-
-\conflictofinterests{Conflicts of Interest}
-
-The authors declare no conflict of interest.
-
-\bibliography{library}
+leading to inadequate or inappropriate spacing or \hl {layouts}.
+%please define CAREER, PI, NSF-CBET
+
+\vspace{6pt}
+\acknowledgments{\textbf{Acknowledgments:} The authors would like to acknowledge funding through a National Science
+Foundation CAREER %please define
+ award (PI %please define
+  Wosnik, NSF-CBET%please define
+ 1150797, program manager 
+Gregory L. Rorrer), a grant through the \linebreak Leslie S. Hubbard Marine Program
+Endowment to purchase acoustic flow measurement instrumentation and a grant for
+laboratory infrastructure upgrades through the U.S. Department of Energy.}
+
+\authorcontributions{\textbf{Author Contributions:} \hl{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.}}
+
+\conflictofinterests{\textbf{Conflicts of Interest:} The authors declare no conflict of interest.}
+
 \bibliographystyle{mdpi}
+\renewcommand\bibname{References}
+%\bibliography{library}
+%\bibliographystyle{mdpi}
+%
+%\begin{thebibliography}{999}
+%%\providecommand{\natexlab}[1]{#1}
+%
+%\end{thebibliography}
+
+\begin{thebibliography}{999}
+\providecommand{\natexlab}[1]{#1}
+
+\bibitem[Acheson(1990)]{Acheson1990}
+Acheson, D.
+\newblock {\em Elementary Fluid Dynamics}; Oxford University Press: \hl{City, Country,} 1990.
+
+\bibitem[Baker and Brockie(1991)]{Baker1991}
+Baker, C.; Brockie, N.
+\newblock Wind tunnel tests to obtain train aerodynamic drag coefficients:
+ {R}eynolds number and ground simulation effects.
+\newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf
+ 1991}, {\em 38},~23--28.
+
+\bibitem[Paraschivoiu(2002)]{Para2002}
+Paraschivoiu, I.
+\newblock {\em Wind Turbine Design with Emphasis on Darrieus Concept}, 1st ed.;
+ Polytechnic International: Montreal, QC, Canada, 2002.
+
+\bibitem[Templin(1974)]{Templin1974}
+Templin, R.J.
+\newblock \emph{Aerodynamic Performance Theory for the NRC Vertical Axis Wind
+ Turbine};
+\newblock Technical Report LTR-LA-160; \hl {NRC (Please define)} of Canada: \hl{City, Country,} 1974.
+
+\bibitem[Fiedler and Tullis(2009)]{Fiedler2009}
+Fiedler, A.J.; Tullis, S.
+\newblock Blade offset and pitch effects on a high solidity vertical axis wind
+ turbine.
+\newblock {\em Wind~Eng.} {\bf 2009}, {\em 33},~237--246.
+
+\bibitem[Migliore \em{et~al.}(1980)Migliore, Wolfe, and Fanucci]{Migliore1980}
+Migliore, P.G.; Wolfe, W.P.; Fanucci, J.B.
+\newblock Flow curvature effects on {d}arrieus turbine blade aerodynamics.
+\newblock {\em J.~Energy} {\bf 1980}, {\em 4},~49--55.
+
+\bibitem[Blackwell \em{et~al.}(1976)Blackwell, Sheldahl, and
+ Feltz]{Blackwell1976}
+Blackwell, B.; Sheldahl, R.; Feltz, L.
+\newblock \emph{Wind Tunnel Performance Data for the {Darrieus} Wind Turbine with
+ {NACA} 0012 Blades};
+\newblock Report SAND76-0130; Sandia National Laboratories: Albuquerque, NM, USA,
+ 1976.
+
+\bibitem[Polagye \em{et~al.}(2013)Polagye, Cavagnaro, and
+ Niblick]{Polagye2013b}
+Polagye, B.L.; Cavagnaro, R.J.; Niblick, A.L.
+\newblock Micropower from tidal turbines.
+\newblock In Proceedings of the ASME Fluids Engineering Division Summer Meeting, \hl{City, Country,} \hl{Day Month} 2013.
+
+\bibitem[Bravo \em{et~al.}(2007)Bravo, Tullis, and Ziada]{Bravo2007}
+Bravo, R.; Tullis, S.; Ziada, S.
+\newblock Performance testing of a small vertical-axis wind turbine.
+\newblock In Proceedings of the 21st Canadian Congress of Applied Mechanics
+ CANCAM, \hl{City, Country,} \hl{Day Month} 2007.
+
+\bibitem[Bachant and Wosnik(2014)]{Bachant2014}
+Bachant, P.; Wosnik, M.
+\newblock Reynolds number dependence of cross-flow turbine performance and
+ near-wake characteristics.
+\newblock In Proceedings of the 2nd Marine Energy Technology Symposium METS2014, \hl{City, Country,} \hl{Day Month} 2014.
+
+\bibitem[Brochier \em{et~al.}(1986)Brochier, Fraunie, Beguier, and
+ Paraschivoiu]{Brochier1986}
+Brochier, G.; Fraunie, P.; Beguier, C.; Paraschivoiu, I.
+\newblock Water channel experiments of dynamic stall on darrieus wind turbine
+ blades.
+\newblock {\em AIAA J. Propuls. Power} {\bf 1986}, {\em
+ 2},~46--510.
+
+\bibitem[Fujisawa and Shibuya(2001)]{Fujisawa2001}
+Fujisawa, N.; Shibuya, S.
+\newblock Observations of dynamic stall on Darrieus wind turbine blades.
+\newblock {\em J. Wind Eng. Ind.~Aerodyn.} {\bf
+ 2001}, {\em 89},~201--214.
+
+\bibitem[Tescione \em{et~al.}(2014)Tescione, Ragni, He, Ferreira, and
+ G.J]{Tescione2014}
+Tescione, G.; Ragni, D.; He, C.; Ferreira, C.S.; G.J.
+\newblock Near wake flow analysis of a vertical axis wind turbine by
+ stereoscopic particle image velocimetery.
+\newblock {\em Renew. Energy} {\bf 2014}, {\em 70},~47--61.
+
+\bibitem[Krogstad and Adaramola(2012)]{Krogstad2012a}
+Krogstad, P.; Adaramola, M.S.
+\newblock Performance and near wake measurements of a model horizontal axis
+ wind turbine.
+\newblock {\em Wind Energy} {\bf 2012}, {\em 15},~743--756.
+
+\bibitem[Walker \em{et~al.}(2014)Walker, Flack, Lust, and Luznik]{Walker2014}
+Walker, J.M.; Flack, K.A.; Lust, E.E.; Luznik, M.P.S.L.
+\newblock Experimental and numerical studies of blade roughness and fouling on
+ marine current turbine performance.
+\newblock {\em Renew. Energy} {\bf 2014}, {\em 66},~257--267.
+
+\bibitem[McTavish \em{et~al.}(2013)McTavish, Feszty, and
+ Nitzsche]{McTavish2013}
+McTavish, S.; Feszty, D.; Nitzsche, F.
+\newblock Evaluating {R}eynolds number effects in small-scale wind turbine
+ experiments.
+\newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf 2013}, {\em 120},~81--90.
+
+\bibitem[Chamorro \em{et~al.}(2011)Chamorro, Arndt, and
+ Sotiropoulos]{Chamorro2011}
+Chamorro, L.P.; Arndt, R.E.A.; Sotiropoulos, F.
+\newblock Turbulent flow properties around a staggered wind farm.
+\newblock {\em Bound. Layer Meteorol.} {\bf 2011}, \emph{141}, 349--367.
+
+\bibitem[Chamorro and Porte-Agel(2011)]{Chamorro2011b}
+Chamorro, L.P.; Porte-Agel, F.
+\newblock Turbulent flow inside and above a wind farm: A wind-tunnel study.
+\newblock {\em Energies} {\bf 2011}, {\em 4}, 1916--1936.
+
+\bibitem[Goldenberg and Fekete(1983)]{Goldenberg1983}
+Goldenberg, J.; Fekete, G.
+\newblock Mean flow energy content of boundary layer downstream of
+ vertical-axis wind-turbine simulators.
+\newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf
+ 1983}, {\em 12},~1--14.
+
+\bibitem[Dabiri(2011)]{Dabiri2011}
+Dabiri, J.
+\newblock Potential order-of-magnitude enhancement of wind farm power density
+ via counter-rotating vertical-axis wind turbine arrays.
+\newblock {\em J. Renew. Sustain. Energy} {\bf 2011}, {\em
+ 3},~1--13.
+
+\bibitem[Kinzel \em{et~al.}(2012)Kinzel, Mulligan, and Dabiri]{Kinzel2012}
+Kinzel, M.; Mulligan, Q.; Dabiri, J.O.
+\newblock Energy exchange in an array of vertical-axis wind turbines.
+\newblock {\em J. Turbul.} {\bf 2012}, {\em 13},~1--13.
+
+\bibitem[Bachant and Wosnik(2015)]{Bachant2015-JoT}
+Bachant, P.; Wosnik, M.
+\newblock Characterising the near-wake of a cross-flow turbine.
+\newblock {\em J. Turbul.} {\bf 2015}, {\em 16},~392--410.
+
+\bibitem[Jacobs and Sherman(1937)]{Jacobs1937}
+Jacobs, E.N.; Sherman, A.
+\newblock \emph{Airfoil Section Characteristics as Affected by Variation of the
+ Reynolds Number};
+\newblock Technical Report; National Advisory Committee for Aeronautics: \hl{City, Country}, 1937.
+
+\bibitem[Lissaman(1983)]{Lissaman1983}
+Lissaman, P.B.S.
+\newblock Low-{R}eynolds-number airfoils.
+\newblock {\em Ann. Rev. Fluid Mech.} {\bf 1983}, {\em
+ 15},~223--239.
+
+\bibitem[McMasters and Henderson(1980)]{McMasters1980}
+McMasters, J.H.; Henderson, M.L.
+\newblock Low-speed single-element airfoil synthesis.
+\newblock {\em Tech. Soar.} {\bf 1980}, {\em 6},~1--21.
+
+\bibitem[Bedon \em{et~al.}(2014)Bedon, Antonini, Betta, Castelli, and
+ Benini]{Bedon2014}
+Bedon, G.; Antonini, E.G.; Betta, S.D.; Castelli, M.R.; Benini, E.
+\newblock Evaluation of the different aerodynamic databases for vertical axis
+ wind turbine simulations.
+\newblock {\em Renew. Sustain. Energy Rev.} {\bf 2014}, {\em 40}, 386--399.
+
+\bibitem[Bousman(2000)]{Bousman2000-evaluation}
+Bousman, W.G.
+\newblock Evaluation of airfoil dynamic stall characteristics for
+ maneuverability.
+\newblock In Proceedings of the 26th {E}uropean Rotorcraft Forum, \hl{City, Country,} \hl{Day Month} 2000.
+
+\bibitem[Singleton and Yeager(2000)]{Singleton2000}
+Singleton, J.D.; Yeager, W.T., Jr.
+\newblock Important scaling parameters for testing model-scale helicopter
+ rotors.
+\newblock {\em J.~Aircr.} {\bf 2000}, {\em 37},~396--402.
+
+\bibitem[Yoon \em{et~al.}(2005)Yoon, Hill, Balachandar, Adrian, and
+ Ha]{Yoon2005}
+Yoon, H.; Hill, D.; Balachandar, S.; Adrian, R.; Ha, M.
+\newblock Reynolds number scaling of flow in a Rushton turbine stirred tank.
+ Part I—Mean flow, circular jet and tip vortex scaling.
+\newblock {\em Chem. Eng. Sci.} {\bf 2005}, {\em 60},~3169--3183.
+
+\bibitem[Chamorro \em{et~al.}(2012)Chamorro, Arndt, and
+ Sotiropolous]{Chamorro2012}
+Chamorro, L.P.; Arndt, R.E.A.; Sotiropolous, F.
+\newblock Reynolds number dependence of turbulence statistics in the wake of
+ wind turbines.
+\newblock {\em Wind Energy} {\bf 2012}, {\em 15},~733--742.
+
+\bibitem[Neary \em{et~al.}(2014)Neary, Previsic, Jepsen, Lawson, Yu, Copping,
+ Fontaine, Hallett, and Murray]{Neary2014}
+Neary, V.S.; Previsic, M.; Jepsen, R.A.; Lawson, M.J.; Yu, Y.H.; Copping, A.E.;
+ Fontaine, A.A.; Hallett, K.C.; Murray, D.K.
+\newblock \emph{Methodology for Design and Economic Analysis of Marine Energy
+ Conversion ({MEC}) Technologies};
+\newblock Technical Report SAND2014-9040; Sandia National Laboratories: \hl{City, Country}, 2014.
+
+\bibitem[Cavagnaro and Polagye(2014)]{Cavagnaro2014}
+Cavagnaro, R.; Polagye, B.
+\newblock An evaluation of blockage corrections for a helical cross-flow
+ turbine.
+\newblock In~Proceedings of the 3rd Oxford Tidal Energy Workshop, \hl{City, Country,} \hl{Day Month} 2014.
+
+\bibitem[Bachant and Wosnik(2014)]{Bachant2014-RVAT-CAD}
+Bachant, P.; Wosnik, M.
+\newblock {UNH-RVAT CAD Models}.
+\newblock \hl {Fig\textbf{share}}. 
+%Is it necessary to make the type of share bold?
+Available online: 
+ \url{http://dx.doi.org/10.6084/m9.figshare.1062009} (accessed on \hl{Day Month} 2014).
+
+\bibitem[Bachant and Wosnik(2015)]{Bachant2015-RVAT-Re-dep-data}
+Bachant, P.; Wosnik, M.
+\newblock {UNH-RVAT} {R}eynolds Number Dependence Experiment: Reduced Dataset
+ and Processing Code.
+\newblock Fig\textbf{share}. Available online: 
+ \url{http://dx.doi.org/10.6084/m9.figshare.1286960} (accessed on \hl{Day Month} 2015).
+
+\bibitem[Coleman and Steele(2009)]{ColemanSteele}
+Coleman, H.W.; Steele, W.G.
+\newblock {\em Experimentation, Validation, and Uncertainty Analysis for
+ Engineers}, 3rd ed.; John~ Wiley \& Sons: \hl{ Hoboken, NJ, USA,} 2009.
+
+\bibitem[Drela(1989)]{Drela1989}
+Drela, M.
+\newblock XFOIL: An Analysis and Design System for Low Reynolds Number
+ Airfoils. In {\em Low Reynolds Number Aerodynamics}; Lecture Notes in Engineering; Springer Verlag: \hl{Berlin, Germany,} 1989;
+ Volume~54.
+
+\bibitem[Castelli \em{et~al.}(2011)Castelli, Garbo, and Benini]{Castelli2011}
+Castelli, M.R.; Garbo, F.; Benini, E.
+\newblock Numerical investigation of laminar to turbulent boundary layer
+ transition on a NACA 0012 airfoil for vertical-axis wind turbine
+ applications.
+\newblock {\em Wind Eng.} {\bf 2011}, {\em 35},~661--686.
+
+\bibitem[Marten \em{et~al.}(2013)Marten, Wendler, Pechlivanoglou, Nayeri, and
+ Paschereit]{Marten2013}
+Marten, D.; Wendler, J.; Pechlivanoglou, G.; Nayeri, C.; Paschereit, C.
+\newblock {QB}lade: An open source tool for design and simulation of horizontal
+ and vertical axis wind turbines.
+\newblock {\em Int. J. Emerg. Technol. Adv.
+ Eng.} {\bf 2013}, {\em 3},~264--269.
+
+\bibitem[Bachant and Wosnik(2015)]{Bachant2015-NACAXX20-XFOIL}
+Bachant, P.; Wosnik, M.
+\newblock NACA XX20 XFOIL Coefficients.
+\newblock Zenodo. Available online: \url{http://dx.doi.org/10.5281/zenodo.32054} (accessed on \hl{Day Month} 2015).
+
+\bibitem[Carmichael(1981)]{Carmichael1981}
+Carmichael, B.H.
+\newblock \emph{Low {R}eynolds Number Airfoil Survey Volume {I}}.
+\newblock Technical Report, National Aeronautics and Space Administration: \hl{City, Country},
+ 1981.
+
+\end{thebibliography}
 

### 2016-01-21 paper.tex
Add ZIP code
Estimator for predicting --> predictor

@@
  Center for Ocean Renewable Energy, University of New Hampshire, 24
-Colovos Rd., \mbox{Durham, NH \hl{zip code}, USA}}
+Colovos Rd., \mbox{Durham, NH 03824, USA}}
 
@@
  foil data and cross-flow turbine kinematics was shown to be a reasonable
- estimator for predicting the Reynolds number dependence of an actual cross-flow
+ predictor for Reynolds number dependence of an actual cross-flow
  turbine operating under dynamic conditions. Mean velocity and turbulence

### 2016-01-21 paper.tex
Use equals instead of equiv

@@
 Scaled physical models are often used in science and engineering to approximate
-real-world systems. The principle of dynamic similarity allows for geometrically-scaled systems to be dynamically identical if certain nondimensional physical
-parameters are matched. In the case of fluid systems, the most important
-nondimensional parameter is often the Reynolds number, $Re$, which quantifies
-the importance of inertial forces over viscous forces on the flow physics
-\cite{Acheson1990}: \hl {$Re \equiv Ul/\nu$, } 
-%please check the symbol ``\equiv '' and make sure it is correct
-where $U$ and $l$ are characteristic
-velocity and length scales, respectively, and $\nu$ is the fluid kinematic
-viscosity. The~other common dynamical scale, for systems with a free
-surface, is the Froude number \hl {$Fr \equiv U/\sqrt{gl}$,}
-%please check the symbol ``\equiv '' and make sure it is correct
-  where $g$ is the
-gravitational acceleration. Matching both the Froude and Reynolds number is not
-possible for a scaled model in a given fluid, which is illustrated by the
-relation $Re = l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$,
-or the geometric scale. In general, the geometric scale of a physical model is
-not necessarily the same as its dynamical scale. As such, hereafter, we will use
-the word ``scale'' to refer to this dynamical scale or the Reynolds number and
-assume the Froude number is small enough to be~ negligible.
+real-world systems. The principle of dynamic similarity allows for
+geometrically-scaled systems to be dynamically identical if certain
+nondimensional physical parameters are matched. In the case of fluid systems,
+the most important nondimensional parameter is often the Reynolds number, $Re$,
+which quantifies the importance of inertial forces over viscous forces on the
+flow physics \cite{Acheson1990}: $Re = Ul/\nu$, where $U$ and $l$ are
+characteristic velocity and length scales, respectively, and $\nu$ is the fluid
+kinematic viscosity. The~other common dynamical scale, for systems with a free
+surface, is the Froude number $Fr = U/\sqrt{gl}$, where $g$ is the gravitational
+acceleration. Matching both the Froude and Reynolds number is not possible for a
+scaled model in a given fluid, which is illustrated by the relation $Re =
+l^{3/2} g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$, or the geometric
+scale. In general, the geometric scale of a physical model is not necessarily
+the same as its dynamical scale. As such, hereafter, we will use the word
+``scale'' to refer to this dynamical scale or the Reynolds number and assume the
+Froude number is small enough to be negligible.
 
@@
 frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
-of blades, strut drag and operational parameters, such as tip speed ratio, {
- and} on the Reynolds number \cite{Para2002}. Note that an average blade chord
+of blades, strut drag and operational parameters, such as tip speed ratio---and
+on the Reynolds number \cite{Para2002}. Note that an average blade chord
 Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be
-expressed in terms of tip speed ratio \hl {$\lambda \equiv \omega R/ U_\infty$, } 
-%please check the symbol ``\equiv '' and make sure it is correct
-where
+expressed in terms of tip speed ratio $\lambda = \omega R/ U_\infty$, where
 $U_\infty$ is the free stream velocity, $c$ is the blade chord length, $\nu$ is
@@
 \cite{Templin1974}, which allows for the use of a simpler Reynolds number based
-on turbine diameter $Re_D \equiv U_\infty D/\nu$.
+on turbine diameter $Re_D = U_\infty D/\nu$.
 
@@
 
-Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of Reynolds
-number on the performance of a 2-m diameter Darrieus vertical-axis cross-flow
-wind turbine with NACA 0012 blades. By varying wind tunnel speed, the turbine
-was made to operate at an approximately constant blade chord Reynolds number $Re_c$
-ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime, the turbine
-power coefficient $C_P$ was shown to be sensitive to $Re_c$, with sensitivity
-diminishing at the higher Reynolds numbers, especially for turbines of lower
-solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
-maximum radius). More recently, Polagye and Cavagnaro \cite{Polagye2013b}
-observed significant Reynolds number sensitivity for a high solidity helical
-cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007} observed the power
-coefficient of a straight-bladed VAWT to become Reynolds number independent at
-$Re_c \approx 4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} observed
-Reynolds number independence of the power coefficient at the optimal tip speed ratio for
-a high solidity cross-flow turbine at $Re_D \approx 10^6$ or
-$Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.
+Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of
+Reynolds number on the performance of a 2-m diameter Darrieus vertical-axis
+cross-flow wind turbine with NACA 0012 blades. By varying wind tunnel speed, the
+turbine was made to operate at an approximately constant blade chord Reynolds
+number $Re_c$ ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime,
+the turbine power coefficient $C_P$ was shown to be sensitive to $Re_c$, with
+sensitivity diminishing at the higher Reynolds numbers, especially for turbines
+of lower solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the
+turbine's maximum radius). More recently, Polagye and Cavagnaro
+\cite{Polagye2013b} observed significant Reynolds number sensitivity for a high
+solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
+observed the power coefficient of a straight-bladed VAWT to become Reynolds
+number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
+\cite{Bachant2014} observed Reynolds number independence of the power
+coefficient at the optimal tip speed ratio for a high solidity cross-flow
+turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.
 
@@
     profile.
+
     \item For azimuthal angles of 0 to 180 degrees and a given tip speed ratio,
     calculate the geometric angle of attack and relative velocity magnitude from
- the blade and undisturbed free stream velocity \hl {($\equiv 1$)} vectors.
- %please check this symbol
+    the blade and undisturbed free stream velocity (taken as unity) vectors.
+
     \item With arrays of geometric angle of attack and relative velocity
@@
     coefficient from Equations~(\ref{eq:ct}) and (\ref{eq:cc}), respectively.
+
     \item Extract the maximum value of $C_T$, and repeat this process for static

### 2016-01-21 paper.tex
Add Martin's phone number
Use semicolon rather than em-dash

@@
 % there are more than one corresponding author.)
-\corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.); martin.wosnik@unh.edu (M.W.); \linebreak Tel.: \hl{(P.B.); (M.W.)}}
+\corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.);
+    martin.wosnik@unh.edu (M.W.); \linebreak Tel.: +1-603-862-1891 (M.W.)}
 
@@
 frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
-of blades, strut drag and operational parameters, such as tip speed ratio---and
+of blades, strut drag and operational parameters, such as tip speed ratio; and
 on the Reynolds number \cite{Para2002}. Note that an average blade chord

### 2016-01-21 paper.tex
RPM --> angular velocity
lbf --> pound-force

@@
   permanent magnet servo \hl {motor (Manufacturer, City, Country)} with a 20:1 gearbox,
-providing precise control over shaft \hl {RPM}. This servo was controlled by the tow
+providing precise control over shaft angular velocity. This servo was controlled by the tow
 tank's main motion controller for high synchronization with the carriage motion,
@@
 guides, such that the total streamwise drag force was transferred to a pair of
-Sentran \hl {ZB3 500-lbf} %please define/check the unit
+Sentran ZB3 500 pound-force
  capacity S-beam \hl {load cells (Manufacturer, City, Country)}, providing the rotor drag

### 2016-01-21 paper.tex
Do not put dash before unit abbreviations

@@
 Blackwell \emph{et al.} \cite{Blackwell1976} investigated the effects of
-Reynolds number on the performance of a 2-m diameter Darrieus vertical-axis
+Reynolds number on the performance of a 2 m diameter Darrieus vertical-axis
 cross-flow wind turbine with NACA 0012 blades. By varying wind tunnel speed, the
@@
 Adaramola \cite{Krogstad2012a} observed Reynolds number independence of the
-performance of a 0.9-m diameter axial-flow turbine rotor at $Re_D \approx 5
+performance of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5
 \times 10^5$ in wind tunnel tests. Their turbine blades had an S826 profile
@@
 \caption{\hl {University of New Hampshire (UNH)}-Reference Vertical Axis Turbine (RVAT) turbine model. Turbine blades and struts made from NACA 0020
- profiles with 0.140-m chord. Note that the upper and lower mounting flanges have
+ profiles with 0.140 m chord. Note that the upper and lower mounting flanges have
  been excluded, as these were included in the tare drag measurements.}
@@
 thereby giving precise measurement and control of the tip speed ratio. An Interface
-T8 200-Nm capacity rotary torque \hl {transducer (Manufacturer, City, Country)} was installed inline between the
+T8 200 Nm capacity rotary torque \hl {transducer (Manufacturer, City, Country)} was installed inline between the
 servo and the turbine, and the servo was also mounted on a slewing ring bearing,
@@
 Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler \hl {velocimeter
-(ADV)  (Manufacturer, City, Country)}, which has an approximately 6-mm diameter sampling volume and sampled at
+(ADV)  (Manufacturer, City, Country)}, which has an approximately 6 mm diameter sampling volume and sampled at
 200 Hz. The probe was mounted on an automated positioning system, also
@@
 tip speed ratio was held constant, ranging from 0.1 to 3.1 in 0.1 increments. Full
-performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2-m/s increments, for which the turbine diameter and approximate blade chord Reynolds
+performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2 m/s increments, for which the turbine diameter and approximate blade chord Reynolds
 number are presented in Table~\ref{tab:Re}. Performance was also measured for

### 2016-01-21 paper.tex
Add author contributions

@@
 
-\authorcontributions{\textbf{Author Contributions:} \hl{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.}}
+\authorcontributions{\textbf{Author Contributions:} Peter Bachant conducted the
+    experiments and analyzed the data. Both authors contributed equally to the
+    writing and editing of the manuscript.}
 

### 2016-01-21 paper.tex
Remove 'the' before prototype scale

@@
 Scaled models have the benefit of being significantly less expensive; however, a
-key question is whether or not an acceptable scale mismatch exists. Similarly,~it is questionable whether numerical models validated with physical model data
-obtained orders of magnitude away from the prototype scale should be considered
-validated at all. An example of the errors that can result from the
+key question is whether or not an acceptable scale mismatch exists.
+Similarly,~it is questionable whether numerical models validated with physical
+model data obtained orders of magnitude away from prototype scale should be
+considered validated at all. An example of the errors that can result from the
 extrapolation of small-scale experiments can be found in \cite{Baker1991}.

### 2016-01-21 paper.tex
Add comma

@@
 frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
-of blades, strut drag and operational parameters, such as tip speed ratio; and
+of blades, strut drag, operational parameters, such as tip speed ratio; and
 on the Reynolds number \cite{Para2002}. Note that an average blade chord

### 2016-01-21 paper.tex
Re-insert Oxford comma

@@
 frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
-of blades, strut drag, operational parameters, such as tip speed ratio; and
-on the Reynolds number \cite{Para2002}. Note that an average blade chord
-Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be
-expressed in terms of tip speed ratio $\lambda = \omega R/ U_\infty$, where
-$U_\infty$ is the free stream velocity, $c$ is the blade chord length, $\nu$ is
-the fluid kinematic viscosity, $\omega$ is the rotor's angular velocity and $R$
-is the rotor radius. The value of $\lambda$ at which a turbine reaches peak
-performance in general decreases with turbine solidity $Nc/(\pi D)$
-\cite{Templin1974}, which allows for the use of a simpler Reynolds number based
-on turbine diameter $Re_D = U_\infty D/\nu$.
+of blades, strut drag, operational parameters, such as tip speed ratio; and on
+the Reynolds number \cite{Para2002}. Note that an average blade chord Reynolds
+number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be expressed
+in terms of tip speed ratio $\lambda = \omega R/ U_\infty$, where $U_\infty$ is
+the free stream velocity, $c$ is the blade chord length, $\nu$ is the fluid
+kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$ is the
+rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
+in general decreases with turbine solidity $Nc/(\pi D)$ \cite{Templin1974},
+which allows for the use of a simpler Reynolds number based on turbine diameter
+$Re_D = U_\infty D/\nu$.
 

### 2016-01-21 paper.tex
The optimal --> its optimal

@@
 Reynolds numbers: $Re_D = 10^4$ and $10^3$, respectively. Tescione \emph{et al.}
-\cite{Tescione2014} studied the wake of a vertical-axis wind turbine at the optimal
+\cite{Tescione2014} studied the wake of a vertical-axis wind turbine at its optimal
 tip speed ratio in a wind tunnel using stereo PIV at an approximate blade chord

### 2016-01-21 paper.tex
Diameter based --> based on diameter

@@
 therefore important to realize the limitations of evaluating both the power
-output of turbines and the wake recovery when the Reynolds number is very far from
-full scale. Sometimes,~the models used are not turbines, but wake-generating objects,
-e.g., porous disks, that are meant to replicate the wakes of real turbines
-\cite{Goldenberg1983}. In this case, it is of interest to determine at what
-Reynolds number one might be able to realistically study wake flows in an array
-and also to evaluate the effectiveness of a wake generator. In other words, a
-wake generator may do a fine job simulating a scaled turbine, but how well can
-it simulate a full-scale device? Note that for a porous disk, the Reynolds
-number of interest is diameter based, since it is the scaling of the far-wake
-dynamics that~ matters.
+output of turbines and the wake recovery when the Reynolds number is very far
+from full scale. Sometimes,~the models used are not turbines, but
+wake-generating objects, e.g., porous disks, that are meant to replicate the
+wakes of real turbines \cite{Goldenberg1983}. In this case, it is of interest to
+determine at what Reynolds number one might be able to realistically study wake
+flows in an array and also to evaluate the effectiveness of a wake generator. In
+other words, a wake generator may do a fine job simulating a scaled turbine, but
+how well can it simulate a full-scale device? Note that for a porous disk, the
+Reynolds number of interest is based on diameter, since it is the scaling of the
+far-wake dynamics that~ matters.
 

### 2016-01-21 paper.tex
Text wrapping

@@
 we seek to replicate those same momentum and energy balance considerations at
-multiple Reynolds numbers, to examine the implications on how scaled, \emph{i.e.}, low
-Reynolds number experiments, may be used to study flows in turbine arrays.
+multiple Reynolds numbers, to examine the implications on how scaled,
+\emph{i.e.}, low Reynolds number experiments, may be used to study flows in
+turbine arrays.
 

### 2016-01-21 paper.tex
We are looking at --> it is of interest to examine
Text wrapping

@@
 
-We are looking at how Reynolds number scaling affects both the blade loading,
-\emph{i.e.}, turbine performance, and the near-wake. Typically, static airfoil data
-show that the static stall angle increases with blade chord Reynolds number
-$Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil
-behavior is presented in \cite{Lissaman1983}. In general, airfoil
+It is of interest to examine how Reynolds number scaling affects both the blade
+loading, \emph{i.e.}, turbine performance, and the near-wake. Typically, static
+airfoil data show that the static stall angle increases with blade chord
+Reynolds number $Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on
+airfoil behavior is presented in \cite{Lissaman1983}. In general, airfoil
 performance, often characterized by the profile's lift-to-drag ratio, is
@@
 
-\textls[-10]{Static foil performance does not tell the whole story for a cross-flow turbine.
-The azimuthal, and} therefore, temporal, variation of $\alpha$ in a cross-flow
-turbine implies dynamic loading, encountering \textls[-10]{dynamic stall for tip speed ratios
-near and below those of maximum rotor torque \cite{Para2002}. Bousman
-\cite{Bousman2000-evaluation}} states that dynamic stall is relatively
-insensitive to Reynolds number for $Re=1.0 \times 10^5$ to $2.5 \times
-10^5$, judging from measurements on a pitching VR-7 foil in a wind and water
-tunnel, since the loading is dominated by vortex shedding. However, Singleton
-and Yeager Jr. \cite{Singleton2000} state that the effect of Reynolds number on
-dynamic stall remains an unsolved question.
+\textls[-10]{Static foil performance does not tell the whole story for a
+    cross-flow turbine. The azimuthal, and} therefore, temporal, variation of
+$\alpha$ in a cross-flow turbine implies dynamic loading, encountering
+\textls[-10]{dynamic stall for tip speed ratios near and below those of maximum
+    rotor torque \cite{Para2002}. Bousman \cite{Bousman2000-evaluation}} states that
+dynamic stall is relatively insensitive to Reynolds number for $Re=1.0 \times
+10^5$ to $2.5 \times 10^5$, judging from measurements on a pitching VR-7 foil in
+a wind and water tunnel, since the loading is dominated by vortex shedding.
+However, Singleton and Yeager Jr. \cite{Singleton2000} state that the effect of
+Reynolds number on dynamic stall remains an unsolved question.
 

### 2016-01-21 paper.tex
Define AFT

@@
 
-Chamorro \emph{et al.} \cite{Chamorro2012} showed that turbulence statistics in an AFT%please define
-wake became $Re$-independent at $Re_D \approx 1 \times 10^5$ and that mean
-velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
-\times 10^4$. Note that in this study, the tip chord Reynolds number was not
-reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
-10^5$. It is therefore our objective to observe similar scaling relationships
-for a cross-flow turbine near-wake.
+Chamorro \emph{et al.} \cite{Chamorro2012} showed that turbulence statistics in
+an horizontal-axis or axial-flow wind turbine wake became $Re$-independent at
+$Re_D \approx 1 \times 10^5$ and that mean velocity profiles became
+$Re$-independent slightly earlier at $Re_D \approx 5 \times 10^4$. Note that in
+this study, the tip chord Reynolds number was not reported, but is estimated to
+be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times 10^5$. It is therefore our
+objective to observe similar scaling relationships for a cross-flow turbine
+near-wake.
 

### 2016-01-21 paper.tex
Define acronyms

@@
 Experiments were performed in a turbine test bed specifically designed for
-cross-flow turbines. The test bed was integrated as part of the \hl {University of New Hampshire (UNH)} 
- tow tank,
-which is a 36 m-long facility with a 3.66 m-wide and 2.44 m-deep cross-section.
-The turbine model used in this study was the UNH Reference Vertical Axis Turbine
-(RVAT), which was designed to be a generic case for numerical model testing,
-similar to the Sandia National Labs/U.S. Department of Energy RM2%please define
- River Turbine
-\cite{Neary2014}, but with a higher solidity or blade chord-to-radius ratio.
+cross-flow turbines. The test bed was integrated as part of the University of
+New Hampshire (UNH) tow tank, which is a 36 m-long facility with a 3.66 m-wide
+and 2.44 m-deep cross-section. The turbine model used in this study was the UNH
+Reference Vertical Axis Turbine (RVAT), which was designed to be a generic case
+for numerical model testing, similar to the Sandia National Labs/U.S. Department
+of Energy Reference Model 2 (RM2) River Turbine \cite{Neary2014}, but with a
+higher solidity or blade chord-to-radius ratio.
 

### 2016-01-21 paper.tex
Add note about NACA profiles
Define CFT earlier

@@
 The performance and wake characteristics of cross-flow (vertical-axis) turbines
-depend on turbine solidity, blade profile (lift/drag, dynamic stall at reduced
-frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
-of blades, strut drag, operational parameters, such as tip speed ratio; and on
-the Reynolds number \cite{Para2002}. Note that an average blade chord Reynolds
-number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be expressed
-in terms of tip speed ratio $\lambda = \omega R/ U_\infty$, where $U_\infty$ is
-the free stream velocity, $c$ is the blade chord length, $\nu$ is the fluid
-kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$ is the
-rotor radius. The value of $\lambda$ at which a turbine reaches peak performance
-in general decreases with turbine solidity $Nc/(\pi D)$ \cite{Templin1974},
-which allows for the use of a simpler Reynolds number based on turbine diameter
-$Re_D = U_\infty D/\nu$.
+(CFTs) depend on turbine solidity, blade profile (lift/drag, dynamic stall at
+reduced frequency of turbine rotation, symmetry, thickness, camber), blade
+pitch, number of blades, strut drag, operational parameters, such as tip speed
+ratio; and on the Reynolds number \cite{Para2002}. Note that an average blade
+chord Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$,
+can be expressed in terms of tip speed ratio $\lambda = \omega R/ U_\infty$,
+where $U_\infty$ is the free stream velocity, $c$ is the blade chord length,
+$\nu$ is the fluid kinematic viscosity, $\omega$ is the rotor's angular
+velocity, and $R$ is the rotor radius. The value of $\lambda$ at which a turbine
+reaches peak performance in general decreases with turbine solidity $Nc/(\pi D)$
+\cite{Templin1974}, which allows for the use of a simpler Reynolds number based
+on turbine diameter $Re_D = U_\infty D/\nu$.
 
@@
 
-The UNH-RVAT turbine has three blades made from \hl {NACA} 0020 profiles
-%please define NACA and add the information of Manufacturer, City, Country
-  with a 0.14 m
+% Note below: NACA is not a manufacturer, but a very common type of foil profile
+
+The UNH-RVAT turbine has three blades made from NACA 0020 profiles with a 0.14 m
 chord. The~blades are mounted at mid-chord w.r.t. the turbine axis, and the
@@
 Figure~\ref{fig:turbine}. The blockage ratio produced by the rotor's frontal
-area is 11\%, which we decided not to correct for, as reliable methods are
-not yet agreed upon for \hl {CFTs} %please define
-\cite{Cavagnaro2014}. As such, blockage should be
-taken into account when using these data for model validation, \emph{i.e.}, the
+area is 11\%, which we decided not to correct for, as reliable methods are not
+yet agreed upon for CFTs \cite{Cavagnaro2014}. As such, blockage should be taken
+into account when using these data for model validation, \emph{i.e.}, the
 experimental domain should be mimicked. The rotor has a relatively high solidity
-$Nc/(\pi D) = 0.13$ and a large chord-to-radius ratio $c/R = 0.28$. A CAD model of
-the turbine is available from \cite{Bachant2014-RVAT-CAD}.
+$Nc/(\pi D) = 0.13$ and a large chord-to-radius ratio $c/R = 0.28$. A CAD model
+of the turbine is available from \cite{Bachant2014-RVAT-CAD}.
 

### 2016-01-21 paper.tex
Add instrumentation mfg info

@@
 
-\caption{\hl {University of New Hampshire (UNH)}-Reference Vertical Axis Turbine (RVAT) turbine model. Turbine blades and struts made from NACA 0020
- profiles with 0.140 m chord. Note that the upper and lower mounting flanges have
- been excluded, as these were included in the tare drag measurements.}
+\caption{University of New Hampshire Reference Vertical Axis Turbine (UNH-RVAT)
+    turbine model. Turbine blades and struts made from NACA 0020 profiles with
+    0.14~m chord. Note that the upper and lower mounting flanges have been excluded,
+    as these were included in the tare drag measurements.}
 
@@
 \end{figure}
+
+% Note below: AKM is the model number, not an acronym
 
@@
 Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
-coupled to a Kollmorgen \hl {AKM}
-%please define AKM and RPM
-  permanent magnet servo \hl {motor (Manufacturer, City, Country)} with a 20:1 gearbox,
-providing precise control over shaft angular velocity. This servo was controlled by the tow
-tank's main motion controller for high synchronization with the carriage motion,
-thereby giving precise measurement and control of the tip speed ratio. An Interface
-T8 200 Nm capacity rotary torque \hl {transducer (Manufacturer, City, Country)} was installed inline between the
-servo and the turbine, and the servo was also mounted on a slewing ring bearing,
-which allowed a redundant measurement of torque via an arm and load cell used to
-counteract the turbine moment. The frame was mounted to the carriage via linear
-guides, such that the total streamwise drag force was transferred to a pair of
-Sentran ZB3 500 pound-force
- capacity S-beam \hl {load cells (Manufacturer, City, Country)}, providing the rotor drag
-measurements, after a separately measured tare drag was subtracted in post-processing. Similarly, a tare torque was measured by rotating the turbine shaft
-in air. Turbine angular and tow carriage linear position were measured using
-quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10 ${\mu}$m resolution for the carriage position. These~signals, along with
-the torque and drag signals, were sampled at 2 kHz.
+coupled to a Kollmorgen AKM permanent magnet servo motor (Kollmorgen, Radford,
+VA, USA) with a 20:1 gearbox, providing precise control over shaft angular
+velocity. This servo was controlled by the tow tank's main motion controller for
+high synchronization with the carriage motion, thereby giving precise
+measurement and control of the tip speed ratio. An Interface T8 200 Nm capacity
+rotary torque transducer (Interface, Scottsdale, AZ, USA) was installed inline
+between the servo and the turbine, and the servo was also mounted on a slewing
+ring bearing, which allowed a redundant measurement of torque via an arm and
+load cell used to counteract the turbine moment. The frame was mounted to the
+carriage via linear guides, such that the total streamwise drag force was
+transferred to a pair of Sentran ZB3 500 pound-force capacity S-beam load cells
+(Sentran, Santa Ana, CA, USA), providing the rotor drag measurements, after a
+separately measured tare drag was subtracted in post-processing. Similarly, a
+tare torque was measured by rotating the turbine shaft in air. Turbine angular
+and tow carriage linear position were measured using quadrature encoder signals,
+with $10^5$ counts-per-rev for the turbine and 10 ${\mu}$m resolution for the
+carriage position. These~signals, along with the torque and drag signals, were
+sampled at 2 kHz.
 

### 2016-01-21 paper.tex
Add Nortek info

@@
 
-Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler \hl {velocimeter
-(ADV)  (Manufacturer, City, Country)}, which has an approximately 6 mm diameter sampling volume and sampled at
+Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
+(ADV) (Nortek AS, Rud, Norway)}, which has an approximately 6~mm diameter sampling volume and sampled at
 200 Hz. The probe was mounted on an automated positioning system, also

### 2016-01-21 paper.tex
Spell out data acquisition

@@
 Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
-(ADV) (Nortek AS, Rud, Norway)}, which has an approximately 6~mm diameter sampling volume and sampled at
-200 Hz. The probe was mounted on an automated positioning system, also
-controlled by the tow tank's main motion controller. The ADV and \hl {DAQ (please define)} systems'
-sampling times were synchronized by triggering the start of data acquisition via
-a pulse sent from the motion controller. Additional details of the turbine and
-experimental setup are described in \cite{Bachant2015-JoT}.
+(ADV) (Nortek AS, Rud, Norway), which has an approximately 6~mm diameter
+sampling volume and sampled at 200 Hz. The probe was mounted on an automated
+positioning system, also controlled by the tow tank's main motion controller.
+The ADV and data acquisition systems' sampling times were synchronized by
+triggering the start of data acquisition via a pulse sent from the motion
+controller. Additional details of the turbine and experimental setup are
+described in \cite{Bachant2015-JoT}.
 

### 2016-01-21 paper.tex
Put a and b labels on exp setup figure
Clean up ADV positioning paragraph

@@
 \usepackage{microtype}
+\usepackage{caption}
+\usepackage{subcaption}
 \def \p{\partial}
@@
     \centering
-
+    \begin{subfigure}[t]{\textwidth}
+        \centering
         \includegraphics[width=0.7\textwidth]{figures/exp-setup-photo}
-
+        \caption{}
+        \label{fig:exp-setup-photo}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
         \includegraphics[width=0.7\textwidth]{figures/exp_setup_drawing}
-
- \caption{\hl {Experimental} setup: turbine test bed installed in the UNH tow tank.}
- %please mark (a) and (b) in the figure and label (a) and (b) in the figure caption.
+        \caption{}
+        \label{fig:exp-setup-dwg}
+    \end{subfigure}
+
+    \caption{Experimental setup photo (a) and drawing (b): turbine test bed
+        installed in the UNH tow tank.}
 
@@
 performance curve consisted of 31~tows, where during each tow, the mean turbine
-tip speed ratio was held constant, ranging from 0.1 to 3.1 in 0.1 increments. Full
-performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2 m/s increments, for which the turbine diameter and approximate blade chord Reynolds
-number are presented in Table~\ref{tab:Re}. Performance was also measured for
-$\lambda=1.9$ at tow speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows
-each.
-
-Each wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
-Velocimeter (ADV), measuring three components of velocity at a 200-Hz sampling
-frequency, at 270 different locations, varied in the cross-stream and vertical
-directions at one turbine diameter downstream ($x/D=1$). These locations have
-vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
-the cross-stream direction $y/R = \pm 3$. These locations are shown in
-Figure~\ref{fig:wake-locations}.
+tip speed ratio was held constant, ranging from 0.1 to 3.1 in 0.1 increments.
+Full performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in
+0.2 m/s increments, for which the turbine diameter and approximate blade chord
+Reynolds number are presented in Table~\ref{tab:Re}. Performance was also
+measured for $\lambda=1.9$ at tow speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s
+for two tows each.
+
+Each wake map was generated by positioning the ADV, which measured three
+components of velocity at a 200 Hz sampling frequency, at 270 different
+locations, varied in the cross-stream and vertical directions at one turbine
+diameter downstream ($x/D=1$). These locations had vertical coordinates from the
+turbine centerline up to $z/H=0.625$, ranging in the cross-stream direction $y/R
+= \pm 3$. These locations are shown in Figure~\ref{fig:wake-locations}.
 

### 2016-01-21 scripts/proof-diff.sh
Add script to create proof diff

@@ -1,0 +1,6 @@
+#!/usr/bin/env bash
+# This script creates a diff with colored words in HTML. It requires
+# http://www.pixelbeat.org/scripts/ansi2html.sh
+
+git show 71bd282ca2 --word-diff --color > archive/proof-diff.diff
+cat archive/proof-diff.diff | ansi2html.sh > archive/proof-diff.html

### 2016-01-21 getfigs.py
Use local figs dir

@@ -7,8 +7,7 @@ import os
 import sys
 import shutil
 
-figdir = os.path.join(os.path.expanduser("~"), "Research",
-                      "Experiments", "RVAT Re dep", "Figures")
+figdir = os.path.join("RVAT-Re-dep", "Figures")
 
 figlist = ["cp_curves.pdf",
            "cd_curves.pdf",
@@ -17,8 +16,6 @@ figlist = ["cp_curves.pdf",
            "mom_bar_graph.pdf",
            "wake_spectra.pdf",
            "wake_trans_totals.pdf",
-           "meancontquiv_10.pdf",
-           "k_contours_10.pdf",
            "mean_u_k_profiles.pdf"]
 
 for fig in figlist:

### 2016-01-21 paper.tex
Remove empty line
Full --> complete
Use color on performance curve figures

@@
 
-Full-power and rotor drag (also known as thrust) coefficient curves for various
-Reynolds numbers are plotted in Figures~\ref{fig:cp-curves} and
-\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
-with Reynolds number. The power coefficient curves also show a slight downward
-shift in the optimal tip speed ratio (peak~performance) with increasing Reynolds
+Complete power and rotor drag (also known as thrust) coefficient curves for
+various Reynolds numbers are plotted in Figures~\ref{fig:cp-curves} and
+\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases with
+Reynolds number. The power coefficient curves also show a slight downward shift
+in the optimal tip speed ratio (peak~performance) with increasing Reynolds
 number, from about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall
@@
 
-\caption{\hl {Mean} power coefficient curves plotted for multiple Reynolds numbers.}
-%is it better to mark different color to the lines?
+    \caption{Mean power coefficient curves plotted for multiple Reynolds
+        numbers.}
+
     \label{fig:cp-curves}
@@
 
-\caption{\hl {Mean} rotor drag coefficient curves plotted for multiple Reynolds
+    \caption{Mean rotor drag coefficient curves plotted for multiple Reynolds
         numbers.}
- % is it better to mark different color to the lines?
 

### 2016-01-21 getfigs.py
meancontquiv and kcont figures should be copied

@@ -16,6 +16,8 @@ figlist = ["cp_curves.pdf",
            "mom_bar_graph.pdf",
            "wake_spectra.pdf",
            "wake_trans_totals.pdf",
+           "meancontquiv_10.pdf",
+           "k_contours_10.pdf",
            "mean_u_k_profiles.pdf"]
 
 for fig in figlist:

### 2016-01-21 scripts/makefigs.sh
Add script for making figures

@@ -1,0 +1,6 @@
+#!/usr/bin/env bash
+
+cd RVAT-Re-dep
+python plot.py --save --noshow "$@"
+cd ..
+python getfigs.py

### 2016-01-21 paper.tex
Clarify Re dependence
Mark (a) and (b) in perf-re-dep figure

@@
 
-Mean power and drag coefficients at $\lambda=1.9$ are plotted \textit{versus} the Reynolds
-number in Figure~\ref{fig:perf-Re-dep}. Note that the large error bars for $C_P$
-at low $Re$ are dominated by systematic error estimates for the torque
-transducer, since torque values are at the lower end of its measurement range.
-However, the uncertainty due to random error or repeatability remains relatively
-low. There is a drastic improvement in $C_P$ with increasing Reynolds number at
-lower values of the Reynolds number. The power coefficient then becomes essentially
-$Re$-independent at $Re_D = 0.8 \times 10^6$, which corresponds to an
-approximate average blade chord Reynolds number $Re_{c, \mathrm{ave}} = 2.1
+Mean power and drag coefficients at $\lambda=1.9$ are plotted \textit{versus}
+the Reynolds number in Figure~\ref{fig:perf-Re-dep}. Note that the large error
+bars for $C_P$ at low $Re$ are dominated by systematic error estimates for the
+torque transducer, since torque values are at the lower end of its measurement
+range. However, the uncertainty due to random error or repeatability remains
+relatively low. There is a drastic improvement in $C_P$ with increasing Reynolds
+number at the lower end of the $Re$ range. The power coefficient then becomes
+essentially $Re$-independent at $Re_D = 0.8 \times 10^6$, which corresponds to
+an approximate average blade chord Reynolds number $Re_{c, \mathrm{ave}} = 2.1
 \times 10^5$. This threshold is consistent with the behavior of the blade
@@
 
-\caption{UNH-RVAT measured mean power (\hl {left}) and drag (right) coefficients at
- $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error bars indicate expanded
- uncertainty estimates for 95\% confidence, which for $C_P$ is dominated by
- systematic error estimates from the torque transducer operating at the lower end
- of its measurement range.}
-
-%please use (a) and (b) to instead left and right and mark (a) and (b) in the figure.
+\caption{UNH-RVAT measured mean power \textbf{(a)} and drag \textbf{(b)}
+    coefficients at $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error
+    bars indicate expanded uncertainty estimates for 95\% confidence, which for
+    $C_P$ is dominated by systematic error estimates from the torque transducer
+    operating at the lower end of its measurement range.}
 

### 2016-01-21 paper.tex
Text wrapping
Whitespace
Fix up first paragraph of XFOIL section

@@
 where increases in blade drag at lower $Re$ somewhat offset the reduction in
-lift, causing the total streamwise force to vary less than the rotor torque. The~tendency for $C_D$ to continue increasing with $Re$ may also include the effects
-of an increasing Froude number (from faster tow speeds), which increases free
-surface deformation and wave drag during towing without increasing flow through
-the turbine.
+lift, causing the total streamwise force to vary less than the rotor torque.
+The~tendency for $C_D$ to continue increasing with $Re$ may also include the
+effects of an increasing Froude number (from faster tow speeds), which increases
+free surface deformation and wave drag during towing without increasing flow
+through the turbine.
 
 \subsubsection{Relation to Static Foil Characteristics}
+
+% \texttt is used below since this is a software variable
 
@@
 performance, a series of static foil coefficient datasets were computed with the
-\hl {XFOIL (please define)} viscous panel method code \cite{Drela1989}, a commonly-used tool for
+viscous panel method code XFOIL \cite{Drela1989}, a commonly-used tool for
 airfoil analysis, e.g., \cite{Castelli2011, Walker2014}, implemented as part of
 the open source turbine design software QBlade \cite{Marten2013}. Simulations
-were run for an angle of attack range of 0$^{\circ}$
- to 40$^{\circ}$, in increments of 0.5$^{\circ}$. Solver parameters used were 100 panels, a fixed speed, zero Mach
-number, \texttt{NCrit = 9} %please check the use of \texttt here
- (default $\mathrm{e}^n$ transition criteria parameter
-for an average wind tunnel) and no forced boundary layer transition.
-Characteristics were computed for the approximate average blade chord Reynolds
-numbers encountered in the tow tank experiments.
+were run for an angle of attack range of 0$^{\circ}$ to 40$^{\circ}$, in
+increments of 0.5$^{\circ}$. Solver parameters used were 100 panels, a fixed
+speed, zero Mach number, \texttt{NCrit = 9} (default $\mathrm{e}^n$ transition
+criteria parameter for an average wind tunnel) and no forced boundary layer
+transition. Characteristics were computed for the approximate average blade
+chord Reynolds numbers encountered in the tow tank experiments.
 

### 2016-01-21 paper.tex
Re-insert Oxford comma

@@
 circular path \cite{Migliore1980}, the XFOIL calculations were performed for
-20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520) and 4\% (NACA 4520)
+20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520), and 4\% (NACA 4520)
 camber about their half-chord location. A 4\% camber approximates the maximum

### 2016-01-21 paper.tex
Clarify why foil characteristics were normalized

@@
 Figure~\ref{fig:foil-Re-dep}, where values of the maximum lift coefficient
-$C_{l_{\max}}$, minimum drag coefficient $C_{d_{\min}}$ and maximum
-lift-to-drag ratio are normalized (for the comparison between foils) and plotted
-\textit{versus} $Re_c$. In general, larger camber is associated with decreased foil
-performance at lower Reynolds number. The data and processing code for these
-calculations is available from \cite{Bachant2015-NACAXX20-XFOIL}.
+$C_{l_{\max}}$, minimum drag coefficient $C_{d_{\min}}$, and maximum
+lift-to-drag ratio are normalized (to visualize relative differences in scaling
+between the foils) and plotted \textit{versus} $Re_c$. In general, larger camber
+is associated with decreased foil performance at lower Reynolds number. The data
+and processing code for these calculations is available from
+\cite{Bachant2015-NACAXX20-XFOIL}.
 

### 2016-01-21 getfigs.py
Use submodule for XFOIL figures

@@ -24,8 +24,7 @@ for fig in figlist:
     shutil.copy2(os.path.join(figdir, fig), os.path.join("figures", fig))
 
 
-figdir = os.path.join(os.path.expanduser("~"), "Google Drive", "Research",
-                      "Foils", "Data", "NACAXX20_QBlade", "figures")
+figdir = os.path.join("NACAXX20-XFOIL", "figures")
 
 figlist = ["all_foils_re_dep.pdf",
            "cft_re_dep_foils.pdf",

### 2016-01-22 paper.tex
Clean up foil characteristics fig caption
Clean up paragraph about induction

@@
     \centering
+
     \includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
-\caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
+
+    \caption{Normalized maximum lift coefficient \textbf{(a)}, drag coefficient
+        \textbf{(b)}, and lift-to-drag ratio \textbf{(c)} computed by XFOIL at
+        various $Re_c$  for each profile.}
+
     \label{fig:foil-Re-dep}
@@
 of the free stream velocity and the opposite of the blade tangential velocity.
-Note that this neglects any induction or slowing of the free stream by the
+Note that this neglects any induction, i.e., slowing of the free stream by the
 turbine forces, which would be present in a momentum/streamtube model. Since~the
-goal of this approach was not to predict absolute performance, but rather gain
-insight into relative changes with $Re$, this method was deemed acceptable, as it
-is extremely simple and fast to~compute.
+goal of this approach was not to predict absolute performance, but rather to
+gain insight into relative changes with $Re$, this method was deemed acceptable,
+as it is extremely simple and fast to~compute.
 

### 2016-01-22 paper.tex
Angle of stall --> static stall angle

@@
 
-Values for the blade angle of attack, relative velocity and torque coefficient
+Values for the blade angle of attack, relative velocity, and torque coefficient
 are plotted in Figure~\ref{fig:blade-kinematics} for the upstream half of
@@
 torque coefficient plot, and by the time the angle of attack has decreased below
-that of the stall, the relative velocity is so low that there is not much
+the static stall angle, the relative velocity is so low that there is not much
 contribution to the torque coefficient.

### 2016-01-22 paper.tex
Clean up blade kinematics figure caption

@@
 
-\caption{\hl {Geometric} angle of attack, relative velocity and torque coefficient
- calculated with a NACA 0020 foil operating at $\lambda=1.9$.}
-%please mark (a) and (b), (c) in the figure and label them in the caption
+    \caption{Geometric angle of attack \textbf{(a)}, relative velocity
+        \textbf{(b)},  and torque coefficient \textbf{(c)} calculated with a NACA
+        0020 foil operating at $\lambda=1.9$.}
+
     \label{fig:blade-kinematics}

### 2016-01-22 paper.tex
Text wrapping
Text wrapping; can't fix exponential labels

@@
 Results for the normalized peak torque coefficient for each foil are plotted
-\textit{versus} $Re_c$ in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the
-convergence of $C_{T_\mathrm{max}}$ with increasing Reynolds number is more
-dramatic than any of the common foil performance characteristics plotted in
+\textit{versus} $Re_c$ in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting
+that the convergence of $C_{T_\mathrm{max}}$ with increasing Reynolds number is
+more dramatic than any of the common foil performance characteristics plotted in
 Figure~\ref{fig:foil-Re-dep}, meaning that the cross-flow turbine's unique
@@
 regime for cross-flow turbine performance using only 2D static airfoil
-characteristics, \emph{i.e.}, the Reynolds number scaling of the peak $C_T$ computed this
-way behaves much like that of the measured turbine performance.
+characteristics, \emph{i.e.}, the Reynolds number scaling of the peak $C_T$
+computed this way behaves much like that of the measured turbine performance.
 
@@
 
-\caption{\hl {Reynolds} number dependence of the normalized peak torque coefficient
- calculated from static foil coefficients and blade kinematics, compared to
- experimental data from a cross-flow turbine. Note that the experimental data
- represents the mean torque coefficient, not the maximum.}
- %please pay attention to the value ``1e5'' in the figure.
-is it ``\times 10^5''.
-If yes, please revise it to \times 10 ^5 in the figure.
-meanwhile, is it better to make different color with the lines?
+    \caption{Reynolds number dependence of the normalized peak torque
+        coefficient calculated from static foil coefficients and blade kinematics,
+        compared to experimental data from a cross-flow turbine. Note that the
+        experimental data represents the mean torque coefficient, not the maximum.}
 

### 2016-01-22 paper.tex
Clarity wrt JoT paper
Re-insert Oxford comma

@@
 
-The near-wake of this turbine at $x/D=1$, $\lambda=1.9$ and momentum and
+The near-wake of this turbine at $x/D=1$, $\lambda=1.9$, along with momentum and
 kinetic energy balances at a $Re_D = 1.0 \times 10^6$ were discussed in
@@
 These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times
-10^6$, $0.8 \times 10^6$ and $1.2 \times 10^6$. Qualitatively, these look very
+10^6$, $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look very
 similar, so they have not been plotted here, though profiles at $z/H=0.0$ are

### 2016-01-22 paper.tex
Re-insert Oxford commas
Fix up wake profiles figure caption

@@
 
- \caption{Mean velocity field at $x/D=1$, $\lambda=1.9$ and $Re_D=1.0 \times
+ \caption{Mean velocity field at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0 \times
  10^6$.}
@@
 
- \caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$ and $Re_D=1.0
+ \caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0
  \times 10^6$.}
@@
 
- \caption{Mean streamwise velocity \hl {(left)} and turbulence kinetic energy
- (right) profiles at $z/H=0.0$. The legend indicates $Re_D$.}
- %please use (a) and (b) to instead left and right and mark (a) and (b) in the figure.
+     \caption{Mean streamwise velocity \textbf{(a)} and turbulence kinetic
+         energy \textbf{(b)} profiles at $z/H=0.0$. Turbine diameter Reynolds number
+         $Re_D$ is indicated by the legend.}
+
     \label{fig:profiles}

### 2016-01-22 paper.tex
Text wrapping
Whitespace
Define abbreviations in K equation
Bring back comma

@@
 
-Spectra of the cross-stream velocity normalized by the free stream were
-computed using a fast Fourier transform-based method, applying a Hanning window
-and averaging over four adjacent frequency bands to decrease confidence
-intervals. These spectra are plotted in Figure~\ref{fig:wake-spectra} for
-regions on either side of the turbine. On the $-y$ side of the turbine, there is
-broadband turbulence produced by blade stall, and on the $+y$ side, there is a
-clear peak in the spectra caused by the blade passage. We see that on both sides,
-there is higher spectral energy at lower Reynolds numbers. On the $+y$ side of
-the turbine, we notice higher spectral energy at the blade passage frequency's
-first harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's
-shed vorticity being less stable at higher $Re$.
+Spectra of the cross-stream velocity normalized by the free stream were computed
+using a fast Fourier transform-based method, applying a Hanning window and
+averaging over four adjacent frequency bands to decrease confidence intervals.
+These spectra are plotted in Figure~\ref{fig:wake-spectra} for regions on either
+side of the turbine. On the $-y$ side of the turbine, there is broadband
+turbulence produced by blade stall, and on the $+y$ side, there is a clear peak
+in the spectra caused by the blade passage. We see that on both sides, there is
+higher spectral energy at lower Reynolds numbers. On the $+y$ side of the
+turbine, we notice higher spectral energy at the blade passage frequency's first
+harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's shed
+vorticity being less stable at higher $Re$.
 
@@
 
-\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$,
- $y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (\textbf{a}) and $y/R=1.5$
- ($+0.75$ m in Figure~\ref{fig:wake-locations}) (\textbf{b}). Dashed vertical lines
- indicate $[1, 3, 6, 9]$ times the turbine rotational frequency. Shaded regions
- indicate 95\% confidence intervals assuming a $\chi^2$~distribution.}
+    \caption{Cross-stream velocity (normalized by $U_\infty$) spectra at
+        $z/H=0.25$, $y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations})
+        (\textbf{a}) and $y/R=1.5$ ($+0.75$ m in Figure~\ref{fig:wake-locations})
+        (\textbf{b}). Dashed vertical lines indicate $[1, 3, 6, 9]$ times the
+        turbine rotational frequency. Shaded regions indicate 95\% confidence
+        intervals assuming a $\chi^2$~distribution.}
 
@@
 interest~\cite{Bachant2015-JoT}, \emph{i.e.}, momentum:
-
 \begin{equation}
@@
 		}}
-		\bigg{]}.
+		\bigg{]},
 	\label{eq:K-full}
@@
 \end{equation}
-
-Note that Equations~(\ref{eq:momentum}) and (\ref{eq:K-full}) only differ from the
-typical forms of the mean momentum and mean kinetic energy equations by a factor
-of $1/U$, but importantly, the other convective terms ( $\partial U/\partial y$,
-$\partial U/\partial z$ and $\partial K/\partial y$, $\partial K/\partial z$ )
-are moved to the right-hand sides of the equations to illustrate with what sign
-they add to the streamwise recovery of streamwise momentum or mean kinetic
-energy.
+where advection is abbreviated by ``adv.,'' turbulent transport by ``turb.
+trans.,'' production by ``prod.,'' and dissipation by ``diss.''
+
+Note that Equations~(\ref{eq:momentum}) and (\ref{eq:K-full}) only differ from
+the typical forms of the mean momentum and mean kinetic energy equations by a
+factor of $1/U$, but importantly, the other convective terms ( $\partial
+U/\partial y$, $\partial U/\partial z$, and $\partial K/\partial y$, $\partial
+K/\partial z$ ) are moved to the right-hand sides of the equations to illustrate
+with what sign they add to the streamwise recovery of streamwise momentum or
+mean kinetic energy.
 

### 2016-01-22 paper.tex
i.e. should use emph

@@
 of the free stream velocity and the opposite of the blade tangential velocity.
-Note that this neglects any induction, i.e., slowing of the free stream by the
-turbine forces, which would be present in a momentum/streamtube model. Since~the
-goal of this approach was not to predict absolute performance, but rather to
-gain insight into relative changes with $Re$, this method was deemed acceptable,
-as it is extremely simple and fast to~compute.
+Note that this neglects any induction, \emph{i.e.}, slowing of the free stream
+by the turbine forces, which would be present in a momentum/streamtube model.
+Since~the goal of this approach was not to predict absolute performance, but
+rather to gain insight into relative changes with $Re$, this method was deemed
+acceptable, as it is extremely simple and fast to~compute.
 

### 2016-01-22 paper.tex
Text wrapping
Text wrapping and whitespace
Minor wording in first paragraph of conclusion

@@
 
-Results from the terms that can be computed from the experimental data (\emph{i.e.},
-excluding $\partial / \partial x$) from the right-hand sides of
+Results from the terms that can be computed from the experimental data
+(\emph{i.e.}, excluding $\partial / \partial x$) from the right-hand sides of
 Equations~(\ref{eq:momentum}) and (\ref{eq:K-full}) are plotted in
@@
 Reynolds numbers, which is consistent with the physical meaning of the Reynolds
-number itself. From these results, one might expect that the viscous effects will
-become significant to the wake dynamics as the Reynolds number is decreased below
-$Re_D \sim 10^4$.
+number itself. From these results, one might expect that the viscous effects
+will become significant to the wake dynamics as the Reynolds number is decreased
+below $Re_D \sim 10^4$.
 
@@
 Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph} are plotted in
-Figure~\ref{fig:wake-trans-totals}. We see that in general, the wake transport of
-both mean streamwise momentum and kinetic energy is enhanced at lower Reynolds
-numbers and levels off consistent with the behavior of the turbine power
-coefficient. This is an important consideration if studying sub-scale models of
-turbine arrays, where increased levels of wake recovery could motivate different
-ideal array configurations when compared to full-scale turbines if the scale
-model Reynolds number is too low.
-
-\begin{figure}[H]
-\centering
+Figure~\ref{fig:wake-trans-totals}. We see that in general, the wake transport
+of both mean streamwise momentum and kinetic energy is enhanced at lower
+Reynolds numbers and levels off consistent with the behavior of the turbine
+power coefficient. This is an important consideration if studying sub-scale
+models of turbine arrays, where increased levels of wake recovery could motivate
+different ideal array configurations when compared to full-scale turbines if the
+scale model Reynolds number is too low.
+
+\begin{figure}[H]
+    \centering
+
     \includegraphics[width=0.65\textwidth]{figures/wake_trans_totals}
-\caption{Normalized transport totals from Figures~\ref{fig:mom-bar-graph} and
-\ref{fig:K-bar-graph} plotted \textit{versus} the Reynolds number.}
+
+    \caption{Normalized transport totals from Figures~\ref{fig:mom-bar-graph}
+        and \ref{fig:K-bar-graph} plotted \textit{versus} the Reynolds number.}
+
     \label{fig:wake-trans-totals}
@@
 ($c/R=0.28$) cross-flow turbine becomes essentially $Re$-independent at a
-Reynolds number based on a turbine diameter of $Re_D \approx 10^6$ or an
-approximate Reynolds number based on a blade chord of $Re_c \approx 2 \times
-10^5$. The power coefficient values on the left-hand side of the
-$C_P$-$\lambda$ curve are reduced more than those on the right-hand side at low
-$Re$ due to deeper stalling. The $Re$-independent threshold corresponds to that
-at which the foil suction surface boundary layer becomes turbulent before having
-to recover static pressure against the adverse pressure gradient, highlighted in
+Reynolds number based on the turbine diameter $Re_D \approx 10^6$ or an
+approximate Reynolds number based on the blade chord $Re_c \approx 2 \times
+10^5$. The power coefficient values on the left-hand side of the $C_P$-$\lambda$
+curve are reduced more than those on the right-hand side at low $Re$ due to
+deeper stalling. The $Re$-independent threshold corresponds to that at which the
+foil suction surface boundary layer becomes turbulent before having to recover
+static pressure against the adverse pressure gradient, highlighted in
 \cite{Lissaman1983, McMasters1980, Carmichael1981}.

### 2016-01-22 paper.tex
Clarify wording

@@
 It was shown that the peak torque coefficient computed this way shows similar
-Reynolds number sensitivity, a better predictor than conventional
-quantifications of airfoil performance, e.g., the lift-to-drag ratio. Note that this
-method is not restricted to standard symmetrical NACA foils, though those
-evaluated here with higher camber were more sensitive at lower $Re$, but had
-smaller slopes in the linear~regime.
+Reynolds number sensitivity as the experimental results for an actual turbine,
+making it a better predictor than conventional quantifications of airfoil
+performance, e.g., the lift-to-drag ratio. Note that this method is not
+restricted to standard symmetrical NACA foils, though those evaluated here with
+higher camber were more sensitive at lower $Re$, but had smaller slopes in the
+linear~regime.
 

### 2016-01-22 paper.tex
Text wrapping and space between citation

@@
 
-From these results, we recommend that physical model tests of cross-flow turbines
-be performed at $Re_D \sim 10^6$ or greater to provide reasonable predictions of
-full-scale performance. This threshold also applies to the validation of
-predictive engineering models, especially for high-fidelity CFD models where the
-boundary layer is to be resolved, since transition to turbulence plays an
-important role in overall blade loading. Note that in our case, blockage, though
-reasonably low, may be increasing flow through the turbine compared to a free
-case, which could mean $Re$ thresholds for free flows could be slightly higher,
-though a reliable blockage correction algorithm has not yet been agreed upon for
-CFTs~ \cite{Cavagnaro2014}.
+From these results, we recommend that physical model tests of cross-flow
+turbines be performed at $Re_D \sim 10^6$ or greater to provide reasonable
+predictions of full-scale performance. This threshold also applies to the
+validation of predictive engineering models, especially for high-fidelity CFD
+models where the boundary layer is to be resolved, since transition to
+turbulence plays an important role in overall blade loading. Note that in our
+case, blockage, though reasonably low, may be increasing flow through the
+turbine compared to a free case, which could mean $Re$ thresholds for free flows
+could be slightly higher, though a reliable blockage correction algorithm has
+not yet been agreed upon for CFTs~\cite{Cavagnaro2014}.
 

### 2016-01-22 paper.tex
Remove highlight

@@
 model studies of turbine arrays may see exaggerated levels of wake recovery,
-leading to inadequate or inappropriate spacing or \hl {layouts}.
-%please define CAREER, PI, NSF-CBET
+leading to inadequate or inappropriate spacing or layouts.
 

### 2016-01-22 paper.tex
Clean up acknowledgements

@@
 
-\vspace{6pt}
-\acknowledgments{\textbf{Acknowledgments:} The authors would like to acknowledge funding through a National Science
-Foundation CAREER %please define
- award (PI %please define
-  Wosnik, NSF-CBET%please define
- 1150797, program manager 
-Gregory L. Rorrer), a grant through the \linebreak Leslie S. Hubbard Marine Program
-Endowment to purchase acoustic flow measurement instrumentation and a grant for
-laboratory infrastructure upgrades through the U.S. Department of Energy.}
+% Note for acknowledgements: CAREER is a type of award from the National Science
+% Foundation, and is not an acronym.
+
+\vspace{6pt} \acknowledgments{\textbf{Acknowledgments:} The authors would like
+    to acknowledge funding through a National Science Foundation CAREER award
+    (principal investigator Martin Wosnik, NSF 1150797, Energy for Sustainability,
+    program manager Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine
+    Program Endowment to purchase acoustic flow measurement instrumentation, and a
+    grant for laboratory infrastructure upgrades through the U.S. Department of
+    Energy.}
 

### 2016-01-22 paper.tex
Text wrapping
Add address for Carmichael1981
Add accessed date for XFOIL coeffs

@@
 
-\conflictofinterests{\textbf{Conflicts of Interest:} The authors declare no conflict of interest.}
+\conflictofinterests{\textbf{Conflicts of Interest:} The authors declare no
+    conflict of interest.}
 
@@
 \bibitem[Marten \em{et~al.}(2013)Marten, Wendler, Pechlivanoglou, Nayeri, and
- Paschereit]{Marten2013}
-Marten, D.; Wendler, J.; Pechlivanoglou, G.; Nayeri, C.; Paschereit, C.
-\newblock {QB}lade: An open source tool for design and simulation of horizontal
- and vertical axis wind turbines.
-\newblock {\em Int. J. Emerg. Technol. Adv.
- Eng.} {\bf 2013}, {\em 3},~264--269.
-
-\bibitem[Bachant and Wosnik(2015)]{Bachant2015-NACAXX20-XFOIL}
-Bachant, P.; Wosnik, M.
-\newblock NACA XX20 XFOIL Coefficients.
-\newblock Zenodo. Available online: \url{http://dx.doi.org/10.5281/zenodo.32054} (accessed on \hl{Day Month} 2015).
-
-\bibitem[Carmichael(1981)]{Carmichael1981}
-Carmichael, B.H.
-\newblock \emph{Low {R}eynolds Number Airfoil Survey Volume {I}}.
-\newblock Technical Report, National Aeronautics and Space Administration: \hl{City, Country},
- 1981.
+Paschereit]{Marten2013} Marten, D.; Wendler, J.; Pechlivanoglou, G.; Nayeri, C.;
+Paschereit, C. \newblock {QB}lade: An open source tool for design and simulation
+of horizontal and vertical axis wind turbines. \newblock {\em Int. J. Emerg.
+    Technol. Adv. Eng.} {\bf 2013}, {\em 3},~264--269.
+
+\bibitem[Bachant and Wosnik(2015)]{Bachant2015-NACAXX20-XFOIL} Bachant, P.;
+Wosnik, M. \newblock NACA XX20 XFOIL Coefficients. \newblock Zenodo. Available
+online: \url{http://dx.doi.org/10.5281/zenodo.32054} (accessed on 1 November
+2015).
+
+\bibitem[Carmichael(1981)]{Carmichael1981} Carmichael, B.H. \newblock \emph{Low
+    {R}eynolds Number Airfoil Survey Volume {I}}. \newblock Technical Report,
+National Aeronautics and Space Administration: Capistrano Beach, California,
+USA, 1981.
 

### 2016-01-22 paper.tex
Update XFOIL repo information

@@
 
-\bibitem[Bachant and Wosnik(2015)]{Bachant2015-NACAXX20-XFOIL} Bachant, P.;
+\bibitem[Bachant and Wosnik(2016)]{Bachant2015-NACAXX20-XFOIL} Bachant, P.;
 Wosnik, M. \newblock NACA XX20 XFOIL Coefficients. \newblock Zenodo. Available
-online: \url{http://dx.doi.org/10.5281/zenodo.32054} (accessed on 1 November
-2015).
+online: \url{http://dx.doi.org/10.5281/zenodo.45099} (accessed on 22 January
+2016).
 

### 2016-01-22 paper.tex
Update figshare access dates

@@
 
-\bibitem[Bachant and Wosnik(2014)]{Bachant2014-RVAT-CAD}
-Bachant, P.; Wosnik, M.
-\newblock {UNH-RVAT CAD Models}.
-\newblock \hl {Fig\textbf{share}}. 
-%Is it necessary to make the type of share bold?
-Available online: 
- \url{http://dx.doi.org/10.6084/m9.figshare.1062009} (accessed on \hl{Day Month} 2014).
-
-\bibitem[Bachant and Wosnik(2015)]{Bachant2015-RVAT-Re-dep-data}
-Bachant, P.; Wosnik, M.
-\newblock {UNH-RVAT} {R}eynolds Number Dependence Experiment: Reduced Dataset
- and Processing Code.
-\newblock Fig\textbf{share}. Available online: 
- \url{http://dx.doi.org/10.6084/m9.figshare.1286960} (accessed on \hl{Day Month} 2015).
-
-\bibitem[Coleman and Steele(2009)]{ColemanSteele}
-Coleman, H.W.; Steele, W.G.
+\bibitem[Bachant and Wosnik(2014)]{Bachant2014-RVAT-CAD} Bachant, P.; Wosnik, M.
+\newblock {UNH-RVAT CAD Models}. \newblock Fig\textbf{share}. Available online:
+\url{http://dx.doi.org/10.6084/m9.figshare.1062009} (accessed on 20 June 2014).
+
+\bibitem[Bachant and Wosnik(2016)]{Bachant2015-RVAT-Re-dep-data} Bachant, P.;
+Wosnik, M. \newblock {UNH-RVAT} {R}eynolds Number Dependence Experiment: Reduced
+Dataset and Processing Code. \newblock Fig\textbf{share}. Available online:
+\url{http://dx.doi.org/10.6084/m9.figshare.1286960} (accessed on 22 January
+2015).
+
+\bibitem[Coleman and Steele(2009)]{ColemanSteele} Coleman, H.W.; Steele, W.G.
 \newblock {\em Experimentation, Validation, and Uncertainty Analysis for
- Engineers}, 3rd ed.; John~ Wiley \& Sons: \hl{ Hoboken, NJ, USA,} 2009.
-
-\bibitem[Drela(1989)]{Drela1989}
-Drela, M.
-\newblock XFOIL: An Analysis and Design System for Low Reynolds Number
- Airfoils. In {\em Low Reynolds Number Aerodynamics}; Lecture Notes in Engineering; Springer Verlag: \hl{Berlin, Germany,} 1989;
- Volume~54.
+    Engineers}, 3rd ed.; John~ Wiley \& Sons: Hoboken, NJ, USA, 2009.
+
+\bibitem[Drela(1989)]{Drela1989} Drela, M. \newblock XFOIL: An Analysis and
+Design System for Low Reynolds Number Airfoils. In {\em Low Reynolds Number
+    Aerodynamics}; Lecture Notes in Engineering; Springer Verlag: Berlin, Germany,
+1989; Volume~54.
 

### 2016-01-22 paper.tex
Add remaining bib data

@@
 
-\bibitem[Acheson(1990)]{Acheson1990}
-Acheson, D.
-\newblock {\em Elementary Fluid Dynamics}; Oxford University Press: \hl{City, Country,} 1990.
-
-\bibitem[Baker and Brockie(1991)]{Baker1991}
-Baker, C.; Brockie, N.
-\newblock Wind tunnel tests to obtain train aerodynamic drag coefficients:
- {R}eynolds number and ground simulation effects.
-\newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf
- 1991}, {\em 38},~23--28.
-
-\bibitem[Paraschivoiu(2002)]{Para2002}
-Paraschivoiu, I.
-\newblock {\em Wind Turbine Design with Emphasis on Darrieus Concept}, 1st ed.;
- Polytechnic International: Montreal, QC, Canada, 2002.
-
-\bibitem[Templin(1974)]{Templin1974}
-Templin, R.J.
-\newblock \emph{Aerodynamic Performance Theory for the NRC Vertical Axis Wind
- Turbine};
-\newblock Technical Report LTR-LA-160; \hl {NRC (Please define)} of Canada: \hl{City, Country,} 1974.
-
-\bibitem[Fiedler and Tullis(2009)]{Fiedler2009}
-Fiedler, A.J.; Tullis, S.
+\bibitem[Acheson(1990)]{Acheson1990} Acheson, D. \newblock {\em Elementary Fluid
+    Dynamics}; Oxford University Press: Oxford, UK, 1990.
+
+\bibitem[Baker and Brockie(1991)]{Baker1991} Baker, C.; Brockie, N. \newblock
+Wind tunnel tests to obtain train aerodynamic drag coefficients: {R}eynolds
+number and ground simulation effects. \newblock {\em J. Wind Eng. Ind. Aerodyn.}
+{\bf 1991}, {\em 38},~23--28.
+
+\bibitem[Paraschivoiu(2002)]{Para2002} Paraschivoiu, I. \newblock {\em Wind
+    Turbine Design with Emphasis on Darrieus Concept}, 1st ed.; Polytechnic
+International: Montreal, QC, Canada, 2002.
+
+\bibitem[Templin(1974)]{Templin1974} Templin, R.J. \newblock \emph{Aerodynamic
+    Performance Theory for the NRC Vertical Axis Wind Turbine}; \newblock Technical
+Report LTR-LA-160; National Research Council Canada: Ottawa, Canada 1974.
+
+\bibitem[Fiedler and Tullis(2009)]{Fiedler2009} Fiedler, A.J.; Tullis, S.
 \newblock Blade offset and pitch effects on a high solidity vertical axis wind
- turbine.
-\newblock {\em Wind~Eng.} {\bf 2009}, {\em 33},~237--246.
+turbine. \newblock {\em Wind~Eng.} {\bf 2009}, {\em 33},~237--246.
 
 \bibitem[Migliore \em{et~al.}(1980)Migliore, Wolfe, and Fanucci]{Migliore1980}
-Migliore, P.G.; Wolfe, W.P.; Fanucci, J.B.
-\newblock Flow curvature effects on {d}arrieus turbine blade aerodynamics.
-\newblock {\em J.~Energy} {\bf 1980}, {\em 4},~49--55.
+Migliore, P.G.; Wolfe, W.P.; Fanucci, J.B. \newblock Flow curvature effects on
+{d}arrieus turbine blade aerodynamics. \newblock {\em J.~Energy} {\bf 1980},
+{\em 4},~49--55.
 
 \bibitem[Blackwell \em{et~al.}(1976)Blackwell, Sheldahl, and
- Feltz]{Blackwell1976}
-Blackwell, B.; Sheldahl, R.; Feltz, L.
-\newblock \emph{Wind Tunnel Performance Data for the {Darrieus} Wind Turbine with
- {NACA} 0012 Blades};
-\newblock Report SAND76-0130; Sandia National Laboratories: Albuquerque, NM, USA,
- 1976.
-
-\bibitem[Polagye \em{et~al.}(2013)Polagye, Cavagnaro, and
- Niblick]{Polagye2013b}
-Polagye, B.L.; Cavagnaro, R.J.; Niblick, A.L.
-\newblock Micropower from tidal turbines.
-\newblock In Proceedings of the ASME Fluids Engineering Division Summer Meeting, \hl{City, Country,} \hl{Day Month} 2013.
-
-\bibitem[Bravo \em{et~al.}(2007)Bravo, Tullis, and Ziada]{Bravo2007}
-Bravo, R.; Tullis, S.; Ziada, S.
-\newblock Performance testing of a small vertical-axis wind turbine.
-\newblock In Proceedings of the 21st Canadian Congress of Applied Mechanics
- CANCAM, \hl{City, Country,} \hl{Day Month} 2007.
-
-\bibitem[Bachant and Wosnik(2014)]{Bachant2014}
-Bachant, P.; Wosnik, M.
+Feltz]{Blackwell1976} Blackwell, B.; Sheldahl, R.; Feltz, L. \newblock
+\emph{Wind Tunnel Performance Data for the {Darrieus} Wind Turbine with {NACA}
+    0012 Blades}; \newblock Report SAND76-0130; Sandia National Laboratories:
+Albuquerque, NM, USA, 1976.
+
+\bibitem[Polagye \em{et~al.}(2013)Polagye, Cavagnaro, and Niblick]{Polagye2013b}
+Polagye, B.L.; Cavagnaro, R.J.; Niblick, A.L. \newblock Micropower from tidal
+turbines. \newblock In Proceedings of the ASME Fluids Engineering Division
+Summer Meeting, Incline Village, NV, USA, 7--11 July 2013.
+
+\bibitem[Bravo \em{et~al.}(2007)Bravo, Tullis, and Ziada]{Bravo2007} Bravo, R.;
+Tullis, S.; Ziada, S. \newblock Performance testing of a small vertical-axis
+wind turbine. \newblock In Proceedings of the 21st Canadian Congress of Applied
+Mechanics CANCAM, Toronto, Ontario, Canada, 3--7 June 2007.
+
+\bibitem[Bachant and Wosnik(2014)]{Bachant2014} Bachant, P.; Wosnik, M.
 \newblock Reynolds number dependence of cross-flow turbine performance and
- near-wake characteristics.
-\newblock In Proceedings of the 2nd Marine Energy Technology Symposium METS2014, \hl{City, Country,} \hl{Day Month} 2014.
+near-wake characteristics. \newblock In Proceedings of the 2nd Marine Energy
+Technology Symposium METS2014, Seattle, WA, USA, 15--18 April 2014.
 
 \bibitem[Brochier \em{et~al.}(1986)Brochier, Fraunie, Beguier, and
- Paraschivoiu]{Brochier1986}
-Brochier, G.; Fraunie, P.; Beguier, C.; Paraschivoiu, I.
-\newblock Water channel experiments of dynamic stall on darrieus wind turbine
- blades.
-\newblock {\em AIAA J. Propuls. Power} {\bf 1986}, {\em
- 2},~46--510.
-
-\bibitem[Fujisawa and Shibuya(2001)]{Fujisawa2001}
-Fujisawa, N.; Shibuya, S.
+Paraschivoiu]{Brochier1986} Brochier, G.; Fraunie, P.; Beguier, C.;
+Paraschivoiu, I. \newblock Water channel experiments of dynamic stall on
+darrieus wind turbine blades. \newblock {\em AIAA J. Propuls. Power} {\bf 1986},
+{\em 2},~46--510.
+
+\bibitem[Fujisawa and Shibuya(2001)]{Fujisawa2001} Fujisawa, N.; Shibuya, S.
 \newblock Observations of dynamic stall on Darrieus wind turbine blades.
-\newblock {\em J. Wind Eng. Ind.~Aerodyn.} {\bf
- 2001}, {\em 89},~201--214.
+\newblock {\em J. Wind Eng. Ind.~Aerodyn.} {\bf 2001}, {\em 89},~201--214.
 
 \bibitem[Tescione \em{et~al.}(2014)Tescione, Ragni, He, Ferreira, and
- G.J]{Tescione2014}
-Tescione, G.; Ragni, D.; He, C.; Ferreira, C.S.; G.J.
+G.J]{Tescione2014} Tescione, G.; Ragni, D.; He, C.; Ferreira, C.S.; G.J.
 \newblock Near wake flow analysis of a vertical axis wind turbine by
- stereoscopic particle image velocimetery.
-\newblock {\em Renew. Energy} {\bf 2014}, {\em 70},~47--61.
-
-\bibitem[Krogstad and Adaramola(2012)]{Krogstad2012a}
-Krogstad, P.; Adaramola, M.S.
-\newblock Performance and near wake measurements of a model horizontal axis
- wind turbine.
-\newblock {\em Wind Energy} {\bf 2012}, {\em 15},~743--756.
+stereoscopic particle image velocimetery. \newblock {\em Renew. Energy} {\bf
+    2014}, {\em 70},~47--61.
+
+\bibitem[Krogstad and Adaramola(2012)]{Krogstad2012a} Krogstad, P.; Adaramola,
+M.S. \newblock Performance and near wake measurements of a model horizontal axis
+wind turbine. \newblock {\em Wind Energy} {\bf 2012}, {\em 15},~743--756.
 
 \bibitem[Walker \em{et~al.}(2014)Walker, Flack, Lust, and Luznik]{Walker2014}
-Walker, J.M.; Flack, K.A.; Lust, E.E.; Luznik, M.P.S.L.
-\newblock Experimental and numerical studies of blade roughness and fouling on
- marine current turbine performance.
-\newblock {\em Renew. Energy} {\bf 2014}, {\em 66},~257--267.
-
-\bibitem[McTavish \em{et~al.}(2013)McTavish, Feszty, and
- Nitzsche]{McTavish2013}
-McTavish, S.; Feszty, D.; Nitzsche, F.
-\newblock Evaluating {R}eynolds number effects in small-scale wind turbine
- experiments.
-\newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf 2013}, {\em 120},~81--90.
+Walker, J.M.; Flack, K.A.; Lust, E.E.; Luznik, M.P.S.L. \newblock Experimental
+and numerical studies of blade roughness and fouling on marine current turbine
+performance. \newblock {\em Renew. Energy} {\bf 2014}, {\em 66},~257--267.
+
+\bibitem[McTavish \em{et~al.}(2013)McTavish, Feszty, and Nitzsche]{McTavish2013}
+McTavish, S.; Feszty, D.; Nitzsche, F. \newblock Evaluating {R}eynolds number
+effects in small-scale wind turbine experiments. \newblock {\em J. Wind Eng.
+    Ind. Aerodyn.} {\bf 2013}, {\em 120},~81--90.
 
 \bibitem[Chamorro \em{et~al.}(2011)Chamorro, Arndt, and
- Sotiropoulos]{Chamorro2011}
-Chamorro, L.P.; Arndt, R.E.A.; Sotiropoulos, F.
-\newblock Turbulent flow properties around a staggered wind farm.
-\newblock {\em Bound. Layer Meteorol.} {\bf 2011}, \emph{141}, 349--367.
-
-\bibitem[Chamorro and Porte-Agel(2011)]{Chamorro2011b}
-Chamorro, L.P.; Porte-Agel, F.
-\newblock Turbulent flow inside and above a wind farm: A wind-tunnel study.
-\newblock {\em Energies} {\bf 2011}, {\em 4}, 1916--1936.
-
-\bibitem[Goldenberg and Fekete(1983)]{Goldenberg1983}
-Goldenberg, J.; Fekete, G.
-\newblock Mean flow energy content of boundary layer downstream of
- vertical-axis wind-turbine simulators.
-\newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf
- 1983}, {\em 12},~1--14.
-
-\bibitem[Dabiri(2011)]{Dabiri2011}
-Dabiri, J.
-\newblock Potential order-of-magnitude enhancement of wind farm power density
- via counter-rotating vertical-axis wind turbine arrays.
-\newblock {\em J. Renew. Sustain. Energy} {\bf 2011}, {\em
- 3},~1--13.
+Sotiropoulos]{Chamorro2011} Chamorro, L.P.; Arndt, R.E.A.; Sotiropoulos, F.
+\newblock Turbulent flow properties around a staggered wind farm. \newblock {\em
+    Bound. Layer Meteorol.} {\bf 2011}, \emph{141}, 349--367.
+
+\bibitem[Chamorro and Porte-Agel(2011)]{Chamorro2011b} Chamorro, L.P.;
+Porte-Agel, F. \newblock Turbulent flow inside and above a wind farm: A
+wind-tunnel study. \newblock {\em Energies} {\bf 2011}, {\em 4}, 1916--1936.
+
+\bibitem[Goldenberg and Fekete(1983)]{Goldenberg1983} Goldenberg, J.; Fekete, G.
+\newblock Mean flow energy content of boundary layer downstream of vertical-axis
+wind-turbine simulators. \newblock {\em J. Wind Eng. Ind. Aerodyn.} {\bf 1983},
+{\em 12},~1--14.
+
+\bibitem[Dabiri(2011)]{Dabiri2011} Dabiri, J. \newblock Potential
+order-of-magnitude enhancement of wind farm power density via counter-rotating
+vertical-axis wind turbine arrays. \newblock {\em J. Renew. Sustain. Energy}
+{\bf 2011}, {\em 3},~1--13.
 
 \bibitem[Kinzel \em{et~al.}(2012)Kinzel, Mulligan, and Dabiri]{Kinzel2012}
-Kinzel, M.; Mulligan, Q.; Dabiri, J.O.
-\newblock Energy exchange in an array of vertical-axis wind turbines.
-\newblock {\em J. Turbul.} {\bf 2012}, {\em 13},~1--13.
-
-\bibitem[Bachant and Wosnik(2015)]{Bachant2015-JoT}
-Bachant, P.; Wosnik, M.
-\newblock Characterising the near-wake of a cross-flow turbine.
-\newblock {\em J. Turbul.} {\bf 2015}, {\em 16},~392--410.
-
-\bibitem[Jacobs and Sherman(1937)]{Jacobs1937}
-Jacobs, E.N.; Sherman, A.
+Kinzel, M.; Mulligan, Q.; Dabiri, J.O. \newblock Energy exchange in an array of
+vertical-axis wind turbines. \newblock {\em J. Turbul.} {\bf 2012}, {\em
+    13},~1--13.
+
+\bibitem[Bachant and Wosnik(2015)]{Bachant2015-JoT} Bachant, P.; Wosnik, M.
+\newblock Characterising the near-wake of a cross-flow turbine. \newblock {\em
+    J. Turbul.} {\bf 2015}, {\em 16},~392--410.
+
+\bibitem[Jacobs and Sherman(1937)]{Jacobs1937} Jacobs, E.N.; Sherman, A.
 \newblock \emph{Airfoil Section Characteristics as Affected by Variation of the
- Reynolds Number};
-\newblock Technical Report; National Advisory Committee for Aeronautics: \hl{City, Country}, 1937.
-
-\bibitem[Lissaman(1983)]{Lissaman1983}
-Lissaman, P.B.S.
-\newblock Low-{R}eynolds-number airfoils.
-\newblock {\em Ann. Rev. Fluid Mech.} {\bf 1983}, {\em
- 15},~223--239.
-
-\bibitem[McMasters and Henderson(1980)]{McMasters1980}
-McMasters, J.H.; Henderson, M.L.
-\newblock Low-speed single-element airfoil synthesis.
-\newblock {\em Tech. Soar.} {\bf 1980}, {\em 6},~1--21.
+    Reynolds Number}; \newblock Technical Report; National Advisory Committee for
+Aeronautics: Langley Field, VA, United States, 1937.
+
+\bibitem[Lissaman(1983)]{Lissaman1983} Lissaman, P.B.S. \newblock
+Low-{R}eynolds-number airfoils. \newblock {\em Ann. Rev. Fluid Mech.} {\bf
+    1983}, {\em 15},~223--239.
+
+\bibitem[McMasters and Henderson(1980)]{McMasters1980} McMasters, J.H.;
+Henderson, M.L. \newblock Low-speed single-element airfoil synthesis. \newblock
+{\em Tech. Soar.} {\bf 1980}, {\em 6},~1--21.
 
 \bibitem[Bedon \em{et~al.}(2014)Bedon, Antonini, Betta, Castelli, and
- Benini]{Bedon2014}
-Bedon, G.; Antonini, E.G.; Betta, S.D.; Castelli, M.R.; Benini, E.
-\newblock Evaluation of the different aerodynamic databases for vertical axis
- wind turbine simulations.
-\newblock {\em Renew. Sustain. Energy Rev.} {\bf 2014}, {\em 40}, 386--399.
-
-\bibitem[Bousman(2000)]{Bousman2000-evaluation}
-Bousman, W.G.
-\newblock Evaluation of airfoil dynamic stall characteristics for
- maneuverability.
-\newblock In Proceedings of the 26th {E}uropean Rotorcraft Forum, \hl{City, Country,} \hl{Day Month} 2000.
-
-\bibitem[Singleton and Yeager(2000)]{Singleton2000}
-Singleton, J.D.; Yeager, W.T., Jr.
-\newblock Important scaling parameters for testing model-scale helicopter
- rotors.
-\newblock {\em J.~Aircr.} {\bf 2000}, {\em 37},~396--402.
+Benini]{Bedon2014} Bedon, G.; Antonini, E.G.; Betta, S.D.; Castelli, M.R.;
+Benini, E. \newblock Evaluation of the different aerodynamic databases for
+vertical axis wind turbine simulations. \newblock {\em Renew. Sustain. Energy
+    Rev.} {\bf 2014}, {\em 40}, 386--399.
+
+\bibitem[Bousman(2000)]{Bousman2000-evaluation} Bousman, W.G. \newblock
+Evaluation of airfoil dynamic stall characteristics for maneuverability.
+\newblock In Proceedings of the 26th {E}uropean Rotorcraft Forum, The Hague,
+Netherlands, 26--29 September 2000.
+
+\bibitem[Singleton and Yeager(2000)]{Singleton2000} Singleton, J.D.; Yeager,
+W.T., Jr. \newblock Important scaling parameters for testing model-scale
+helicopter rotors. \newblock {\em J.~Aircr.} {\bf 2000}, {\em 37},~396--402.
 
 \bibitem[Yoon \em{et~al.}(2005)Yoon, Hill, Balachandar, Adrian, and
- Ha]{Yoon2005}
-Yoon, H.; Hill, D.; Balachandar, S.; Adrian, R.; Ha, M.
-\newblock Reynolds number scaling of flow in a Rushton turbine stirred tank.
- Part I—Mean flow, circular jet and tip vortex scaling.
-\newblock {\em Chem. Eng. Sci.} {\bf 2005}, {\em 60},~3169--3183.
+Ha]{Yoon2005} Yoon, H.; Hill, D.; Balachandar, S.; Adrian, R.; Ha, M. \newblock
+Reynolds number scaling of flow in a Rushton turbine stirred tank. Part I—Mean
+flow, circular jet and tip vortex scaling. \newblock {\em Chem. Eng. Sci.} {\bf
+    2005}, {\em 60},~3169--3183.
 
 \bibitem[Chamorro \em{et~al.}(2012)Chamorro, Arndt, and
- Sotiropolous]{Chamorro2012}
-Chamorro, L.P.; Arndt, R.E.A.; Sotiropolous, F.
+Sotiropolous]{Chamorro2012} Chamorro, L.P.; Arndt, R.E.A.; Sotiropolous, F.
 \newblock Reynolds number dependence of turbulence statistics in the wake of
- wind turbines.
-\newblock {\em Wind Energy} {\bf 2012}, {\em 15},~733--742.
+wind turbines. \newblock {\em Wind Energy} {\bf 2012}, {\em 15},~733--742.
 
 \bibitem[Neary \em{et~al.}(2014)Neary, Previsic, Jepsen, Lawson, Yu, Copping,
- Fontaine, Hallett, and Murray]{Neary2014}
-Neary, V.S.; Previsic, M.; Jepsen, R.A.; Lawson, M.J.; Yu, Y.H.; Copping, A.E.;
- Fontaine, A.A.; Hallett, K.C.; Murray, D.K.
-\newblock \emph{Methodology for Design and Economic Analysis of Marine Energy
- Conversion ({MEC}) Technologies};
-\newblock Technical Report SAND2014-9040; Sandia National Laboratories: \hl{City, Country}, 2014.
-
-\bibitem[Cavagnaro and Polagye(2014)]{Cavagnaro2014}
-Cavagnaro, R.; Polagye, B.
+Fontaine, Hallett, and Murray]{Neary2014} Neary, V.S.; Previsic, M.; Jepsen,
+R.A.; Lawson, M.J.; Yu, Y.H.; Copping, A.E.; Fontaine, A.A.; Hallett, K.C.;
+Murray, D.K. \newblock \emph{Methodology for Design and Economic Analysis of
+    Marine Energy Conversion ({MEC}) Technologies}; \newblock Technical Report
+SAND2014-9040; Sandia National Laboratories: Albuquerque, NM, USA, 2014.
+
+\bibitem[Cavagnaro and Polagye(2014)]{Cavagnaro2014} Cavagnaro, R.; Polagye, B.
 \newblock An evaluation of blockage corrections for a helical cross-flow
- turbine.
-\newblock In~Proceedings of the 3rd Oxford Tidal Energy Workshop, \hl{City, Country,} \hl{Day Month} 2014.
+turbine. \newblock In~Proceedings of the 3rd Oxford Tidal Energy Workshop,
+Oxford, UK, 7--8 April 2014.
 
@@
 \url{http://dx.doi.org/10.6084/m9.figshare.1286960} (accessed on 22 January
-2015).
+2016).
 

### 2016-01-22 scripts/zip.py
Ignore more files in zip

@@ -11,9 +11,11 @@ import os
 version = check_output(["git", "describe"]).strip().decode()
 
 figures = ["figures/" + f for f in os.listdir("figures")
-           if f not in ["SolidWorks", "McMasters-Henderson-1980"]]
+           if f not in ["SolidWorks",
+                        "McMasters-Henderson-1980",
+                        "McMasters-Henderson-1980.PNG"]]
 
-files = ["paper.tex", "library.bib", "mdpi.bst", "mdpi.cls"]
+files = ["paper.tex", "mdpi.bst", "mdpi.cls"]
 
 with ZipFile("archive/paper-{}.zip".format(version), "w") as f:
     for fig in figures:

### 2016-01-22 scripts/zip.py
Docstring formatting

@@ -1,7 +1,5 @@
 #!/usr/bin/env python
-"""
-This script zips up all the necessary source files to compile the paper.
-"""
+"""This script zips up all the necessary source files to compile the paper."""
 
 from zipfile import ZipFile
 from subprocess import check_output

### 2016-01-24 paper.tex
Wrapping
Wosnik edit to abstract

@@
     the performance of the cross-flow turbine becomes essentially $Re$-independent
- at a Reynolds number based on a turbine diameter of $Re_D \approx 10^6$ or an
- approximate average Reynolds number based on a blade chord of $Re_c \approx 2
- \times 10^5$.~A simple model that calculates the peak torque coefficient from static
- foil data and cross-flow turbine kinematics was shown to be a reasonable
- predictor for Reynolds number dependence of an actual cross-flow
+    at a Reynolds number based on turbine diameter of $Re_D \approx 10^6$ or an
+    approximate average Reynolds number based on a blade chord length of $Re_c
+    \approx 2 \times 10^5$.~A simple model that calculates the peak torque
+    coefficient from static foil data and cross-flow turbine kinematics was shown to
+    be a reasonable predictor for Reynolds number dependence of an actual cross-flow
     turbine operating under dynamic conditions. Mean velocity and turbulence

### 2016-01-24 paper.tex
Semicolon to comma

@@
 pitch, number of blades, strut drag, operational parameters, such as tip speed
-ratio; and on the Reynolds number \cite{Para2002}. Note that an average blade
+ratio, and on the Reynolds number \cite{Para2002}. Note that an average blade
 chord Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$,

### 2016-01-24 paper.tex
CFTs are often vertical-axis

@@
 
-The performance and wake characteristics of cross-flow (vertical-axis) turbines
-(CFTs) depend on turbine solidity, blade profile (lift/drag, dynamic stall at
-reduced frequency of turbine rotation, symmetry, thickness, camber), blade
-pitch, number of blades, strut drag, operational parameters, such as tip speed
-ratio, and on the Reynolds number \cite{Para2002}. Note that an average blade
-chord Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$,
-can be expressed in terms of tip speed ratio $\lambda = \omega R/ U_\infty$,
-where $U_\infty$ is the free stream velocity, $c$ is the blade chord length,
-$\nu$ is the fluid kinematic viscosity, $\omega$ is the rotor's angular
+The performance and wake characteristics of cross-flow (often vertical-axis)
+turbines (CFTs) depend on turbine solidity, blade profile (lift/drag, dynamic
+stall at reduced frequency of turbine rotation, symmetry, thickness, camber),
+blade pitch, number of blades, strut drag, operational parameters, such as tip
+speed ratio, and on the Reynolds number \cite{Para2002}. Note that an average
+blade chord Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/
+\nu$, can be expressed in terms of tip speed ratio $\lambda = \omega R/
+U_\infty$, where $U_\infty$ is the free stream velocity, $c$ is the blade chord
+length, $\nu$ is the fluid kinematic viscosity, $\omega$ is the rotor's angular
 velocity, and $R$ is the rotor radius. The value of $\lambda$ at which a turbine

### 2016-01-24 paper.tex
Transport looks at streamwise momentum
Add info to transport total fig caption

@@
 
- \caption{Normalized momentum transport quantities computed as weighted
- averages from Equation~(\ref{eq:momentum}).}
+ \caption{Normalized streamwise momentum transport quantities computed as
+     weighted averages from Equation~(\ref{eq:momentum}).}
 
@@
     \caption{Normalized transport totals from Figures~\ref{fig:mom-bar-graph}
-        and \ref{fig:K-bar-graph} plotted \textit{versus} the Reynolds number.}
+        (streamwise momentum, $U$) and \ref{fig:K-bar-graph} (mean kinetic energy,
+        $K$) plotted \textit{versus} the Reynolds number.}
 

### 2016-01-24 paper.tex
Clarify abstract wording a little better

@@
     the performance of the cross-flow turbine becomes essentially $Re$-independent
-    at a Reynolds number based on turbine diameter of $Re_D \approx 10^6$ or an
-    approximate average Reynolds number based on a blade chord length of $Re_c
+    at a Reynolds number based on the rotor diameter $Re_D \approx 10^6$ or an
+    approximate average Reynolds number based on the blade chord length $Re_c
     \approx 2 \times 10^5$.~A simple model that calculates the peak torque

### 2016-01-24 scripts/zip.py
Add images to zip

@@ -13,7 +13,8 @@ figures = ["figures/" + f for f in os.listdir("figures")
                         "McMasters-Henderson-1980",
                         "McMasters-Henderson-1980.PNG"]]
 
-files = ["paper.tex", "mdpi.bst", "mdpi.cls"]
+files = ["paper.tex", "mdpi.bst", "mdpi.cls", "logo-ccby.png", "logo-mdpi.png",
+         "energies-logo.png"]
 
 with ZipFile("archive/paper-{}.zip".format(version), "w") as f:
     for fig in figures:

### 2016-01-25 paper.tex
Modifications from editor, round 2

@@
 \corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.);
-    martin.wosnik@unh.edu (M.W.); \linebreak Tel.: +1-603-862-1891 (M.W.)}
+    martin.wosnik@unh.edu (M.W.); \linebreak Tel.: +1-603-862-1891 \hl {(M.W.)}}
+    %please also provide the author (P.B.)'s office number. and please also pay attention to the keywords section, can we define VAWT as vertical-axis wind turbine?
 
@@
 \keyword{\textls[-10]{Reynolds number; cross-flow turbine; turbine performance; marine
-hydrokinetic energy}; wind energy; {VAWT}; scale model}
+hydrokinetic energy}; wind energy; {vertical-axis wind turbine (VAWT)}; scale model}
 
@@
 ``scale'' to refer to this dynamical scale or the Reynolds number and assume the
-Froude number is small enough to be negligible.
+Froude number is small enough to be~ negligible.
 
@@
 solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
-observed the power coefficient of a straight-bladed VAWT to become Reynolds
+observed the power coefficient of a straight-bladed \hl {vertical-axis wind turbine (VAWT)} to become Reynolds
 number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{figures/exp-setup-photo}
+        \includegraphics[width=0.6\textwidth]{figures/exp-setup-photo}
         \caption{}
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{figures/exp_setup_drawing}
+        \includegraphics[width=0.6\textwidth]{figures/exp_setup_drawing}
         \caption{}
@@
     \caption{Experimental setup photo (a) and drawing (b): turbine test bed
-        installed in the UNH tow tank.}
+        installed in the UNH tow~ tank.}
 
@@
 
-\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}
+\includegraphics[width=0.8\textwidth]{figures/turbine_coordinate_system}
 
@@
 \includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
-
-\caption{UNH-RVAT measured mean power \textbf{(a)} and drag \textbf{(b)}
+%Please use scientific notation in the figure. for instance, please revise 1e6 to \times 10^6
+\caption{\hl {UNH-RVAT} measured mean power (\textbf{a}) and drag (\textbf{b})
     coefficients at $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error
@@
 increments of 0.5$^{\circ}$. Solver parameters used were 100 panels, a fixed
-speed, zero Mach number, \texttt{NCrit = 9} (default $\mathrm{e}^n$ transition
+speed, zero Mach number, \linebreak \texttt{NCrit = 9} (default $\mathrm{e}^n$ transition
 criteria parameter for an average wind tunnel) and no forced boundary layer
@@
     \includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
-
-    \caption{Normalized maximum lift coefficient \textbf{(a)}, drag coefficient
-        \textbf{(b)}, and lift-to-drag ratio \textbf{(c)} computed by XFOIL at
+ % Please use scientific notation in the figure. for instance, please revise 1e5 to \times 10^5 
+    \caption{\hl {Normalized maximum} lift coefficient (\textbf{a}), drag coefficient
+        (\textbf{b}), and lift-to-drag ratio (\textbf{c}) computed by XFOIL at
         various $Re_c$  for each profile.}
@@
 
-    \caption{Geometric angle of attack \textbf{(a)}, relative velocity
-        \textbf{(b)},  and torque coefficient \textbf{(c)} calculated with a NACA
+    \caption{Geometric angle of attack (\textbf{a}), relative velocity
+        (\textbf{b}),  and torque coefficient (\textbf{c}) calculated with a NACA
         0020 foil operating at $\lambda=1.9$.}
@@
     \includegraphics[width=0.8\textwidth]{figures/cft_re_dep_foils}
-
-    \caption{Reynolds number dependence of the normalized peak torque
+%Please use scientific notation in the figure. for instance, please revise 1e6 to \times 10^6
+    \caption{\hl {Reynolds} number dependence of the normalized peak torque
         coefficient calculated from static foil coefficients and blade kinematics,
@@
 
-     \caption{Mean streamwise velocity \textbf{(a)} and turbulence kinetic
-         energy \textbf{(b)} profiles at $z/H=0.0$. Turbine diameter Reynolds number
+     \caption{Mean streamwise velocity (\textbf{a}) and turbulence kinetic
+         energy (\textbf{b}) profiles at $z/H=0.0$. Turbine diameter Reynolds number
          $Re_D$ is indicated by the legend.}
@@
     Program Endowment to purchase acoustic flow measurement instrumentation, and a
-    grant for laboratory infrastructure upgrades through the U.S. Department of
+    grant for laboratory infrastructure upgrades through the U.S. Department of~
     Energy.}

### 2016-01-25 paper.tex
Martin is only corresponding author
Approve definition of VAWT

@@
 % Authors (Add full first names)
-\Author{Peter Bachant * and Martin Wosnik *}
+\Author{Peter Bachant and Martin Wosnik *}
 
@@
 % there are more than one corresponding author.)
-\corres[2]{\hspace{-1em} Correspondence: pxL3@unh.edu (P.B.);
-    martin.wosnik@unh.edu (M.W.); \linebreak Tel.: +1-603-862-1891 \hl {(M.W.)}}
-    %please also provide the author (P.B.)'s office number. and please also pay attention to the keywords section, can we define VAWT as vertical-axis wind turbine?
+\corres[1]{\hspace{-1em} Correspondence: martin.wosnik@unh.edu;
+    \linebreak Tel.: +1-603-862-1891}
 
@@
 solidity helical cross-flow turbine, and Bravo \emph{et al.} \cite{Bravo2007}
-observed the power coefficient of a straight-bladed \hl {vertical-axis wind turbine (VAWT)} to become Reynolds
-number independent at $Re_c \approx 4 \times 10^5$. Bachant and Wosnik
-\cite{Bachant2014} observed Reynolds number independence of the power
-coefficient at the optimal tip speed ratio for a high solidity cross-flow
-turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2 \times 10^5$.
+observed the power coefficient of a straight-bladed vertical-axis wind turbine
+(VAWT) to become Reynolds number independent at $Re_c \approx 4 \times 10^5$.
+Bachant and Wosnik \cite{Bachant2014} observed Reynolds number independence of
+the power coefficient at the optimal tip speed ratio for a high solidity
+cross-flow turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2
+\times 10^5$.
 

### 2016-01-25 scripts/zip.py
Add graphical abstract to zip

@@ -14,7 +14,7 @@ figures = ["figures/" + f for f in os.listdir("figures")
                         "McMasters-Henderson-1980.PNG"]]
 
 files = ["paper.tex", "mdpi.bst", "mdpi.cls", "logo-ccby.png", "logo-mdpi.png",
-         "energies-logo.png"]
+         "energies-logo.png", "graphical-abstract.png"]
 
 with ZipFile("archive/paper-{}.zip".format(version), "w") as f:
     for fig in figures:

### 2016-01-25 scripts/zip.py
Do not zip up summary dir

@@ -11,7 +11,8 @@ version = check_output(["git", "describe"]).strip().decode()
 figures = ["figures/" + f for f in os.listdir("figures")
            if f not in ["SolidWorks",
                         "McMasters-Henderson-1980",
-                        "McMasters-Henderson-1980.PNG"]]
+                        "McMasters-Henderson-1980.PNG",
+                        "summary"]]
 
 files = ["paper.tex", "mdpi.bst", "mdpi.cls", "logo-ccby.png", "logo-mdpi.png",
          "energies-logo.png", "graphical-abstract.png"]

### 2016-01-25 paper.tex
Remove comment
No new line after correspondence email address
Remove highlights

@@
 % there are more than one corresponding author.)
-\corres[1]{\hspace{-1em} Correspondence: martin.wosnik@unh.edu;
-    \linebreak Tel.: +1-603-862-1891}
+\corres[1]{\hspace{-1em} Correspondence: martin.wosnik@unh.edu; Tel.:
+    +1-603-862-1891}
 
@@
 \includegraphics[width=0.9\textwidth]{figures/perf_re_dep}
-%Please use scientific notation in the figure. for instance, please revise 1e6 to \times 10^6
-\caption{\hl {UNH-RVAT} measured mean power (\textbf{a}) and drag (\textbf{b})
+
+\caption{UNH-RVAT measured mean power (\textbf{a}) and drag (\textbf{b})
     coefficients at $\lambda=1.9$ plotted \textit{versus} the Reynolds number. Error
@@
     \includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
- % Please use scientific notation in the figure. for instance, please revise 1e5 to \times 10^5 
-    \caption{\hl {Normalized maximum} lift coefficient (\textbf{a}), drag coefficient
+
+    \caption{Normalized maximum lift coefficient (\textbf{a}), drag coefficient
         (\textbf{b}), and lift-to-drag ratio (\textbf{c}) computed by XFOIL at
@@
     \includegraphics[width=0.8\textwidth]{figures/cft_re_dep_foils}
-%Please use scientific notation in the figure. for instance, please revise 1e6 to \times 10^6
-    \caption{\hl {Reynolds} number dependence of the normalized peak torque
+
+    \caption{Reynolds number dependence of the normalized peak torque
         coefficient calculated from static foil coefficients and blade kinematics,
